### 2014-09-18 2014-06-04-paper_v4-ianbefore.tex
initial commit

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{jheppub}

\newcommand{\IGNORE}[1]{}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bea}{\begin{aligned}}
\newcommand{\eea}{\end{aligned}}

\title{Gibbons-Hawking-York Boundary Term in Causal Sets}
 \author[a]{M. Buck}
 \author[a,b]{\!, F. Dowker}
 \author[a]{and I. Jubb\,}
\affiliation[a]{Theoretical Physics Group, Blackett Laboratory, Imperial College, London, SW7 2AZ, U.K.}
\affiliation[b]{Institute for Quantum Computing, University of Waterloo, ON, N2L 2Y5, Canada}

\abstract{ 
We show some stuff.
}

\begin{document}

\maketitle

\noindent NOTES:\\
1. define s as the RV, not as its mean

\section{Introduction}

%[MB1: USEFUL QUOTE FROM RAFAEL: In the quantum theory, however, these boundary terms are important. They are essential in order that the quantum mechanical amplitudes satisfy the correct composition law and in order that these amplitudes have the correct classical limit. In a number of examples [5] they give a contribution to the partition function which is important for the agreement with calculations based on straightforward thermo- dynamics. In this note we shall derive the boundary terms in the action for the Regge calculus]

In furthering causal set theory it is crucial that we understand the kinematics of the theory. The action of a given causal set is a crucial piece of the kinematics that would be extremely useful to know and understand. Proposals for the action of a causal set are available \cite{Benincasa_Dowker:The_Scalar_Curvature_of_a_Causal_Set} and these hold analytically in some cases and numerically in many more. These cases being when the causal set is embeddable in some existing spacetime. One can show that the action of the causal set then agrees with the Einstein-Hilbert (EH) action of the spacetime in some continuum limit. How this limiting procedure is carried out will be described in more detail below.

The EH action, however, is not the complete story in the continuum. If one has a spacetime manifold with a boundary, such as Minkowski space but cut off at some finite time, then one needs to include the Gibbons-Hawking-York boundary term (GHYBT) in the EH action \cite{Gibbons_Hawking_Boundary}. The boundary of the manifold will be some hypersurface of codimension-1. The GHYBT is an integral over this hypersurface, of the extrinsic curvature that it has. This term is needed in the action to keep the variational principle well defined. In this paper we turn our attention to this term, and provide the causal set analogue for spacelike hypersurfaces.

We define a \textit{causal set} (or \textit{causet}) as a locally finite partial order. This means it is a pair $(\mathcal{C},\preceq)$ where $\mathcal{C}$ is the set of points and $\preceq$ is a partial order relation on $\mathcal{C}$ that has the following properties. It is (i) reflexive: $x\preceq x$, (ii) acyclic: $x\preceq y\preceq x \Rightarrow x=y$, and (iii) transitive: $x\preceq y\preceq z \Rightarrow x\preceq z$ for all points $x, y, z \in \mathcal{C}$. We define an inclusive order interval as the set $I(x,y):=\lbrace z\in\mathcal{C}|x\preceq z\preceq y\rbrace$ for any $x, y\in\mathcal{C}$. The \textit{locally finite} condition is simply that the cardinality of any order interval is finite, that is $|I(x,y)|<\infty$. This condition ensures we are dealing with a discrete structure, as all the other properties of the relation would apply to an order relation between points on a continuous manifold.

In order to say we have a causal set analogue for a continuous expression we need a well defined procedure for relating the discrete theory to the continuum. This procedure, or tool, is called a \textit{Poisson sprinkling}, or just a \textit{sprinkling}. It is a Poisson process which provides a way to generate a causet from a $d$-dimensional Lorentzian manifold $(\mathcal{M},g)$ by selecting points in $\mathcal{M}$ to be the elements of $\mathcal{C}$, with an order relation given by the causal order of the manifold. The number of points chosen in a region of spacetime volume, $V$, is a Poisson random variable. This means that the expected number of points in some region will be $\rho V$, where $\rho$ is the density of the sprinkling. The density is related to the discreteness scale, $l$, by $\rho=l^{-d}$ in $d$ spacetime dimensions. It is called a sprinkling as one can envisage the point selection process as a 'sprinkling' of points into the manifold. If a causet, $\mathcal{C}$, can be generated with relatively high probability by a sprinkling into the manifold, $(\mathcal{M},g)$, then we say that the manifold is a good approximation for the causal set.

\section{The Claim}

Consider a sufficiently well-behaved $d$-dimensional spacetime $(M,g)$ and Cauchy surface $\Sigma$ in $M$. The causal past and future sets $M^\pm=J^\pm(\Sigma)$ form a partition of $M$ and $\partial M^\pm = \Sigma$. The Gibbons-Hawking-York boundary term for $M^\pm$ is in this case given by
\be\label{eq:GHYBT_in_continuum}
S_{GHY} = \pm \frac{1}{l_p^{d-2}}\int_{\Sigma} d^{d-1}x \sqrt{h} K
\ee
where $l_p$ is the rationalised Planck length and $K$ is the trace of the extrinsic curvature $K_{\mu\nu}=h_{\mu}^\rho h_\nu^\sigma \nabla_\rho n_\sigma$ of $\Sigma$ defined with future-pointing timelike unit normal $n_{\mu}=\partial_\mu S/\sqrt{g^{\mu\nu}\partial_\mu S\partial_\nu S}$.

The extrinsic curvature of a spacelike hypersurface can be identified with the volume gradient across that hypersurface [MB1: HOW SO?]. On the causal set, the analogue of a spacelike hypersurface is a maximal anti-chain, i.e. a subset $\mathcal{A}\subseteq\mathcal{C}$ in which all elements are unrelated. These unrelated elements are maximal in the sense that there exist no elements $y\in\mathcal{C}\setminus\mathcal{A}$ such that $x\preceq y$, $\forall x\in\mathcal{A}$. The intuitive analogue of the boundary term would then be the rate of change of the number of causal set elements below and above the antichain. We shall see that this intuitive idea indeed bears out.

Consider a causal set $\mathcal C$ obtained by a Poisson sprinkling at density $\rho=1/l^d$ into a $d$-dimensional spacetime $(M,g)$. A Cauchy surface $\Sigma\subset M$ induces a partition $\mathcal C = \mathcal C^+ \cup\, \mathcal C^-$ of the sprinkling, where $\mathcal C^\pm$ denotes the restriction of $\mathcal C$ to the points sprinkled to the causal future/past of $\Sigma$. Let us denote the number of maximal elements in $\mathcal C^-$ and the number of minimal elements in $\mathcal C^+$ by $N_{max}$ and $N_{min}$, respectively. We propose the following definition for the discrete Gibbons-Hawking-York boundary term:
\be\label{GH_boundary_to_causet}
S^{(d)}_{GHY}[\mathcal C]=\left(l/l_p\right)^{d-2}c_{d}(N_{max}-N_{min})
\ee
where the constant $c_{d}$ only depends on the spacetime dimension and is given by
\be\label{Cn}
c_{d}=\frac{d(d+1)}{2(d+2)\Gamma\left(\frac{2}{d}\right)}\left[\frac{A_{d-2}}{d(d-1)}\right]^{\frac{2}{d}},
\ee
$A_d$ denoting the volume of the unit $d$-sphere. To support this proposal we show that in the limit of infinite sprinkling density into a $d$-dimensional spacetime we obtain
\be
\lim_{\rho\rightarrow\infty}\left\langle S^{(d)}_{GHY}[\mathcal C] \right\rangle= \frac{1}{l_p^{d-2}}\int_{\Sigma} d^{d-1}x \sqrt{h} K\label{eq:mainconjecture}
\ee
where $\left\langle\cdot\right\rangle$ denotes the mean over sprinklings. The limit $\rho\rightarrow\infty$ is equivalent to setting the discreteness scale to zero, $l\rightarrow 0$. Figure \ref{fig:Nmin_Nmax} shows an illustration of the idea behind this conjecture. The minimal and maximal elements have been highlighted and the shaded regions correspond to the regions whose volumes expressions will be needed in the proof.
%MB1: (too vague) During the calculation approximations will be made. These approximations are justified on the grounds that if certain orders are retained throughout the calculation they will in fact vanish in the limit $\rho \rightarrow \infty$
\begin{figure}
  \centering
    \includegraphics[width=\textwidth]{minmaxplot}
     \caption{An illustration of the idea. Pictured here is a causal set obtained by a sprinkling into a spacetime with a partitioning spacelike hypersurface. The minimal and maximal points about the surface are highlighted and the shaded regions illustrate the regions whose volumes are needed for the calculation.}
     \label{fig:Nmin_Nmax}
\end{figure}

\section{The Proof}

\subsection{Poisson Sprinklings for $\langle N_{max}\rangle$ and $\langle N_{min}\rangle$}

In order to prove~\eqref{eq:mainconjecture} let us choose a set of synchronous or Gaussian Normal Coordinates (GNC) adapted to $\Sigma$ such that in a neighbourhood $U_\Sigma$ of $\Sigma$ the line element is
\be
ds^2 = -dt^2 + h_{ij}(t,\mathbf x) dx^i dx^j.
\ee
In these coordinates, the surface $\Sigma$ corresponds to $t=0$, and the coordinate $t$ measures the proper time elapsed along geodesics whose tangent vector is proportional to the surface normal on $\Sigma$.
%\be\label{GNC_metric}
%g_{\mu\nu}(x)=
%\begin{pmatrix}
 %-1&0 \\
% 0&h_{ij}(x)
%\end{pmatrix}.
%\ee

%In order to find $N_{max}$, the number of maximal points below the surface, one has to use the fact that the points have been sprinkled with a Poisson distribution. 
For a sprinkling into $(M,g)$, the probability that a sprinkled point $x$ below the surface is \emph{maximal} is given by the probability that the sprinkling contains no points in the intersection $J^{+}(x)\cap J^{-}(\Sigma)$ of the causal future of $p$ with the causal past of the surface $\Sigma$. This region will in general be some sort of curvy $d$-dimensional cone. The Poisson distribution assigns a probability
\be\label{Poisson}
\mathbb P\left(\text{no points in }J^{+}(x)\cap J^{-}(\Sigma)\right)=e^{-\rho V(x,\Sigma)}
\ee
to this event, where $V(x,\Sigma):=V(J^{+}(x)\cap J^{-}(\Sigma))$ is the spacetime volume of the region $J^{+}(x)\bigcap J^{-}(\Sigma)$. The probability of sprinkling an element into an infinitesimal volume element $d^dx$ at $x\in M$ is $\rho\sqrt{-g(x)}d^dx$, where $\rho$ is the density of the sprinkling, and so the total expected number of maximal elements below $\Sigma$ is

\be\label{eq:nmax}
\left\langle N_{max}\right\rangle =\rho\int_{J^{-}(\Sigma)}d^{d}x\:\sqrt{-g}\\ e^{-\rho V(x,\Sigma)}
\ee

Similarly the expected number of minimal elements above $\Sigma$ is
\be\label{eq:nmin}
\left\langle N_{min}\right\rangle =\rho\int_{J^{+}(\Sigma)}d^{n}x\:\sqrt{-g}\ \ e^{-\rho V(\Sigma,x)}
\ee
where $V(\Sigma,x):=V(J^{+}(\Sigma)\cap J^{-}(x))$.

In the limit $\rho\rightarrow\infty$, both quantities will diverge, but if their difference $\langle N_{max}\rangle - \langle N_{min}\rangle = \langle N_{max} - N_{min}\rangle$ grows slower than or at order $\rho^{1-\frac2d}$, the proposed action~\eqref{GH_boundary_to_causet} will tend to a finite value in the continuum limit.

If we pick a point $x_0=(0,\mathbf{x})\in \Sigma$ on the surface that has the same spatial coordinates as $x=(t,\mathbf{x})$ then the coordinate time, $t$, is the proper time elapsed along the unique geodesic between $x_0$ and $x$. The volume $V(\Sigma,x)$ then only depends on $t$. Now as $\rho\rightarrow\infty$, we may pick any $\epsilon>0$ such that the contribution to \eqref{eq:nmax} and \eqref{eq:nmin} from points with $|t|>\epsilon$ will be negligible. We assume now that for large enough $\rho$, the region $\left\{|t|<\epsilon\right\}$ is entirely contained in the neighbourhood $U_\Sigma$ in which the GNC are valid.
Equations \eqref{eq:nmax} and \eqref{eq:nmin} can then be simplified as any time, $t$, which is far away from the surface will give rise to a large volume, which will then be exponentially suppressed as $\rho \rightarrow \infty$. More technically, if we choose some finite coordinate distance $\varepsilon$ away from the surface, either to the past or future, then any volume $V(p,\Sigma)$ or $V(\Sigma,q)$ with $|t|>\varepsilon$ will contribute nothing to the integral as $\rho \rightarrow \infty$. This $\varepsilon$ can be chosen arbitrarily close to $0$ allowing one to expand metric in $t$ about $t=0$. This gives

\begin{align}\label{eq:nmax_and_eq:nmin}
\left\langle N_{max}\right\rangle & =\int_{\Sigma}d^{d-1}x\int_{-\varepsilon}^{0}dt\:
h^{\frac{1}{2}}\left(1+
\frac{1}{2}\frac{\dot{h}}{h}t+O(t^2)\right)
 \rho\ e^{-\rho V(-t,\mathbf{x};0,\mathbf{x})}
\\
\left\langle N_{min}\right\rangle & =\int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
h^{\frac{1}{2}}\left(1+
\frac{1}{2}\frac{\dot{h}}{h}t+O(t^2)\right) \rho\ e^{-\rho V(0,\mathbf{x};t,\mathbf{x})}
\end{align}
where $h\equiv det\left(h_{ij}(0,\mathbf{x})\right)$ and $\dot{}\equiv \frac{\partial}{\partial t}$. The metric determinant has been in expanded in small $t$. The volume functions have been rewritten as follows: $V(x,\Sigma)=V(x,x_0)=V(-t,\mathbf{x};0,\mathbf{x})$ and $V(\Sigma,q)=V(x_0,x)=V(0,\mathbf{x};t,\mathbf{x})$

The only curvy cones that will contribute will have small volumes, as we are restricting ourselves close to the surface. This means that the volumes can be approximated by the volumes of less curvy cones, and it can be shown that higher order corrections, in the approximation scheme, vanish in the limit of $\rho \rightarrow \infty$. This will all be made more precise below. The approximation procedure will be outlined below for a cone to the future of $\Sigma$, as the process for the past cone is nearly identical. Riemann normal coordinates are integral to the following calculation so it is worth noting the relevant formulas before jumping in.

\subsection{Riemann Normal Coordinates}

One can always find Riemann normal coordinates (RNC), about a point $p$, by changing from coordinates $x^{\alpha}$ to coordinates $y^{\overline{\alpha}}$ such that in these new coordinates the metric and the Christoffel symbols at $p$ are that of flat space. These conditions  can be expressed as follows

\be\label{eq:RNCMetricTransAtPAndChris}
g_{\overline{\alpha} \overline{\beta}}(p)=\eta_{\overline{\alpha} \overline{\beta}}=A^{\alpha}_{\;\overline{\alpha}}A^{\beta}_{\;\overline{\beta}}g_{\alpha\beta}(p)\;,\;\;\;\;\Gamma^{\overline{\alpha}}_{\;\overline{\beta}\overline{\gamma}}(p)=0
\ee
The $A^{\alpha}_{\;\overline{\alpha}}$ govern the coordinate transformation to linear order, but $O(x^2)$ corrections may be required. To second order the coordinate transformation is given by

\be\label{eq:RNCtotaltrans}
y^{\overline{\alpha}}=A^{\overline{\alpha}}_{\;\beta}x^\beta+\frac{1}{2}A^{\overline{\alpha}}_{\;\alpha}\Gamma^{\alpha}_{\;\beta\gamma}(p)x^\beta x^\gamma+O(x^3)
\ee
and the inverse transformation is just

\be\label{eq:RNCinversetrans}
x^{\alpha}=A^{\alpha}_{\;\overline{\beta}}y^{\overline{\beta}}+O(y^2)
\ee
One finds that the $A^{\alpha}_{\;\overline{\alpha}}$ also satisfy

\be\label{eq:RNCeqnforA}
A^{\overline{\alpha}}_{\;\mu}A^{\mu}_{\;\overline{\beta}}=\delta^{\overline{\alpha}}_{\;\overline{\beta}}\;,\;\;\;\;A^{\alpha}_{\;\overline{\mu}}A^{\overline{\mu}}_{\;\beta}=\delta^{\alpha}_{\;\beta}
\ee
These relations for RNC will cover all that will be needed in this discussion.

\subsection{Lightcone Volumes}

The volumes of the truncated lightcones can be found as follows. Consider the past lightcone emanating at the point $p_0=(T,\mathbf x_0)$. There is a unique point $q_0=(0,\mathbf x_0)$ on $\Sigma$ associated with $p_0$, separated from $p_0$ by a proper time $T$.

The volume of this cone is given by

\be\label{eq:VolumeWithNoSimplifications}
V(0,\mathbf{x}_0;T,\mathbf{x}_0)=\int_{\mathcal{X}} d^d x\;\sqrt{-g}
\ee
where the integration region, $\mathcal{X}:= J^-(p_0)\bigcap J^+(\Sigma)$ will be some complicated thing. As we are dealing with points close to the surface we can use the  Riemann normal coordinates $y^{\overline{\mu}}$, defined as above, about the point $q_0$ (the point $q_0$ being the origin of our RNC system). One finds that $A^0_{\;\overline{0}}=1$, $A^0_{\;i}=0$ and $\delta_{\overline{\emph\i}\overline{\emph\j}}=A^i_{\;\overline{\emph\i}}A^j_{\;\overline{\emph\j}}h_{ij}(0,\mathbf{x}_0)$. It can be shown that the volume integral reduces to \cite{Khetrapal_Sumati:Causal_Diamond_Volume}

\be\label{eq:VolumeWithRNC}
V(0,\mathbf{x}_0;T,\mathbf{x}_0) =\int_{\mathcal{X}}d^dy+\int_{\mathcal{X}_0}d^dy\left(-\frac{1}{6}R_{\overline{\mu}\overline{\nu}}(q_0)y^{\overline{\mu}}y^{\overline{\nu}} \right)+O(T^{d+3})
\ee
where $\mathcal{X}_0=\left\lbrace y^{\overline{\mu}} \mid 0\leq y^{\overline{0}}=\overline{t}\leq T\; ,\; \sqrt{(y^{\overline{1}})^2+(y^{\overline{2}})^2+...+(y^{\overline{d-1}})^2}\leq T-\overline{t} \right\rbrace$ is what we call the \textit{flat cone}, and $R_{\overline{\mu}\overline{\nu}}(q_0)$ is the Ricci tensor in RNC evaluated at $q_0$. The arguments of the volume function need not be changed as the volume only depends on the points in the manifold, $q_0$ and $p_0$. The \textit{flat cone} term comes in at $O(T^{d+2})$ so the volume we have to calculate has reduced to

\be\label{eq:VolumeToLowestOrder}
V(0,\mathbf{x}_0;T,\mathbf{x}_0) =\int_{\mathcal{X}}d^dy+O(T^{d+2})
\ee
Terms of $O(T^{d+2})$ can be retained till the end, but in the limit, $\rho\rightarrow\infty$, they vanish. This will be proved below.

The boundaries of $\mathcal{X}$ must now be figured out in order to do this integral. First we look at the top part of the cone. From \cite{Khetrapal_Sumati:Causal_Diamond_Volume} it can be shown that the first curvature correction to the top part of the cone comes in at $O(T^{d+2})$ and so can be ignored for our purposes. This means that the top can be treated as a that of a \textit{flat cone} and so points on the boundary obey the equation $\sqrt{(y^{\overline{1}})^2+(y^{\overline{2}})^2+...+(y^{\overline{d-1}})} = T-\overline{t}$.

The bottom surface of the cone was defined in terms of GNC to be $t=0$, so we can use (\ref{eq:RNCtotaltrans}) to find the equation for the surface in RNC. Equation (\ref{eq:RNCtotaltrans}) gives

\be\label{eq:BottomSurfaceWithGNC}
y^{\overline{0}}=\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij}(0,\mathbf{x}_0)x^i x^j+O(x^3)
\ee
The first part on the right of (\ref{eq:RNCtotaltrans}) vanishes as $A^{\overline{0}}_{\;\mu}x^{\mu}=x^0$ (as $A^{\overline{0}}_{\;i}=0$ and $A^{\overline{0}}_{\;0}=1$) and $x^0=t=0$ for the bottom surface. Using the inverse RNC relation (\ref{eq:RNCinversetrans}) one can find the equation for the bottom surface entirely in RNC.

\be\label{eq:BottomSurface}
\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij}(0,\mathbf{x}_0)A^{i}_{\;\overline{\emph\i}}A^{j}_{\;\overline{\emph\j}}y^{\overline{\emph\i}} y^{\overline{\emph\j}}+O(y^3)
\ee

This equation can be written in terms of the radial coordinate of the cone, $r=\sqrt{(y^{\overline{1}})^2+...+(y^{\overline{d-1}})^2}$, as

\be\label{eq:RadialBottomSurface}
\overline{t}=\frac{1}{2}\left(\Gamma^{0}_{\;ij}(0,\mathbf{x}_0)A^{i}_{\;\overline{\emph\i}}A^{j}_{\;\overline{\emph\j}}\frac{y^{\overline{\emph\i}} y^{\overline{\emph\j}}}{r^2}\right)r^2+O(y^3)=\frac{1}{2}f(\mathbf{x}_0,\phi)r^2+O(y^3)
\ee
where $\phi$ stands for all the angular coordinates $\phi_1,..,\phi_{d-2}$. The function $f(\mathbf{x}_0,\phi)$ depends on $\mathbf{x}_0$ as $\Gamma^{0}_{\;ij}$ and $A^{i}_{\;\overline{\emph\i}}$ will, in general, be different for a different point on the surface. One can see that $f(\mathbf{x}_0,\phi)$ depends on the angles, $\phi$, by substituting the relations below into (\ref{eq:RadialBottomSurface})

\begin{equation}
\begin{aligned}\label{eq:SphericalCoords}
y^{\overline{1}} &= r \cos(\phi_1)  \\
y^{\overline{2}} &= r \sin(\phi_1) \cos(\phi_2)  \\
y^{\overline{2}} &= r \sin(\phi_1) \sin(\phi_2) \cos(\phi_3)  \\
    &\vdots  \\
y^{\overline{d-2}} &= r \sin(\phi_1) \cdots \sin(\phi_{d-3}) \cos(\phi_{d-2})  \\
y^{\overline{d-1}} &= r \sin(\phi_1) \cdots \sin(\phi_{d-3}) \sin(\phi_{d-2})
\end{aligned}
\end{equation}

With the boundaries of the integration region in place, we can now write down the integral explicitly, in spherical coordinates:

\be\label{eq:VolumeIntegralSpherical}
V(0,\mathbf{x}_0;T,\mathbf{x}_0)=\int_{S^{d-2}}
d\Omega_{d-2}
\int_{0}^{r_{max}(\phi)}r^{d-2}dr
\int_{\frac{1}{2}f(\mathbf{x}_0,\phi)r^2}^{-r+T}
d\overline{t}+O(T^{d+2})
\ee
where $r_{max}(\phi)$ is the radial coordinate of where the bottom surface meets the top part of the cone at an angle $\phi$, shown in \ref{fig:cone_plot}. To find this one needs to solve $\frac{1}{2}f(\mathbf{x}_0,\phi){r_{max}}^2(\phi)=-r_{max}(\phi)+T$ and take the positive solution. The solution can be expanded in $T$ and is simply $r_{max}=T+O(T^2)$, with angular dependent terms contributing at $O(T^2)$. The $O(T^2)$ term will contribute at $O(T^{d+2})$ in the integral and so can be ignored. Substituting $r_{max}=T$ into (\ref{eq:VolumeIntegralSpherical}) means it can be evaluated to give
\begin{figure}[t]
  \centering
    \includegraphics[scale=0.5]{coneplot}
     \caption{The size of the region inside the top and bottom bounding surfaces is the volume we want to calculate.}
     \label{fig:cone_plot}
\end{figure}
\be\label{eq:VolumeNoK}
V(0,\mathbf{x}_0;T,\mathbf{x}_0)
=\frac{A_{d-2}}{d(d-1)}T^d\left(1-\frac{d}{2(d+1)}\Gamma^{0}_{\;ij}(0,\mathbf{x}_0)A^{i}_{\;\overline{\emph\i}}A^{j}_{\;\overline{\emph\j}}\delta^{\overline{\emph\i}\overline{\emph\j}}T\right)
+O(T^{d+2})
\ee
where $A_{d-2}$ is the $d$-$2$ dimensional volume of a $d$-$2$ dimensional sphere and the $\delta^{\overline{\emph\i}\overline{\emph\j}}$ has come from the fact that cross terms ($\overline{\emph\i}\neq \overline{\emph\j}$) vanish under the angular integration. The defining relations for $A^{i}_{\;\overline{\emph\i}}$ can be rearranged to give $A^{i}_{\;\overline{\emph\i}}A^{j}_{\;\overline{\emph\j}}\delta^{\overline{\emph\i}\overline{\emph\j}}=h^{ij}(0,\mathbf{x}_0)$. In GNC it can be shown that the extrinsic curvature is given by

\be\label{eq:K}
K(0,\mathbf{x}_0)
=g^{\mu\nu }\nabla_{\mu}n_{\nu}
=-\Gamma^{0}_{\;ij}(0,\mathbf{x}_0)h^{ij}(0,\mathbf{x}_0)=-\frac{1}{2}\frac{\dot{h}(0,\mathbf{x}_0)}{h(0,\mathbf{x}_0)}
\ee
which means that the first correction to the volume of the cone depends on $K(0,\mathbf{x}_0)$, and thus the approximate volume of the cone is

\begin{align}
V(0,\mathbf{x}_0;T,\mathbf{x}_0)
&=\frac{A_{d-2}}{d(d-1)}T^d\left(1+\frac{d}{2(d+1)}K(0,\mathbf{x}_0)T\right)
+O(T^{d+2}) \label{eq:TopVolumeWithK}\\
V(-T,\mathbf{x}_0;0,\mathbf{x}_0)
&=\frac{A_{d-2}}{d(d-1)}T^d\left(1-\frac{d}{2(d+1)}K(0,\mathbf{x}_0)T\right)
+O(T^{d+2}) \label{eq:BottomVolumeWithK}
\end{align}
where (\ref{eq:BottomVolumeWithK}) is the volume of the cone with its tip below the surface, and can be found from similar arguments.

\subsection{Final Stretch of the Proof}

Let us now use these volume functions in the expressions for $\left\langle N_{max}\right\rangle$ and $\left\langle N_{min}\right\rangle$ to calculate the $\left\langle N_{max}-N_{min}\right\rangle$ term in equation (\ref{GH_boundary_to_causet}) as $\rho \rightarrow \infty$. We have
\begin{gather}\label{eq:NmaxNminStart}
\begin{aligned}
\lim_{\rho \to \infty}\rho^{\frac{2}{d}-1}\left\langle N_{max}-N_{min} \right\rangle &= \\
\lim_{\rho \to \infty}\rho^{\frac{2}{d}}
\int_{\Sigma}d^{d-1}x & \int_{-\varepsilon}^{0}dt\
h^{\frac{1}{2}}\left(1+
\frac{1}{2}\frac{\dot{h}}{h}t-\rho B(-1)^{d}t^{d+1}+\rho O(t^{d+2})\right)e^{-\rho A(-1)^{d}t^{d}} \\
-\rho^{\frac{2}{d}}\int_{\Sigma}d^{d-1}x &
\int_{0}^{\varepsilon}dt\
h^{\frac{1}{2}}\left(1+
\frac{1}{2}\frac{\dot{h}}{h}t-\rho Bt^{d+1}+\rho O(t^{d+2})\right)e^{-\rho At^{d}}
\end{aligned}
\end{gather}
where we have defined
\begin{gather}\label{A_and_B_defn}
\begin{aligned}
A & \equiv \frac{A_{d-2}}{d(d-1)} \\
B & \equiv \frac{A_{d-2}}{2(d-1)(d+1)}K(0,\mathbf{x})
\end{aligned}
\end{gather}
and we have Taylor expanded the $O(t^{d+1})$ parts of the volume functions, in small $t$, from the exponents. We have also included a pre-factor of $\rho^{\frac{2}{d}-1}$ to account for the factors of $l$ in (\ref{GH_boundary_to_causet}). With this pre-factor the limit of the above equation is well defined. If we change to $u=-t$ in the first part on the right hand side, and $u=t$ on the second part then the integral can be simplified to
\be\label{eq:NmaxNminSimplified}
\lim_{\rho \to \infty}\rho^{\frac{2}{d}}
\int_{\Sigma}d^{d-1}x \int_{0}^{\varepsilon}du\
h^{\frac{1}{2}}\left(-\frac{\dot{h}}{h}u+2\rho B u^{d+1}+\rho O(u^{d+2}) \right)e^{-\rho Au^{d}}
\ee
We now show that $O(u^{d+2})$, or higher, terms vanish as $\rho \rightarrow\infty$. All higher order corrections that we have ignored so far contribute at $O(u^{d+2})$ or higher, so proving that these contributions vanish proves that our approximations are valid. Taking only the part we are interested in from (\ref{eq:NmaxNminSimplified}) leaves us with the task of showing that the following integral vanishes.

\be\label{eq:VanishingOrderIntegral}
\lim_{\rho \to \infty}\rho^{\frac{2}{d}+1}\int_{0}^{\varepsilon}du\
u^{d+n}e^{-\rho Au^{d}}
\ee
for $n>1$, where $n\in\mathbb{N}$. The additional factor of $\rho$ has come from inside the integral in (\ref{eq:NmaxNminSimplified}). We make the substitution $z=\rho Au^{d}$ and find that (\ref{eq:VanishingOrderIntegral}) is proportional to

\be\label{eq:GammaVanishingOrderIntegral}
\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\int_{0}^{\rho A\varepsilon^d}dz\;z^{\frac{n+1}{d}}e^{-z}=\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\int_{0}^{\infty}dz\;z^{\left[\left(\frac{n+1}{d}+1\right)-1 \right]}e^{-z}
\ee
where $\rho$ has been taken to $\infty$ in the limits on the RHS, and the power of $z$ has been written in this way so that the integral is in the form of a gamma function, $\Gamma(t)=\int_{0}^{\infty}z^{t-1}e^{-z}dz$. Using the fact that for gamma functions, $\Gamma(t+1)=t\Gamma(t)$, we find that (\ref{eq:GammaVanishingOrderIntegral}) is then proportional to

\be\label{eq:GammaVanishingOrder}
\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\Gamma\left(\frac{n+1}{d}\right)
\ee
One can see from above that if $n>1$ then this term will vanish, thus proving that terms of $O(u^{d+2})$ or higher will go to $0$ in the limit $\rho\rightarrow\infty$.

The non-vanishing terms of (\ref{eq:NmaxNminSimplified}) can also be put into the form of gamma functions. This allows the integral over $u$ to be evaluated in (\ref{eq:NmaxNminSimplified}). One also sees that we have terms like $\frac{\dot{h}}{h}$, and from (\ref{eq:K}) we know that these can be written in terms of the extrinsic curvature, $K(0,\mathbf{x})$. We find then that (\ref{eq:NmaxNminSimplified}) reduces to

\be\label{eq:nmax_eq:nmin_end}
\lim_{\rho \to \infty}\rho^{\frac{2}{d}-1}\left\langle N_{max}-N_{min} \right\rangle=
\frac{2(d+2)}{d(d+1)}
\left[\frac{A_{d-2}}{d(d-1)}\right]^{-\frac{2}{d}}
\Gamma\left(\frac{2}{d}\right)\int_{\Sigma}d^{d-1}x\
\sqrt{h}K(0,\mathbf{x}).
\ee
$K(0,\mathbf{x})$ is evaluated at the surface and is therefore the same $K$ as in equation (\ref{GH_boundary_to_causet}). We can therefore conclude that to make (\ref{eq:mainconjecture}) true we need a factor $l_{p}^{2-d}$ to make our expression dimensionless, and a value of $c_d$ equal to that of equation (\ref{Cn}). This concludes the proof of the conjectured equality in (\ref{eq:mainconjecture}).

\section{Fluctuations}
So far we have only talked about the mean of the causal set boundary action. Let us now turn to its fluctuations (the standard deviation) $\sigma[S^{(d)}_{GHY}(\mathcal C)]=\text{Var}[S^{(d)}_{GHY}(\mathcal C)]^\frac12$.\footnote{Whenever we say fluctuations we refer to the standard deviation of the random variable, not to its variance.} We can make a heuristic argument to estimate the dependence of fluctuations on $\rho$. In any spacetime region of fixed volume $V$ the number of causal set elements $N$ experiences Poisson fluctuations of order $\sqrt N$. Since $N_{max}$ and $N_{min}$ are quantities associated with a codimension-1 surface (counting elements ``near" $\Sigma$), we may expect them to inherit fluctuations of order $N^\frac{d-1}{2d}$. Then the action $S^{(d)}_{GHY}$, being proportional to $\rho^\frac{d-2}{d}$ times the difference of two independent\footnote{The independence is as a feature of the Poisson process in $M$: the number of elements $N(R)$ in any subregion $R\subset M$ is a Poisson variable itself, and for any two disjoint regions $R_1$, $R_2$ the numbers $N(R_1)$ and $N(R_2)$ are independent random variables.} random variables with standard deviation $N^\frac{d-1}{2d} = (\rho V)^\frac{d-1}{2d}$ should see fluctuations of order $\rho^\frac{2-d}{d}\rho^\frac{d-1}{2d}=\rho^\frac{d-3}{2d}$. This suggests that for $d=2$ these fluctuations should grow as $\rho\rightarrow\infty$, for $d=3$ they should be constant, and for $d>3$ they should be damped.

In order to test this heuristic argument one may like to write down the integral expression for $\text{Var}[S^{(d)}_{GHY}(\mathcal C)]^\frac12$ but it is complicated enough even in finite regions of flat space that it is not very illuminating to reproduce it here. Instead we will show the results of computer simulations that support results of the heuristic argument.

The simulations were carried out as follows. Denote the discreteness scale by $l$. Take a $d$-cube $[0,L]^d$ in $d$-dimensional Minkowski space with metric $ds^2=-dt^2+d{\mathbf x}^2$ and define the hypersurface $\Sigma: t=L/2$, which partitions the cube into two halves. Sprinkle at density $\rho=l^{-d}$ into the cube, which in any given run will place $N$ points inside the cube where $N$ is a Poisson random number with mean $\left\langle N\right\rangle = \rho V=  (L/l)^d$. Evaluate $S^{(d)}_{GHY}=c_d\rho^\frac{d-2}{d}(N_{min} - N_{max})$ for $\Sigma$ by counting the minimal/maximal elements in the upper/lower half of the cube, and repeat.

\bibliographystyle{utphys}

\bibliography{references}{}

\end{document}

%%
JHEPPUB:

\title{Gibbons-Hawking-York Boundary Term in Causal Sets}
 \author[a]{M. Buck}
 \author[a,b]{\!, F. Dowker}
 \author[a]{and I. Jubb\,}
\affiliation[a]{Theoretical Physics Group, Blackett Laboratory, Imperial College, London, SW7 2AZ, U.K.}
\affiliation[b]{Institute for Quantum Computing, University of Waterloo, ON, N2L 2Y5, Canada}

\abstract{ 
We show some stuff.
}

\begin{document}

\maketitle

\pagebreak

\noindent NOTES:\\
1. define s as the RV, not as its mean

\section{Introduction}

%%IOPART

\begin{document}

\title{Gibbons-Hawking-York Boundary Term in Causal Sets}

\author{Michel Buck Fay Dowker, Ian Jubb}
\address{Blackett Laboratory, Imperial College, London, SW7 2AZ, U.K.}

\begin{abstract}

We show some stuff.

\end{abstract}
%\pacs{03.67.-a, 03.65.Ta, 03.70.+k}

\maketitle
\section{Introduction}
%%

### 2014-09-18 2014-06-04-paper_v4.tex
initial commit

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{jheppub}
\usepackage{marginnote,xparse,changepage,caption}

\newcommand{\IGNORE}[1]{}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bea}{\begin{aligned}}
\newcommand{\eea}{\end{aligned}}
\newcommand{\mb}[1]{\marginnote{\texttt{\small MB:\,#1}}}
\renewcommand{\ij}[1]{\marginnote{\texttt{\small IJ:\,#1}}}
\newcommand{\ibar}{{\overline{\emph\i}}}
\newcommand{\jbar}{{\overline{\emph\j}}}

\title{Boundary terms in the action for Causal Sets}
 \author[a]{M. Buck}
 \author[a,b]{\!, F. Dowker}
 \author[a]{and I. Jubb\,}
\affiliation[a]{Theoretical Physics Group, Blackett Laboratory, Imperial College, London, SW7 2AZ, U.K.}
\affiliation[b]{Institute for Quantum Computing, University of Waterloo, ON, N2L 2Y5, Canada}

\abstract{ 
We propose a formula for the boundary term in the action of a causal set that is well-approximated by a continuum manifold with spacelike boundary. The boundary term is proportional to the difference in the number of elements immediately to future and the number of elements immediately to the past of the surface. We show that in the continuum limit one recovers the Gibbons-Hawking-York boundary term in the mean.
}

\begin{document}

\maketitle

\noindent NOTES:\\
1. define s as the RV, not as its mean

\section{Introduction}

%[MB1: USEFUL QUOTE FROM RAFAEL: In the quantum theory, however, these boundary terms are important. They are essential in order that the quantum mechanical amplitudes satisfy the correct composition law and in order that these amplitudes have the correct classical limit. In a number of examples [5] they give a contribution to the partition function which is important for the agreement with calculations based on straightforward thermodynamics. In this note we shall derive the boundary terms in the action for the Regge calculus]

In furthering causal set theory it is crucial that we understand the kinematics of the theory. The action of a given causal set is a crucial piece of the kinematics that would be extremely useful to know and understand. Proposals for the action of a causal set are available \cite{Benincasa_Dowker:The_Scalar_Curvature_of_a_Causal_Set} and these hold analytically in some cases and numerically in many more.\mb{rephrase} These cases being when the causal set is embeddable in some existing spacetime. One can show that the action of the causal set then agrees with the Einstein-Hilbert action of the spacetime in some continuum limit. How this limiting procedure is carried out will be described in more detail below.

It is well known that the Einstein-Hilbert action is not the full story in the continuum. In the presence of spacetime boundaries the gravitational action must include a boundary term $S_{GHY}$, the Gibbons-Hawking-York action, in order to yield a well-defined variational principle~\cite{Gibbons_Hawking_Boundary}. The contributions of this term play an essential role in particular in the quantum theory. For instance, in the calculation of the black hole entropy via the Euclidean path-integral, it is the boundary terms that produce the answer $A/4l_p^2$ necessary for the unification of black hole mechanics with thermodynamics. To give another example, in Regge calculus boundary terms are necessary in order that the quantum mechanical amplitudes satisfy the correct composition law and that they have the correct classical limit\cite{boundarybook,hartlesorkin}.

MB1: paragraph about boundary term in causets. In this paper we discuss a proposal for the  causal set analogue of the boundary action for \emph{spacelike} hypersurfaces.

We define a \textit{causal set} (or \textit{causet}) as a locally finite partial order. This means it is a pair $(\mathcal{C},\preceq)$ where $\mathcal{C}$ is the set of points and $\preceq$ is a partial order relation on $\mathcal{C}$ that has the following properties. It is (i) reflexive: $x\preceq x$, (ii) acyclic: $x\preceq y\preceq x \Rightarrow x=y$, and (iii) transitive: $x\preceq y\preceq z \Rightarrow x\preceq z$ for all points $x, y, z \in \mathcal{C}$. We define an inclusive order interval as the set $I(x,y)\equiv \lbrace z\in\mathcal{C}|x\preceq z\preceq y\rbrace$ for any $x, y\in\mathcal{C}$. The \textit{locally finite} condition is simply that the cardinality of any order interval is finite, that is $|I(x,y)|<\infty$. This condition ensures we are dealing with a discrete structure, as all the other properties of the relation would apply to an order relation between points on a continuous manifold.

In order to say we have a causal set analogue for a continuous expression we need a well defined procedure for relating the discrete theory to the continuum. This procedure, or tool, is called a \textit{Poisson sprinkling}, or just a \textit{sprinkling}. It is a Poisson process which provides a way to generate a causet from a $d$-dimensional Lorentzian manifold $(\mathcal{M},g)$ by selecting points in $\mathcal{M}$ to be the elements of $\mathcal{C}$, with an order relation given by the causal order of the manifold. The number of points chosen in a region of spacetime volume, $V$, is a Poisson random variable. This means that the expected number of points in some region will be $\rho V$, where $\rho$ is the density of the sprinkling. The density is related to the discreteness scale, $l$, by $\rho=l^{-d}$ in $d$ spacetime dimensions. It is called a sprinkling as one can envisage the point selection process as a 'sprinkling' of points into the manifold. If a causet, $\mathcal{C}$, can be generated with relatively high probability by a sprinkling into the manifold, $(\mathcal{M},g)$, then we say that the manifold is a good approximation for the causal set.

\section{The Claim}

Consider a sufficiently well-behaved $d$-dimensional spacetime $(M,g)$ and Cauchy surface $\Sigma$ in $M$. The causal past and future sets $M^\pm=J^\pm(\Sigma)$ form a partition of $M$ and $\partial M^\pm = \Sigma$. The Gibbons-Hawking-York boundary term for $M^\pm$ is in this case given by
\be\label{eq:GHYBT_in_continuum}
S_{GHY} = \pm \frac{1}{l_p^{d-2}}\int_{\Sigma} K d\Sigma
\mb{lose $\pm$?}
\ee
where $l_p$ is the rationalised Planck length and $K$ is the trace of the extrinsic curvature $K_{\mu\nu}=h_{\mu}^\rho h_\nu^\sigma \nabla_\rho n_\sigma$ of $\Sigma$ defined with future-pointing timelike unit normal $n_{\mu}=\partial_\mu S/\sqrt{g^{\mu\nu}\partial_\mu S\partial_\nu S}$.

Now observe that the integral in~\eqref{eq:GHYBT_in_continuum} can be thought of as the ``volume gradient'' across the surface $\Sigma$. More formally
\be
\int K d\Sigma = \frac{\partial}{\partial n}\int d\Sigma,
\ee
where the right hand side is the derivative of the area $\int d\Sigma$ as each point of $\Sigma$ is moved an equal distance along the outward unit normal $n$. \mb{does the antichain defn really work?} In the causal set, the analogue of a spacelike hypersurface is an anti-chain, i.e. a subset $\mathcal{A}\subseteq\mathcal{C}$ in which all elements are unrelated. The intuitive analogue of the boundary term would then be the rate of change of the number of causal set elements below and above the antichain. We shall see that this intuitive idea indeed bears out.\mb{Write some stuff about antichains being a bit pathological and all that.} %This suggests a very simple analogy for a causal set with a ``spacelike hypersurface''. \mb{under construction} The analogy These unrelated elements are maximal in the sense that there exist no elements $y\in\mathcal{C}\setminus\mathcal{A}$ such that $x\preceq y$, $\forall x\in\mathcal{A}$.

Consider a causal set $\mathcal C$ obtained by a Poisson sprinkling at density $\rho=1/l^d$ into a $d$-dimensional spacetime $(M,g)$. A Cauchy surface $\Sigma\subset M$ induces a partition $\mathcal C = \mathcal C^+ \cup\, \mathcal C^-$ of the sprinkling, where $\mathcal C^+$ and $\mathcal C^-$ denote the restrictions of $\mathcal C$ to the points sprinkled to the future and past of $\Sigma$, respectively. Spacetime volume in the causal set is obtained simply by counting the number of elements, and hence the volume gradient corresponds to the difference in the number of ``immediate neighbours'' to the future and to the past of the surface. The most natural definition for the nearest neighbours to the future/past of $\Sigma$ in the sprinkling is to take the minimal/maximal elements in $\mathcal C^\pm$, respectively. Thus $x\in \mathcal C^-$ is a maximal element in $\mathcal C^-$ if it is maximal in the precedence relation of the causal set, i.e. if there exists no element $y\in\mathcal C^-$ such that $x\preceq y$. Figure \ref{fig:Nmin_Nmax} shows an illustration of the idea. The minimal and maximal elements have been highlighted.

Let us denote the number of maximal elements in $\mathcal C^-$ by $N_{max}$ and the number of minimal elements in $\mathcal C^+$  by $N_{min}$. We propose the following definition for the discrete Gibbons-Hawking-York boundary term:\mb{$S[\mathcal C, \Sigma]$?}
\be\label{GH_boundary_to_causet}
S^{(d)}_{GHY}[\mathcal C]=\left(l/l_p\right)^{d-2}c_{d}\left(N_{max}-N_{min}\right)
\ee
where the constant $c_{d}$ only depends on the spacetime dimension and is given by
\be\label{Cn}
c_{d}=\frac{d(d+1)}{2(d+2)\Gamma\left(\frac{2}{d}\right)}\left[\frac{A_{d-2}}{d(d-1)}\right]^{\frac{2}{d}},
\ee
$A_d=2\pi^\frac{d+1}2/\Gamma\left(\frac{d+1}{2}\right)$ denoting the volume of the unit $d$-sphere. To support this proposal we show that in the continuum limit $l\rightarrow 0$ (i.e. $\rho\rightarrow\infty$) we obtain
%of infinite sprinkling density into a $d$-dimensional spacetime we obtain
\be
\lim_{l\rightarrow0}\left\langle S^{(d)}_{GHY}[\mathcal C] \right\rangle= \frac{1}{l_p^{d-2}}\int_{\Sigma} d^{d-1}x \sqrt{h} K\label{eq:mainconjecture}
\ee
where $\left\langle\cdot\right\rangle$ denotes the mean over sprinklings.
%MB1: (too vague) During the calculation approximations will be made. These approximations are justified on the grounds that if certain orders are retained throughout the calculation they will in fact vanish in the limit $\rho \rightarrow \infty$
\begin{figure}
  \centering
   {\includegraphics[width=\textwidth]{minmaxplot}}
     \caption{An illustration of the idea. Pictured here is a causal set obtained by a sprinkling into a spacetime partitioned by a spacelike hypersurface. The minimal and maximal points about the surface are highlighted and the shaded regions illustrate the regions whose volumes $V_\blacktriangle$ and $V_\blacktriangledown$ are needed in the proof.}
     \label{fig:Nmin_Nmax}
\end{figure}

\section{The Proof}

\subsection{Poisson Sprinklings for $\langle N_{max}\rangle$ and $\langle N_{min}\rangle$}

In order to prove~\eqref{eq:mainconjecture} let us first derive an expression for the mean value of $S^{(d)}_{GHY}[\mathcal C]$. For any instance of the sprinkling, the probability that a sprinkled point $p\in M$ below the surface $\Sigma$ is \emph{maximal} is given by the probability that no point of the sprinkling lies in the region $J^{+}(p)\cap J^{-}(\Sigma)$, the intersection of the causal future of $p$ with the causal past of the surface $\Sigma$.\footnote{For notational convenience we use the symbol $x$ to refer to the causal element $x\in\mathcal C$, to its embedding in the manifold $M$, and to its coordinates in some chart on $M$.} This region will in general be the interior of some curvy $d$-dimensional cone truncated by the surface $\Sigma$, as illustrated in Figure~\ref{fig:Nmin_Nmax}. We will refer to these regions as \emph{truncated cones} below. The Poisson process assigns a probability
\be\label{Poisson}
\mathbb P\left(\text{no points in }J^{+}(p)\cap J^{-}(\Sigma)\right)=e^{-\rho V_\blacktriangledown(p)}
\ee
to this event, where $V_\blacktriangledown(p)\equiv V(J^{+}(p)\cap J^{-}(\Sigma))$ is the spacetime volume of the region $J^{+}(p)\cap J^{-}(\Sigma)$. The probability of sprinkling an element into an infinitesimal four-volume $dV_p$ at $p$ is $\rho dV_p$ where $\rho=l^{-d}$ is the sprinkling density, and so the total expected number of maximal elements below $\Sigma$ is
\be\label{eq:nmax}
\left\langle N_{max}\right\rangle =\rho\int_{J^{-}(\Sigma)}dV_p\; e^{-\rho V_\blacktriangledown(p)}
\ee
Similarly the expected number of minimal elements above $\Sigma$ is
\be\label{eq:nmin}
\left\langle N_{min}\right\rangle =\rho\int_{J^{+}(\Sigma)}dV_p\; e^{-\rho V_\blacktriangle(p)}
\ee
where $V_\blacktriangle(p)\equiv V(J^{+}(\Sigma)\cap J^{-}(p))$.

In the limit $\rho\rightarrow\infty$, both quantities will diverge, but if their difference $\langle N_{max}\rangle - \langle N_{min}\rangle = \langle N_{max} - N_{min}\rangle$ grows slower than or at order $\rho^{1-\frac2d}$, the proposed action~\eqref{GH_boundary_to_causet} will tend to a finite value in the continuum limit.

Consider a set of synchronous or Gaussian Normal Coordinates (GNC) $x^\mu=(t,\mathbf x)$ adapted to $\Sigma$ such that in a neighbourhood $U_\Sigma$ of $\Sigma$ the line element is
\be
ds^2 = -dt^2 + h_{ij}(t,\mathbf x) dx^i dx^j.
\ee
In these coordinates, the surface $\Sigma$ corresponds to $t=0$, and the coordinate $t$ measures the proper time elapsed along geodesics whose tangent vector is proportional to the surface normal on $\Sigma$.
%\be\label{GNC_metric}
%g_{\mu\nu}(x)=
%\begin{pmatrix}
 %-1&0 \\
% 0&h_{ij}(x)
%\end{pmatrix}.
%\ee
%In order to find $N_{max}$, the number of maximal points below the surface, one has to use the fact that the points have been sprinkled with a Poisson distribution. 
%If we pick a point $x_0=(0,\mathbf{x})\in \Sigma$ on the surface that has the same spatial coordinates as $x=(t,\mathbf{x})$ then the coordinate time, $t$, is the proper time elapsed along the unique geodesic between $x_0$ and $x$. The volume $V(\Sigma,x)$ then only depends on $t$.~\mb{error?}  
The integrals~\eqref{eq:nmax} and ~\eqref{eq:nmin} seem intractable as they stand, since the integration is over the entire causal past/future of the surface and the volume expressions will be complicated in the presence of curvature. However, for large $\rho$ (small $l$), the integrands $\exp(-\rho V)$ will be exponentially suppressed unless the volumes are small. Consider the spacetime region $U_\epsilon=\left\{|t|<\epsilon\right\}$ around the surface $\Sigma$ for some $\epsilon>0$ (with $\epsilon$ small enough such that the GNC system is valid throughout $U_\epsilon$). We will assume that for any such $\epsilon$, the contribution from points with $|t|>\epsilon$ to the integrals~\eqref{eq:nmax} and \eqref{eq:nmin} can be made arbitrarily small by setting $\rho$ large enough, since the volumes of the truncated past/future lightcones for such points will be too large. Hence, as $\rho\rightarrow\infty$, the integration ranges in \eqref{eq:nmax} and \eqref{eq:nmin} can be cut off at finite time $t=\pm\epsilon$ with $\epsilon$ arbitrarily small, which allows us to expand the time integration-variable about zero. These assumptions of course impose certain regularity conditions on the surface $\Sigma$, but we will assume that they are reasonable conditions.\mb{say more about this? we can't have an asymptotically null surface i suppose} The integrals then simplify to

\begin{align}\label{eq:nmax_and_eq:nmin}
\left\langle N_{max}\right\rangle & =\int_{\Sigma}d^{d-1}x\int_{-\varepsilon}^{0}dt\:
h^{\frac{1}{2}}\left(1+
\frac{1}{2}\frac{\dot{h}}{h}t+O(t^2)\right)
 \rho\ e^{-\rho V_\blacktriangledown(t,\mathbf x)}
\\
\left\langle N_{min}\right\rangle & =\int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
h^{\frac{1}{2}}\left(1+
\frac{1}{2}\frac{\dot{h}}{h}t+O(t^2)\right) \rho\ e^{-\rho V_\blacktriangle(t,\mathbf x)}
\end{align}
where $h\equiv det\left(h_{ij}(0,\mathbf{x})\right)$ and $\dot{}\equiv \frac{\partial}{\partial t}$ and the metric determinant has been in expanded in small $t$.

The only truncated cones that contribute will have small volumes, as we are restricting ourselves close to the surface.\mb{this sounds circular w.r.t. last paragraph} This means that the volumes can be approximated by the volumes of less curvy cones, and it can be shown that higher order corrections, in the approximation scheme, vanish in the limit of $\rho \rightarrow \infty$. This will all be made more precise below. The approximation procedure will be outlined below for a cone to the future of $\Sigma$, as the process for the past cone is nearly identical.

\subsection{Lightcone Volumes}

In order to evaluate the volume $V_\blacktriangledown(p)$ or $V_\blacktriangle(p)$ of a truncated cone we will perform a coordinate transform to Riemann Normal Coordinates (RNCs) in a neighbourhood containing the cone. The discussion for the two volume integrals is identical so we will outline that for $V_\blacktriangle(p)$ (i.e. for points to the future of $\Sigma$).

Fix $p\in M$ and denote its coordinate values in GNCs by $x^\mu_p=(t_p,\mathbf x_p)$. It will be convenient to use RNCs centered not at the tip $p$ of the cone but instead at the point $p_0$ where the unique geodesic through $p$ whose tangent is normal to $\Sigma$ intersects $\Sigma$. In GNCs this simply corresponds to the point $p_0$ with coordinates $x_0^\mu=x^\mu(p_0)=(0,\mathbf x_p)$. RNCs centered at $p_0$ will be given the symbol $y^{\overline{\mu}}$.
\mb{let's put all facts about $x^\mu$ and $y^{\overline\mu}$ for $p$ and $p_0$ here and expansion of det in RNC.} We need to assume that for any $p$ in $U_\epsilon$, the Riemann normal neighbourhood $U_p\subset M$ (throughout which the RNC system centered at $p_0$ is well-defined) contains the truncated cone $J^-(p)\cap J^+(\Sigma)$. This assumption seems reasonable given the that $U_\epsilon$ can be made arbitrarily ``thin'' as $\rho\rightarrow\infty$. In RNCs the metric and the Christoffel symbols at $p_0$ are those of flat space:
\be\label{eq:RNCMetricTransAtPAndChris}
g_{\overline{\mu} \overline{\nu}}(p_0)=\eta_{\overline{\mu} \overline{\nu}}=A^{\mu}_{\;\overline{\mu}}A^{\nu}_{\;\overline{\nu}}g_{\mu\nu}(p_0)\;,\;\;\;\;\Gamma^{\overline{\mu}}_{\;\overline{\nu}\overline{\rho}}(p_0)=0
\ee
The $A^{\mu}_{\;\overline{\mu}}$ govern the coordinate transformation from GNCs to RNCs to linear order. %but $O(x^2)$ corrections may be required. 
To second order the coordinate transformation is given by
\be\label{eq:RNCtotaltrans}
y^{\overline{\mu}}=A^{\overline{\mu}}_{\;\nu}x^\nu+\frac{1}{2}A^{\overline{\mu}}_{\;\mu}\Gamma^{\mu}_{\;\nu\rho}(p_0)x^\nu x^\rho+O((x-x_0)^3).
\ee
The inverse relation to first order is 
\be\label{eq:RNCinversetrans}
x^{\mu}=A^{\mu}_{\;\overline{\nu}}y^{\overline{\nu}}+O(y^2)
\ee
and one finds that the $A^{\mu}_{\;\overline{\mu}}$ satisfy
\be\label{eq:RNCeqnforA}
A^{\overline{\mu}}_{\;\mu}A^{\mu}_{\;\overline{\nu}}=\delta^{\overline{\mu}}_{\;\overline{\nu}}\;,\;\;\;\;A^{\mu}_{\;\overline{\mu}}A^{\overline{\mu}}_{\;\nu}=\delta^{\mu}_{\;\nu}
\ee
These relations for RNC will cover all that will be needed in this discussion.

Now the volume of the truncated cone can be found as follows. 
%Consider the past lightcone emanating at the point $p_0=(T,\mathbf x_0)$ truncated by $\Sigma$. There is a unique point $q_0=(0,\mathbf x_0)$ on $\Sigma$ associated with $p_0$, separated from $p_0$ by proper time $T$. 
The volume of the cone is given by

\be\label{eq:VolumeWithNoSimplifications}
V_\blacktriangle(p)=\int_{\mathcal{X}_p} d^d x\;\sqrt{-g}
\ee
where the integration region, $\mathcal{X}_p\equiv  J^-(p)\cap J^+(\Sigma)$ will be a complicated expression. As we are dealing with points close to the surface we can use the  RNCs $y^{\overline{\mu}}$, defined as above, about the point $p_0$. Let us denote the time-coordinate of $p$ in GNCs by $T=x^0_p=t_p$. For the transformation from GNCs to RNCs at $p_0$ one finds that $A^{\overline 0}_{\;0}=1$, $A^{\overline 0}_{\;i}=0$ and $\delta_{\ibar\jbar}=A^i_{\;\ibar}A^j_{\;\jbar}h_{ij}(p_0)$, which in particular implies $y^{\overline{0}}_p=\overline{t}_p=t_p=T$. \mb{more detail needed i think}It can then be shown that the volume integral reduces to~\cite{Khetrapal_Sumati:Causal_Diamond_Volume}

\be\label{eq:VolumeWithRNC}
V_\blacktriangle(p) =\int_{\mathcal{X}}d^dy+\int_{\mathcal{X}_0}d^dy\left(-\frac{1}{6}R_{\overline{\mu}\overline{\nu}}(p_0)y^{\overline{\mu}}y^{\overline{\nu}} \right)+O(T^{d+3})
\ee
where $\mathcal{X}_0=\left\lbrace y^{\overline{\mu}} \mid 0 \leq \overline{t}\leq T\; ,\; (y^{\overline{1}})^2+(y^{\overline{2}})^2+...+(y^{\overline{d-1}})^2\leq (T-\overline{t})^2 \right\rbrace$ is what we call the \textit{flat cone}, and $R_{\overline{\mu}\overline{\nu}}(p_0)$ is the Ricci tensor in RNC evaluated at $p_0$.\mb{justify beforehand which orders to keep?}
%The arguments of the volume function need not be changed as the volume only depends on the points in the manifold, $q_0$ and $p_0$. 
The flat cone term comes in at $O(T^{d+2})$ so the volume we have to calculate has reduced to

\be\label{eq:VolumeToLowestOrder}
V_\blacktriangle(p) =\int_{\mathcal{X}}d^dy+O(T^{d+2})
\ee
Terms of $O(T^{d+2})$ can be retained till the end, but in the limit, $\rho\rightarrow\infty$, they vanish. This will be proved below.

This is now a simple volume integral but we need to find the boundaries of $\mathcal{X}$ for the correct integration limits. There are two boundaries to the (solid) truncated cone: the lightcone emanating at $p$ and the base, given by the intersection of $\Sigma$ with the interior of the lightcone. First we look at the lightcone. Following \cite{Khetrapal_Sumati:Causal_Diamond_Volume} it can be shown that the first curvature correction to the lightcone comes in at $O(T^{d+2})$ and so can be ignored for our purposes. This means that the lightcone can be treated as effectively flat in RNCs and thus corresponding to the set $(y^{\overline{1}})^2+(y^{\overline{2}})^2+...+(y^{\overline{d-1}})^2= T-\overline{t}$.\mb{what limit on $\bar t$?}

The base of the cone in GNCs is simply a part of the surface $t=0$, so we can use (\ref{eq:RNCtotaltrans}) to find the equation for the surface in RNCs. Equation (\ref{eq:RNCtotaltrans}) gives

\be\label{eq:BottomSurfaceWithGNC}
\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij}(p_0)x^i x^j+O( x^3)\ee

The first part on the right of (\ref{eq:RNCtotaltrans}) vanishes as $A^{\overline{0}}_{\;\mu}x^{\mu}=x^0$ (as $A^{\overline{0}}_{\;i}=0$ and $A^{\overline{0}}_{\;0}=1$) and $x^0=t=0$ for the bottom surface. Using the inverse RNC relation (\ref{eq:RNCinversetrans}) one can find the equation for the bottom surface in RNCs:

\be\label{eq:BottomSurface}
\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij}(p_0)A^{i}_{\;\ibar}A^{j}_{\;\jbar}y^{\ibar} y^{\jbar}+O(y^3)
\ee

Let us rewrite this equation in spherically symmetric coordinates, i.e. define $r=\sqrt{\delta_{\ibar\jbar}y^\ibar y^\jbar}$ and the usual angular coordinates $\phi_1,..,\phi_{d-2}$ in terms of the spatial coordinates $y^{\overline{1}} = r \cos(\phi_1),\ldots, y^{\overline{d-1}} = r \sin(\phi_1) \cdots \sin(\phi_{d-3}) \sin(\phi_{d-2})$. Then

\be\label{eq:RadialBottomSurface}
\overline{t}=\frac{1}{2}\left(\Gamma^{0}_{\;ij}(p_0)A^{i}_{\;\ibar}A^{j}_{\;\jbar}\frac{y^{\ibar} y^{\jbar}}{r^2}\right)r^2+O(y^3)=\frac{1}{2}f(\mathbf{x}_p,{\boldsymbol\phi})r^2+O(y^3)
\ee
where $\boldsymbol\phi$ stands collectively for all the angular coordinates $\phi_1,..,\phi_{d-2}$. The function $f(\mathbf{x}_p,\boldsymbol\phi)$ depends on $\mathbf{x}_p$ since $\Gamma^{0}_{\;ij}$ and $A^{i}_{\;\ibar}$ depend on $p_0$.
%One can see that $f(\mathbf{x}_0,\boldsymbol\phi)$ depends on the angles, $\phi$, by substituting the relations below into (\ref{eq:RadialBottomSurface})
%\begin{equation}
%\begin{aligned}\label{eq:SphericalCoords}
%y^{\overline{1}} &= r \cos(\phi_1)  \\
%y^{\overline{2}} &= r \sin(\phi_1) \cos(\phi_2)  \\
%y^{\overline{2}} &= r \sin(\phi_1) \sin(\phi_2) \cos(\phi_3)  \\
%    &\vdots  \\
%y^{\overline{d-2}} &= r \sin(\phi_1) \cdots \sin(\phi_{d-3}) \cos(\phi_{d-2})  \\
%y^{\overline{d-1}} &= r \sin(\phi_1) \cdots \sin(\phi_{d-3}) \sin(\phi_{d-2})
%\end{aligned}
%\end{equation}

With the boundaries of the integration region in place, we can now write down the integral explicitly in spherical coordinates:

\be\label{eq:VolumeIntegralSpherical}
V_\blacktriangle(p)=\int_{S^{d-2}}
d\Omega_{d-2}
\int_{0}^{r_{max}(\phi)}r^{d-2}dr
\int_{\frac{1}{2}f(\mathbf{x}_p,\phi)r^2}^{-r+T}
d\overline{t}+O(T^{d+2})
\ee
where $r_{max}(\boldsymbol\phi)$ is the value of the radial coordinate for which the surface $\Sigma$ intersects the lightcone at angle $\boldsymbol\phi$, as shown in Figure \ref{fig:cone_plot}. To find this value one needs to solve 
\be
\frac{1}{2}f(\mathbf{x}_p,\boldsymbol\phi){r_{max}}^2(\boldsymbol\phi)=-r_{max}(\boldsymbol\phi)+T
\ee 
for $r_{max}(\boldsymbol\phi)$ and take the positive solution. The solution can be expanded in $T$ and is simply $r_{max}=T+O(T^2)$, with angular dependent terms contributing at $O(T^2)$. The $O(T^2)$ term will contribute at $O(T^{d+2})$ in the integral and so can be ignored. Substituting $r_{max}=T$ into (\ref{eq:VolumeIntegralSpherical}) lets us evaluate the integral and we find 
\begin{figure}[t]
  \centering
    {\includegraphics[scale=0.5]{coneplot}}
     \caption{The size of the region inside the top and bottom bounding surfaces is the volume we want to calculate.}
     \label{fig:cone_plot}
\end{figure}
\be\label{eq:VolumeNoK}
V_\blacktriangle(p)
=\frac{A_{d-2}}{d(d-1)}T^d\left(1-\frac{d}{2(d+1)}\Gamma^{0}_{\;ij}(p_0)A^{i}_{\;\ibar}A^{j}_{\;\jbar}\delta^{\ibar\jbar}T\right)
+O(T^{d+2})
\ee
where $A_{d-2}$ is the volume of the unit $(d-2)$-sphere and the $\delta^{\ibar\jbar}$ comes from the fact that cross terms ($\ibar\neq \jbar$) vanish under the angular integration. The defining relations for $A^{i}_{\;\ibar}$ can now be rearranged to give $A^{i}_{\;\ibar}A^{j}_{\;\jbar}\delta^{\ibar\jbar}=h^{ij}(p_0)$. Now in GNCs the extrinsic curvature on the surface is given by
\be\label{eq:K}
K
=g^{\mu\nu }\nabla_{\mu}n_{\nu}
=-\Gamma^{0}_{\;ij}h^{ij}=-\frac{1}{2}\frac{\dot{h}}{h}.
\ee
Substituting this into~\eqref{eq:VolumeNoK} with we obtain 
\begin{align}
V_\blacktriangle(T,\mathbf x)
&=\frac{A_{d-2}}{d(d-1)}T^d\left(1+\frac{d}{2(d+1)}K(0,\mathbf{x})T\right)
+O(T^{d+2}) \label{eq:TopVolumeWithK}\\
V_\blacktriangledown(-T,\mathbf x)
&=\frac{A_{d-2}}{d(d-1)}T^d\left(1-\frac{d}{2(d+1)}K(0,\mathbf{x})T\right)
+O(T^{d+2}), \label{eq:BottomVolumeWithK}
\end{align}
having dropped the subscript $p$. Given the volume expressions in GNCs we now proceed to evaluate the integrals for $\langle N_{min}\rangle $ and $\langle N_{max}\rangle$.

\subsection{The mean of $S^{(d)}_{GHY}[\mathcal C]$}

Using the previous equation we obtain for (\ref{GH_boundary_to_causet}) in the limit $\rho \rightarrow \infty$
\begin{gather}\label{eq:NmaxNminStart}
\begin{aligned}
\lim_{\rho \to \infty}\rho^{\frac{2}{d}-1}\left\langle N_{max}-N_{min} \right\rangle &= \\
\lim_{\rho \to \infty}\rho^{\frac{2}{d}}
\int_{\Sigma}d^{d-1}x & \int_{-\varepsilon}^{0}dt\
h^{\frac{1}{2}}\left(1+
\frac{1}{2}\frac{\dot{h}}{h}t+\rho B(-t)^{d+1}+\rho O(t^{d+2})\right)e^{-\rho A(-t)^{d}} \\
-\rho^{\frac{2}{d}}\int_{\Sigma}d^{d-1}x &
\int_{0}^{\varepsilon}dt\
h^{\frac{1}{2}}\left(1+
\frac{1}{2}\frac{\dot{h}}{h}t-\rho Bt^{d+1}+\rho O(t^{d+2})\right)e^{-\rho At^{d}}
\end{aligned}
\end{gather}
where we have defined
\begin{gather}\label{A_and_B_defn}
\begin{aligned}
A & \equiv \frac{A_{d-2}}{d(d-1)} \\
B & \equiv \frac{A_{d-2}}{2(d-1)(d+1)}K(0,\mathbf{x})
\end{aligned}
\end{gather}
and we have Taylor expanded the $O(t^{d+1})$ parts of the volume functions, in small $t$, from the exponents. The pre-factor $\rho^{\frac{2}{d}-1}$ accounts for the factors of $l$ in (\ref{GH_boundary_to_causet}). With this pre-factor the limit of the above equation is well defined. Swapping the integration variable to $t\rightarrow -t$ in the first term and combining the two integrals we obtain
\be\label{eq:NmaxNminSimplified}
\lim_{\rho \to \infty}\rho^{\frac{2}{d}}
\int_{\Sigma}d^{d-1}x \int_{0}^{\varepsilon}dt\
h^{\frac{1}{2}}\left(-\frac{\dot{h}}{h}t+2\rho B t^{d+1}+\rho O(t^{d+2}) \right)e^{-\rho At^{d}}
\ee
We now show that $O(t^{d+2})$ and higher terms vanish as $\rho \rightarrow\infty$. All corrections that we have ignored so far contribute at $O(t^{d+2})$ or higher, so proving that these contributions vanish proves that our approximations are justified. Inspecting (\ref{eq:NmaxNminSimplified}) we see that all we need to show is that for integer $n>1$,  the following integral vanishes:

\be\label{eq:VanishingOrderIntegral}
\lim_{\rho \to \infty}\rho^{\frac{2}{d}+1}\int_{0}^{\varepsilon}dt\
t^{d+n}e^{-\rho At^{d}}.
\ee
The additional factor of $\rho$ has come from inside the integral in (\ref{eq:NmaxNminSimplified}). We make the substitution $z=\rho At^{d}$ and find that (\ref{eq:VanishingOrderIntegral}) is proportional to

\be\label{eq:GammaVanishingOrderIntegral}
\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\int_{0}^{\rho A\varepsilon^d}dz\;z^{\frac{n+1}{d}}e^{-z}=\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\int_{0}^{\infty}dz\;z^{\left[\left(\frac{n+1}{d}+1\right)-1 \right]}e^{-z}
\ee
where $\rho$ has been taken to $\infty$ in the limits on the RHS, and the power of $z$ has been written in this way so that the integral is in the form of a gamma function, $\Gamma(t)=\int_{0}^{\infty}z^{t-1}e^{-z}dz$. Using $\Gamma(t+1)=t\Gamma(t)$ we find that (\ref{eq:GammaVanishingOrderIntegral}) is proportional to
\be\label{eq:GammaVanishingOrder}
\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\Gamma\left(\frac{n+1}{d}\right)
\ee
One can see from above that if $n>1$ then this term will vanish, thus proving that terms of $O(t^{d+2})$ or higher will go to $0$ in the limit $\rho\rightarrow\infty$.

The non-vanishing terms of (\ref{eq:NmaxNminSimplified}) can also be put into the form of gamma functions. This allows the integral over $t$ to be evaluated in (\ref{eq:NmaxNminSimplified}). One also sees that we have terms like $\frac{\dot{h}}{h}$, and from (\ref{eq:K}) we know that these can be written in terms of the extrinsic curvature $K$. We find then that (\ref{eq:NmaxNminSimplified}) reduces to

\be\label{eq:nmax_eq:nmin_end}
\lim_{\rho \to \infty}\rho^{\frac{2}{d}-1}\left\langle N_{max}-N_{min} \right\rangle=
\frac{2(d+2)}{d(d+1)}
\left[\frac{A_{d-2}}{d(d-1)}\right]^{-\frac{2}{d}}
\Gamma\left(\frac{2}{d}\right)\int_{\Sigma}d^{d-1}x\
\sqrt{h}K(0,\mathbf{x}).
\ee
$K(0,\mathbf{x})$ is evaluated at the surface and is therefore the same $K$ as in equation (\ref{GH_boundary_to_causet}). We can therefore conclude that to make (\ref{eq:mainconjecture}) true we need a factor $l_{p}^{2-d}$ to make our expression dimensionless, and a value of $c_d$ equal to that of equation (\ref{Cn}). This concludes the proof of the conjectured equality in (\ref{eq:mainconjecture}).

\section{Fluctuations}
So far we have only talked about the mean of the causal set boundary action. Let us now turn to its fluctuations (its standard deviation) $\sigma[S^{(d)}_{GHY}(\mathcal C)]=\text{Var}[S^{(d)}_{GHY}(\mathcal C)]^\frac12$.\footnote{Whenever we say fluctuations we refer to the standard deviation of the random variable, not to its variance.} We can make a heuristic argument to estimate the dependence of fluctuations on $\rho=l^{-d}$. In any spacetime region of fixed volume $V$ the number of causal set elements $N$ experiences Poisson fluctuations of order $\sqrt N$. Since $N_{max}$ and $N_{min}$ are quantities associated with a codimension-1 surface (counting elements ``near" $\Sigma$), we may expect them to inherit fluctuations of order $N^\frac{d-1}{2d}$. Then the action $S^{(d)}_{GHY}$, being proportional to $\rho^\frac{d-2}{d}$ times the difference of two independent\footnote{The independence is as a feature of the Poisson process in $M$: the number of elements $N(R)$ in any subregion $R\subset M$ is a Poisson variable itself, and for any two disjoint regions $R_1$, $R_2$ the numbers $N(R_1)$ and $N(R_2)$ are independent random variables.} random variables with standard deviation $N^\frac{d-1}{2d} = (\rho V)^\frac{d-1}{2d}$ should see fluctuations of order $\rho^\frac{2-d}{d}\rho^\frac{d-1}{2d}=\rho^\frac{d-3}{2d}$. This suggests that for $d=2$ these fluctuations should grow as $\rho\rightarrow\infty$, for $d=3$ they should be constant, and for $d>3$ they should be damped.

In order to test this heuristic argument one may like to write down the integral expression for $\text{Var}[S^{(d)}_{GHY}(\mathcal C)]^\frac12$ but it is complicated enough even in finite regions of flat space that it is not very illuminating to reproduce it here. It also seems far less tractable than the expression for the mean because it involves terms of the type $\int d^dx\int d^dy \exp\left[-\rho V(J^+(x)\cap J^+(y))\right]$ for which the above methods of cutting off the time-integration range and Taylor expanding will not follow through. Instead we will show the results of computer simulations that support results of the heuristic argument.

The simulations were carried out as follows. Denote the discreteness scale by $l$. Take a $d$-cube $[0,L]^d$ in $d$-dimensional Minkowski space with metric $ds^2=-dt^2+d{\mathbf x}^2$ and define the hypersurface $\Sigma: t=L/2$, which partitions the cube into two halves. Sprinkle at density $\rho=l^{-d}$ into the cube, which in any given run will place $N$ points inside the cube where $N$ is a Poisson random number with mean $\left\langle N\right\rangle = \rho V=  (L/l)^d$. Evaluate $S^{(d)}_{GHY}=c_d\rho^\frac{d-2}{d}(N_{min} - N_{max})$ for $\Sigma$ by counting the minimal/maximal elements in the upper/lower half of the cube, and repeat for $R$ runs.

\newpage

\bibliographystyle{jhep}

\bibliography{references}

\end{document}

%%
JHEPPUB:

\title{Gibbons-Hawking-York Boundary Term in Causal Sets}
 \author[a]{M. Buck}
 \author[a,b]{\!, F. Dowker}
 \author[a]{and I. Jubb\,}
\affiliation[a]{Theoretical Physics Group, Blackett Laboratory, Imperial College, London, SW7 2AZ, U.K.}
\affiliation[b]{Institute for Quantum Computing, University of Waterloggedo, ON, N2L 2Y5, Canada}

\abstract{ 
We show some stuff.
}

\begin{document}

\maketitle

\pagebreak

\noindent NOTES:\\
1. define s as the RV, not as its mean

\section{Introduction}

%%IOPART

\begin{document}

\title{Gibbons-Hawking-York Boundary Term in Causal Sets}

\author{Michel Buck Fay Dowker, Ian Jubb}
\address{Blackett Laboratory, Imperial College, London, SW7 2AZ, U.K.}

\begin{abstract}

We show some stuff.

\end{abstract}
%\pacs{03.67.-a, 03.65.Ta, 03.70.+k}

\maketitle
\section{Introduction}
%%

### 2014-09-18 paper_v2.tex
initial commit

--- 2014-09-18 2014-06-04-paper_v4-ianbefore.tex
@@
 \newcommand{\ee}{\end{equation}}
-\newcommand{\bea}{\begin{aligned}}
-\newcommand{\eea}{\end{aligned}}
 
@@
 
-\noindent NOTES:\\
-1. define s as the RV, not as its mean
+\pagebreak
 
@@
 
-%[MB1: USEFUL QUOTE FROM RAFAEL: In the quantum theory, however, these boundary terms are important.
-They are essential in order that the quantum mechanical amplitudes satisfy the correct composition law and in order that these amplitudes have the correct classical limit.
-In a number of examples [5] they give a contribution to the partition function which is important for the agreement with calculations based on straightforward thermo- dynamics.
-In this note we shall derive the boundary terms in the action for the Regge calculus]
+First and foremost one must draw attention to the fact that Michel Buck (on of the authors of this paper) is in fact a smelly son of bitch.
+Gerrrl, he so ugly he look like someone tried to put a fire out on his face with a fork!
 
 In furthering causal set theory it is crucial that we understand the kinematics of the theory.
-The action of a given causal set is a crucial piece of the kinematics that would be extremely useful to know and understand.
-Proposals for the action of a causal set are available \cite{Benincasa_Dowker:The_Scalar_Curvature_of_a_Causal_Set} and these hold analytically in some cases and numerically in many more.
+The action of a given causal set is one such kinematical property that would be extremely useful to know and understand.
+Proposals for the action of a causal set are available [[REFERENCE]] and these hold analytically in some cases and numerically in many more.
 These cases being when the causal set is embeddable in some existing spacetime.
-One can show that the action of the causal set then agrees with the Einstein-Hilbert (EH) action of the spacetime in some continuum limit.
+One can then show that the action of the causal set then agrees with the Einstein-Hilbert action of the spacetime in some continuum limit.
 How this limiting procedure is carried out will be described in more detail below.
@@
 The EH action, however, is not the complete story in the continuum.
-If one has a spacetime manifold with a boundary, such as Minkowski space but cut off at some finite time, then one needs to include the Gibbons-Hawking-York boundary term (GHYBT) in the EH action \cite{Gibbons_Hawking_Boundary}.
-The boundary of the manifold will be some hypersurface of codimension-1.
-The GHYBT is an integral over this hypersurface, of the extrinsic curvature that it has.
-This term is needed in the action to keep the variational principle well defined.
-In this paper we turn our attention to this term, and provide the causal set analogue for spacelike hypersurfaces.
 
-We define a \textit{causal set} (or \textit{causet}) as a locally finite partial order.
-This means it is a pair $(\mathcal{C},\preceq)$ where $\mathcal{C}$ is the set of points and $\preceq$ is a partial order relation on $\mathcal{C}$ that has the following properties.
-It is (i) reflexive: $x\preceq x$, (ii) acyclic: $x\preceq y\preceq x \Rightarrow x=y$, and (iii) transitive: $x\preceq y\preceq z \Rightarrow x\preceq z$ for all points $x, y, z \in \mathcal{C}$.
-We define an inclusive order interval as the set $I(x,y):=\lbrace z\in\mathcal{C}|x\preceq z\preceq y\rbrace$ for any $x, y\in\mathcal{C}$.
-The \textit{locally finite} condition is simply that the cardinality of any order interval is finite, that is $|I(x,y)|<\infty$.
-This condition ensures we are dealing with a discrete structure, as all the other properties of the relation would apply to an order relation between points on a continuous manifold.
-
-In order to say we have a causal set analogue for a continuous expression we need a well defined procedure for relating the discrete theory to the continuum.
-This procedure, or tool, is called a \textit{Poisson sprinkling}, or just a \textit{sprinkling}.
-It is a Poisson process which provides a way to generate a causet from a $d$-dimensional Lorentzian manifold $(\mathcal{M},g)$ by selecting points in $\mathcal{M}$ to be the elements of $\mathcal{C}$, with an order relation given by the causal order of the manifold.
-The number of points chosen in a region of spacetime volume, $V$, is a Poisson random variable.
-This means that the expected number of points in some region will be $\rho V$, where $\rho$ is the density of the sprinkling.
-The density is related to the discreteness scale, $l$, by $\rho=l^{-d}$ in $d$ spacetime dimensions.
-It is called a sprinkling as one can envisage the point selection process as a 'sprinkling' of points into the manifold.
-If a causet, $\mathcal{C}$, can be generated with relatively high probability by a sprinkling into the manifold, $(\mathcal{M},g)$, then we say that the manifold is a good approximation for the causal set.
+If one has a spacetime manifold with a boundary, such as Minkowski space but cut off at some finite time, then one needs to include the Gibbons-Hawking-York boundary term (GHYBT) in the Einstein-Hilbert action.
+This term is needed to keep the variational principle well defined.
+In this section we prove a conjecture for the form of this boundary term in causal set theory.
 
@@
 
-Consider a sufficiently well-behaved $d$-dimensional spacetime $(M,g)$ and Cauchy surface $\Sigma$ in $M$.
+Consider a sufficiently well-behaved spacetime $(M,g)$ and Cauchy surface $\Sigma$ in $M$.
 The causal past and future sets $M^\pm=J^\pm(\Sigma)$ form a partition of $M$ and $\partial M^\pm = \Sigma$.
 The Gibbons-Hawking-York boundary term for $M^\pm$ is in this case given by
-\be\label{eq:GHYBT_in_continuum}
-S_{GHY} = \pm \frac{1}{l_p^{d-2}}\int_{\Sigma} d^{d-1}x \sqrt{h} K
-\ee
-where $l_p$ is the rationalised Planck length and $K$ is the trace of the extrinsic curvature $K_{\mu\nu}=h_{\mu}^\rho h_\nu^\sigma \nabla_\rho n_\sigma$ of $\Sigma$ defined with future-pointing timelike unit normal $n_{\mu}=\partial_\mu S/\sqrt{g^{\mu\nu}\partial_\mu S\partial_\nu S}$.
+$$
+S_{GHY} = \pm\frac1{\kappa^2} \int_{\Sigma} d^{d-1}x \sqrt{h} K
+$$
+where $K$ is the trace of the extrinsic curvature $K_{\mu\nu}=h_{\mu}^\rho h_\nu^\sigma \nabla_\rho n_\sigma$ of $\Sigma$ defined with future-pointing timelike unit normal $n_{\mu}=\partial_\mu S/\sqrt{g^{\mu\nu}\partial_\mu S\partial_\nu S}$.
 
-The extrinsic curvature of a spacelike hypersurface can be identified with the volume gradient across that hypersurface [MB1: HOW SO?].
+The extrinsic curvature of a spacelike hypersurface can be identified with the volume gradient across that hypersurface.
 On the causal set, the analogue of a spacelike hypersurface is a maximal anti-chain, i.e.
-a subset $\mathcal{A}\subseteq\mathcal{C}$ in which all elements are unrelated.
-These unrelated elements are maximal in the sense that there exist no elements $y\in\mathcal{C}\setminus\mathcal{A}$ such that $x\preceq y$, $\forall x\in\mathcal{A}$.
+a subset $A$ of $C$, no two elements of which are related.
 The intuitive analogue of the boundary term would then be the rate of change of the number of causal set elements below and above the antichain.
@@
 
-Consider a causal set $\mathcal C$ obtained by a Poisson sprinkling at density $\rho=1/l^d$ into a $d$-dimensional spacetime $(M,g)$.
+Consider a causal set $\mathcal C$ obtained by a Poisson sprinkling of density $\rho$ into a spacetime $(M,g)$.
 A Cauchy surface $\Sigma\subset M$ induces a partition $\mathcal C = \mathcal C^+ \cup\, \mathcal C^-$ of the sprinkling, where $\mathcal C^\pm$ denotes the restriction of $\mathcal C$ to the points sprinkled to the causal future/past of $\Sigma$.
-Let us denote the number of maximal elements in $\mathcal C^-$ and the number of minimal elements in $\mathcal C^+$ by $N_{max}$ and $N_{min}$, respectively.
-We propose the following definition for the discrete Gibbons-Hawking-York boundary term:
+Let us denote the number of maximal elements in $\mathcal C^-$ and the number of minimal elements in $\mathcal C^+$ by $N_{min}$ and $N_{max}$, respectively.
+We propose the following definition for the discrete Gibbons-Hawking-York boundary term for a causal set $\mathcal C$ in $d$ spacetime dimensions:
 \be\label{GH_boundary_to_causet}
-S^{(d)}_{GHY}[\mathcal C]=\left(l/l_p\right)^{d-2}c_{d}(N_{max}-N_{min})
+S^{(d)}_{GHY}[\mathcal C]=\rho^{\frac{2}{d}-1}c_{d}\left\langle N_{max}-N_{min}\right\rangle.
 \ee
-where the constant $c_{d}$ only depends on the spacetime dimension and is given by
+The constant $c_{d}$ only depends on the dimension and is given by
+
 \be\label{Cn}
-c_{d}=\frac{d(d+1)}{2(d+2)\Gamma\left(\frac{2}{d}\right)}\left[\frac{A_{d-2}}{d(d-1)}\right]^{\frac{2}{d}},
+c_{d}=\frac{d(d+1)}{2(d+2)\Gamma\left(\frac{2}{d}\right)}\left[\frac{A_{d-2}}{d(d-1)}\right]^{\frac{2}{d}}.
 \ee
-$A_d$ denoting the volume of the unit $d$-sphere. To support this proposal we show that in the limit of infinite sprinkling density into a $d$-dimensional spacetime we obtain
+
+To support this proposal we show that in the limit of infinite sprinkling density we obtain
 \be
-\lim_{\rho\rightarrow\infty}\left\langle S^{(d)}_{GHY}[\mathcal C] \right\rangle= \frac{1}{l_p^{d-2}}\int_{\Sigma} d^{d-1}x \sqrt{h} K\label{eq:mainconjecture}
+\lim_{\rho\rightarrow\infty}S^{(d)}_{GHY}[\mathcal C] = \frac1{\kappa^2} \int_{\Sigma} d^{d-1}x \sqrt{h} K.\label{eq:mainconjecture}
 \ee
-where $\left\langle\cdot\right\rangle$ denotes the mean over sprinklings.
-The limit $\rho\rightarrow\infty$ is equivalent to setting the discreteness scale to zero, $l\rightarrow 0$.
-Figure \ref{fig:Nmin_Nmax} shows an illustration of the idea behind this conjecture.
-The minimal and maximal elements have been highlighted and the shaded regions correspond to the regions whose volumes expressions will be needed in the proof.
 %MB1: (too vague) During the calculation approximations will be made.
 These approximations are justified on the grounds that if certain orders are retained throughout the calculation they will in fact vanish in the limit $\rho \rightarrow \infty$
-\begin{figure}
-  \centering
-    \includegraphics[width=\textwidth]{minmaxplot}
-     \caption{An illustration of the idea.
-Pictured here is a causal set obtained by a sprinkling into a spacetime with a partitioning spacelike hypersurface.
-The minimal and maximal points about the surface are highlighted and the shaded regions illustrate the regions whose volumes are needed for the calculation.}
-     \label{fig:Nmin_Nmax}
-\end{figure}
 
-\section{The Proof}
+\section{Poisson Sprinklings for $\langle N_{max}\rangle$ and $\langle N_{min}\rangle$}
 
-\subsection{Poisson Sprinklings for $\langle N_{max}\rangle$ and $\langle N_{min}\rangle$}
-
-In order to prove~\eqref{eq:mainconjecture} let us choose a set of synchronous or Gaussian Normal Coordinates (GNC) adapted to $\Sigma$ such that in a neighbourhood $U_\Sigma$ of $\Sigma$ the line element is
+In order to prove~\eqref{eq:mainconjecture} we choose a set of synchronous, or Gaussian Normal Coordinates (GNC) adapted to $\Sigma$ such that in a neighbourhood $U_\Sigma$ of $\Sigma$ the line element is
 \be
@@
 \ee
-In these coordinates, the surface $\Sigma$ corresponds to $t=0$, and the coordinate $t$ measures the proper time elapsed along geodesics whose tangent vector is proportional to the surface normal on $\Sigma$.
+In these coordinates, the surface $\Sigma$ corresponds to the set $t=0$. 
 %\be\label{GNC_metric}
@@
 to this event, where $V(x,\Sigma):=V(J^{+}(x)\cap J^{-}(\Sigma))$ is the spacetime volume of the region $J^{+}(x)\bigcap J^{-}(\Sigma)$.
-The probability of sprinkling an element into an infinitesimal volume element $d^dx$ at $x\in M$ is $\rho\sqrt{-g(x)}d^dx$, where $\rho$ is the density of the sprinkling, and so the total expected number of maximal elements below $\Sigma$ is
+The probability of sprinkling an element into an infinitesimal volume element $d^dx$ at $x\in M$ is $\rho\sqrt{-g(x)}d^dx$, where $\rho$ is the density of the sprinkling, and so the expected number of maximal points below $\Sigma$ is
 
 \be\label{eq:nmax}
-\left\langle N_{max}\right\rangle =\rho\int_{J^{-}(\Sigma)}d^{d}x\:\sqrt{-g}\\ e^{-\rho V(x,\Sigma)}
+\left\langle N_{max}\right\rangle =\int_{J^{-}(\Sigma)}d^{d}x\:\sqrt{-g}\ \rho\ e^{-\rho V(x,\Sigma)}
 \ee
 
-Similarly the expected number of minimal elements above $\Sigma$ is
+Similarly the expected number of minimal points above $\Sigma$ is
 \be\label{eq:nmin}
-\left\langle N_{min}\right\rangle =\rho\int_{J^{+}(\Sigma)}d^{n}x\:\sqrt{-g}\ \ e^{-\rho V(\Sigma,x)}
+\left\langle N_{min}\right\rangle =\int_{J^{+}(\Sigma)}d^{n}x\:\sqrt{-g}\ \rho\ e^{-\rho V(\Sigma,x)}
 \ee
@@
 
-In the limit $\rho\rightarrow\infty$, both quantities will diverge, but if their difference $\langle N_{max}\rangle - \langle N_{min}\rangle = \langle N_{max} - N_{min}\rangle$ grows slower than or at order $\rho^{1-\frac2d}$, the proposed action~\eqref{GH_boundary_to_causet} will tend to a finite value in the continuum limit.
+Both quantities will diverge, but if their difference $\langle N_{max}\rangle - \langle N_{min}\rangle = \langle N_{max} - N_{min}\rangle$ grows slower than or at order $\rho^{1-\frac2d}$, the proposed action~\eqref{GH_boundary_to_causet} will tend to a finite value in the continuum limit.
+Figure \ref{fig:Nmin_Nmax} shows a 2-dimensional spacetime with an extrinsically curved spacelike surface (a line).
+The minimal and maximal points have been highlighted and the shaded regions correspond to those that  we wish to find the volumes of.
+This is a simple task here as the example is only 2D and the spacetime is flat.
+
+\begin{figure}
+  \centering
+    \includegraphics[width=\textwidth]{minmaxplot}
+     \caption{Minimal and maximal points about a surface in 2D Minkowski space. The shaded regions illustrate the regions that one must calculate the volume of.}
+     \label{fig:Nmin_Nmax}
+\end{figure}
 
@@
 
-\subsection{Riemann Normal Coordinates}
+\section{Riemann Normal Coordinates}
 
@@
 
-\subsection{Lightcone Volumes}
+\section{Lightcone Volumes}
 
@@
 As we are dealing with points close to the surface we can use the  Riemann normal coordinates $y^{\overline{\mu}}$, defined as above, about the point $q_0$ (the point $q_0$ being the origin of our RNC system).
-One finds that $A^0_{\;\overline{0}}=1$, $A^0_{\;i}=0$ and $\delta_{\overline{\emph\i}\overline{\emph\j}}=A^i_{\;\overline{\emph\i}}A^j_{\;\overline{\emph\j}}h_{ij}(0,\mathbf{x}_0)$.
+One finds that $A^0_{\;\overline{0}}=1$, $A^0_{\;i}=0$ and $\delta_{\overline{i}\overline{j}}=A^i_{\;\overline{i}}A^j_{\;\overline{j}}h_{ij}(0,\mathbf{x}_0)$.
 It can be shown that the volume integral reduces to \cite{Khetrapal_Sumati:Causal_Diamond_Volume}
@@
 \ee
-The first part on the right of (\ref{eq:RNCtotaltrans}) vanishes as $A^{\overline{0}}_{\;\mu}x^{\mu}=x^0$ (as $A^{\overline{0}}_{\;i}=0$ and $A^{\overline{0}}_{\;0}=1$) and $x^0=t=0$ for the bottom surface.
+The first part on the right vanishes as $A^{\overline{0}}_{\;\mu}x^{\mu}=x^0$ (as $A^{\overline{0}}_{\;i}=0$ and $A^{\overline{0}}_{\;0}=1$) and $x^0=t=0$ for the bottom surface.
 Using the inverse RNC relation (\ref{eq:RNCinversetrans}) one can find the equation for the bottom surface entirely in RNC.
@@
 \be\label{eq:BottomSurface}
-\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij}(0,\mathbf{x}_0)A^{i}_{\;\overline{\emph\i}}A^{j}_{\;\overline{\emph\j}}y^{\overline{\emph\i}} y^{\overline{\emph\j}}+O(y^3)
+\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij}(0,\mathbf{x}_0)A^{i}_{\;\overline{i}}A^{j}_{\;\overline{j}}y^{\overline{i}} y^{\overline{j}}+O(y^3)
 \ee
@@
 \be\label{eq:RadialBottomSurface}
-\overline{t}=\frac{1}{2}\left(\Gamma^{0}_{\;ij}(0,\mathbf{x}_0)A^{i}_{\;\overline{\emph\i}}A^{j}_{\;\overline{\emph\j}}\frac{y^{\overline{\emph\i}} y^{\overline{\emph\j}}}{r^2}\right)r^2+O(y^3)=\frac{1}{2}f(\mathbf{x}_0,\phi)r^2+O(y^3)
+\overline{t}=\frac{1}{2}\left(\Gamma^{0}_{\;ij}(0,\mathbf{x}_0)A^{i}_{\;\overline{i}}A^{j}_{\;\overline{j}}\frac{y^{\overline{i}} y^{\overline{j}}}{r^2}\right)r^2+O(y^3)=\frac{1}{2}f(\mathbf{x}_0,\phi)r^2+O(y^3)
 \ee
 where $\phi$ stands for all the angular coordinates $\phi_1,..,\phi_{d-2}$.
-The function $f(\mathbf{x}_0,\phi)$ depends on $\mathbf{x}_0$ as $\Gamma^{0}_{\;ij}$ and $A^{i}_{\;\overline{\emph\i}}$ will, in general, be different for a different point on the surface.
+The function $f(\mathbf{x}_0,\phi)$ depends on $\mathbf{x}_0$ as $\Gamma^{0}_{\;ij}$ and $A^{i}_{\;\overline{i}}$ will, in general, be different for a different point on the surface.
 One can see that $f(\mathbf{x}_0,\phi)$ depends on the angles, $\phi$, by substituting the relations below into (\ref{eq:RadialBottomSurface})
 
-\begin{equation}
-\begin{aligned}\label{eq:SphericalCoords}
-y^{\overline{1}} &= r \cos(\phi_1)  \\
-y^{\overline{2}} &= r \sin(\phi_1) \cos(\phi_2)  \\
-y^{\overline{2}} &= r \sin(\phi_1) \sin(\phi_2) \cos(\phi_3)  \\
-    &\vdots  \\
-y^{\overline{d-2}} &= r \sin(\phi_1) \cdots \sin(\phi_{d-3}) \cos(\phi_{d-2})  \\
+\begin{align}\label{eq:SphericalCoords}
+y^{\overline{1}} &= r \cos(\phi_1) \nonumber \\
+y^{\overline{2}} &= r \sin(\phi_1) \cos(\phi_2) \nonumber \\
+y^{\overline{2}} &= r \sin(\phi_1) \sin(\phi_2) \cos(\phi_3) \nonumber \\
+    &\vdots \nonumber \\
+y^{\overline{d-2}} &= r \sin(\phi_1) \cdots \sin(\phi_{d-3}) \cos(\phi_{d-2}) \nonumber \\
 y^{\overline{d-1}} &= r \sin(\phi_1) \cdots \sin(\phi_{d-3}) \sin(\phi_{d-2})
-\end{aligned}
-\end{equation}
+\end{align}
 
@@
 Substituting $r_{max}=T$ into (\ref{eq:VolumeIntegralSpherical}) means it can be evaluated to give
-\begin{figure}[t]
+
+\begin{figure}
   \centering
-    \includegraphics[scale=0.5]{coneplot}
+    \includegraphics[scale=0.6]{coneplot}
      \caption{The size of the region inside the top and bottom bounding surfaces is the volume we want to calculate.}
@@
 \end{figure}
+
 \be\label{eq:VolumeNoK}
 V(0,\mathbf{x}_0;T,\mathbf{x}_0)
-=\frac{A_{d-2}}{d(d-1)}T^d\left(1-\frac{d}{2(d+1)}\Gamma^{0}_{\;ij}(0,\mathbf{x}_0)A^{i}_{\;\overline{\emph\i}}A^{j}_{\;\overline{\emph\j}}\delta^{\overline{\emph\i}\overline{\emph\j}}T\right)
+=\frac{A_{d-2}}{d(d-1)}T^d\left(1-\frac{d}{2(d+1)}\Gamma^{0}_{\;ij}(0,\mathbf{x}_0)A^{i}_{\;\overline{i}}A^{j}_{\;\overline{j}}\delta^{\overline{i}\overline{j}}T\right)
 +O(T^{d+2})
 \ee
-where $A_{d-2}$ is the $d$-$2$ dimensional volume of a $d$-$2$ dimensional sphere and the $\delta^{\overline{\emph\i}\overline{\emph\j}}$ has come from the fact that cross terms ($\overline{\emph\i}\neq \overline{\emph\j}$) vanish under the angular integration.
-The defining relations for $A^{i}_{\;\overline{\emph\i}}$ can be rearranged to give $A^{i}_{\;\overline{\emph\i}}A^{j}_{\;\overline{\emph\j}}\delta^{\overline{\emph\i}\overline{\emph\j}}=h^{ij}(0,\mathbf{x}_0)$.
+where $A_{d-2}$ is the $d$-$2$ dimensional volume of a $d$-$2$ dimensional sphere and the $\delta^{\overline{i}\overline{j}}$ has come from the fact that cross terms ($\overline{i}\neq \overline{j}$) vanish under the angular integration.
+The defining relations for $A^{i}_{\;\overline{i}}$ can be rearranged to give $A^{i}_{\;\overline{i}}A^{j}_{\;\overline{j}}\delta^{\overline{i}\overline{j}}=h^{ij}(0,\mathbf{x}_0)$.
 In GNC it can be shown that the extrinsic curvature is given by
@@
 
-\subsection{Final Stretch of the Proof}
+\section{Final Stretch of the Proof}
 
@@
 We have
+
 \begin{gather}\label{eq:NmaxNminStart}
 \begin{aligned}
-\lim_{\rho \to \infty}\rho^{\frac{2}{d}-1}\left\langle N_{max}-N_{min} \right\rangle &= \\
-\lim_{\rho \to \infty}\rho^{\frac{2}{d}}
+\lim_{\rho \to \infty}\left\langle N_{max}-N_{min} \right\rangle &= \\
+\lim_{\rho \to \infty}\rho
 \int_{\Sigma}d^{d-1}x & \int_{-\varepsilon}^{0}dt\
@@
 \frac{1}{2}\frac{\dot{h}}{h}t-\rho B(-1)^{d}t^{d+1}+\rho O(t^{d+2})\right)e^{-\rho A(-1)^{d}t^{d}} \\
--\rho^{\frac{2}{d}}\int_{\Sigma}d^{d-1}x &
+-\rho\int_{\Sigma}d^{d-1}x &
 \int_{0}^{\varepsilon}dt\
@@
 where we have defined
+
 \begin{gather}\label{A_and_B_defn}
@@
 and we have Taylor expanded the $O(t^{d+1})$ parts of the volume functions, in small $t$, from the exponents.
-We have also included a pre-factor of $\rho^{\frac{2}{d}-1}$ to account for the factors of $l$ in (\ref{GH_boundary_to_causet}).
-With this pre-factor the limit of the above equation is well defined.
 If we change to $u=-t$ in the first part on the right hand side, and $u=t$ on the second part then the integral can be simplified to
+
 \be\label{eq:NmaxNminSimplified}
-\lim_{\rho \to \infty}\rho^{\frac{2}{d}}
+\lim_{\rho \to \infty}\left\langle N_{max}-N_{min} \right\rangle=
+\lim_{\rho \to \infty}\rho
 \int_{\Sigma}d^{d-1}x \int_{0}^{\varepsilon}du\
-h^{\frac{1}{2}}\left(-\frac{\dot{h}}{h}u+2\rho B u^{d+1}+\rho O(u^{d+2}) \right)e^{-\rho Au^{d}}
+h^{\frac{1}{2}}\left(-\frac{\dot{h}}{h}u+2\rho B u^{d+1}+\rho Cu^{d+2}+\rho O(u^{d+3}) \right)e^{-\rho Au^{d}}
 \ee
-We now show that $O(u^{d+2})$, or higher, terms vanish as $\rho \rightarrow\infty$.
-All higher order corrections that we have ignored so far contribute at $O(u^{d+2})$ or higher, so proving that these contributions vanish proves that our approximations are valid.
-Taking only the part we are interested in from (\ref{eq:NmaxNminSimplified}) leaves us with the task of showing that the following integral vanishes.
+where $C$ is the coefficient of $O(u^{d+2})$ terms.
+This term has been included as showing that it, and higher powers, vanish as $\rho \rightarrow\infty$ will prove our claim that no higher order terms are needed.
+To prove this we need to show that the following integral vanishes
 
 \be\label{eq:VanishingOrderIntegral}
-\lim_{\rho \to \infty}\rho^{\frac{2}{d}+1}\int_{0}^{\varepsilon}du\
+\lim_{\rho \to \infty}\rho^{\frac{2}{d}-1}\rho\int_{0}^{\varepsilon}du\
+\rho u^{d+n}e^{-\rho Au^{d}}=\lim_{\rho \to \infty}\rho^{\frac{2}{d}+1}\int_{0}^{\varepsilon}du\
 u^{d+n}e^{-\rho Au^{d}}
 \ee
-for $n>1$, where $n\in\mathbb{N}$.
-The additional factor of $\rho$ has come from inside the integral in (\ref{eq:NmaxNminSimplified}).
-We make the substitution $z=\rho Au^{d}$ and find that (\ref{eq:VanishingOrderIntegral}) is proportional to
+where $n\in\mathbb{N}$.
+We have left the powers of $\rho$ separated on the left side to show where they have come from.
+The first $\rho^{\frac{2}{d}-1}$ comes from (\ref{GH_boundary_to_causet}) and is needed to balance dimensions.
+This must be included in order to take the limit properly.
+We then make the substitution $z=\rho Au^{d}$ to transform the integral into that of an incomplete gamma function, which in the limit becomes a complete gamma function
 
 \be\label{eq:GammaVanishingOrderIntegral}
-\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\int_{0}^{\rho A\varepsilon^d}dz\;z^{\frac{n+1}{d}}e^{-z}=\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\int_{0}^{\infty}dz\;z^{\left[\left(\frac{n+1}{d}+1\right)-1 \right]}e^{-z}
+\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\frac{A^{\frac{n+1}{d}+1}}{d}\int_{0}^{\rho A\varepsilon^d}dz\
+z^{\frac{n+1}{d}}e^{-z}\sim\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\int_{0}^{\infty}dz\
+z^{\left[\left(\frac{n+1}{d}+1\right)-1 \right]}e^{-z}\sim\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\Gamma\left(\frac{n+1}{d}\right)
 \ee
-where $\rho$ has been taken to $\infty$ in the limits on the RHS, and the power of $z$ has been written in this way so that the integral is in the form of a gamma function, $\Gamma(t)=\int_{0}^{\infty}z^{t-1}e^{-z}dz$.
-Using the fact that for gamma functions, $\Gamma(t+1)=t\Gamma(t)$, we find that (\ref{eq:GammaVanishingOrderIntegral}) is then proportional to
-
-\be\label{eq:GammaVanishingOrder}
-\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\Gamma\left(\frac{n+1}{d}\right)
-\ee
+The $\rho$ in the limit of the second expression has been taken to $\infty$.
+The power of $z$ in this integral has been written in this manner so it can be more easily recognised as being of gamma function form, $\Gamma(t)=\int_{0}^{\infty}z^{t-1}e^{-z}dz$.
 One can see from above that if $n>1$ then this term will vanish, thus proving that terms of $O(u^{d+2})$ or higher will go to $0$ in the limit $\rho\rightarrow\infty$.
 
-The non-vanishing terms of (\ref{eq:NmaxNminSimplified}) can also be put into the form of gamma functions.
-This allows the integral over $u$ to be evaluated in (\ref{eq:NmaxNminSimplified}).
+We use this fact and the method of putting the integrals into a gamma function form to evaluate (\ref{eq:NmaxNminSimplified}).
 One also sees that we have terms like $\frac{\dot{h}}{h}$, and from (\ref{eq:K}) we know that these can be written in terms of the extrinsic curvature, $K(0,\mathbf{x})$.
@@
 \be\label{eq:nmax_eq:nmin_end}
-\lim_{\rho \to \infty}\rho^{\frac{2}{d}-1}\left\langle N_{max}-N_{min} \right\rangle=
-\frac{2(d+2)}{d(d+1)}
+\lim_{\rho \to \infty}\left\langle N_{max}-N_{min} \right\rangle=\lim_{\rho \to \infty}
+\rho^{1-\frac{2}{d}}\frac{2(d+2)}{d(d+1)}
 \left[\frac{A_{d-2}}{d(d-1)}\right]^{-\frac{2}{d}}
 \Gamma\left(\frac{2}{d}\right)\int_{\Sigma}d^{d-1}x\
-\sqrt{h}K(0,\mathbf{x}).
+\sqrt{h}K(0,\mathbf{x})
 \ee
 $K(0,\mathbf{x})$ is evaluated at the surface and is therefore the same $K$ as in equation (\ref{GH_boundary_to_causet}).
-We can therefore conclude that to make (\ref{eq:mainconjecture}) true we need a factor $l_{p}^{2-d}$ to make our expression dimensionless, and a value of $c_d$ equal to that of equation (\ref{Cn}).
-This concludes the proof of the conjectured equality in (\ref{eq:mainconjecture}).
-
-\section{Fluctuations}
-So far we have only talked about the mean of the causal set boundary action.
-Let us now turn to its fluctuations (the standard deviation) $\sigma[S^{(d)}_{GHY}(\mathcal C)]=\text{Var}[S^{(d)}_{GHY}(\mathcal C)]^\frac12$.\footnote{Whenever we say fluctuations we refer to the standard deviation of the random variable, not to its variance.} We can make a heuristic argument to estimate the dependence of fluctuations on $\rho$.
-In any spacetime region of fixed volume $V$ the number of causal set elements $N$ experiences Poisson fluctuations of order $\sqrt N$.
-Since $N_{max}$ and $N_{min}$ are quantities associated with a codimension-1 surface (counting elements ``near" $\Sigma$), we may expect them to inherit fluctuations of order $N^\frac{d-1}{2d}$.
-Then the action $S^{(d)}_{GHY}$, being proportional to $\rho^\frac{d-2}{d}$ times the difference of two independent\footnote{The independence is as a feature of the Poisson process in $M$: the number of elements $N(R)$ in any subregion $R\subset M$ is a Poisson variable itself, and for any two disjoint regions $R_1$, $R_2$ the numbers $N(R_1)$ and $N(R_2)$ are independent random variables.} random variables with standard deviation $N^\frac{d-1}{2d} = (\rho V)^\frac{d-1}{2d}$ should see fluctuations of order $\rho^\frac{2-d}{d}\rho^\frac{d-1}{2d}=\rho^\frac{d-3}{2d}$.
-This suggests that for $d=2$ these fluctuations should grow as $\rho\rightarrow\infty$, for $d=3$ they should be constant, and for $d>3$ they should be damped.
-
-In order to test this heuristic argument one may like to write down the integral expression for $\text{Var}[S^{(d)}_{GHY}(\mathcal C)]^\frac12$ but it is complicated enough even in finite regions of flat space that it is not very illuminating to reproduce it here.
-Instead we will show the results of computer simulations that support results of the heuristic argument.
-
-The simulations were carried out as follows.
-Denote the discreteness scale by $l$.
-Take a $d$-cube $[0,L]^d$ in $d$-dimensional Minkowski space with metric $ds^2=-dt^2+d{\mathbf x}^2$ and define the hypersurface $\Sigma: t=L/2$, which partitions the cube into two halves.
-Sprinkle at density $\rho=l^{-d}$ into the cube, which in any given run will place $N$ points inside the cube where $N$ is a Poisson random number with mean $\left\langle N\right\rangle = \rho V=  (L/l)^d$.
-Evaluate $S^{(d)}_{GHY}=c_d\rho^\frac{d-2}{d}(N_{min} - N_{max})$ for $\Sigma$ by counting the minimal/maximal elements in the upper/lower half of the cube, and repeat.
-
-\bibliographystyle{utphys}
-
-\bibliography{references}{}
+We can therefore conclude that to make (\ref{GH_boundary_to_causet}) true we need a factor $\rho^{\frac{2}{d}-1}$ to cancel the factors of $\rho$ above, and a value of $c_d$ equal to that of equation (\ref{Cn}).
+The factor of $\rho$ needed is exactly that required on dimensional grounds, thus we have proved that (\ref{GH_boundary_to_causet}) is in fact true in general.
 
 \end{document}
-
-%%
-JHEPPUB:
-
-\title{Gibbons-Hawking-York Boundary Term in Causal Sets}
- \author[a]{M. Buck}
- \author[a,b]{\!, F. Dowker}
- \author[a]{and I. Jubb\,}
-\affiliation[a]{Theoretical Physics Group, Blackett Laboratory, Imperial College, London, SW7 2AZ, U.K.}
-\affiliation[b]{Institute for Quantum Computing, University of Waterloo, ON, N2L 2Y5, Canada}
-
-\abstract{ 
-We show some stuff.
-}
-
-\begin{document}
-
-\maketitle
-
-\pagebreak
-
-\noindent NOTES:\\
-1. define s as the RV, not as its mean
-
-\section{Introduction}
-
-%%IOPART
-
-\begin{document}
-
-\title{Gibbons-Hawking-York Boundary Term in Causal Sets}
-
-\author{Michel Buck Fay Dowker, Ian Jubb}
-\address{Blackett Laboratory, Imperial College, London, SW7 2AZ, U.K.}
-
-\begin{abstract}
-
-We show some stuff.
-
-\end{abstract}
-%\pacs{03.67.-a, 03.65.Ta, 03.70.+k}
-
-\maketitle
-\section{Introduction}
-%%

### 2014-09-18 paper_v2_backup_20140520.tex
initial commit

--- 2014-09-18 paper_v2.tex
@@
+% CURRENTLY BEING EDITED OFF-LINE BY: IAN
+
 \documentclass[12pt]{article}
@@
 
-\section{Introduction}
-
-First and foremost one must draw attention to the fact that Michel Buck (on of the authors of this paper) is in fact a smelly son of bitch.
-Gerrrl, he so ugly he look like someone tried to put a fire out on his face with a fork!
-
-In furthering causal set theory it is crucial that we understand the kinematics of the theory.
-The action of a given causal set is one such kinematical property that would be extremely useful to know and understand.
-Proposals for the action of a causal set are available [[REFERENCE]] and these hold analytically in some cases and numerically in many more.
-These cases being when the causal set is embeddable in some existing spacetime.
-One can then show that the action of the causal set then agrees with the Einstein-Hilbert action of the spacetime in some continuum limit.
-How this limiting procedure is carried out will be described in more detail below.
-
-The EH action, however, is not the complete story in the continuum.
-
-If one has a spacetime manifold with a boundary, such as Minkowski space but cut off at some finite time, then one needs to include the Gibbons-Hawking-York boundary term (GHYBT) in the Einstein-Hilbert action.
-This term is needed to keep the variational principle well defined.
-In this section we prove a conjecture for the form of this boundary term in causal set theory.
-
-\section{The Claim}
-
+\section{The main result for spacelike surfaces}
+\begin{enumerate}
+\item MB1: need to justify why higher order terms can be ignored.
+\item MB1: why the threefold coordinate transformation? why no just RNCs at the base?
+\item MB1: 
+\end{enumerate}
 Consider a sufficiently well-behaved spacetime $(M,g)$ and Cauchy surface $\Sigma$ in $M$.
@@
 $$
-S_{GHY} = \pm\frac1{\kappa^2} \int_{\Sigma} d^{d-1}x \sqrt{h} K
+S_{GHY} = \pm\frac1{\kappa^2} \int_{\Sigma} d^{d-1}x \sqrt{-h} K
 $$
-where $K$ is the trace of the extrinsic curvature $K_{\mu\nu}=h_{\mu}^\rho h_\nu^\sigma \nabla_\rho n_\sigma$ of $\Sigma$ defined with future-pointing timelike unit normal $n_{\mu}=\partial_\mu S/\sqrt{g^{\mu\nu}\partial_\mu S\partial_\nu S}$.
+where $K$ is the trace of the extrinsic curvature $K_{\mu\nu}=h_{\mu}^\rho h_\nu^\sigma \nabla_\rho n_\sigma$ of $\Sigma$ defined with future-pointing timelike unit normal $n$ \IGNORE{\partial_\mu S/\sqrt{g^{\mu\nu}\partial_\mu S\partial_\nu S}$}.
 
-The extrinsic curvature of a spacelike hypersurface can be identified with the volume gradient across that hypersurface.
+The extrinsic curvature of a spacelike hypersurface can be identified with the volume gradient across that hypersurface [[MB1: HOW?]].
 On the causal set, the analogue of a spacelike hypersurface is a maximal anti-chain, i.e.
@@
 Let us denote the number of maximal elements in $\mathcal C^-$ and the number of minimal elements in $\mathcal C^+$ by $N_{min}$ and $N_{max}$, respectively.
+%\IGNORE{\be
+%\begin{aligned}
+%N_{min}=|\left\{x\in\mathcal C^+:\nexists\;y\in\mathcal C^+ s.t. \right\}|
+%\end{aligned}
+%\ee} 
 We propose the following definition for the discrete Gibbons-Hawking-York boundary term for a causal set $\mathcal C$ in $d$ spacetime dimensions:
@@
 \be
-\lim_{\rho\rightarrow\infty}S^{(d)}_{GHY}[\mathcal C] = \frac1{\kappa^2} \int_{\Sigma} d^{d-1}x \sqrt{h} K.\label{eq:mainconjecture}
+\lim_{\rho\rightarrow\infty}S^{(d)}_{GHY}[\mathcal C] = \frac1{\kappa^2} \int_{\Sigma} d^{d-1}x \sqrt{-h} K.\label{eq:mainconjecture}
 \ee
@@
 These approximations are justified on the grounds that if certain orders are retained throughout the calculation they will in fact vanish in the limit $\rho \rightarrow \infty$
-
-\section{Poisson Sprinklings for $\langle N_{max}\rangle$ and $\langle N_{min}\rangle$}
 
@@
 
+[[IJ: NEED A FIGURE FOR THIS NEXT BIT]]
+
 Both quantities will diverge, but if their difference $\langle N_{max}\rangle - \langle N_{min}\rangle = \langle N_{max} - N_{min}\rangle$ grows slower than or at order $\rho^{1-\frac2d}$, the proposed action~\eqref{GH_boundary_to_causet} will tend to a finite value in the continuum limit.
-Figure \ref{fig:Nmin_Nmax} shows a 2-dimensional spacetime with an extrinsically curved spacelike surface (a line).
-The minimal and maximal points have been highlighted and the shaded regions correspond to those that  we wish to find the volumes of.
-This is a simple task here as the example is only 2D and the spacetime is flat.
-
-\begin{figure}
-  \centering
-    \includegraphics[width=\textwidth]{minmaxplot}
-     \caption{Minimal and maximal points about a surface in 2D Minkowski space. The shaded regions illustrate the regions that one must calculate the volume of.}
-     \label{fig:Nmin_Nmax}
-\end{figure}
 
@@
 The metric determinant has been in expanded in small $t$.
-The volume functions have been rewritten as follows: $V(x,\Sigma)=V(x,x_0)=V(-t,\mathbf{x};0,\mathbf{x})$ and $V(\Sigma,q)=V(x_0,x)=V(0,\mathbf{x};t,\mathbf{x})$
+The volume functions have been rewritten as follows: $V(x,\Sigma)=V(p,p_0)=V(-t,\mathbf{x};0,\mathbf{x})$ and $V(\Sigma,q)=V(q_0,q)=V(0,\mathbf{x};t,\mathbf{x})$
 
-The only curvy cones that will contribute will have small volumes, as we are restricting ourselves close to the surface.
-This means that the volumes can be approximated by the volumes of less curvy cones, and it can be shown that higher order corrections, in the approximation scheme, vanish in the limit of $\rho \rightarrow \infty$.
-This will all be made more precise below.
-The approximation procedure will be outlined below for a cone to the future of $\Sigma$, as the process for the past cone is nearly identical.
+[[OLD SECTION THAT MB WANTED TO REWRITE]]
+[[MB: Need to rewrite this paragraph]]. 
+The volumes will depend on the proper time elapsed along the unique geodesic from a point in $\Sigma$ to a point $p$ such that the spatial coordinates remain unchanged.
+This is shown in Fig.
+.
+The coordinate time, $t$, is in fact the same as the proper time in the GNC we have chosen.
+Equations (\ref{eq:nmax}) and (\ref{eq:nmin}) can then be simplified as any time, $t$, which is far away from the surface will give rise to a large volume, which will then be exponentially suppressed as $\rho \rightarrow \infty$.
+We can choose a distance $\varepsilon$ away from the surface such that beyond it all volumes will be suppressed in the final limit.
+(\ref{eq:nmax}) and (\ref{eq:nmin}) are now given by
+
+\begin{align}
+\left\langle N_{max}\right\rangle & =\int_{\Sigma}d^{d-1}x\int_{-\varepsilon}^{0}dt\:
+h^{\frac{1}{2}}\left(1+
+\frac{1}{2}\frac{\dot{h}}{h}t+O(t^2)\right)
+ \rho\ e^{-\rho V(t,\mathbf{x};0,\mathbf{x})}
+\\
+\left\langle N_{min}\right\rangle & =\int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
+h^{\frac{1}{2}}\left(1+
+\frac{1}{2}\frac{\dot{h}}{h}t+O(t^2)\right) \rho\ e^{-\rho V(0,\mathbf{x};t,\mathbf{x})}
+\end{align}
+where $h\equiv det\left(h_{ij}(0,\mathbf{x})\right)$ and $\dot{}\equiv \frac{\partial}{\partial t}$. The metric determinant has been in expanded in small $t$.
+[[END OF OLD SECTION]]
+
+The volumes can also be approximated using the fact that the tips of the cones are close to the surface, as $|t|\leq\varepsilon$ in the integral limits.
+The process for utilising the possible simplifications will be outlined below for the cone to the future of $\Sigma$, as the process for the past cone is nearly identical.
 Riemann normal coordinates are integral to the following calculation so it is worth noting the relevant formulas before jumping in.
 
-\section{Riemann Normal Coordinates}
+The volumes of the truncated lightcones can be found as follows.
+Consider the past lightcone emanating at the point $p_0=(t_0,\mathbf x_0)$.
+There is a unique point $q_0=(0,\mathbf x_0)$ on $\Sigma$ associated with $p_0$, separated from $p_0$ by a proper time $t_0$.
+Now change  spatial coordinates to primed coordinates $(t',\mathbf{x} ')$, such that the spatial metric at $P$ is flat.
+This is equivalent to using Riemann normal coordinates (RNC), but just in the spatial dimensions.
+We require
+\be\label{metric_flat_conditions}
+h_{i'j'}(p_0)=\delta_{ij}  \;\;\text{and}\;\; \Gamma^{i'}_{j'k'}(p_0)=0.
+\ee
+This of course still allows for $\Gamma^{0}_{ij}(P)\neq0$.
+The surface $\Sigma$ still corresponds to $t'=t=0$ after this coordinate change.
+We now change from these coordinates to full spacetime RNC to calculate the volume.
+The new coordinates, $z^\mu$, are related to the previous coordinates via
 
-One can always find Riemann normal coordinates (RNC), about a point $p$, by changing from coordinates $x^{\alpha}$ to coordinates $y^{\overline{\alpha}}$ such that in these new coordinates the metric and the Christoffel symbols at $p$ are that of flat space.
-These conditions  can be expressed as follows
+\be\label{RNC_to_GNC_flat}
+z^\mu=\delta^\mu_\nu y^\nu+\frac{1}{2}\delta^{\tilde{\mu}}_{\mu '}\Gamma^{\mu '}_{\alpha '\beta '}(P)x'^{\alpha '}x'^{\beta '}+O(x'^3)
+\ee
+This is the standard transformation to RNC. The inverse relations are given by
 
-\be\label{eq:RNCMetricTransAtPAndChris}
-g_{\overline{\alpha} \overline{\beta}}(p)=\eta_{\overline{\alpha} \overline{\beta}}=A^{\alpha}_{\;\overline{\alpha}}A^{\beta}_{\;\overline{\beta}}g_{\alpha\beta}(p)\;,\;\;\;\;\Gamma^{\overline{\alpha}}_{\;\overline{\beta}\overline{\gamma}}(p)=0
-\ee
-The $A^{\alpha}_{\;\overline{\alpha}}$ govern the coordinate transformation to linear order, but $O(x^2)$ corrections may be required. To second order the coordinate transformation is given by
-
-\be\label{eq:RNCtotaltrans}
-y^{\overline{\alpha}}=A^{\overline{\alpha}}_{\;\beta}x^\beta+\frac{1}{2}A^{\overline{\alpha}}_{\;\alpha}\Gamma^{\alpha}_{\;\beta\gamma}(p)x^\beta x^\gamma+O(x^3)
-\ee
-and the inverse transformation is just
-
-\be\label{eq:RNCinversetrans}
-x^{\alpha}=A^{\alpha}_{\;\overline{\beta}}y^{\overline{\beta}}+O(y^2)
-\ee
-One finds that the $A^{\alpha}_{\;\overline{\alpha}}$ also satisfy
-
-\be\label{eq:RNCeqnforA}
-A^{\overline{\alpha}}_{\;\mu}A^{\mu}_{\;\overline{\beta}}=\delta^{\overline{\alpha}}_{\;\overline{\beta}}\;,\;\;\;\;A^{\alpha}_{\;\overline{\mu}}A^{\overline{\mu}}_{\;\beta}=\delta^{\alpha}_{\;\beta}
-\ee
-These relations for RNC will cover all that will be needed in this discussion.
-
-\section{Lightcone Volumes}
-
-The volumes of the truncated lightcones can be found as follows.
-Consider the past lightcone emanating at the point $p_0=(T,\mathbf x_0)$.
-There is a unique point $q_0=(0,\mathbf x_0)$ on $\Sigma$ associated with $p_0$, separated from $p_0$ by a proper time $T$.
-
-The volume of this cone is given by
-
-\be\label{eq:VolumeWithNoSimplifications}
-V(0,\mathbf{x}_0;T,\mathbf{x}_0)=\int_{\mathcal{X}} d^d x\;\sqrt{-g}
-\ee
-where the integration region, $\mathcal{X}:= J^-(p_0)\bigcap J^+(\Sigma)$ will be some complicated thing.
-As we are dealing with points close to the surface we can use the  Riemann normal coordinates $y^{\overline{\mu}}$, defined as above, about the point $q_0$ (the point $q_0$ being the origin of our RNC system).
-One finds that $A^0_{\;\overline{0}}=1$, $A^0_{\;i}=0$ and $\delta_{\overline{i}\overline{j}}=A^i_{\;\overline{i}}A^j_{\;\overline{j}}h_{ij}(0,\mathbf{x}_0)$.
-It can be shown that the volume integral reduces to \cite{Khetrapal_Sumati:Causal_Diamond_Volume}
-
-\be\label{eq:VolumeWithRNC}
-V(0,\mathbf{x}_0;T,\mathbf{x}_0) =\int_{\mathcal{X}}d^dy+\int_{\mathcal{X}_0}d^dy\left(-\frac{1}{6}R_{\overline{\mu}\overline{\nu}}(q_0)y^{\overline{\mu}}y^{\overline{\nu}} \right)+O(T^{d+3})
-\ee
-where $\mathcal{X}_0=\left\lbrace y^{\overline{\mu}} \mid 0\leq y^{\overline{0}}=\overline{t}\leq T\; ,\; \sqrt{(y^{\overline{1}})^2+(y^{\overline{2}})^2+...+(y^{\overline{d-1}})^2}\leq T-\overline{t} \right\rbrace$ is what we call the \textit{flat cone}, and $R_{\overline{\mu}\overline{\nu}}(q_0)$ is the Ricci tensor in RNC evaluated at $q_0$.
-The arguments of the volume function need not be changed as the volume only depends on the points in the manifold, $q_0$ and $p_0$.
-The \textit{flat cone} term comes in at $O(T^{d+2})$ so the volume we have to calculate has reduced to
-
-\be\label{eq:VolumeToLowestOrder}
-V(0,\mathbf{x}_0;T,\mathbf{x}_0) =\int_{\mathcal{X}}d^dy+O(T^{d+2})
-\ee
-Terms of $O(T^{d+2})$ can be retained till the end, but in the limit, $\rho\rightarrow\infty$, they vanish. This will be proved below.
-
-The boundaries of $\mathcal{X}$ must now be figured out in order to do this integral.
-First we look at the top part of the cone.
-From \cite{Khetrapal_Sumati:Causal_Diamond_Volume} it can be shown that the first curvature correction to the top part of the cone comes in at $O(T^{d+2})$ and so can be ignored for our purposes.
-This means that the top can be treated as a that of a \textit{flat cone} and so points on the boundary obey the equation $\sqrt{(y^{\overline{1}})^2+(y^{\overline{2}})^2+...+(y^{\overline{d-1}})} = T-\overline{t}$.
-
-The bottom surface of the cone was defined in terms of GNC to be $t=0$, so we can use (\ref{eq:RNCtotaltrans}) to find the equation for the surface in RNC. Equation (\ref{eq:RNCtotaltrans}) gives
-
-\be\label{eq:BottomSurfaceWithGNC}
-y^{\overline{0}}=\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij}(0,\mathbf{x}_0)x^i x^j+O(x^3)
-\ee
-The first part on the right vanishes as $A^{\overline{0}}_{\;\mu}x^{\mu}=x^0$ (as $A^{\overline{0}}_{\;i}=0$ and $A^{\overline{0}}_{\;0}=1$) and $x^0=t=0$ for the bottom surface.
-Using the inverse RNC relation (\ref{eq:RNCinversetrans}) one can find the equation for the bottom surface entirely in RNC.
-
-\be\label{eq:BottomSurface}
-\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij}(0,\mathbf{x}_0)A^{i}_{\;\overline{i}}A^{j}_{\;\overline{j}}y^{\overline{i}} y^{\overline{j}}+O(y^3)
+\be\label{inverse_trans}
+x'^{\mu '}=\delta^{\mu '}_{\tilde{\mu}}\tilde{x}^{\tilde{\mu}}+O(\tilde{x}^2)
 \ee
 
-This equation can be written in terms of the radial coordinate of the cone, $r=\sqrt{(y^{\overline{1}})^2+...+(y^{\overline{d-1}})^2}$, as
+Using (\ref{inverse_trans}) in (\ref{RNC_to_GNC_flat})
+ allows one to find the equation for the surface in RNC. It is given by
 
-\be\label{eq:RadialBottomSurface}
-\overline{t}=\frac{1}{2}\left(\Gamma^{0}_{\;ij}(0,\mathbf{x}_0)A^{i}_{\;\overline{i}}A^{j}_{\;\overline{j}}\frac{y^{\overline{i}} y^{\overline{j}}}{r^2}\right)r^2+O(y^3)=\frac{1}{2}f(\mathbf{x}_0,\phi)r^2+O(y^3)
+\be\label{RNC_surface}
+\tilde{t}=\frac{1}{2}\Gamma^{t'}_{i'j'}(P)\delta^{i'}_{\tilde{i}}\delta^{j'}_{\tilde{j}}
+\tilde{x}^{\tilde{i}}\tilde{x}^{\tilde{j}}+O(\tilde{x}^3)
 \ee
-where $\phi$ stands for all the angular coordinates $\phi_1,..,\phi_{d-2}$.
-The function $f(\mathbf{x}_0,\phi)$ depends on $\mathbf{x}_0$ as $\Gamma^{0}_{\;ij}$ and $A^{i}_{\;\overline{i}}$ will, in general, be different for a different point on the surface.
-One can see that $f(\mathbf{x}_0,\phi)$ depends on the angles, $\phi$, by substituting the relations below into (\ref{eq:RadialBottomSurface})
+when $t=t'=0$. It is a quadratic in the spatial variables.
 
-\begin{align}\label{eq:SphericalCoords}
-y^{\overline{1}} &= r \cos(\phi_1) \nonumber \\
-y^{\overline{2}} &= r \sin(\phi_1) \cos(\phi_2) \nonumber \\
-y^{\overline{2}} &= r \sin(\phi_1) \sin(\phi_2) \cos(\phi_3) \nonumber \\
-    &\vdots \nonumber \\
-y^{\overline{d-2}} &= r \sin(\phi_1) \cdots \sin(\phi_{d-3}) \cos(\phi_{d-2}) \nonumber \\
-y^{\overline{d-1}} &= r \sin(\phi_1) \cdots \sin(\phi_{d-3}) \sin(\phi_{d-2})
-\end{align}
+Following (Sumati paper) we find the equation for the top part of the cone [[MB: need to define what you mean]] is simply
 
-With the boundaries of the integration region in place, we can now write down the integral explicitly, in spherical coordinates:
+\be\label{top_of_cone_eqn}
+\tilde{t}=-\tilde{r}+T
+\ee
+to the order we require.
+If we kept higher orders the top part would have an elliptic cross section.
+The volume would then acquire terms of $O(T^{n+2})$ which are suppressed in the final limit.
+The shape of the cone has been altered from the curvy cone to the one shown in Fig.
 
-\be\label{eq:VolumeIntegralSpherical}
-V(0,\mathbf{x}_0;T,\mathbf{x}_0)=\int_{S^{d-2}}
-d\Omega_{d-2}
-\int_{0}^{r_{max}(\phi)}r^{d-2}dr
-\int_{\frac{1}{2}f(\mathbf{x}_0,\phi)r^2}^{-r+T}
-d\overline{t}+O(T^{d+2})
+We now have our integral boundaries and we can attempt the integral. From (Sumati paper) the volume of the top cone can be written as [[MB1: integration region is wrong]]
+
+\begin{gather}\label{volume_from_sumati}
+\begin{aligned}
+V(0,\mathbf{x}_0;T,\mathbf{x}_0) & =\int_{J^{+}(\Sigma)\bigcap J^{-}(T,\mathbf{\tilde{x}}_0)}d^n\tilde{x}\;  \sqrt{-g} \\
+& =\int_{J^{+}(0,\mathbf{\tilde{x}})\bigcap J^{-}(T,\mathbf{\tilde{x}}_0)}d^n\tilde{x}\;\\
+ & +\int_{flat\;cone}d^n\tilde{x}\left(-\frac{1}{6}\tilde{x}^{\tilde{c}}\tilde{x}^{\tilde{d}}
+R_{\tilde{c}\tilde{d}}(0,\mathbf{\tilde{x}}_0)\right)
+\end{aligned}
+\end{gather}
+where the flat cone is the cone you would have in a flat spacetime [[MB1: need to specify what you mean]].
+The flat cone term comes in at $O(T^{n+2})$ and so can be ignored.
+The origins of the coordinate systems agree, as do the points at the tips of the cones in each coordinate system, thus the volume can be a function of the original GNC.
+The integral boundaries of the first term can be approximated from equations (\ref{RNC_surface}) and (\ref{top_of_cone_eqn}).
+Equation (\ref{top_of_cone_eqn}) can be written in the radial coordinate $\tilde{r}^2={\tilde{x}_{1}}^2+...+{\tilde{x}_{d-1}}^2$ as
+
+\be\label{RNC_surface_in_r}
+\tilde{t}=\frac{1}{2}\left(\Gamma^{t'}_{i'j'}(0,\mathbf{x}_0)\delta^{i'}_{\tilde{i}}
+\delta^{j'}_{\tilde{j}}
+\frac{\tilde{x}^{\tilde{i}}
+\tilde{x}^{\tilde{j}}}{{\tilde{r}}^2}\right)
+{\tilde{r}}^2
+=\frac{1}{2}f(\tilde{\phi}){\tilde{r}}^2
 \ee
-where $r_{max}(\phi)$ is the radial coordinate of where the bottom surface meets the top part of the cone at an angle $\phi$, shown in \ref{fig:cone_plot}.
-To find this one needs to solve $\frac{1}{2}f(\mathbf{x}_0,\phi){r_{max}}^2(\phi)=-r_{max}(\phi)+T$ and take the positive solution.
-The solution can be expanded in $T$ and is simply $r_{max}=T+O(T^2)$, with angular dependent terms contributing at $O(T^2)$.
-The $O(T^2)$ term will contribute at $O(T^{d+2})$ in the integral and so can be ignored.
-Substituting $r_{max}=T$ into (\ref{eq:VolumeIntegralSpherical}) means it can be evaluated to give
+where $f(\tilde{\phi})$ only depends on the angular coordinates $\tilde{\phi}_1,..,\tilde{\phi}_{n-2}$.
 
-\begin{figure}
-  \centering
-    \includegraphics[scale=0.6]{coneplot}
-     \caption{The size of the region inside the top and bottom bounding surfaces is the volume we want to calculate.}
-     \label{fig:cone_plot}
-\end{figure}
+The volume then becomes
 
-\be\label{eq:VolumeNoK}
+\be\label{volume_with_approx}
+V(0,\mathbf{x}_0;T,\mathbf{x}_0)=\int_{S^{n-2}}
+d\tilde{\Omega}_{n-2}
+\int_{0}^{T}\tilde{r}^{n-2}dr
+\int_{\frac{1}{2}f(\tilde{\phi})\tilde{r}^2}^{-\tilde{r}+T}
+d\tilde{t}
+\ee
+where the max limit on $\tilde{r}$ comes from solving the equation $\frac{1}{2}f(\tilde{\phi})\tilde{r}^2_{max}=
+-\tilde{r}_{max}+T$ and expanding in $T$ to the lowest order necessary. Evaluating this integral gives
+
+\be\label{volume_no_K}
 V(0,\mathbf{x}_0;T,\mathbf{x}_0)
-=\frac{A_{d-2}}{d(d-1)}T^d\left(1-\frac{d}{2(d+1)}\Gamma^{0}_{\;ij}(0,\mathbf{x}_0)A^{i}_{\;\overline{i}}A^{j}_{\;\overline{j}}\delta^{\overline{i}\overline{j}}T\right)
-+O(T^{d+2})
+=\frac{A_{n-2}}{n(d-1)}T^n\left(1-\frac{n}{2(n+1)}\Gamma^{t'}_{i'j'}(0,\mathbf{x}_0)\delta^{i'j'}T\right)
++O(T^{n+2})
 \ee
-where $A_{d-2}$ is the $d$-$2$ dimensional volume of a $d$-$2$ dimensional sphere and the $\delta^{\overline{i}\overline{j}}$ has come from the fact that cross terms ($\overline{i}\neq \overline{j}$) vanish under the angular integration.
-The defining relations for $A^{i}_{\;\overline{i}}$ can be rearranged to give $A^{i}_{\;\overline{i}}A^{j}_{\;\overline{j}}\delta^{\overline{i}\overline{j}}=h^{ij}(0,\mathbf{x}_0)$.
-In GNC it can be shown that the extrinsic curvature is given by
+We observe that $\Gamma^{t'}_{i'j'}(0,\mathbf{x}_0)\delta^{i'j'}$ is written in GNC with spatial flatness at $P$ (the $primed$ indices), although it is a scalar and so is independent of coordinate choice.
+The extrinsic curvature of the surface in these coordinates is given by
 
-\be\label{eq:K}
+\be\label{K_eqn}
 K(0,\mathbf{x}_0)
-=g^{\mu\nu }\nabla_{\mu}n_{\nu}
-=-\Gamma^{0}_{\;ij}(0,\mathbf{x}_0)h^{ij}(0,\mathbf{x}_0)=-\frac{1}{2}\frac{\dot{h}(0,\mathbf{x}_0)}{h(0,\mathbf{x}_0)}
+=g^{\mu '\nu '}\nabla_{\mu '}n_{\nu '}
+=-\Gamma^{t'}_{i'j'}(0,\mathbf{x}_0)\delta^{i'j'}
 \ee
-which means that the first correction to the volume of the cone depends on $K(0,\mathbf{x}_0)$, and thus the approximate volume of the cone is
+where $n_{\nu '}=(1,0,0)$ is the surface normal. Thus we can directly substitute in for $K$ in (\ref{volume_no_K}) to obtain
 
@@
 V(0,\mathbf{x}_0;T,\mathbf{x}_0)
-&=\frac{A_{d-2}}{d(d-1)}T^d\left(1+\frac{d}{2(d+1)}K(0,\mathbf{x}_0)T\right)
-+O(T^{d+2}) \label{eq:TopVolumeWithK}\\
+&=\frac{A_{n-2}}{n(d-1)}T^n\left(1+\frac{n}{2(n+1)}K(0,\mathbf{x}_0)T\right)
++O(T^{n+2}) \label{top_volume_with_K}\\
 V(-T,\mathbf{x}_0;0,\mathbf{x}_0)
-&=\frac{A_{d-2}}{d(d-1)}T^d\left(1-\frac{d}{2(d+1)}K(0,\mathbf{x}_0)T\right)
-+O(T^{d+2}) \label{eq:BottomVolumeWithK}
+&=\frac{A_{n-2}}{n(d-1)}T^n\left(1-\frac{n}{2(n+1)}K(0,\mathbf{x}_0)T\right)
++O(T^{n+2}) \label{bottom_volume_with_K}
 \end{align}
-where (\ref{eq:BottomVolumeWithK}) is the volume of the cone with its tip below the surface, and can be found from similar arguments.
-
-\section{Final Stretch of the Proof}
+where (\ref{bottom_volume_with_K}) is the volume for the bottom cone, and can be found from similar arguments.
+These results do not disagree with (Sumati paper or Myrhiem), which states that the first correction to a small causal diamond is $O(T^{n+2})$.
+The total volume for a causal diamond would be a sum of (\ref{top_volume_with_K}) and (\ref{bottom_volume_with_K}), and therefore $O(T^{n+1})$ terms would cancel.
 
@@
 
-\begin{gather}\label{eq:NmaxNminStart}
+\begin{gather}\label{eq:nmax_eq:nmin_start}
 \begin{aligned}
@@
 h^{\frac{1}{2}}\left(1+
-\frac{1}{2}\frac{\dot{h}}{h}t-\rho B(-1)^{d}t^{d+1}+\rho O(t^{d+2})\right)e^{-\rho A(-1)^{d}t^{d}} \\
+\frac{1}{2}\frac{\dot{h}}{h}t-\rho B(-1)^{n}t^{n+1}\right)e^{-\rho A(-1)^{n}t^{n}} \\
 -\rho\int_{\Sigma}d^{d-1}x &
@@
 h^{\frac{1}{2}}\left(1+
-\frac{1}{2}\frac{\dot{h}}{h}t-\rho Bt^{d+1}+\rho O(t^{d+2})\right)e^{-\rho At^{d}}
+\frac{1}{2}\frac{\dot{h}}{h}t-\rho Bt^{n+1}\right)e^{-\rho At^{n}}
 \end{aligned}
@@
 \begin{aligned}
-A & \equiv \frac{A_{d-2}}{d(d-1)} \\
-B & \equiv \frac{A_{d-2}}{2(d-1)(d+1)}K(0,\mathbf{x})
+A & \equiv \frac{A_{n-2}}{n(d-1)} \\
+B & \equiv \frac{A_{n-2}}{2(d-1)(d+1)}K(0,\mathbf{x})
 \end{aligned}
 \end{gather}
-and we have Taylor expanded the $O(t^{d+1})$ parts of the volume functions, in small $t$, from the exponents.
-If we change to $u=-t$ in the first part on the right hand side, and $u=t$ on the second part then the integral can be simplified to
-
-\be\label{eq:NmaxNminSimplified}
-\lim_{\rho \to \infty}\left\langle N_{max}-N_{min} \right\rangle=
-\lim_{\rho \to \infty}\rho
-\int_{\Sigma}d^{d-1}x \int_{0}^{\varepsilon}du\
-h^{\frac{1}{2}}\left(-\frac{\dot{h}}{h}u+2\rho B u^{d+1}+\rho Cu^{d+2}+\rho O(u^{d+3}) \right)e^{-\rho Au^{d}}
-\ee
-where $C$ is the coefficient of $O(u^{d+2})$ terms.
-This term has been included as showing that it, and higher powers, vanish as $\rho \rightarrow\infty$ will prove our claim that no higher order terms are needed.
-To prove this we need to show that the following integral vanishes
-
-\be\label{eq:VanishingOrderIntegral}
-\lim_{\rho \to \infty}\rho^{\frac{2}{d}-1}\rho\int_{0}^{\varepsilon}du\
-\rho u^{d+n}e^{-\rho Au^{d}}=\lim_{\rho \to \infty}\rho^{\frac{2}{d}+1}\int_{0}^{\varepsilon}du\
-u^{d+n}e^{-\rho Au^{d}}
-\ee
-where $n\in\mathbb{N}$.
-We have left the powers of $\rho$ separated on the left side to show where they have come from.
-The first $\rho^{\frac{2}{d}-1}$ comes from (\ref{GH_boundary_to_causet}) and is needed to balance dimensions.
-This must be included in order to take the limit properly.
-We then make the substitution $z=\rho Au^{d}$ to transform the integral into that of an incomplete gamma function, which in the limit becomes a complete gamma function
-
-\be\label{eq:GammaVanishingOrderIntegral}
-\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\frac{A^{\frac{n+1}{d}+1}}{d}\int_{0}^{\rho A\varepsilon^d}dz\
-z^{\frac{n+1}{d}}e^{-z}\sim\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\int_{0}^{\infty}dz\
-z^{\left[\left(\frac{n+1}{d}+1\right)-1 \right]}e^{-z}\sim\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\Gamma\left(\frac{n+1}{d}\right)
-\ee
-The $\rho$ in the limit of the second expression has been taken to $\infty$.
-The power of $z$ in this integral has been written in this manner so it can be more easily recognised as being of gamma function form, $\Gamma(t)=\int_{0}^{\infty}z^{t-1}e^{-z}dz$.
-One can see from above that if $n>1$ then this term will vanish, thus proving that terms of $O(u^{d+2})$ or higher will go to $0$ in the limit $\rho\rightarrow\infty$.
-
-We use this fact and the method of putting the integrals into a gamma function form to evaluate (\ref{eq:NmaxNminSimplified}).
-One also sees that we have terms like $\frac{\dot{h}}{h}$, and from (\ref{eq:K}) we know that these can be written in terms of the extrinsic curvature, $K(0,\mathbf{x})$.
-We find then that (\ref{eq:NmaxNminSimplified}) reduces to
+and we have expanded the $O(t^{n+1})$ parts of the exponentials.
+This is allowed as the higher power of $t$ means they are suppressed more in the final limiting procedure.
+Using the fact that in GNCs, $K(0,\mathbf{x})=-\frac{1}{2}\frac{\dot{h}}{h}$, the integral can be found to be
 
@@
 \lim_{\rho \to \infty}\left\langle N_{max}-N_{min} \right\rangle=\lim_{\rho \to \infty}
-\rho^{1-\frac{2}{d}}\frac{2(d+2)}{d(d+1)}
-\left[\frac{A_{d-2}}{d(d-1)}\right]^{-\frac{2}{d}}
-\Gamma\left(\frac{2}{d}\right)\int_{\Sigma}d^{d-1}x\
-\sqrt{h}K(0,\mathbf{x})
+\rho^{1-\frac{2}{n}}\frac{2(n+2)}{n(n+1)}
+\left[\frac{A_{n-2}}{n(d-1)}\right]^{-\frac{2}{n}}
+\Gamma\left(\frac{2}{n}\right)\int_{\Sigma}d^{d-1}x\
+h^{\frac{1}{2}}K(0,\mathbf{x})
 \ee
+The gamma function appears as a result of taking integral limits to $\infty$ as $\rho\rightarrow\infty$, while still retaining the $\rho$ outside the integral.
 $K(0,\mathbf{x})$ is evaluated at the surface and is therefore the same $K$ as in equation (\ref{GH_boundary_to_causet}).
-We can therefore conclude that to make (\ref{GH_boundary_to_causet}) true we need a factor $\rho^{\frac{2}{d}-1}$ to cancel the factors of $\rho$ above, and a value of $c_d$ equal to that of equation (\ref{Cn}).
-The factor of $\rho$ needed is exactly that required on dimensional grounds, thus we have proved that (\ref{GH_boundary_to_causet}) is in fact true in general.
+We can therefore conclude that to make (\ref{GH_boundary_to_causet}) true we need a factor $\rho^{\frac{2}{n}-1}$ to cancel the factors of $\rho$ above and a value of $C_n$ equal to that of equation (\ref{Cn}).
 

### 2014-09-18 paper_v3.tex
initial commit

--- 2014-09-18 2014-06-04-paper_v4-ianbefore.tex
@@
 They are essential in order that the quantum mechanical amplitudes satisfy the correct composition law and in order that these amplitudes have the correct classical limit.
-In a number of examples [5] they give a contribution to the partition function which is important for the agreement with calculations based on straightforward thermo- dynamics.
+In a number of examples [5] they give a contribution to the partition function which is important for the agreement with calculations based on straightforward thermodynamics.
 In this note we shall derive the boundary terms in the action for the Regge calculus]
@@
 The action of a given causal set is a crucial piece of the kinematics that would be extremely useful to know and understand.
-Proposals for the action of a causal set are available \cite{Benincasa_Dowker:The_Scalar_Curvature_of_a_Causal_Set} and these hold analytically in some cases and numerically in many more.
+Proposals for the action of a causal set are available [[REFERENCE]] and these hold analytically in some cases and numerically in many more.
 These cases being when the causal set is embeddable in some existing spacetime.
@@
 The EH action, however, is not the complete story in the continuum.
-If one has a spacetime manifold with a boundary, such as Minkowski space but cut off at some finite time, then one needs to include the Gibbons-Hawking-York boundary term (GHYBT) in the EH action \cite{Gibbons_Hawking_Boundary}.
+If one has a spacetime manifold with a boundary, such as Minkowski space but cut off at some finite time, then one needs to include the Gibbons-Hawking-York boundary term (GHYBT) in the EH action.
 The boundary of the manifold will be some hypersurface of codimension-1.
@@
 
-In order to say we have a causal set analogue for a continuous expression we need a well defined procedure for relating the discrete theory to the continuum.
+[[MB1: we need to say what $\rho$ is in terms of the discreteness scale $l$]] In order to say we have a causal set analogue for a continuous expression we need a well defined procedure for relating the discrete theory to the continuum.
 This procedure, or tool, is called a \textit{Poisson sprinkling}, or just a \textit{sprinkling}.
@@
 This means that the expected number of points in some region will be $\rho V$, where $\rho$ is the density of the sprinkling.
-The density is related to the discreteness scale, $l$, by $\rho=l^{-d}$ in $d$ spacetime dimensions.
 It is called a sprinkling as one can envisage the point selection process as a 'sprinkling' of points into the manifold.
@@
 The Gibbons-Hawking-York boundary term for $M^\pm$ is in this case given by
-\be\label{eq:GHYBT_in_continuum}
+$$
 S_{GHY} = \pm \frac{1}{l_p^{d-2}}\int_{\Sigma} d^{d-1}x \sqrt{h} K
-\ee
+$$
 where $l_p$ is the rationalised Planck length and $K$ is the trace of the extrinsic curvature $K_{\mu\nu}=h_{\mu}^\rho h_\nu^\sigma \nabla_\rho n_\sigma$ of $\Sigma$ defined with future-pointing timelike unit normal $n_{\mu}=\partial_\mu S/\sqrt{g^{\mu\nu}\partial_\mu S\partial_\nu S}$.
@@
 A Cauchy surface $\Sigma\subset M$ induces a partition $\mathcal C = \mathcal C^+ \cup\, \mathcal C^-$ of the sprinkling, where $\mathcal C^\pm$ denotes the restriction of $\mathcal C$ to the points sprinkled to the causal future/past of $\Sigma$.
-Let us denote the number of maximal elements in $\mathcal C^-$ and the number of minimal elements in $\mathcal C^+$ by $N_{max}$ and $N_{min}$, respectively.
+Let us denote the number of maximal elements in $\mathcal C^-$ and the number of minimal elements in $\mathcal C^+$ by $N_{min}$ and $N_{max}$, respectively.
 We propose the following definition for the discrete Gibbons-Hawking-York boundary term:
 \be\label{GH_boundary_to_causet}
-S^{(d)}_{GHY}[\mathcal C]=\left(l/l_p\right)^{d-2}c_{d}(N_{max}-N_{min})
+S^{(d)}_{GHY}[\mathcal C]=\left(l/l_p\right)^{2-d}c_{d}(N_{max}-N_{min})
 \ee
@@
 where $\left\langle\cdot\right\rangle$ denotes the mean over sprinklings.
-The limit $\rho\rightarrow\infty$ is equivalent to setting the discreteness scale to zero, $l\rightarrow 0$.
-Figure \ref{fig:Nmin_Nmax} shows an illustration of the idea behind this conjecture.
+Figure \ref{fig:Nmin_Nmax} shows an illustration of this idea.
 The minimal and maximal elements have been highlighted and the shaded regions correspond to the regions whose volumes expressions will be needed in the proof.
@@
 \section{The Proof}
+To reduce the clutter in equations below it will be more convenient to work in natural units, so we set $l_p=1$ for now and recover it later.
 
@@
 and we have Taylor expanded the $O(t^{d+1})$ parts of the volume functions, in small $t$, from the exponents.
-We have also included a pre-factor of $\rho^{\frac{2}{d}-1}$ to account for the factors of $l$ in (\ref{GH_boundary_to_causet}).
-With this pre-factor the limit of the above equation is well defined.
 If we change to $u=-t$ in the first part on the right hand side, and $u=t$ on the second part then the integral can be simplified to
-\be\label{eq:NmaxNminSimplified}
-\lim_{\rho \to \infty}\rho^{\frac{2}{d}}
+\be\bea\label{eq:NmaxNminSimplified}
+\lim_{\rho \to \infty}&\rho^{\frac{2}{d}-1}\left\langle N_{max}-N_{min} \right\rangle=\\
+&\lim_{\rho \to \infty}\rho^{\frac{2}{d}}
 \int_{\Sigma}d^{d-1}x \int_{0}^{\varepsilon}du\
-h^{\frac{1}{2}}\left(-\frac{\dot{h}}{h}u+2\rho B u^{d+1}+\rho O(u^{d+2}) \right)e^{-\rho Au^{d}}
-\ee
-We now show that $O(u^{d+2})$, or higher, terms vanish as $\rho \rightarrow\infty$.
-All higher order corrections that we have ignored so far contribute at $O(u^{d+2})$ or higher, so proving that these contributions vanish proves that our approximations are valid.
-Taking only the part we are interested in from (\ref{eq:NmaxNminSimplified}) leaves us with the task of showing that the following integral vanishes.
+h^{\frac{1}{2}}\left(-\frac{\dot{h}}{h}u+2\rho B u^{d+1}+\rho Cu^{d+2}+\rho O(u^{d+3}) \right)e^{-\rho Au^{d}}
+\eea\ee
+where $C$ is the coefficient of $O(u^{d+2})$ terms.
+This term has been included as showing that it, and higher powers, vanish as $\rho \rightarrow\infty$ will prove our claim that no higher order terms are needed.
+To prove this we need to show that the following integral vanishes
 
 \be\label{eq:VanishingOrderIntegral}
-\lim_{\rho \to \infty}\rho^{\frac{2}{d}+1}\int_{0}^{\varepsilon}du\
+\lim_{\rho \to \infty}\rho^{\frac{2}{d}-1}\rho\int_{0}^{\varepsilon}du\
+\rho u^{d+n}e^{-\rho Au^{d}}=\lim_{\rho \to \infty}\rho^{\frac{2}{d}+1}\int_{0}^{\varepsilon}du\
 u^{d+n}e^{-\rho Au^{d}}
 \ee
-for $n>1$, where $n\in\mathbb{N}$.
-The additional factor of $\rho$ has come from inside the integral in (\ref{eq:NmaxNminSimplified}).
-We make the substitution $z=\rho Au^{d}$ and find that (\ref{eq:VanishingOrderIntegral}) is proportional to
+where $n\in\mathbb{N}$.
+We have left the powers of $\rho$ separated on the left side to show where they have come from.
+The first $\rho^{\frac{2}{d}-1}$ comes from (\ref{GH_boundary_to_causet}) and is needed to balance dimensions.
+This must be included in order to take the limit properly.
+We then make the substitution $z=\rho Au^{d}$ to transform the integral into that of an incomplete gamma function, which in the limit becomes a complete gamma function
 
-\be\label{eq:GammaVanishingOrderIntegral}
-\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\int_{0}^{\rho A\varepsilon^d}dz\;z^{\frac{n+1}{d}}e^{-z}=\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\int_{0}^{\infty}dz\;z^{\left[\left(\frac{n+1}{d}+1\right)-1 \right]}e^{-z}
+\be\begin{aligned}\label{eq:GammaVanishingOrderIntegral}
+\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\frac{A^{\frac{n+1}{d}+1}}{d}\int_{0}^{\rho A\varepsilon^d}dz\
+z^{\frac{n+1}{d}}e^{-z}
+&\sim\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\int_{0}^{\infty}dz
+z^{\left[\left(\frac{n+1}{d}+1\right)-1 \right]}e^{-z}\\
+&\sim\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\Gamma\left(\frac{n+1}{d}\right)
+\end{aligned}
 \ee
-where $\rho$ has been taken to $\infty$ in the limits on the RHS, and the power of $z$ has been written in this way so that the integral is in the form of a gamma function, $\Gamma(t)=\int_{0}^{\infty}z^{t-1}e^{-z}dz$.
-Using the fact that for gamma functions, $\Gamma(t+1)=t\Gamma(t)$, we find that (\ref{eq:GammaVanishingOrderIntegral}) is then proportional to
-
-\be\label{eq:GammaVanishingOrder}
-\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\Gamma\left(\frac{n+1}{d}\right)
-\ee
+The $\rho$ in the limit of the second expression has been taken to $\infty$.
+The power of $z$ in this integral has been written in this manner so it can be more easily recognised as being of gamma function form, $\Gamma(t)=\int_{0}^{\infty}z^{t-1}e^{-z}dz$.
 One can see from above that if $n>1$ then this term will vanish, thus proving that terms of $O(u^{d+2})$ or higher will go to $0$ in the limit $\rho\rightarrow\infty$.
 
-The non-vanishing terms of (\ref{eq:NmaxNminSimplified}) can also be put into the form of gamma functions.
-This allows the integral over $u$ to be evaluated in (\ref{eq:NmaxNminSimplified}).
+We use this fact and the method of putting the integrals into a gamma function form to evaluate (\ref{eq:NmaxNminSimplified}).
 One also sees that we have terms like $\frac{\dot{h}}{h}$, and from (\ref{eq:K}) we know that these can be written in terms of the extrinsic curvature, $K(0,\mathbf{x})$.
@@
 $K(0,\mathbf{x})$ is evaluated at the surface and is therefore the same $K$ as in equation (\ref{GH_boundary_to_causet}).
-We can therefore conclude that to make (\ref{eq:mainconjecture}) true we need a factor $l_{p}^{2-d}$ to make our expression dimensionless, and a value of $c_d$ equal to that of equation (\ref{Cn}).
-This concludes the proof of the conjectured equality in (\ref{eq:mainconjecture}).
+We can therefore conclude that to make (\ref{GH_boundary_to_causet}) true we need a factor $\rho^{\frac{2}{d}-1}$ to cancel the factors of $\rho$ above, and a value of $c_d$ equal to that of equation (\ref{Cn}).
+The factor of $\rho$ needed is exactly that required on dimensional grounds, thus we have proved that (\ref{GH_boundary_to_causet}) is in fact true in general.
 
@@
 In order to test this heuristic argument one may like to write down the integral expression for $\text{Var}[S^{(d)}_{GHY}(\mathcal C)]^\frac12$ but it is complicated enough even in finite regions of flat space that it is not very illuminating to reproduce it here.
-Instead we will show the results of computer simulations that support results of the heuristic argument.
+Instead we will show the results of computer simulations that support the conclusion of the heuristic argument.
 
@@
 Evaluate $S^{(d)}_{GHY}=c_d\rho^\frac{d-2}{d}(N_{min} - N_{max})$ for $\Sigma$ by counting the minimal/maximal elements in the upper/lower half of the cube, and repeat.
-
-\bibliographystyle{utphys}
-
-\bibliography{references}{}
 

### 2014-09-18 paper_v4.tex
initial commit

--- 2014-09-18 2014-06-04-paper_v4.tex
@@
 The density is related to the discreteness scale, $l$, by $\rho=l^{-d}$ in $d$ spacetime dimensions.
-It is called a sprinkling as one can envisage the point selection process as a 'sprinkling' of points into the manifold.
+It is called a sprinkling as one can envisage the point selection process as a `sprinkling' of points into the manifold.
 If a causet, $\mathcal{C}$, can be generated with relatively high probability by a sprinkling into the manifold, $(\mathcal{M},g)$, then we say that the manifold is a good approximation for the causal set.
 
-\section{The Claim}
+\section{The Claims}
 
@@
 where $l_p$ is the rationalised Planck length and $K$ is the trace of the extrinsic curvature $K_{\mu\nu}=h_{\mu}^\rho h_\nu^\sigma \nabla_\rho n_\sigma$ of $\Sigma$ defined with future-pointing timelike unit normal $n_{\mu}=\partial_\mu S/\sqrt{g^{\mu\nu}\partial_\mu S\partial_\nu S}$.
+This translates into a past-pointing normal vector $n^{\mu}$.
 
@@
 \ee
-where the right hand side is the derivative of the area $\int d\Sigma$ as each point of $\Sigma$ is moved an equal distance along the outward unit normal $n$.
+where the right hand side is the derivative of the area $\int d\Sigma$ as each point of $\Sigma$ is moved an equal distance along the unit normal vector, $n^{\mu}$, which corresponds to a rate of change of the surface volume backwards in time, as $n^{\mu}$ is past-pointing.
 \mb{does the antichain defn really work?} In the causal set, the analogue of a spacelike hypersurface is an anti-chain, i.e.
@@
 Figure \ref{fig:Nmin_Nmax} shows an illustration of the idea.
-The minimal and maximal elements have been highlighted.
+The minimal and maximal elements have been highlighted in COLOUR.
 
-Let us denote the number of maximal elements in $\mathcal C^-$ by $N_{max}$ and the number of minimal elements in $\mathcal C^+$  by $N_{min}$.
-We propose the following definition for the discrete Gibbons-Hawking-York boundary term:\mb{$S[\mathcal C, \Sigma]$?}
+We will also show that it is possible to restrict oneself to either $C^+$ or $C^-$ and still calculate the GHY term.
+To do this requires us to define the $k$-next-to-maximal/minimal elements in $C^-$ or $C^+$ respectively.
+These are the elements of the causal set that have $k$ elements to their future or past.
+This means that the $0$-to-maximal/minimal elements are the maximal/minimal elements.
+Elements with $1$ element to their past/future have been highlighted in COLOUR in Figure \ref{fig:Nmin_Nmax}, and those with $2$ have been highlighted in COLOUR.
+
+Let us denote the number of $k$-next-to-maximal elements in $\mathcal C^-$ by $N_{max}^{(k)}$ and the number of $k$-next-to-minimal elements in $\mathcal C^+$  by $N_{min}^{(k)}$.
+
+We propose the following definitions for the discrete Gibbons-Hawking-York boundary term (one can adopt the appropriate definition to best suit their calculation needs):\mb{$S[\mathcal C, \Sigma]$?}
+
 \be\label{GH_boundary_to_causet}
-S^{(d)}_{GHY}[\mathcal C]=\left(l/l_p\right)^{d-2}c_{d}\left(N_{max}-N_{min}\right)
+ \mathcal{S}^{(d)}_{GHY}[\mathcal C]=\left(l/l_p\right)^{d-2}\frac{c_{d}}{2\: \Gamma\left(\frac{2}{d} \right)}\times
+  \begin{cases}
+   N_{max}^{(0)}-N_{min}^{(0)} \\
+   d\: N_{max}^{(1)}-N_{max}^{(0)} \\
+   N_{min}^{(0)}-d\: N_{min}^{(1)}
+  \end{cases}
 \ee
@@
 \be\label{Cn}
-c_{d}=\frac{d(d+1)}{2(d+2)\Gamma\left(\frac{2}{d}\right)}\left[\frac{A_{d-2}}{d(d-1)}\right]^{\frac{2}{d}},
+c_{d}=\frac{d(d+1)}{(d+2)}\left[\frac{A_{d-2}}{d(d-1)}\right]^{\frac{2}{d}},
 \ee
-$A_d=2\pi^\frac{d+1}2/\Gamma\left(\frac{d+1}{2}\right)$ denoting the volume of the unit $d$-sphere.
-To support this proposal we show that in the continuum limit $l\rightarrow 0$ (i.e.
-$\rho\rightarrow\infty$) we obtain
+$A_d=2\pi^\frac{d+1}2/\Gamma\left(\frac{d+1}{2}\right)$ and denotes the volume of the unit $d$-sphere.
+In fact, one can write down a general expression for the boundary term as a sum over $k$ of $N_{max/min}^{\: (k)}$ terms with particular coefficients.
+This can be written as follows:
+
+\begin{gather}\label{general_boundary_sum}
+\begin{aligned}
+\mathcal{S}^{(d)}_{GHY}[\mathcal C]=\left(l/l_p\right)^{d-2} & c_{d}\left( \sum_m p_m \frac{\Gamma\left(\frac{2}{d}+m \right)}{\Gamma\left(\frac{1}{d}+m \right)}  - \sum_n q_n\frac{\Gamma\left(\frac{2}{d}+n \right)}{\Gamma\left(\frac{1}{d}+n \right)} \right)^{-1} \\
+& \times \left( \sum_m p_m \frac{m!}{\Gamma\left(\frac{1}{d}+m \right)} N_{max}^{(m)}
++  \sum_n q_n \frac{n!}{\Gamma\left(\frac{1}{d}+n \right)} N_{min}^{(n)}\right)
+\end{aligned}
+\end{gather}
+The coefficients, $p_m , q_n \in \mathbb{R}$, must satisfy the following relation:
+
+\be\label{coefficient_relation}
+\sum_m p_m + \sum_n q_n = 0
+\ee
+The different definitions of the boundary term in (\ref{GH_boundary_to_causet}) are just simple cases of this general formula, and the objects that one would preferably use in a calculation.
+From above one can see that at least two terms are needed in the sum to find the boundary term.
+The condition~\eqref{coefficient_relation} means that there will be one free parameter if two terms are present\footnote{For example you could include only the maximal and minimal terms (like the first case in~\eqref{GH_boundary_to_causet}) with $p_0=p$ and $q_0=-p$ where $p\in\mathbb{R}$.
+This would satisfy~\eqref{coefficient_relation} and would leave $p$ as a free parameter.
+In this case you would get a factor of $p^{-1}$ from the term after $c_d$ in~\eqref{general_boundary_sum} which will cancel the $p$ from the last term.}.
+This free parameter will be cancelled by the term after $c_d$ in~\eqref{general_boundary_sum}.
+Thus, with only two terms, the form of the discrete boundary term is fixed.
+If one chooses to include more than two terms the coefficients of the $N_{{\,}_{max/min}}^{\:(k)}$ terms are no longer unique.
+
+To support this proposal we show that in the continuum limit of $l\rightarrow 0$ (i.e. $\rho\rightarrow\infty$) we obtain
 %of infinite sprinkling density into a $d$-dimensional spacetime we obtain
 \be
-\lim_{l\rightarrow0}\left\langle S^{(d)}_{GHY}[\mathcal C] \right\rangle= \frac{1}{l_p^{d-2}}\int_{\Sigma} d^{d-1}x \sqrt{h} K\label{eq:mainconjecture}
+\lim_{l\rightarrow0}\left\langle S^{(d)}_{GHY}[\mathcal C] \right\rangle= \frac{1}{l_p^{d-2}}\int_{\Sigma} d^{d-1}x\: \sqrt{h}\: K\label{eq:mainconjecture}
 \ee
-where $\left\langle\cdot\right\rangle$ denotes the mean over sprinklings.
+for the most general definition above. Here $\left\langle\cdot\right\rangle$ denotes the mean over sprinklings.
+
+We also propose a discrete definition for the spatial volume of the surface:
+
+\be\label{eq:surface_volume}
+\mathcal{A}^{(d)}[\mathcal{C}]=\left(l/l_p\right)^{d-1}\frac{b_{d}}{\Gamma\left(\frac{1}{d}\right)}\: N_{{\,}_{max/min}}^{\:(0)}
+\ee
+where the constant $b_d$ is given by
+
+\be\label{constant_b_d}
+b_d=d\left[\frac{A_{d-2}}{d(d-1)}\right]^{\frac{1}{d}}
+\ee
+Again on can find a general expression written as a sum of different terms.
+
+\be\label{general_area_sum}
+\mathcal{A}^{(d)}[\mathcal{C}]=\left(l/l_p\right)^{d-1}b_{d}\sum_m p_m \frac{m!}{\Gamma\left(\frac{1}{d}+m \right)}\: N_{{\,}_{max/min}}^{\:(m)}
+\ee
+where the coefficients, $p_m$, must satisfy $\sum_m p_m =1$.
+This time we see that only one term is needed to find the surface volume, and that when more than one term is added the coefficients will not be unique.
+We support this claim in the same way as the boundary term, and show that
+
+\be\label{eq:conjecture_for_area}
+\lim_{l\rightarrow0}\left\langle \mathcal{A}^{(d)}[\mathcal{C}] \right\rangle= \frac{1}{l_p^{d-1}}\int_{\Sigma} d^{d-1}x\: \sqrt{h}
+\ee
+where the right hand side is the spatial volume of the surface $\Sigma$.
+
 %MB1: (too vague) During the calculation approximations will be made.
@@
 
-\subsection{Poisson Sprinklings for $\langle N_{max}\rangle$ and $\langle N_{min}\rangle$}
+\subsection{Poisson Sprinklings for $\langle N_{max}^{(m)}\rangle$ and $\langle N_{min}^{(m)}\rangle$}
 
-In order to prove~\eqref{eq:mainconjecture} let us first derive an expression for the mean value of $S^{(d)}_{GHY}[\mathcal C]$.
-For any instance of the sprinkling, the probability that a sprinkled point $p\in M$ below the surface $\Sigma$ is \emph{maximal} is given by the probability that no point of the sprinkling lies in the region $J^{+}(p)\cap J^{-}(\Sigma)$, the intersection of the causal future of $p$ with the causal past of the surface $\Sigma$.\footnote{For notational convenience we use the symbol $x$ to refer to the causal element $x\in\mathcal C$, to its embedding in the manifold $M$, and to its coordinates in some chart on $M$.} This region will in general be the interior of some curvy $d$-dimensional cone truncated by the surface $\Sigma$, as illustrated in Figure~\ref{fig:Nmin_Nmax}.
+In order to prove~\eqref{eq:mainconjecture} and~\eqref{eq:conjecture_for_area} let us first derive expressions for the mean values of $N_{max}^{(m)}$ and $N_{min}^{(m)}$.
+For any instance of the sprinkling, the probability that a sprinkled point $p\in M$ below the surface $\Sigma$ is $k$\emph{-next-to-maximal} is given by the probability that $k$ points of the sprinkling lie in the region $J^{+}(p)\cap J^{-}(\Sigma)$, the intersection of the causal future of $p$ with the causal past of the surface $\Sigma$.\footnote{For notational convenience we use the symbol $x$ to refer to the causal element $x\in\mathcal C$, to its embedding in the manifold $M$, and to its coordinates in some chart on $M$.} This region will in general be the interior of some curvy $d$-dimensional cone truncated by the surface $\Sigma$, as illustrated in Figure~\ref{fig:Nmin_Nmax}.
 We will refer to these regions as \emph{truncated cones} below.
@@
 \be\label{Poisson}
-\mathbb P\left(\text{no points in }J^{+}(p)\cap J^{-}(\Sigma)\right)=e^{-\rho V_\blacktriangledown(p)}
+\mathbb P\left(\text{k points in }J^{+}(p)\cap J^{-}(\Sigma)\right)=\frac{\left(\rho\: V_\blacktriangledown(p)\right)^k}{k!}e^{-\rho V_\blacktriangledown(p)}
 \ee
 to this event, where $V_\blacktriangledown(p)\equiv V(J^{+}(p)\cap J^{-}(\Sigma))$ is the spacetime volume of the region $J^{+}(p)\cap J^{-}(\Sigma)$.
-The probability of sprinkling an element into an infinitesimal four-volume $dV_p$ at $p$ is $\rho dV_p$ where $\rho=l^{-d}$ is the sprinkling density, and so the total expected number of maximal elements below $\Sigma$ is
+The probability of sprinkling an element into an infinitesimal four-volume $dV_p$ at $p$ is $\rho dV_p$ where $\rho=l^{-d}$ is the sprinkling density, and so the total expected number of $k$-next-to-maximal elements below $\Sigma$ is
 \be\label{eq:nmax}
-\left\langle N_{max}\right\rangle =\rho\int_{J^{-}(\Sigma)}dV_p\; e^{-\rho V_\blacktriangledown(p)}
+\left\langle N_{max}^{(k)}\right\rangle =\rho\int_{J^{-}(\Sigma)}dV_p\; \frac{\left(\rho\: V_\blacktriangledown(p)\right)^k}{k!}e^{-\rho V_\blacktriangledown(p)}
 \ee
-Similarly the expected number of minimal elements above $\Sigma$ is
+Similarly the expected number of $k$-next-to-minimal elements above $\Sigma$ is
 \be\label{eq:nmin}
-\left\langle N_{min}\right\rangle =\rho\int_{J^{+}(\Sigma)}dV_p\; e^{-\rho V_\blacktriangle(p)}
+\left\langle N_{min}^{(k)}\right\rangle =\rho\int_{J^{+}(\Sigma)}dV_p\; \frac{\left(\rho\: V_\blacktriangle(p)\right)^k}{k!}e^{-\rho V_\blacktriangle(p)}
 \ee
@@
 
-In the limit $\rho\rightarrow\infty$, both quantities will diverge, but if their difference $\langle N_{max}\rangle - \langle N_{min}\rangle = \langle N_{max} - N_{min}\rangle$ grows slower than or at order $\rho^{1-\frac2d}$, the proposed action~\eqref{GH_boundary_to_causet} will tend to a finite value in the continuum limit.
+In the limit $\rho\rightarrow\infty$, both quantities will diverge, but if some combination of maximal and minimal terms grows slower than or at order $\rho^{1-\frac2d}$, the proposed action~\eqref{general_boundary_sum} will tend to a finite value in the continuum limit.
 
@@
 we can't have an asymptotically null surface i suppose} The integrals then simplify to
-
-\begin{align}\label{eq:nmax_and_eq:nmin}
-\left\langle N_{max}\right\rangle & =\int_{\Sigma}d^{d-1}x\int_{-\varepsilon}^{0}dt\:
+\begin{gather}\label{eq:nmax_and_eq:nmin}
+\begin{aligned}
+\left\langle N_{max}^{(k)}\right\rangle & =\rho \int_{\Sigma}d^{d-1}x\int_{-\varepsilon}^{0}dt\:
 h^{\frac{1}{2}}\left(1+
 \frac{1}{2}\frac{\dot{h}}{h}t+O(t^2)\right)
- \rho\ e^{-\rho V_\blacktriangledown(t,\mathbf x)}
+ \frac{\left(\rho\: V_\blacktriangledown(p)\right)^k}{k!} e^{-\rho V_\blacktriangledown(t,\mathbf x)}
 \\
-\left\langle N_{min}\right\rangle & =\int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
+\left\langle N_{min}^{(k)}\right\rangle & =\rho \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
 h^{\frac{1}{2}}\left(1+
-\frac{1}{2}\frac{\dot{h}}{h}t+O(t^2)\right) \rho\ e^{-\rho V_\blacktriangle(t,\mathbf x)}
-\end{align}
+\frac{1}{2}\frac{\dot{h}}{h}t+O(t^2)\right) \frac{\left(\rho\: V_\blacktriangle(p)\right)^k}{k!} e^{-\rho V_\blacktriangle(t,\mathbf x)}
+\end{aligned}
+\end{gather}
 where $h\equiv det\left(h_{ij}(0,\mathbf{x})\right)$ and $\dot{}\equiv \frac{\partial}{\partial t}$ and the metric determinant has been in expanded in small $t$.
@@
 This will all be made more precise below.
-The approximation procedure will be outlined below for a cone to the future of $\Sigma$, as the process for the past cone is nearly identical.
 
@@
 For the transformation from GNCs to RNCs at $p_0$ one finds that $A^{\overline 0}_{\;0}=1$, $A^{\overline 0}_{\;i}=0$ and $\delta_{\ibar\jbar}=A^i_{\;\ibar}A^j_{\;\jbar}h_{ij}(p_0)$, which in particular implies $y^{\overline{0}}_p=\overline{t}_p=t_p=T$.
-\mb{more detail needed i think}It can then be shown that the volume integral reduces to~\cite{Khetrapal_Sumati:Causal_Diamond_Volume}
+\mb{more detail needed i think}In RNC one can expand the metric determinant in~\eqref{eq:VolumeWithNoSimplifications} to find
 
 \be\label{eq:VolumeWithRNC}
-V_\blacktriangle(p) =\int_{\mathcal{X}}d^dy+\int_{\mathcal{X}_0}d^dy\left(-\frac{1}{6}R_{\overline{\mu}\overline{\nu}}(p_0)y^{\overline{\mu}}y^{\overline{\nu}} \right)+O(T^{d+3})
+V_\blacktriangle(p) =\int_{\mathcal{X}_p}d^dy+\int_{\mathcal{X}_p}d^dy\left(-\frac{1}{6}R_{\overline{\mu}\overline{\nu}}(p_0)y^{\overline{\mu}}y^{\overline{\nu}} \right)+O(T^{d+3})
 \ee
-where $\mathcal{X}_0=\left\lbrace y^{\overline{\mu}} \mid 0 \leq \overline{t}\leq T\; ,\; (y^{\overline{1}})^2+(y^{\overline{2}})^2+...+(y^{\overline{d-1}})^2\leq (T-\overline{t})^2 \right\rbrace$ is what we call the \textit{flat cone}, and $R_{\overline{\mu}\overline{\nu}}(p_0)$ is the Ricci tensor in RNC evaluated at $p_0$.\mb{justify beforehand which orders to keep?}
+where $R_{\overline{\mu}\overline{\nu}}(p_0)$ is the Ricci tensor in RNC evaluated at $p_0$.\mb{justify beforehand which orders to keep?}
 %The arguments of the volume function need not be changed as the volume only depends on the points in the manifold, $q_0$ and $p_0$. 
-The flat cone term comes in at $O(T^{d+2})$ so the volume we have to calculate has reduced to
+The second term comes in at $O(T^{d+2})$ so the volume we have to calculate has reduced to
 
 \be\label{eq:VolumeToLowestOrder}
-V_\blacktriangle(p) =\int_{\mathcal{X}}d^dy+O(T^{d+2})
+V_\blacktriangle(p) =\int_{\mathcal{X}_p}d^dy+O(T^{d+2})
 \ee
@@
 
-This is now a simple volume integral but we need to find the boundaries of $\mathcal{X}$ for the correct integration limits.
+This is now a simple volume integral but we need to find the boundaries of $\mathcal{X}_p$ to write down the integration limits.
 There are two boundaries to the (solid) truncated cone: the lightcone emanating at $p$ and the base, given by the intersection of $\Sigma$ with the interior of the lightcone.
@@
 
-The first part on the right of (\ref{eq:RNCtotaltrans}) vanishes as $A^{\overline{0}}_{\;\mu}x^{\mu}=x^0$ (as $A^{\overline{0}}_{\;i}=0$ and $A^{\overline{0}}_{\;0}=1$) and $x^0=t=0$ for the bottom surface.
+The linear part on the right of (\ref{eq:RNCtotaltrans}) vanishes as $A^{\overline{0}}_{\;\mu}x^{\mu}=x^0$ (as $A^{\overline{0}}_{\;i}=0$ and $A^{\overline{0}}_{\;0}=1$) and $x^0=t=0$ for the bottom surface.
 Using the inverse RNC relation (\ref{eq:RNCinversetrans}) one can find the equation for the bottom surface in RNCs:
@@
 The solution can be expanded in $T$ and is simply $r_{max}=T+O(T^2)$, with angular dependent terms contributing at $O(T^2)$.
-The $O(T^2)$ term will contribute at $O(T^{d+2})$ in the integral and so can be ignored.
+The $O(T^2)$ term will contribute at $O(T^{d+2})$ in the volume integral and so can be ignored.
 Substituting $r_{max}=T$ into (\ref{eq:VolumeIntegralSpherical}) lets us evaluate the integral and we find
@@
 Now in GNCs the extrinsic curvature on the surface is given by
+
 \be\label{eq:K}
@@
 \end{align}
-having dropped the subscript $p$. Given the volume expressions in GNCs we now proceed to evaluate the integrals for $\langle N_{min}\rangle $ and $\langle N_{max}\rangle$.
+having dropped the subscript $p$. Given the volume expressions in GNCs we now proceed to evaluate the integrals for $\langle N_{min}^{(k)}\rangle $ and $\langle N_{max}^{(k)}\rangle$.
 
@@
 
-Using the previous equation we obtain for (\ref{GH_boundary_to_causet}) in the limit $\rho \rightarrow \infty$
-\begin{gather}\label{eq:NmaxNminStart}
+Using the previous equations for the volumes of the truncated cones we find that~\eqref{eq:nmax_and_eq:nmin} reduces to
+
+\begin{gather}\label{eq:nmax_and_eq:nmin_volume_expanded}
 \begin{aligned}
-\lim_{\rho \to \infty}\rho^{\frac{2}{d}-1}\left\langle N_{max}-N_{min} \right\rangle &= \\
-\lim_{\rho \to \infty}\rho^{\frac{2}{d}}
-\int_{\Sigma}d^{d-1}x & \int_{-\varepsilon}^{0}dt\
+\left\langle N_{max}^{(k)}\right\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{-\varepsilon}^{0}dt\:
 h^{\frac{1}{2}}\left(1+
-\frac{1}{2}\frac{\dot{h}}{h}t+\rho B(-t)^{d+1}+\rho O(t^{d+2})\right)e^{-\rho A(-t)^{d}} \\
--\rho^{\frac{2}{d}}\int_{\Sigma}d^{d-1}x &
-\int_{0}^{\varepsilon}dt\
+\frac{1}{2}\frac{\dot{h}}{h}t\right)
+ \\
+ & \times \Big( A(-t)^d \Big)^k 
+ \Big( 1 - B(-t) \Big)^k
+ e^{-\rho A(-t)^d \left[1-B(-t) \right]} + .\,.\,.
+\\
+\left\langle N_{min}^{(k)}\right\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
 h^{\frac{1}{2}}\left(1+
-\frac{1}{2}\frac{\dot{h}}{h}t-\rho Bt^{d+1}+\rho O(t^{d+2})\right)e^{-\rho At^{d}}
+\frac{1}{2}\frac{\dot{h}}{h}t\right)
+ \\
+ & \times \Big( A\: t^d \Big)^k 
+ \Big( 1 + B\: t \Big)^k
+ e^{-\rho A\: t^d \left[1+B\: t \right]} + .\,.\,.
 \end{aligned}
@@
 \begin{aligned}
-A & \equiv \frac{A_{d-2}}{d(d-1)} \\
-B & \equiv \frac{A_{d-2}}{2(d-1)(d+1)}K(0,\mathbf{x})
+A & := \frac{A_{d-2}}{d(d-1)} \\
+B & := \frac{A_{d-2}}{2(d-1)(d+1)}K(0,\mathbf{x})
 \end{aligned}
 \end{gather}
-and we have Taylor expanded the $O(t^{d+1})$ parts of the volume functions, in small $t$, from the exponents.
-The pre-factor $\rho^{\frac{2}{d}-1}$ accounts for the factors of $l$ in (\ref{GH_boundary_to_causet}).
-With this pre-factor the limit of the above equation is well defined.
-Swapping the integration variable to $t\rightarrow -t$ in the first term and combining the two integrals we obtain
-\be\label{eq:NmaxNminSimplified}
-\lim_{\rho \to \infty}\rho^{\frac{2}{d}}
-\int_{\Sigma}d^{d-1}x \int_{0}^{\varepsilon}dt\
-h^{\frac{1}{2}}\left(-\frac{\dot{h}}{h}t+2\rho B t^{d+1}+\rho O(t^{d+2}) \right)e^{-\rho At^{d}}
+Once again the higher order terms that have been ignored in both equations will be shown to vanish in the limit.
+From~\eqref{eq:K} we can substitute $K$ in for $\frac{1}{2}\frac{\dot{h}}{h}t$.
+We swap the integration variable to $t\rightarrow -t$ in the first equation and expand the $O(t^{d+1})$ part of the exponentials to give
+
+\begin{gather}\label{eq:nmax_and_eq:nmin_volume_expo_expanded}
+\begin{aligned}
+\left\langle N_{max}^{(k)}\right\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
+h^{\frac{1}{2}}\left(1+
+K\: t\right)
+ \\
+ & \times \Big( A\: t^d \Big)^k 
+ \Big( 1 - B\: t \Big)^k
+ \Big( 1 + \rho A B\: t^{d+1} \Big)
+ e^{-\rho A\: t^d} + .\,.\,.
+\\
+\left\langle N_{min}^{(k)}\right\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
+h^{\frac{1}{2}}\left(1-
+Kt\right)
+ \\
+ & \times \Big( A\: t^d \Big)^k 
+ \Big( 1 + B\: t \Big)^k
+  \Big( 1 - \rho A B\: t^{d+1} \Big)
+ e^{-\rho A\: t^d} + .\,.\,.
+\end{aligned}
+\end{gather}
+The reason for expanding the exponentials is that the integrals will have to be put into the form of Gaussian integrals in order to evaluate them in the limit $\rho\rightarrow\infty$.
+The above expressions are almost the same apart from a few sign differences, so from now on we will just work with the expression for $\left\langle N_{min}^{(k)}\right\rangle$.
+By expanding the brackets, and retaining only necessary orders, the integral can be split into two parts of differing powers of $\rho$.
+
+\begin{gather}\label{eq:n_min_different_rho_terms}
+\begin{aligned}
+\left\langle N_{min}^{(k)}\right\rangle & = \frac{\rho^{k+1}A^k}{k!}\int_{\Sigma}d^{d-1}x\: h^{\frac{1}{2}}\int_{0}^{\varepsilon}dt\:
+\left\lbrace \left( t^{dk} + \left(kB-K \right)t^{dk+1}\right) + O\left(t^{dk+2}\right) \right\rbrace e^{-\rho A\: t^d}
+ \\
+ & - \frac{\rho^{k+2}A^{k+1}}{k!}\int_{\Sigma}d^{d-1}x\: h^{\frac{1}{2}}\int_{0}^{\varepsilon}dt\:
+\left\lbrace Bt^{dk+d+1} + O\left(t^{dk+d+2}\right) \right\rbrace e^{-\rho A\: t^d}
+\end{aligned}
+\end{gather}
+To find out how this diverges in the limit of $\rho \rightarrow\infty$, it suffices to find the divergence of the following general expression
+\be\label{eq:general_t_n_integral}
+\lim_{\rho\rightarrow\infty}\rho^{p}\int_{0}^{\varepsilon}dt\
+t^{q}e^{-\rho At^{d}}
 \ee
-We now show that $O(t^{d+2})$ and higher terms vanish as $\rho \rightarrow\infty$.
-All corrections that we have ignored so far contribute at $O(t^{d+2})$ or higher, so proving that these contributions vanish proves that our approximations are justified.
-Inspecting (\ref{eq:NmaxNminSimplified}) we see that all we need to show is that for integer $n>1$,  the following integral vanishes:
+where $p,q \in \mathbb{R}$. We make the substitution $z=\rho At^{d}$ to put~\eqref{eq:general_t_n_integral} into the form of an incomplete gamma function.
 
-\be\label{eq:VanishingOrderIntegral}
-\lim_{\rho \to \infty}\rho^{\frac{2}{d}+1}\int_{0}^{\varepsilon}dt\
-t^{d+n}e^{-\rho At^{d}}.
+\be\label{eq:incomplete_gamma_function}
+\lim_{\rho\rightarrow\infty}\frac{A^{-\left(\frac{q+1}{d} \right)}}{d}\rho^{p-\left(\frac{q+1}{d} \right)}\int_{0}^{\rho A \varepsilon^d}dz\
+z^{\left(\frac{q+1}{d} \right)-1}e^{-z}
 \ee
-The additional factor of $\rho$ has come from inside the integral in (\ref{eq:NmaxNminSimplified}).
-We make the substitution $z=\rho At^{d}$ and find that (\ref{eq:VanishingOrderIntegral}) is proportional to
+The pre-factor has come from the substitution.
+In the limit we can retain the $\rho$ outside the integral while taking the integration limit to $\infty$.
+In doing this the integral becomes a gamma function.
 
-\be\label{eq:GammaVanishingOrderIntegral}
-\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\int_{0}^{\rho A\varepsilon^d}dz\;z^{\frac{n+1}{d}}e^{-z}=\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\int_{0}^{\infty}dz\;z^{\left[\left(\frac{n+1}{d}+1\right)-1 \right]}e^{-z}
+\be\label{eq:gamma_function}
+\lim_{\rho\rightarrow\infty}\frac{A^{-\left(\frac{q+1}{d} \right)}}{d}\rho^{p-\left(\frac{q+1}{d} \right)}\int_{0}^{\infty}dz\
+z^{\left(\frac{q+1}{d} \right)-1}e^{-z}=\lim_{\rho\rightarrow\infty}
+\frac{A^{-\left(\frac{q+1}{d} \right)}}{d}\rho^{p-\left(\frac{q+1}{d} \right)}\Gamma\left( \frac{q+1}{d} \right)
 \ee
-where $\rho$ has been taken to $\infty$ in the limits on the RHS, and the power of $z$ has been written in this way so that the integral is in the form of a gamma function, $\Gamma(t)=\int_{0}^{\infty}z^{t-1}e^{-z}dz$.
-Using $\Gamma(t+1)=t\Gamma(t)$ we find that (\ref{eq:GammaVanishingOrderIntegral}) is proportional to
-\be\label{eq:GammaVanishingOrder}
-\lim_{\rho \to \infty}\rho^{\frac{1-n}{d}}\Gamma\left(\frac{n+1}{d}\right)
-\ee
-One can see from above that if $n>1$ then this term will vanish, thus proving that terms of $O(t^{d+2})$ or higher will go to $0$ in the limit $\rho\rightarrow\infty$.
+We can use this procedure to take the limit of~\eqref{eq:n_min_different_rho_terms} and formulate the answer in terms of gamma functions.
+The limits of both $\left\langle N_{min}^{(k)}\right\rangle$ and $\left\langle N_{min}^{(k)}\right\rangle$ are then given by
 
-The non-vanishing terms of (\ref{eq:NmaxNminSimplified}) can also be put into the form of gamma functions.
-This allows the integral over $t$ to be evaluated in (\ref{eq:NmaxNminSimplified}).
-One also sees that we have terms like $\frac{\dot{h}}{h}$, and from (\ref{eq:K}) we know that these can be written in terms of the extrinsic curvature $K$.
-We find then that (\ref{eq:NmaxNminSimplified}) reduces to
+\begin{gather}\label{eq:nmax_nmin_final}
+\begin{aligned}
+\lim_{\rho\rightarrow\infty}\left\langle N_{max}^{(k)}\right\rangle = \lim_{\rho\rightarrow\infty} & \left\lbrace \rho^{1-\frac{1}{d}} \left(b_d\right)^{-1} \frac{\Gamma\left(\frac{1}{d}+k\right)}{k!}
+\int_{\Sigma}d^{d-1}x\: \sqrt{h} \right.
+ \\
+ &  \left. +\rho^{1-\frac{2}{d}} \left(c_d\right)^{-1} \frac{\Gamma\left(\frac{2}{d}+k\right)}{k!}
+\int_{\Sigma}d^{d-1}x\: \sqrt{h}K + O\left(\rho^{1-\frac{3}{d}} \right) \right\rbrace
+\\
+\lim_{\rho\rightarrow\infty}\left\langle N_{min}^{(k)}\right\rangle = \lim_{\rho\rightarrow\infty} & \left\lbrace \rho^{1-\frac{1}{d}} \left(b_d\right)^{-1} \frac{\Gamma\left(\frac{1}{d}+k\right)}{k!}
+\int_{\Sigma}d^{d-1}x\: \sqrt{h} \right.
+ \\
+ &  \left. -\rho^{1-\frac{2}{d}} \left(c_d\right)^{-1} \frac{\Gamma\left(\frac{2}{d}+k\right)}{k!}
+\int_{\Sigma}d^{d-1}x\: \sqrt{h}K + O\left(\rho^{1-\frac{3}{d}} \right) \right\rbrace
+\\
+\end{aligned}
+\end{gather}
 
-\be\label{eq:nmax_eq:nmin_end}
-\lim_{\rho \to \infty}\rho^{\frac{2}{d}-1}\left\langle N_{max}-N_{min} \right\rangle=
-\frac{2(d+2)}{d(d+1)}
-\left[\frac{A_{d-2}}{d(d-1)}\right]^{-\frac{2}{d}}
-\Gamma\left(\frac{2}{d}\right)\int_{\Sigma}d^{d-1}x\
-\sqrt{h}K(0,\mathbf{x}).
-\ee
-$K(0,\mathbf{x})$ is evaluated at the surface and is therefore the same $K$ as in equation (\ref{GH_boundary_to_causet}).
-We can therefore conclude that to make (\ref{eq:mainconjecture}) true we need a factor $l_{p}^{2-d}$ to make our expression dimensionless, and a value of $c_d$ equal to that of equation (\ref{Cn}).
-This concludes the proof of the conjectured equality in (\ref{eq:mainconjecture}).
+These object diverge as $\rho^{1-\frac{1}{d}}$ in the limit.
+The factor of $\rho$ that must be included in the formula for the surface volume is exactly the inverse of this ($\rho^{\frac{1}{d}-1}$), as can be seen from~\eqref{eq:surface_volume} if one substitutes in $l=\rho^{-\frac{1}{d}}$.
+If we include this factor above and then take the limit we see that the only remaining terms are proportional to the surface volume.
+The other terms all have negative powers of $\rho$ and therefore vanish as $\rho\rightarrow\infty$.
+From this it follows that the mean of any linear combination of terms like $\rho^{\frac{1}{d}-1}\left\langle N_{max}^{(k)}\right\rangle$ or $\rho^{\frac{1}{d}-1}\left\langle N_{min}^{(k)}\right\rangle$ will give something proportional to the surface volume, so long as the terms containing the surface volume don't exactly cancel.
+To simplify the coefficients in this linear sum we take out a factor of $\frac{k!}{\Gamma\left(\frac{1}{d}+k\right)}$.
+The last step in arriving at~\eqref{general_area_sum} is to enforce the condition that the coefficients sum to unity, in order for the linear sum to equal the surface volume and not just be proportional.
+This concludes the proof of the surface volume conjecture.
+
+If we multiply~\eqref{eq:nmax_nmin_final} by the appropriate factor of $\rho$ from the boundary term ($\rho^{\frac{2}{d}-1}$) we find that the above equation still diverges as $\rho^{\frac{1}{d}}$ in the limit.
+This divergence comes from the surface volume term, while the term proportional to the GHY boundary term is finite.
+In order to recover the boundary term we must take a linear combination of these objects such that the term proportional to the surface are vanishes, while the term proportional to the boundary term does not.
+This is where the condition~\eqref{coefficient_relation} comes in.
+This condition for the coefficients is only so simple because we have taken out a factor of $\frac{k!}{\Gamma\left(\frac{1}{d}+k\right)}$ from each coefficient.
+The remaining sum is proportional to the boundary term, and in order to get it exactly we simply have to divide out by the proportionality constant.
+At this point one arrives at~\eqref{general_boundary_sum} and the proof is concluded.
 
@@
 So far we have only talked about the mean of the causal set boundary action.
-Let us now turn to its fluctuations (its standard deviation) $\sigma[S^{(d)}_{GHY}(\mathcal C)]=\text{Var}[S^{(d)}_{GHY}(\mathcal C)]^\frac12$.\footnote{Whenever we say fluctuations we refer to the standard deviation of the random variable, not to its variance.} We can make a heuristic argument to estimate the dependence of fluctuations on $\rho=l^{-d}$.
+Let us now turn to its fluctuations (the standard deviation) $\sigma[S^{(d)}_{GHY}(\mathcal C)]=\text{Var}[S^{(d)}_{GHY}(\mathcal C)]^\frac12$.\footnote{Whenever we say fluctuations we refer to the standard deviation of the random variable, not to its variance.} We can make a heuristic argument to estimate the dependence of fluctuations on $\rho=l^{-d}$.
 In any spacetime region of fixed volume $V$ the number of causal set elements $N$ experiences Poisson fluctuations of order $\sqrt N$.
-Since $N_{max}$ and $N_{min}$ are quantities associated with a codimension-1 surface (counting elements ``near" $\Sigma$), we may expect them to inherit fluctuations of order $N^\frac{d-1}{2d}$.
-Then the action $S^{(d)}_{GHY}$, being proportional to $\rho^\frac{d-2}{d}$ times the difference of two independent\footnote{The independence is as a feature of the Poisson process in $M$: the number of elements $N(R)$ in any subregion $R\subset M$ is a Poisson variable itself, and for any two disjoint regions $R_1$, $R_2$ the numbers $N(R_1)$ and $N(R_2)$ are independent random variables.} random variables with standard deviation $N^\frac{d-1}{2d} = (\rho V)^\frac{d-1}{2d}$ should see fluctuations of order $\rho^\frac{2-d}{d}\rho^\frac{d-1}{2d}=\rho^\frac{d-3}{2d}$.
-This suggests that for $d=2$ these fluctuations should grow as $\rho\rightarrow\infty$, for $d=3$ they should be constant, and for $d>3$ they should be damped.
+We will use the first case of~\eqref{GH_boundary_to_causet} as it is the simplest case in which to form a heuristic argument for the fluctuations.
+Since $N_{max}^{(0)}$ and $N_{min}^{(0)}$ are quantities associated with a codimension-1 surface (they count elements ``near" $\Sigma$), we may expect them to inherit fluctuations of order $N^\frac{d-1}{2d}$.
+This comes from the fact that any one dimension will attribute $N^{\frac{1}{d}}$ points to the total volume $N$ (as $(N^{\frac{1}{d}} )^d=N$).
+Thus, a codimension-1 surface will have some power of $N$ such that multiplying it by an additional dimensional factor of $N^{\frac{1}{d}}$ will give back $N$.
+This gives a surface volume of $N^\frac{d-1}{d}$, and with Poisson fluctuations being the square root of the mean we end up with fluctuations that go like $N^\frac{d-1}{2d}$.
+The action $S^{(d)}_{GHY}$, being proportional to $\rho^\frac{2-d}{d}$ times the difference of two independent\footnote{The independence is as a feature of the Poisson process in $M$: the number of elements $N(R)$ in any subregion $R\subset M$ is a Poisson variable itself, and for any two disjoint regions $R_1$, $R_2$ the numbers $N(R_1)$ and $N(R_2)$ are independent random variables.} random variables with standard deviation $N^\frac{d-1}{2d} = (\rho V)^\frac{d-1}{2d}$ should see fluctuations of order $\rho^\frac{2-d}{d}\rho^\frac{d-1}{2d}=\rho^\frac{3-d}{2d}$.
+This suggests that for $d=2$ these fluctuations should grow like $\rho^{\frac{1}{4}}$ as $\rho\rightarrow\infty$, for $d=3$ they should be constant, and for $d>3$ they should be damped.
 
-In order to test this heuristic argument one may like to write down the integral expression for $\text{Var}[S^{(d)}_{GHY}(\mathcal C)]^\frac12$ but it is complicated enough even in finite regions of flat space that it is not very illuminating to reproduce it here.
+In order to test this heuristic argument one may like to write down the integral expression for $\text{Var}[S^{(d)}_{GHY}(\mathcal C)]^\frac12$ but it is complicated enough even in finite regions of flat space for which it is not very illuminating to reproduce here.
 It also seems far less tractable than the expression for the mean because it involves terms of the type $\int d^dx\int d^dy \exp\left[-\rho V(J^+(x)\cap J^+(y))\right]$ for which the above methods of cutting off the time-integration range and Taylor expanding will not follow through.
@@
 Sprinkle at density $\rho=l^{-d}$ into the cube, which in any given run will place $N$ points inside the cube where $N$ is a Poisson random number with mean $\left\langle N\right\rangle = \rho V=  (L/l)^d$.
-Evaluate $S^{(d)}_{GHY}=c_d\rho^\frac{d-2}{d}(N_{min} - N_{max})$ for $\Sigma$ by counting the minimal/maximal elements in the upper/lower half of the cube, and repeat for $R$ runs.
+Evaluate $S^{(d)}_{GHY}=\rho^\frac{2-d}{d}\frac{c_d}{2\Gamma\left(\frac{2}{d}\right)}}(N_{min} - N_{max})$ for $\Sigma$ by counting the minimal/maximal elements in the upper/lower half of the cube, and repeat for $R$ runs.
 

### 2014-09-18 paper_v5.tex
initial commit

--- 2014-09-18 paper_v4.tex
@@
 \abstract{ 
-We propose a formula for the boundary term in the action of a causal set that is well-approximated by a continuum manifold with spacelike boundary.
-The boundary term is proportional to the difference in the number of elements immediately to future and the number of elements immediately to the past of the surface.
+We propose a family of formulae for the boundary term in the action of a causal set that is well-approximated by a continuum manifold with spacelike boundary.
+%The boundary term is proportional to the difference in the number of elements immediately to future and the number of elements immediately to the past of the surface.
 We show that in the continuum limit one recovers the Gibbons-Hawking-York boundary term in the mean.
@@
 \maketitle
-
-\noindent NOTES:\\
-1. define s as the RV, not as its mean
 
@@
 For instance, in the calculation of the black hole entropy via the Euclidean path-integral, it is the boundary terms that produce the answer $A/4l_p^2$ necessary for the unification of black hole mechanics with thermodynamics.
-To give another example, in Regge calculus boundary terms are necessary in order that the quantum mechanical amplitudes satisfy the correct composition law and that they have the correct classical limit\cite{boundarybook,hartlesorkin}.
-
-MB1: paragraph about boundary term in causets. In this paper we discuss a proposal for the  causal set analogue of the boundary action for \emph{spacelike} hypersurfaces.
+To give another example, in Regge calculus boundary terms are necessary in order that the quantum mechanical amplitudes satisfy the correct composition law and that they have the correct classical limit~\cite{hartlesorkin}.\mb{is this true?}
 
@@
 S_{GHY} = \pm \frac{1}{l_p^{d-2}}\int_{\Sigma} K d\Sigma
-\mb{lose $\pm$?}
 \ee
 where $l_p$ is the rationalised Planck length and $K$ is the trace of the extrinsic curvature $K_{\mu\nu}=h_{\mu}^\rho h_\nu^\sigma \nabla_\rho n_\sigma$ of $\Sigma$ defined with future-pointing timelike unit normal $n_{\mu}=\partial_\mu S/\sqrt{g^{\mu\nu}\partial_\mu S\partial_\nu S}$.
-This translates into a past-pointing normal vector $n^{\mu}$.
+(We will work with a mostly plus convention for the metric, in which this translates into a past-pointing normal vector $n^{\mu}$).\mb{ok or more detail on conventions needed?}
 
@@
 \ee
-where the right hand side is the derivative of the area $\int d\Sigma$ as each point of $\Sigma$ is moved an equal distance along the unit normal vector, $n^{\mu}$, which corresponds to a rate of change of the surface volume backwards in time, as $n^{\mu}$ is past-pointing.
-\mb{does the antichain defn really work?} In the causal set, the analogue of a spacelike hypersurface is an anti-chain, i.e.
+where the right hand side is the derivative of the area $\int d\Sigma$ as each point of $\Sigma$ is moved an equal distance along the unit normal vector, $n^{\mu}$, which corresponds to the rate of change of the surface volume backwards in time, as $n^{\mu}$ is past-pointing.
+\mb{does the antichain analogy really stand?} 
+In the causal set, the analogue of a spacelike hypersurface is an anti-chain, i.e.
 a subset $\mathcal{A}\subseteq\mathcal{C}$ in which all elements are unrelated.
 The intuitive analogue of the boundary term would then be the rate of change of the number of causal set elements below and above the antichain.
-We shall see that this intuitive idea indeed bears out.\mb{Write some stuff about antichains being a bit pathological and all that.} %This suggests a very simple analogy for a causal set with a ``spacelike hypersurface''.
+We shall see that this intuitive idea indeed bears out.
+%This suggests a very simple analogy for a causal set with a ``spacelike hypersurface''.
 \mb{under construction} The analogy These unrelated elements are maximal in the sense that there exist no elements $y\in\mathcal{C}\setminus\mathcal{A}$ such that $x\preceq y$, $\forall x\in\mathcal{A}$.
@@
 Figure \ref{fig:Nmin_Nmax} shows an illustration of the idea.
-The minimal and maximal elements have been highlighted in COLOUR.
+The minimal and maximal elements have been highlighted.
+%in COLOUR.
 
-We will also show that it is possible to restrict oneself to either $C^+$ or $C^-$ and still calculate the GHY term.
-To do this requires us to define the $k$-next-to-maximal/minimal elements in $C^-$ or $C^+$ respectively.
+We will also show that it is possible to restrict oneself to either $\mathcal C^+$ or $\mathcal C^-$ and still calculate the GHY term.
+To do this requires us to define the $k$-next-to-maximal/minimal elements in $\mathcal C^-$ or $\mathcal C^+$ respectively.
 These are the elements of the causal set that have $k$ elements to their future or past.
 This means that the $0$-to-maximal/minimal elements are the maximal/minimal elements.
-Elements with $1$ element to their past/future have been highlighted in COLOUR in Figure \ref{fig:Nmin_Nmax}, and those with $2$ have been highlighted in COLOUR.
+%Elements with $1$ element to their past/future have been highlighted in COLOUR in Figure \ref{fig:Nmin_Nmax}, and those with $2$ have been highlighted in COLOUR.
 
 Let us denote the number of $k$-next-to-maximal elements in $\mathcal C^-$ by $N_{max}^{(k)}$ and the number of $k$-next-to-minimal elements in $\mathcal C^+$  by $N_{min}^{(k)}$.
-
-We propose the following definitions for the discrete Gibbons-Hawking-York boundary term (one can adopt the appropriate definition to best suit their calculation needs):\mb{$S[\mathcal C, \Sigma]$?}
+We propose the following definitions for the discrete Gibbons-Hawking-York boundary term (one can adopt the appropriate definition to best suit their calculation needs):\mb{put $\Sigma$ in argument, $S[\mathcal C, \Sigma]$?}
 
 \be\label{GH_boundary_to_causet}
- \mathcal{S}^{(d)}_{GHY}[\mathcal C]=\left(l/l_p\right)^{d-2}\frac{c_{d}}{2\: \Gamma\left(\frac{2}{d} \right)}\times
+ \mathcal{S}^{(d)}_{GHY}[\mathcal C]=\left(l/l_p\right)^{d-2}\frac{c_{d}}{\Gamma\left(\frac{2}{d} \right)}\times
   \begin{cases}
-   N_{max}^{(0)}-N_{min}^{(0)} \\
+   \frac{1}{2}\left(N_{max}^{(0)}-N_{min}^{(0)}\right) \\
    d\: N_{max}^{(1)}-N_{max}^{(0)} \\
@@
 $A_d=2\pi^\frac{d+1}2/\Gamma\left(\frac{d+1}{2}\right)$ and denotes the volume of the unit $d$-sphere.
-In fact, one can write down a general expression for the boundary term as a sum over $k$ of $N_{max/min}^{\: (k)}$ terms with particular coefficients.
+In fact, one can write down a general expression for the boundary term as a sum over $k$  $N_{\vphantom{i}max}^{\: (k)}$ and/or $N_{min}^{\: (k)}$ terms with particular coefficients.
 This can be written as follows:
@@
 Thus, with only two terms, the form of the discrete boundary term is fixed.
-If one chooses to include more than two terms the coefficients of the $N_{{\,}_{max/min}}^{\:(k)}$ terms are no longer unique.
+If one chooses to include more than two terms the coefficients of the $N_{{max\vphantom{i}}}^{\:(k)}$ and/or $N_{{min}}^{\:(k)}$ terms are no longer unique.
 
@@
 \be
-\lim_{l\rightarrow0}\left\langle S^{(d)}_{GHY}[\mathcal C] \right\rangle= \frac{1}{l_p^{d-2}}\int_{\Sigma} d^{d-1}x\: \sqrt{h}\: K\label{eq:mainconjecture}
+\lim_{l\rightarrow0}\langle S^{(d)}_{GHY}[\mathcal C] \rangle= \frac{1}{l_p^{d-2}}\int_{\Sigma} d^{d-1}x\: \sqrt{h}\: K\label{eq:mainconjecture}
 \ee
-for the most general definition above. Here $\left\langle\cdot\right\rangle$ denotes the mean over sprinklings.
+for the most general definition above. Here $\langle\cdot\rangle$ denotes the mean over sprinklings.
 
@@
 \be\label{eq:conjecture_for_area}
-\lim_{l\rightarrow0}\left\langle \mathcal{A}^{(d)}[\mathcal{C}] \right\rangle= \frac{1}{l_p^{d-1}}\int_{\Sigma} d^{d-1}x\: \sqrt{h}
+\lim_{l\rightarrow0}\langle \mathcal{A}^{(d)}[\mathcal{C}] \rangle= \frac{1}{l_p^{d-1}}\int_{\Sigma} d^{d-1}x\: \sqrt{h}
 \ee
@@
 \be\label{eq:nmax}
-\left\langle N_{max}^{(k)}\right\rangle =\rho\int_{J^{-}(\Sigma)}dV_p\; \frac{\left(\rho\: V_\blacktriangledown(p)\right)^k}{k!}e^{-\rho V_\blacktriangledown(p)}
+\langle N_{max}^{(k)}\rangle =\rho\int_{J^{-}(\Sigma)}dV_p\; \frac{\left(\rho\: V_\blacktriangledown(p)\right)^k}{k!}e^{-\rho V_\blacktriangledown(p)}
 \ee
@@
 \be\label{eq:nmin}
-\left\langle N_{min}^{(k)}\right\rangle =\rho\int_{J^{+}(\Sigma)}dV_p\; \frac{\left(\rho\: V_\blacktriangle(p)\right)^k}{k!}e^{-\rho V_\blacktriangle(p)}
+\langle N_{min}^{(k)}\rangle =\rho\int_{J^{+}(\Sigma)}dV_p\; \frac{\left(\rho\: V_\blacktriangle(p)\right)^k}{k!}e^{-\rho V_\blacktriangle(p)}
 \ee
@@
 However, for large $\rho$ (small $l$), the integrands $\exp(-\rho V)$ will be exponentially suppressed unless the volumes are small.
-Consider the spacetime region $U_\epsilon=\left\{|t|<\epsilon\right\}$ around the surface $\Sigma$ for some $\epsilon>0$ (with $\epsilon$ small enough such that the GNC system is valid throughout $U_\epsilon$).
-We will assume that for any such $\epsilon$, the contribution from points with $|t|>\epsilon$ to the integrals~\eqref{eq:nmax} and \eqref{eq:nmin} can be made arbitrarily small by setting $\rho$ large enough, since the volumes of the truncated past/future lightcones for such points will be too large.
-Hence, as $\rho\rightarrow\infty$, the integration ranges in \eqref{eq:nmax} and \eqref{eq:nmin} can be cut off at finite time $t=\pm\epsilon$ with $\epsilon$ arbitrarily small, which allows us to expand the time integration-variable about zero.
-These assumptions of course impose certain regularity conditions on the surface $\Sigma$, but we will assume that they are reasonable conditions.\mb{say more about this?
-we can't have an asymptotically null surface i suppose} The integrals then simplify to
+Consider the spacetime region $U_\varepsilon=\left\{|t|<\varepsilon\right\}$ around the surface $\Sigma$ for some $\varepsilon>0$ (with $\varepsilon$ small enough such that the GNC system is valid throughout $U_\varepsilon$).
+We will assume that for any such $\varepsilon$, the contribution from points with $|t|>\varepsilon$ to the integrals~\eqref{eq:nmax} and \eqref{eq:nmin} can be made arbitrarily small by setting $\rho$ large enough, since the volumes of the truncated past/future lightcones for such points will be too large.
+Hence, as $\rho\rightarrow\infty$, the integration ranges in \eqref{eq:nmax} and \eqref{eq:nmin} can be cut off at finite time $t=\pm\varepsilon$ with $\varepsilon$ arbitrarily small, which allows us to expand the time integration-variable about zero.
+These assumptions of course impose certain regularity conditions on the surface $\Sigma$, but we will assume that they are reasonable conditions.\mb{do we need to say more about this?
+e.g.
+can't have asymp.
+null $\Sigma$ i guess} The integrals then simplify to
 \begin{gather}\label{eq:nmax_and_eq:nmin}
 \begin{aligned}
-\left\langle N_{max}^{(k)}\right\rangle & =\rho \int_{\Sigma}d^{d-1}x\int_{-\varepsilon}^{0}dt\:
+\langle N_{max}^{(k)}\rangle & =\rho \int_{\Sigma}d^{d-1}x\int_{-\varepsilon}^{0}dt\:
 h^{\frac{1}{2}}\left(1+
@@
 \\
-\left\langle N_{min}^{(k)}\right\rangle & =\rho \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
+\langle N_{min}^{(k)}\rangle & =\rho \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
 h^{\frac{1}{2}}\left(1+
@@
 
-The only truncated cones that contribute will have small volumes, as we are restricting ourselves close to the surface.\mb{this sounds circular w.r.t.
-last paragraph} This means that the volumes can be approximated by the volumes of less curvy cones, and it can be shown that higher order corrections, in the approximation scheme, vanish in the limit of $\rho \rightarrow \infty$.
+The only truncated cones that contribute will have small volumes, as we are restricting ourselves close to the surface.
+%\mb{this sounds circular w.r.t. last paragraph} 
+This means that the volumes can be approximated by the volumes of less curvy cones, and it can be shown that higher order corrections, in the approximation scheme, vanish in the limit of $\rho \rightarrow \infty$.
 This will all be made more precise below.
@@
 RNCs centered at $p_0$ will be given the symbol $y^{\overline{\mu}}$.
-\mb{let's put all facts about $x^\mu$ and $y^{\overline\mu}$ for $p$ and $p_0$ here and expansion of det in RNC.} We need to assume that for any $p$ in $U_\epsilon$, the Riemann normal neighbourhood $U_p\subset M$ (throughout which the RNC system centered at $p_0$ is well-defined) contains the truncated cone $J^-(p)\cap J^+(\Sigma)$.
-This assumption seems reasonable given the that $U_\epsilon$ can be made arbitrarily ``thin'' as $\rho\rightarrow\infty$.
+%\mb{let's put all facts about $x^\mu$ and $y^{\overline\mu}$ for $p$ and $p_0$ here and expansion of det in RNC.} 
+We need to assume that for any $p$ in $U_\varepsilon$, the Riemann normal neighbourhood $U_p\subset M$ (throughout which the RNC system centered at $p_0$ is well-defined) contains the truncated cone $J^-(p)\cap J^+(\Sigma)$.
+This assumption seems reasonable given the that $U_\varepsilon$ can be made arbitrarily ``thin'' as $\rho\rightarrow\infty$.
 In RNCs the metric and the Christoffel symbols at $p_0$ are those of flat space:
@@
 For the transformation from GNCs to RNCs at $p_0$ one finds that $A^{\overline 0}_{\;0}=1$, $A^{\overline 0}_{\;i}=0$ and $\delta_{\ibar\jbar}=A^i_{\;\ibar}A^j_{\;\jbar}h_{ij}(p_0)$, which in particular implies $y^{\overline{0}}_p=\overline{t}_p=t_p=T$.
-\mb{more detail needed i think}In RNC one can expand the metric determinant in~\eqref{eq:VolumeWithNoSimplifications} to find
+In RNC one can expand the metric determinant in~\eqref{eq:VolumeWithNoSimplifications} to find
 
@@
 \begin{aligned}
-\left\langle N_{max}^{(k)}\right\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{-\varepsilon}^{0}dt\:
+\langle N_{max}^{(k)}\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{-\varepsilon}^{0}dt\:
 h^{\frac{1}{2}}\left(1+
@@
 \\
-\left\langle N_{min}^{(k)}\right\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
+\langle N_{min}^{(k)}\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
 h^{\frac{1}{2}}\left(1+
@@
 \begin{aligned}
-\left\langle N_{max}^{(k)}\right\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
+\langle N_{max}^{(k)}\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
 h^{\frac{1}{2}}\left(1+
@@
 \\
-\left\langle N_{min}^{(k)}\right\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
+\langle N_{min}^{(k)}\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
 h^{\frac{1}{2}}\left(1-
@@
 The reason for expanding the exponentials is that the integrals will have to be put into the form of Gaussian integrals in order to evaluate them in the limit $\rho\rightarrow\infty$.
-The above expressions are almost the same apart from a few sign differences, so from now on we will just work with the expression for $\left\langle N_{min}^{(k)}\right\rangle$.
+The above expressions are almost the same apart from a few sign differences, so from now on we will just work with the expression for $\langle N_{min}^{(k)}\rangle$.
 By expanding the brackets, and retaining only necessary orders, the integral can be split into two parts of differing powers of $\rho$.
@@
 \begin{aligned}
-\left\langle N_{min}^{(k)}\right\rangle & = \frac{\rho^{k+1}A^k}{k!}\int_{\Sigma}d^{d-1}x\: h^{\frac{1}{2}}\int_{0}^{\varepsilon}dt\:
+\langle N_{min}^{(k)}\rangle & = \frac{\rho^{k+1}A^k}{k!}\int_{\Sigma}d^{d-1}x\: h^{\frac{1}{2}}\int_{0}^{\varepsilon}dt\:
 \left\lbrace \left( t^{dk} + \left(kB-K \right)t^{dk+1}\right) + O\left(t^{dk+2}\right) \right\rbrace e^{-\rho A\: t^d}
@@
 We can use this procedure to take the limit of~\eqref{eq:n_min_different_rho_terms} and formulate the answer in terms of gamma functions.
-The limits of both $\left\langle N_{min}^{(k)}\right\rangle$ and $\left\langle N_{min}^{(k)}\right\rangle$ are then given by
+The limits of both $\langle N_{min}^{(k)}\rangle$ and $\langle N_{min}^{(k)}\rangle$ are then given by
 
@@
 \begin{aligned}
-\lim_{\rho\rightarrow\infty}\left\langle N_{max}^{(k)}\right\rangle = \lim_{\rho\rightarrow\infty} & \left\lbrace \rho^{1-\frac{1}{d}} \left(b_d\right)^{-1} \frac{\Gamma\left(\frac{1}{d}+k\right)}{k!}
+\lim_{\rho\rightarrow\infty}\langle N_{max}^{(k)}\rangle = \lim_{\rho\rightarrow\infty} & \left\lbrace \rho^{1-\frac{1}{d}} \left(b_d\right)^{-1} \frac{\Gamma\left(\frac{1}{d}+k\right)}{k!}
 \int_{\Sigma}d^{d-1}x\: \sqrt{h} \right.
@@
 \\
-\lim_{\rho\rightarrow\infty}\left\langle N_{min}^{(k)}\right\rangle = \lim_{\rho\rightarrow\infty} & \left\lbrace \rho^{1-\frac{1}{d}} \left(b_d\right)^{-1} \frac{\Gamma\left(\frac{1}{d}+k\right)}{k!}
+\lim_{\rho\rightarrow\infty}\langle N_{min}^{(k)}\rangle = \lim_{\rho\rightarrow\infty} & \left\lbrace \rho^{1-\frac{1}{d}} \left(b_d\right)^{-1} \frac{\Gamma\left(\frac{1}{d}+k\right)}{k!}
 \int_{\Sigma}d^{d-1}x\: \sqrt{h} \right.
@@
 The other terms all have negative powers of $\rho$ and therefore vanish as $\rho\rightarrow\infty$.
-From this it follows that the mean of any linear combination of terms like $\rho^{\frac{1}{d}-1}\left\langle N_{max}^{(k)}\right\rangle$ or $\rho^{\frac{1}{d}-1}\left\langle N_{min}^{(k)}\right\rangle$ will give something proportional to the surface volume, so long as the terms containing the surface volume don't exactly cancel.
+From this it follows that the mean of any linear combination of terms like $\rho^{\frac{1}{d}-1}\langle N_{max}^{(k)}\rangle$ or $\rho^{\frac{1}{d}-1}\langle N_{min}^{(k)}\rangle$ will give something proportional to the surface volume, so long as the terms containing the surface volume don't exactly cancel.
 To simplify the coefficients in this linear sum we take out a factor of $\frac{k!}{\Gamma\left(\frac{1}{d}+k\right)}$.
@@
 In any spacetime region of fixed volume $V$ the number of causal set elements $N$ experiences Poisson fluctuations of order $\sqrt N$.
-We will use the first case of~\eqref{GH_boundary_to_causet} as it is the simplest case in which to form a heuristic argument for the fluctuations.
-Since $N_{max}^{(0)}$ and $N_{min}^{(0)}$ are quantities associated with a codimension-1 surface (they count elements ``near" $\Sigma$), we may expect them to inherit fluctuations of order $N^\frac{d-1}{2d}$.
-This comes from the fact that any one dimension will attribute $N^{\frac{1}{d}}$ points to the total volume $N$ (as $(N^{\frac{1}{d}} )^d=N$).
-Thus, a codimension-1 surface will have some power of $N$ such that multiplying it by an additional dimensional factor of $N^{\frac{1}{d}}$ will give back $N$.
-This gives a surface volume of $N^\frac{d-1}{d}$, and with Poisson fluctuations being the square root of the mean we end up with fluctuations that go like $N^\frac{d-1}{2d}$.
+We will use the first case of~\eqref{GH_boundary_to_causet} as it is the simplest case in which to make this argument for the fluctuations.
+Since $N_{max}^{(0)}$ and $N_{min}^{(0)}$ are quantities associated with a codimension-1 surface (they count elements ``near" $\Sigma$), we may expect them to scale like $N^\frac{d-1}{d}$ --- which agrees with the leading order behaviour of~\eqref{eq:nmax_nmin_final} --- and thus to inherit fluctuations of order $N^\frac{d-1}{2d}$.
+%{This comes from the following observation.
+In a volume corresponding to a thickening of the hypersurface $\Sigma$ by one unit of the discreteness scale $l$ (e.g.
+by Lie dragging the surface along its normal by an amount $l$),....
+}
 The action $S^{(d)}_{GHY}$, being proportional to $\rho^\frac{2-d}{d}$ times the difference of two independent\footnote{The independence is as a feature of the Poisson process in $M$: the number of elements $N(R)$ in any subregion $R\subset M$ is a Poisson variable itself, and for any two disjoint regions $R_1$, $R_2$ the numbers $N(R_1)$ and $N(R_2)$ are independent random variables.} random variables with standard deviation $N^\frac{d-1}{2d} = (\rho V)^\frac{d-1}{2d}$ should see fluctuations of order $\rho^\frac{2-d}{d}\rho^\frac{d-1}{2d}=\rho^\frac{3-d}{2d}$.
@@
 
+\begin{figure}[t!]
+  \centering
+    {\includegraphics[scale=0.6]{GHY-fluctuation_plots}}
+     \caption{A $\log-\log$ plot (base $2$) of the fluctuations in $S^{(d)}_{GHY}$ for a flat ($K=0$) surface in $d=2,3$ and $4$-dimensional Minkowski space.
+Each data point represents the sample standard deviation in $S^{(d)}_{GHY}$ on a sample of $R=100$ runs.
+Black dots, blue triangles and red squares correspond to the simulation results in $d=2,3$ and $4$ dimensions, respectively.
+The corresponding black, blue and red lines have gradients $\frac12$, $0$ and $-\frac18$ and best-fit intercepts of order $1$.}
+     \label{fig:fluctuations}
+\end{figure}
+
 The simulations were carried out as follows.
 Denote the discreteness scale by $l$.
-Take a $d$-cube $[0,L]^d$ in $d$-dimensional Minkowski space with metric $ds^2=-dt^2+d{\mathbf x}^2$ and define the hypersurface $\Sigma: t=L/2$, which partitions the cube into two halves.
-Sprinkle at density $\rho=l^{-d}$ into the cube, which in any given run will place $N$ points inside the cube where $N$ is a Poisson random number with mean $\left\langle N\right\rangle = \rho V=  (L/l)^d$.
-Evaluate $S^{(d)}_{GHY}=\rho^\frac{2-d}{d}\frac{c_d}{2\Gamma\left(\frac{2}{d}\right)}}(N_{min} - N_{max})$ for $\Sigma$ by counting the minimal/maximal elements in the upper/lower half of the cube, and repeat for $R$ runs.
+Take a $d$-cube $[0,L]^d$ in $d$-dimensional Minkowski space with metric $ds^2=-dt^2+d{\mathbf x}^2$ and define the hypersurface $\Sigma: t=L/2$, which partitions the cube into two equal halves.
+Sprinkle at density $\rho=l^{-d}$ into the cube, which in any given run will place $N$ points inside the cube where $N$ is a Poisson random number with mean $\langle N\rangle = \rho V=  (L/l)^d$.
+Evaluate $S^{(d)}_{GHY}=\rho^\frac{2-d}{d}c_d\left(2\Gamma\left(\frac{2}{d}\right)\right)^{-1}(N_{max} - N_{min})$ for $\Sigma$ by counting the minimal/maximal elements in the upper/lower half of the cube, and repeat for $R$ runs.
+The expected value for the mean is zero in all cases, since the extrinsic curvature of the surface is zero.
 
+In Figure~\ref{fig:fluctuations} we display the simulation results for $d=2,3,4$ spacetime dimensions, with $N=\rho$ ranging up to $2^{20}\approx 1$ million, having set $V=L^d=1$.
+Each data point represents the fluctuations (the standard deviation) in a sample of $R=100$ runs.
+The solid lines have been obtained by fitting an arbitrary constant multiplier in the scaling law predicted by the argument above, $\gamma(d)\times N^\frac{3-d}{2d}$, to the data sets.
+The best fit values are all of order $1$: $\gamma(2)=0.63$, $\gamma(3)=0.91$, and $\gamma(4)=1.07$.
+There is clearly a good fit between the data and the scaling predicted by the heuristic argument. Although not visible on the plot, the data points were all consistent with zero mean as well.
+
+We have performed analogous simulations for the second formula in~\eqref{GH_boundary_to_causet}, involving $N^{(1)}_{max}$ and $N^{(0)}_{max}$.
+In this case, heuristic arguments of the type given above are hampered by the fact that the two random variables whose difference is taken are not independent.
+However, the results for the scaling of the fluctuations turn out to be identical to those obtained in the first case.
 \newpage

### 2014-09-18 paper_v6.tex
initial commit

--- 2014-09-18 paper_v5.tex
@@
 \newcommand{\jbar}{{\overline{\emph\j}}}
+\newcommand{\mnote}[1]{\marginpar{\raggedright \scriptsize   #1 } }
 
 \title{Boundary terms in the action for Causal Sets}
- \author[a]{M. Buck}
- \author[a,b]{\!, F. Dowker}
- \author[a]{and I. Jubb\,}
+ \author[a]{Michel Buck}
+ \author[a,b]{\!, Fay Dowker}
+ \author[a]{ Ian Jubb\,}
+ \author[c]{and Sumati Surya}
 \affiliation[a]{Theoretical Physics Group, Blackett Laboratory, Imperial College, London, SW7 2AZ, U.K.}
 \affiliation[b]{Institute for Quantum Computing, University of Waterloo, ON, N2L 2Y5, Canada}
-
+\affiliation[c]{Raman Research Institute, CV Raman Ave, Sadashivanagar, Bangalore 560080, India}
 \abstract{ 
@@
 However, the results for the scaling of the fluctuations turn out to be identical to those obtained in the first case.
+
+\section{A Candidate Boundary Term for an Alexandrov Interval}
+
+Until now our approach has been to reconstruct the  continuum GHY term  from the information in a causal set. 
+The candidate described in the previous sections for a spatial boundary has a natural expression in a causal set  in  terms of its maximal and minimal elements which are the  analogs of a  space-like boundary.
+A suitable analog  of  a time-like boundary  term has been harder to  construct; there is no simple causal set  characterisation of a time-like boundary.
+Indeed, it seems difficult to characterise a causal set in a  spacetime region which is bounded by time-like surfaces.
+On the other hand, since past and future sets are natural in a causal set, such regions do find a natural characterisation in a causal set.
+An important example of such a region is the  Alexandrov interval which is used extensively  in calculations of causal set kinematics.
+In the continuum such a region  is bounded by null hypersurfaces and their spatial joints.
+
+Conversely in the continuum,   while the GHY term  is well defined for both a space-like and time-like boundary,  including their co-dimension 2 joints \cite{harris},  it is an open question what the contribution from a null  boundary  is \cite{ghh}.
+Apart from some sketchy suggestions in the literature \cite{neiman} there is no rigourous construction of such a term; indeed the standard variational route for finding the boundary contribution to the action appears to fail for null boundaries.
+
+Given the lack of a  null GHY term, we can ask if causal set theory  might  in fact guide us in the right direction.
+Our starting point is the $d$-dimensional  Benincasa-Dowker Action \cite{bd,gd} which is constructed from the mean number  $N_i^{(d)}$ \footnote{ $N_i^{(d)}$ is dimension  and geometry dependent, but for our purpose it suffices to retain the dimension label.
+} of $i+1$-element inclusive intervals in the causal set
+\begin{equation} 
+8 \pi G {S_{BD}^d} = -\alpha_d \rho^{\frac{2-d}{d}}\biggl( N+ \frac{\beta_d}{ \alpha_d} \sum_{i=1}^{n_d-1}  C^{(d)}_{i} N _i^{(d)} \biggr) 
+\label{bd} 
+\end{equation}  
+where   
+\begin{equation} 
+ \alpha_d= \left \{ %
+\begin{array} {ccc}
+-\frac{1}{\Gamma(\frac{2+d}{d})}c_d ^{2/d}  & \quad &  \mathrm{d} \, \, \mathrm{odd} \\
+-2 \frac{1}{\Gamma(\frac{2+d}{d})}c_d ^{2/d} & \quad &  \mathrm{d \, \,  even} \\
+\end{array} \right  
+\quad \beta_d = \left \{ 
+\begin{array} {ccc}
+\frac{d+1}{2^{d-1}\Gamma(2/d +1)} c_d^{2/d}  & \quad &  \mathrm{d \, \, odd}\\ 
+\frac{\Gamma(d/2+2)\Gamma(d/2+1)}{\Gamma(2/d)\Gamma(d)}c_d^{2/d}  & \quad & \mathrm{d \, \,  even}\\ 
+\end{array} \right 
+\end{equation} 
+\begin{equation} 
+n_d = \left \{ %
+\begin{array} {ccc}
+\frac{d-1}{2} + 2 & \quad &  \mathrm{d \, \, odd}\\ 
+\frac{d}{2} + 2 & \quad & \mathrm{d \, \,  even}\\ 
+\end{array} \right 
+\end{equation} 
+with $2^{-d/2} c_d$ is the volume of a unit Alexandrov interval $I(x,y)$, i.e., one for which the proper time $\tau(x,y)=1$. 
+%%{ {\bf \color{red} this is not correct! $c_d= vol(S_{d-2}) \frac{ 2^{\frac{2-d}{2}}}{d(d-1)}$ and $vol(S_{d-2})$ the volume of the homogeneous $d-2$ sphere}}.  
+The coefficients  of $N_i^{(d)}$ in the sum are are 
+\begin{equation}
+\label{cid}
+ C_i^d= \left \{ %
+\begin{array} {ccc}
+\sum_{k=0}^{i-1} (-1)^k\binom{i-1}{k} \frac{\Gamma(d/2(k+1)+3/2)}{\Gamma(d/2+3/2) \Gamma(1+dk/2)} & \quad &  \mathrm{d \, \, odd}\\ 
+\sum_{k=0}^{i-1} (-1)^k\binom{i-1}{k} \frac{\Gamma(d/2(k+1)+2)}{\Gamma(d/2+2) \Gamma(1+dk/2)} & \quad & \mathrm{d \, \,  even}\\ 
+\end{array} \right 
+\end{equation} 
+These coefficients can be expressed more compactly as hypergeometric functions 
+\begin{equation}
+C_i^d={}_{q+1}F_{q}(\{ a_1, \ldots a_q, i-1\}, \{b_1, \ldots b_q \} |1)
+\label{chyp} 
+\end{equation} 
+with $q=d/2+1/2, a_i=\frac{d+2i}{d}, b_i=2i/d$ for $d$ odd and $q=d/2, a_i=\frac{d+2i+2}{d}, b_i=2i/d$  for $d$ even.
+
+If $S_{BD}^d$ were precisely the discrete analog of the Einstein Hilbert action {\it without} any boundary contributions, then in Minkowski spacetime  $S_{BD}^d$ should  vanish.
+However, as shown explicitly in \cite{bbdtwo} this is not the case for Alexandrov intervals in 2$d$  Minkowski spacetime.
+Indeed, there is a residual contribution that goes to a constant in the continuum limit $N\rightarrow \infty$.
+When the action is evaluated in rectangular regions in the same spacetime or for Alexandrov intervals in  the flat trousers spacetime, this is no longer the case.
+Thus, it seems plausible that the 2$d$ BD action contains some boundary information.
+Moreover, in the large $N$ limit  the action goes to the same  non-vanishing constant irrespective of the interval size.
+While this might suggest a non-geometric or topological origin, we will now show that it is a part of a universal  behaviour for higher $d$ and in fact  has a geometrical origin.
+
+Consider an Alexandrov interval $I(p,q)=I^+(p) \cap I^-(q)$ for two events $p,q$ in a causal spacetime.
+While the  boundary of $I(p,q)$ may be quite complicated in general, some of these components must necessarily be null.
+In Minkowski spacetime, for example,  the boundary of $I(p,q)$ consists of the two events $p$ and $q$, the subsets $B^\pm$ of the  two null hypersurfaces $\partial J^+(p)$ and $\partial J^-(q)$, and their intersection or ``joint''  $S_{d-2}=\partial J^+(p) \cap \partial J^-(q)$ which is a codimension 2 sphere.
+Evaluating the BD action for an Alexandrov interval in flat spacetime should therefore give us terms that are only dependent on the boundary components.
+In flat spacetime the $S_{d-2}$ joint  has radius $\tau(p,q)/2$, and thus increases with the size of the Alexandrov interval ${\mathrm{vol}}(I(p,q)) = N/\rho$, where $N$ is the mean of the number of causal set elements sprinkled into $I(p,q)$ .
+In what follows we take the continuum limit $\rho, N \rightarrow \infty$ while keeping ${\mathrm{vol}}(I(p,q))$ fixed.
+\mnote{\color{red} SS: It would be nice to have a fancy picture of an Alex.
+interval here..}
+%% If $p$ or $q$ were to be taken to time-like past or future infinity, this volume would go to infinity as expected.
+
+In \cite{glasersurya} a closed form expression was obtained for the mean value of the number of $i+1$ element inclusive intervals  $N_i^d$ contained in an  Alexandrov interval in $d$ dimensional Minkowski  spacetime
+\begin{equation} 
+N_i^d= \frac{\Gamma(d)^2}{\Gamma(i)} \sum_{k=0}^\infty \frac{(-1)^k}{\Gamma(k+1)} \frac{N^{k+i+2}}{(k+i+2)(k+i+1)} \frac{\Gamma(\frac{d(k+i)}{2}+1) \Gamma(\frac{d(k+i+1)}{2}+1)}{\Gamma(\frac{d(k+i)}{2} +d) \Gamma(\frac{d(k+i+1)}{2}+d)},
+\end{equation}
+where $i \geq 1$.   In taking the $N \rightarrow \infty$ limit, it   is convenient to consider the ratio
+\begin{equation} 
+\frac{ 8 \pi G S^d(C)}{vol(S^{d-2})} = \frac{b}{d(d-1) \Gamma(1+2/d) N^{\frac{d-2}{d}}} \left( N + \frac{\beta_d}{\alpha_d} \sum_{i=1}^{n_d} C_i^d N_i^d \right)  
+\label{ratio} 
+\end{equation} 
+where $b=1$ for $d$ odd and $b=2$ for $d$ even. 
+We use Mathematica to evaluate the $N \rightarrow \infty$  limit of this expression for $d=2, \ldots 9$.
+We aid the calculation by expanding the sum  $\sum_{i=1}^{n_d} C_i^d N_i^{(d)} = \sum_{j=1}^\infty A_j^d N^{j+1}$ as a  power series in  $N$.
+After a rearrangement and redefinition of indices we find that
+\begin{equation}
+ A_j^d = \Gamma(d)^2\biggl( \sum_{i=1}^{D+2} (-1)^i \binom{j-1}{i-1}  C_i^d  \biggr) \frac{(-1)^j}{(j+1)!}\frac{\Gamma(\frac{d}{2}(j-1)+1)\Gamma(\frac{d}{2}j+1)}{\Gamma(\frac{d}{2}(j-1)+d) \Gamma(\frac{d}{2}j+d)},
+\end{equation} 
+where $d=2D$ for $d$ even and $d=2D+1$ for $d$ odd.   
+% Using the abbreviated form $C_i^d=\sum_{l=0}^{i-1} \binom{i-1}{l}A(d,l)$ where $A(d,l)$ is defined using \ref{cid} 
+% we find that 
+% \begin{equation}
+% \sum_{i=1}^{D+2} (-1)^i \binom{j-1}{i-1}  C_i^d = (-1)^{j-1} A(d,j-1),  
+% \end{equation}      
+% where we have used the binomial identities  BLAH. Thus 
+% \begin{equation}
+%  \sum_{i=1}^{n_d} C_i^d N_i^{(d)}= ((d-1)!)^2 \sum_{j=1}^\infty(-1)^j B(d,j) \frac{\Gamma(\frac{d}{2}j+1)}{\Gamma(j+2)\Gamma(\frac{d}{2}(j-1)+d)\Gamma(\frac{d}{2}j+d)}N^{j+1}
+% \label{thesum} 
+% \end{equation} 
+% where 
+% \begin{equation} 
+% B(d,j) = \left \{ 
+% \begin{array}{ccc} 
+% \frac{\Gamma(\frac{d}{2}j+3/2)}{\Gamma(\frac{d}{2}+3/2)}  &\quad & d \, \, odd, \\  \frac{\Gamma(\frac{d}{2}j+2)}{\Gamma(\frac{d}{2}+2)}  &\quad & d \, \, even.  \\ 
+% \end{array} 
+% \right
+% \end{equation} 
+An explicit  calculation of $8\pi S_{BD}^d$ in $I(p,q)$ for $d=2, \ldots 9$  gives 
+\begin{equation} 
+\lim_{N \rightarrow \infty} \frac{ 8 \pi G S_{BD}^d(C)}{vol(S^{d-2})} = 1,  
+\label{result} 
+\end{equation}
+
+\mnote{\color{red} SS: There are further simplifications I have that may help mathematica for $n>9$.
+It is still not clear}  Since $N_i^d$ can also be expressed as a hypergeometric function, using  (\ref{chyp}) it   is seems possible to evaluate (\ref{ratio}) using identities involving sums of products of such functions to get a closed form expression for arbitrary $d$.
+In \cite{glasersurya} the asymptotic form of $N_i^d$ was explicitly calculated, and for $d>2$, the leading order was shown to  $~ N^{2-\frac{2}{d}}$ which dominates $N$ dependence for $d>2$.
+To be consistent with  (\ref{result}) this would mean that the leading order contributions from the $N_i^d$ cancel in the action.
+This is indeed explicitly seen to be the case, again for $d=3, \ldots 9$.
+Similarly, for $d=2$, where the leading order contribution is $~ N \log N$.
+Indeed, in all cases the next to leading order terms given in \cite{glasersurya} also do not have the requisite $~N^{\frac{d-2}{d}}$ dependence required by (\ref{result}); in this case, the explicit coefficients are not available to   check that they give a vanishing contribution to (\ref{result}).
+This suggests that  we cannot rely   on the the asymptotic form of the individual terms  in the sum  for an analytic proof for general $d$, which could therefore be quite non-trivial.
+
+The result we have obtained is for flat spacetime.
+One may speculate that the presence of curvature will not modify the essential result, i.e., that the only boundary contribution is from the volume of the spatial $d-2$ sphere at the joint, at least in the case when the boundary consists of the same type of components.
+What will change of course is the fact that the principal radii of the $d-2$ sphere could be different at  every point, hence complicating the relationship between the volume of the interval and the volume of this sphere.
+In  \cite{Khetrapal_Sumati:Causal_Diamond_Volume} it was shown that in an RNC, to first order in curvature,
+the intrinsic boundary geometry of the Alexandrov interval is still flat which would simplify this relationship.
+Using an RNC version of the calculation  one could thus make an explicit calculation possible.
+
+The fact that the BD action does not vanish in flat spacetime suggests  that this action  does not only code information about the bulk Einstein Hilbert action but also contains  extra boundary information in general.
+One might then be tempted to speculate that it contains {\it all} the relevant boundary information suggesting that the null boundary contribution for an Alexandrov interval in any spacetime always vanishes.
+Moreover, in the specific case that there is only a single spatial component coming from the joint, that the only boundary contribution is the volume of this $d-2$ sphere.
+comes from this  that (\ref{result}) gives the GHY contribution for an Alexandrov interval.
+
 \newpage

### 2014-09-18 paper_v7.tex
initial commit

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{jheppub}
\usepackage{marginnote,xparse,changepage,caption,subfigure}

\newcommand{\IGNORE}[1]{}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bea}{\begin{aligned}}
\newcommand{\eea}{\end{aligned}}
\newcommand{\mb}[1]{\marginnote{\texttt{\small MB:\,#1}}}
\renewcommand{\ij}[1]{\marginnote{\texttt{\small IJ:\,#1}}}
\newcommand{\ibar}{{\overline{\emph\i}}}
\newcommand{\jbar}{{\overline{\emph\j}}}

\title{Boundary terms in the action for Causal Sets}
 \author[a]{Michel Buck}
 \author[a,b]{\!, Fay Dowker}
 \author[a]{\!, Ian Jubb\,}
 \author[c]{and Sumati Surya}
\affiliation[a]{Theoretical Physics Group, Blackett Laboratory, Imperial College, London, SW7 2AZ, U.K.}
\affiliation[b]{Institute for Quantum Computing, University of Waterloo, ON, N2L 2Y5, Canada}
\affiliation[c]{Raman Research Institute, CV Raman Ave, Sadashivanagar, Bangalore 560080, India}

\abstract{ 
We propose a family of formulae for the boundary term in the action of a causal set that is well-approximated by a continuum manifold with spacelike boundary. %The boundary term is proportional to the difference in the number of elements immediately to future and the number of elements immediately to the past of the surface. 
We show that in the continuum limit one recovers the Gibbons-Hawking-York boundary term in the mean.
}

\begin{document}

\maketitle

\section{Introduction}

%[MB1: USEFUL QUOTE FROM RAFAEL: In the quantum theory, however, these boundary terms are important. They are essential in order that the quantum mechanical amplitudes satisfy the correct composition law and in order that these amplitudes have the correct classical limit. In a number of examples [5] they give a contribution to the partition function which is important for the agreement with calculations based on straightforward thermodynamics. In this note we shall derive the boundary terms in the action for the Regge calculus]

In furthering causal set theory it is crucial that we understand the kinematics of the theory. The action of a given causal set (also causet) is a crucial piece of the kinematics that would be extremely useful to know and understand. Proposals for the action of a causal set are available \cite{Benincasa_Dowker:The_Scalar_Curvature_of_a_Causal_Set} and these hold analytically in some cases and numerically in many more.\mb{rephrase} These cases being when the causal set is embeddable in some existing spacetime. One can show that the action of the causal set then agrees with the Einstein-Hilbert action of the spacetime in some continuum limit. How this limiting procedure is carried out will be described in more detail below.

It is well known that the Einstein-Hilbert action is not the full story in the continuum. In the presence of spacetime boundaries the gravitational action must include a boundary term $S_{GHY}$, the Gibbons-Hawking-York action, in order to yield a well-defined variational principle~\cite{Gibbons_Hawking_Boundary}. The contributions of this term play an essential role in particular in the quantum theory. For instance, in the calculation of the black hole entropy via the Euclidean path-integral, it is the boundary terms that produce the answer $A/4l_p^2$ necessary for the unification of black hole mechanics with thermodynamics. To give another example, in Regge calculus boundary terms are necessary in order that the quantum mechanical amplitudes satisfy the correct composition law and that they have the correct classical limit~\cite{hartlesorkin}.\mb{is this true?}

We define a \textit{causal set} (or \textit{causet}) as a locally finite partial order. This means it is a pair $ (\mathcal{C},\preceq)$ where $\mathcal{C}$ is the set of points and $\preceq$ is a partial order relation on $\mathcal{C}$ that has the following properties. It is (i) reflexive: $x\preceq x$, (ii) acyclic: $x\preceq y\preceq x \Rightarrow x=y$, and (iii) transitive: $x\preceq y\preceq z \Rightarrow x\preceq z$ for all points $x, y, z \in \mathcal{C}$. We define an inclusive order interval as the set $I (x,y)\equiv \lbrace z\in\mathcal{C}|x\preceq z\preceq y\rbrace$ for any $x, y\in\mathcal{C}$. The \textit{locally finite} condition is simply that the cardinality of any order interval is finite, that is $|I (x,y)|<\infty$. This condition ensures we are dealing with a discrete structure, as all the other properties of the relation would apply to an order relation between points on a continuous manifold.

In order to say we have a causal set analogue for a continuous expression we need a well defined procedure for relating the discrete theory to the continuum. This procedure, or tool, is called a \textit{Poisson sprinkling}, or just a \textit{sprinkling}. It is a Poisson process which provides a way to generate a causet from a $d$-dimensional Lorentzian manifold $ (M,g)$ by selecting certain points in $M$ to be the elements of $\mathcal{C}$, with an order relation given by the causal order of the manifold. The number of points chosen in a region of spacetime volume, $V$, is a Poisson random variable. This means that the expected number of points in some region will be $\rho V$, where $\rho$ is the density of the sprinkling. The density is related to the discreteness scale, $l$, by $\rho=l^{-d}$ in $d$ spacetime dimensions. It is called a sprinkling as one can envisage the point selection process as a `sprinkling' of points into the manifold. If a causet, $\mathcal{C}$, can be generated with relatively high probability by a sprinkling into the manifold, $ (M,g)$, then we say that the manifold is a good approximation for the causal set.

\section{The Claims}

Consider a sufficiently well-behaved $d$-dimensional spacetime $ (M,g)$ and a surface $\Sigma$ in $M$, which for now we take to be a Cauchy surface. The causal past and future sets $M^\pm=J^\pm (\Sigma)$ of the surface then form a partition of $M$, and $\partial M^\pm = \Sigma$. The Gibbons-Hawking-York boundary term for $M^\pm$ is in this case given by
\be\label{eq:GHYBT_in_continuum}
\mathcal{S}_{GHY} = \mp \frac{1}{l_p^{d-2}}\int_{\Sigma} K d\Sigma
\ee
where $l_p$ is the rationalised Planck length and $K$ is the trace of the extrinsic curvature $K_{\mu\nu}=h_{\mu}^\rho h_\nu^\sigma \nabla_\rho n_\sigma$ of $\Sigma$ defined with future-pointing timelike unit normal covector $n_{\mu}=\partial_\mu S/\sqrt{g^{\mu\nu}\partial_\mu S\partial_\nu S}$. (We work with a mostly plus convention for the metric, in which this translates into a past-pointing normal vector $n^{\mu}$).\mb{ok or more detail on conventions needed?}

Now observe that the integral in~\eqref{eq:GHYBT_in_continuum} can be thought of as the ``volume gradient'' across the surface $\Sigma$:
\be\label{eq:normal_deriv_boundary}
\int K d\Sigma = \frac{\partial}{\partial n}\int d\Sigma,
\ee
where the right hand side is the derivative of the area $\int d\Sigma$ as each point of $\Sigma$ is moved an equal distance along the unit normal vector, $n^{\mu}$ (corresponding to the rate of change of the surface volume \emph{backwards} in time, as $n^{\mu}$ is past-pointing). 
\mb{does the antichain analogy really stand?}

Spacetime volume in the causal set is obtained simply by counting the number of elements, and hence the volume gradient intuitively corresponds to the difference in the number of ``immediate neighbours'' to the future and to the past of a hypersurface. What does it mean to be immediate neighbours to a surface in a causal set? In order to make this notion precise, we need to define the causal set analogue of a hypersurface, as well as an appropriate notion of ``closeness''. This turns out to be most natural for the analogue of what in the continuum would correspond to a Cauchy hypersurface.

In the causet, a (partitioning) hypersurface can be described as a particular partition of the causet. \mb{why say this here?}We will only use this definition for causets of finite cardinality.Define a \emph{past} \emph{stem} to be a subset $\mathcal{C}'\subseteq\mathcal{C}$ of the causet that contains its own past\IGNORE{, i.e. $(\exists x\in\mathcal{C}' \text{ such that } y\prec x)\implies y\in\mathcal{C}'$}. That is, any element in the causet that precedes at least one element in $\mathcal{C}'$ is also in $\mathcal{C}'$. A \emph{future stem} is defined analogously. Let us now define a \emph{spacelike partition} as one which partitions the causet  $\mathcal{C}$ into two parts, $\mathcal{C}^-$ and $\mathcal{C}^+$\IGNORE{($\mathcal{C}=\mathcal{C}^-\cup\mathcal{C}^+$)}, where $\mathcal{C}^-$ is a past stem and $\mathcal{C}^+$ is its complement, $\mathcal{C}^+=\mathcal{C}\setminus\mathcal{C}^-$. One can equivalently pick some future stem $\mathcal{C}^+$ and its complement $\mathcal{C}^- = \mathcal{C}\setminus\mathcal{C}^+$ to define a spacelike partition. The reason that we call some partition of this nature ``spacelike" is the following. Given a causet that can be faithfully embedded in some manifold, $M$, then given a spacelike hypersurface, $\Sigma$, we can partition the causet into two bits\mb{only true if $\Sigma$ is Cauchy, no?}, denoting the future (past) part as $\mathcal{C}^+$ ($\mathcal{C}^-$), which is made up of those elements that lie in $M^+$ ($M^-$). The causal relations between the two parts, that can be found from the causal strucutre of the manifold, immediately satisfy our requirement for the partition to be spacelike. We will still refer to a partition as \emph{spacelike} even if the causet is not faithfully embeddable in some manifold. For immediate neighbours to the past of the partition we mean elements of $\mathcal{C}^-$ that have no elements in their causal future in $\mathcal{C}^-$ (these elements being called \textit{maximal} elements in $\mathcal{C}^-$). We assume an analogous definition for immediate neighbours to the $future$ of the partition, which we shall refer to as minimal elements in $\mathcal{C}^-$. We shall see that the intuitive idea for the boundary term as the ``rate of change of spacetime atoms'' across a partition indeed bears out. However, one does not need to know the maximal and minimal elements to calculate the boundary term. It can be found from any elements that lie reasonably close to the surface (in the time coordinate on $M$) but are not necessarily maximal or minimal. This will be made more precise below.

In the continuum the GHY term can be calculated for a spacelike hypersurface, $\Sigma$, that is either a future or past boundary of $M$. We wish to calculate the GHY term for future or past boundaries of causets, but there are a couple issues we must deal with. First, if we can faithfully embed a causet $\mathcal{C}$ into $M$ then the boundary surface of $M$ will not induce a partition of $\mathcal{C}$ as it did previously (every element of $\mathcal{C}$ will lie in the bulk of $M$), thus our previous idea for the boundary term does not work. Second, what does it mean to say a bounding surface of a causet if it is not faithfully embeddable in some manifold? Previously if the causet was not faithfully embeddable we could still define the causal set analogue of the surface as a partition of the causet. We shall see that the first issue is not a problem, and that the GHY term can be found for a future (past) boundary of $M$ (the manifold that the causet has been faithfully embedded in) by taking a difference in the number of certain elements that lie close to the future (past) bounding surface of $M$. Calculating the GHY term in this way must apply whether or not the causet is faithfully embeddable, but this runs us into the second issue above. For what boundary are we calculating our GHY term if the causet is not faithfully embeddable? Given a causet with maximal and/or minimal elements, certain quantities can be calculated for that causet, like the GHY term. We will not provide any physical meaning to a \emph{future} or \emph{past boundary} of a causet, and shall merely state that they are abstract objects for which certain causal set quantities can be calculated. The reason we say that these quantities are being evaluated for the future (past) boundary of the causet is that given a causet that is faithfully embeddable in some manifold with future (past) boundary, the values of the quantities in question will agree with those of certain geometrical quantities for the continuum boundary, in some continuum limit of the causal set. We will henceforth mention future/past boundaries of causets only in the context of some quantity being evaluated for them.

Determining rules for a partition to correspond to a timelike hypersurface in the continuum is extremely difficult and no satisfactory answer has been found yet. The situation for a null hypersurface is as follows. For any finite density sprinkling into a manifold with a null hypersurface, that null hypersurface will partition the causal set in such a way that agrees with the definition of a spacelike partition. We should really refer to such a partition as ``non-timelike", but as we are dealing with spacelike hypersurfaces for most of this paper we shall continue to call such a partition spacelike. To aid in the understanding of why our definition falls short in distinguishing a spacelike and null hypersurface imagine a finite density sprinkling into a patch of $2$-dimensional Minkowski spacetime. Figure~\ref{fig:low_density_sprinkling_with_null_and_spacelike} shows how, given a null surface that partitions the causet into two parts, we can tilt the surface slightly so that it is now spacelike, while still preserving the partition. Figure~\ref{fig:high_density_sprinkling_with_null_and_spacelike} shows that even as we increase the density of the sprinkling we can tilt the surface by a smaller amount such that the partition remains unchanged.

\begin{figure}
\IGNORE{
        \centering
        \begin{subfigure}[b]{0.5\textwidth}
                \includegraphics[width=\textwidth]{low_density_sprinkling_with_null_and_spacelike}
                \caption{low density sprinkling}
                \label{fig:low_density_sprinkling_with_null_and_spacelike}
        \end{subfigure}
        \begin{subfigure}[b]{0.5\textwidth}
                \includegraphics[width=\textwidth]{high_density_sprinkling_with_null_and_spacelike}
                \caption{high density sprinkling}
                \label{fig:high_density_sprinkling_with_null_and_spacelike}
        \end{subfigure}
        }
        \caption{The partition induced by the sprinkling is unchanged after tilting the surface between null and spacelike. With an even higher density, as in (b), one can simply titl the surface by some smaller amount to keep the partition unaltered.}\label{fig:animals}
\end{figure}

Let us define the functions $N_{\vphantom{i}max}^{ (k)}\left[\mathcal{C} \right]$ and $N_{min}^{ (k)}\left[\mathcal{C}^- \right]$ that, for a given finite causet, $\mathcal{C}$, return the number of $k$\emph{-next-to-maximal}/\emph{minimal} elements respectively. A $k$-next-to-maximal (minimal) element is one for which there are $k$ elements to its future (past). This means that the $0$-next-to-maximal (minimal) elements are the maximal (minimal) elements.

Given a causet, $\mathcal{C}$, with a spacelike partition so that $\mathcal C = \mathcal C^+ \cup\, \mathcal C^-$, we propose the following family of functions for calculating the GHY boundary term of this partition:
\be\label{general_boundary_sum}
S^{ (d)}_{GHY}\left[\mathcal{C}^-,\mathcal{C}^+;\vec{p}, \vec{q} \right]= \left (l/l_p\right)^{d-2} c_{d}
\left ( \sum_m p_m N_{\vphantom{i}max}^{ (m)}\left[\mathcal{C}^- \right]
+  \sum_n q_n N_{min}^{ (n)}\left[\mathcal{C}^+ \right]\right)
\ee
where the constant $c_{d}$ only depends on the spacetime dimension and is given by
\be\label{Cn}
c_{d}=\frac{d (d+1)}{ (d+2)}\left[\frac{V_{d-1}}{d}\right]^{\frac{2}{d}},
\ee
and where $V_d=\pi^{\frac{d}{2}}/\Gamma\left (\frac{d}{2}+1\right)$ denotes the volume of the unit $d$-ball ($\Gamma (t)$ is a gamma function, defined as $\Gamma (t):=\int_0^\infty dx\: x^{t-1}\: e^{-x}$). We have used $\vec{p}$ and $\vec{q}$ to stand for infinite dimensional vectors $ (p_0,...,p_m,..)$ and $ (q_0,...,q_n,..)$ respectively, where all the $p_m,\: q_n \in \mathbb{R}$. These vectors contain a finite number of non-zero entries, corresponding to the coefficients of the different $N_{{max\vphantom{i}}}^{\: (m)}$ and $N_{{min}}^{\: (n)}$ terms. We use the summation $\sum_m$ ($\sum_n$) to denote a sum that runs from $0$ to $m_{max}$ ($n_{max}$), where $m_{max}$ ($n_{max}$) is some integer, the value of which depends on the causet $\mathcal{C}^-$ ($\mathcal{C}^+$), for which all $N_{\vphantom{i}max}^{ (m)}\left[\mathcal{C}^- \right]=0$ ($N_{min}^{ (n)}\left[\mathcal{C}^+ \right]=0$) for any $m>m_{max}$ ($n>n_{max}$). A finite value for $m_{max}$ ($n_{max}$) will always exist for any finite causet, as it will not be large enough or be structured in such a way to have any $m$-next-to-maximal ($n$-next-to-minimal) elements for $m>m_{max}$ ($n>n_{max}$). The vectors must be chosen such that they satisfy the following conditions.
\begin{align}\label{coefficient_relation1}
& \sum_m p_m \frac{\Gamma\left (\frac{1}{d}+m \right)}{m!}  + \sum_n q_n\frac{\Gamma\left (\frac{1}{d}+n \right)}{n!}=0
\\
& \label{coefficient_relation2}\sum_m p_m \frac{\Gamma\left (\frac{2}{d}+m \right)}{m!}  - \sum_n q_n\frac{\Gamma\left (\frac{2}{d}+n \right)}{n!}=1
\end{align}
The aforementioned family simply corresponds to all the different vectors, $\vec{p}$ and $\vec{q}$, that satisfy the above criteria.

Given a $d$ dimensional, spatially compact Lorentzian manifold, $M$, the family of functions $S^{ (d)}_{GHY}$ (for different $\vec{p}$ and $\vec{q}$) defines a family of random variables in the following way. The process of sprinkling points into $M$ with density $\rho=l^{-d}$ generates a random causet $\mathcal{C}$, and so the functions $N_{\vphantom{i}max}^{ (k)}$ and $N_{min}^{ (k)}$ acting on this random causet are now the random variables $\textbf{N}_{\vphantom{i}max}^{ (k)}$ and $\textbf{N}_{min}^{ (k)}$. These random variables can be substituted into~\eqref{general_boundary_sum} for $N_{\vphantom{i}max}^{ (k)}$ and $N_{min}^{ (k)}$ to give the family of random variables, $\textbf{S}^{ (d)}_{GHY}$. These random variables will be functions of $\rho$ and the manifold, $M^{\pm}$, where the dependence of $\textbf{N}_{\vphantom{i}max}^{ (k)}$ ($\textbf{N}_{min}^{ (k)}$) on the manifold is restricted to $M^-$ ($M^+$). We claim that in the limit of infinite density the expectation value, in the sprinkling process, of $\textbf{S}^{ (d)}_{GHY}$ (for any vectors $\vec{p}$ and $\vec{q}$ satisfying~\eqref{coefficient_relation1} and ~\eqref{coefficient_relation1}) tends to the continuum GHY boundary term for a surface $\Sigma$ that partitions the manifold into $M^+$ and $M^-$:
\be
\lim_{l\rightarrow0}\left\langle\textbf{S}^{ (d)}_{GHY}[M^-,M^+,\rho;\vec{p} , \vec{q}]\right\rangle= \frac{1}{l_p^{d-2}}\int_{\Sigma} d^{d-1}x\: \sqrt{h}\: K\label{eq:mainconjecture}
\ee
where $\langle\cdot\rangle$ denotes the mean over sprinklings and we have explicitly stated the arguments of $\textbf{S}^{ (d)}_{GHY}$.

There is a huge amount of freedom that remains in the choice of $\vec{p}$ and $\vec{q}$. One can see that at least two terms are necessary in order to satisfy~\eqref{coefficient_relation1} and~\eqref{coefficient_relation2}. For any two terms the coefficients will be unique, and for more than two terms this uniqueness is lost. In a finite difference method one can form a discrete derivative in one direction by taking a difference of two values at different points along that direction, and dividing by the distance between them. We are doing something very similar here. The boundary action (in the continuum) is a derivative in one direction, the normal direction (as in~\eqref{eq:normal_deriv_boundary}), and so a finite difference method should be able to approximate this with just two terms. This agrees with the fact that we are able write the discrete boundary action with only two terms.

There are a few, two-term, members of the family that are worth mentioning, as they have simple forms that make them good candidates for computations. Given a causet, $\mathcal{C}$, with a spacelike partition so that $\mathcal C = \mathcal C^+ \cup\, \mathcal C^-$, we define the function $S^{ (d)}_0$ as a particular member of the family in~\eqref{general_boundary_sum}:
\be\label{eq:min_max_simple_formula}
S^{ (d)}_{0}\left[\mathcal{C}^-,\mathcal{C}^+\right]:=S^{ (d)}_{GHY}\left[\mathcal{C}^-,\mathcal{C}^+;\vec{p}, -\vec{p} \right]=\left (l/l_p\right)^{d-2}\frac{c_{d}}{2\Gamma\left (\frac{2}{d} \right)}\left ( N_{\vphantom{i}max}^{\: (0)}\left[\mathcal{C}^- \right] - N_{min}^{\: (0)}\left[\mathcal{C}^+ \right] \right)
\ee
where $\vec{p}=\left (\left (2\Gamma\left (\frac{2}{d} \right)\right)^{-1},0,0,...\right)$.
This expression is the easiest to employ computationally, and we shall use its random variable counterpart, $\textbf{S}^{ (d)}_{0}\left[M^-,M^+,\rho\right]:=\textbf{S}^{ (d)}_{GHY}\left[M^-,M^+,\rho;\vec{p}, -\vec{p} \right]$, later when we study the fluctuations of the discrete boundary actions numerically.

If one wants to calculate the GHY boundary term for the future (past) boundary of a single causet, $\mathcal{C}$, then one simply inserts $\mathcal{C}$ into the $\mathcal{C}^-$ ($\mathcal{C}^+$) argument of $S^{ (d)}_{GHY}$ and sets $\vec{q}$ ($\vec{p}$) to the zero vector, $\mathbf{0}= (0,0,...)$. This renders the $\mathcal{C}^+$ ($\mathcal{C}^-$) argument unnecessary and so one can insert the empty set, $\emptyset$, for convenience. We define the functions for the future and past boundaries respectively as:
\begin{align}\label{eq:future_past_boundary_terms1}
S^{ (d)}_{+}[\mathcal{C}]:= & S^{ (d)}_{GHY}\left[\mathcal{C},\emptyset;\vec{p},\mathbf{0}\right] =\left (l/l_p\right)^{d-2}\frac{c_{d}}{\Gamma\left (\frac{2}{d} \right)}\left ( d\:N_{\vphantom{i}max}^{\: (1)}\left[\mathcal{C} \right] - N_{\vphantom{i}max}^{\: (0)}\left[\mathcal{C} \right] \right)
\\
\label{eq:future_past_boundary_terms2}
S^{ (d)}_{-}[\mathcal{C}]:= & S^{ (d)}_{GHY}\left[\emptyset,\mathcal{C};\mathbf{0},-\vec{p}\right] =\left (l/l_p\right)^{d-2}\frac{c_{d}}{\Gamma\left (\frac{2}{d} \right)}\left ( N_{min}^{\: (0)}\left[\mathcal{C} \right] - d\: N_{min}^{\: (1)}\left[\mathcal{C} \right] \right)
\end{align}
where $\vec{p}=\left (-\Gamma\left (\frac{2}{d} \right)^{-1},d\:\Gamma\left (\frac{2}{d} \right)^{-1},0,0,...\right)$. In writing down an action for a particular causal set, $\mathcal{C}$, one would write down the Benincasa-Dowker action, $S^{ (d)}_{BD}[\mathcal{C}]$. To complete the story we propose that one must include the boundary terms $S^{ (d)}_{+}[\mathcal{C}]$ and $S^{ (d)}_{-}[\mathcal{C}]$. Explicitly, the total action for a causal set should be given by:
\be\label{eq:total_causet_action}
S^{ (d)}=S^{ (d)}_{BD}+S^{ (d)}_{+}-S^{ (d)}_{-}
\ee
where we have omitted the causet argument, $\mathcal{C}$. There may be contributions from the causal set analogue of timelike boundaries but some results suggest that these are included in $S^{ (d)}_{BD}$. We will see below that $S^{ (d)}_{BD}$ also includes a contribution from a different kind of boundary.

We also propose a family of functions for the causal set analogue of spatial volume of the spacelike partition:
\be\label{general_area_sum}
A^{ (d)}[\mathcal{C}^-,\mathcal{C}^+;\vec{p},\vec{q}]=\left (l/l_p\right)^{d-1}b_{d}\left (\sum_m p_m N_{\vphantom{i}max}^{\: (m)}\left[\mathcal{C}^- \right] + \sum_n q_n N_{min}^{\: (n)}\left[\mathcal{C}^+ \right]\right)
\ee
where the constant $b_d$ is given by
\be\label{constant_b_d}
b_d=d\left[\frac{V_{d-1}}{d}\right]^{\frac{1}{d}}
\ee
and we have used similar definitions for the arguments $\vec{p}$ and $\vec{q}$. This time however, the coefficients must satisfy 
\be\label{area_coefficient_relation}
\sum_m p_m \frac{\Gamma\left (\frac{1}{d}+m \right)}{m!}  + \sum_n q_n\frac{\Gamma\left (\frac{1}{d}+n \right)}{n!}=1
\ee
From this relation we can see that only one term is necessary to determine an expression for the discrete surface volume, and that when more than one term is added the coefficients will cease to be unique. Once again we can define a family of random variables, $\textbf{A}^{ (d)}$ which will depend upon $M^{\pm}$ and $\rho$. We make another claim that, in the limit of infinite density, the expectation value of $\textbf{A}^{ (d)}$ in the sprinkling process tends to the spatial volume of the surface $\Sigma$.
\be\label{eq:conjecture_for_area}
\lim_{l\rightarrow0}\left\langle\textbf{A}^{ (d)}[M^-,M^+,\rho;\vec{p} , \vec{q}]\right\rangle= \frac{1}{l_p^{d-1}}\int_{\Sigma} d^{d-1}x\: \sqrt{h}
\ee
where the right hand side is the spatial volume of the surface $\Sigma$. In proving~\eqref{eq:mainconjecture} we will establish the results necessary to prove this claim as well.

By making the same restrictions on the arguments as we did for $S^{ (d)}_{GHY}$, we can find expressions for the causal set analogue of the spatial volume of the future or past boundaries. We define functions for the future and past boundaries respectively as two of the simplest members of the family:
\begin{align}\label{eq:future_past_spatial_volume}
A^{ (d)}_{+}[\mathcal{C}]:=A^{ (d)}[\mathcal{C},\emptyset;\vec{p},\mathbf{0}]= & \left (l/l_p\right)^{d-1}\frac{b_{d}}{\Gamma\left (\frac{1}{d}\right)}\: N_{\vphantom{i}max}^{\: (0)}[\mathcal{C}]
\\
A^{ (d)}_{-}[\mathcal{C}]:=A^{ (d)}[\emptyset,\mathcal{C};\mathbf{0},\vec{p}]= & \left (l/l_p\right)^{d-1}\frac{b_{d}}{\Gamma\left (\frac{1}{d}\right)}\: N_{min}^{\: (0)}[\mathcal{C}]
\end{align}
where $\vec{p}= (\Gamma\left (\frac{1}{d}\right)^{-1},0,0,...)$.

%MB1: (too vague) During the calculation approximations will be made. These approximations are justified on the grounds that if certain orders are retained throughout the calculation they will in fact vanish in the limit $\rho \rightarrow \infty$
%\begin{figure}
  %\centering
   % {\includegraphics[width=\textwidth]{minmaxplot}}
    % \caption{An illustration of the idea. Pictured here is a causal set obtained by a sprinkling into a spacetime partitioned by a spacelike hypersurface. The minimal and maximal points about the surface are highlighted and the shaded regions illustrate the regions whose volumes $V_\blacktriangle$ and $V_\blacktriangledown$ are needed in the proof.}
     %\label{fig:Nmin_Nmax}
%*\end{figure}

\section{The Proof}

\subsection{Poisson Sprinklings for $\left\langle \textbf{N}_{max}^{\: (m)}\right\rangle$ and $\left\langle \textbf{N}_{min}^{\: (n)}\right\rangle$}

In order to prove~\eqref{eq:mainconjecture} and~\eqref{eq:conjecture_for_area} let us first derive expressions for the mean values of $\textbf{N}_{\vphantom{i}max}^{\: (m)}$ and $\textbf{N}_{min}^{ (n)}$. The proof will will follow almost automatically, as~\eqref{general_boundary_sum} and~\eqref{general_area_sum} are just linear combination of $N_{max}^{\: (m)}$ and $N_{min}^{ (n)}$ terms, and thus the mean values of their random variable counterparts will just be linear combinations of $\left\langle \textbf{N}_{\vphantom{i}max}^{\: (m)}\right\rangle$ and $\left\langle \textbf{N}_{min}^{\: (n)}\right\rangle$ terms.

For any instance of the sprinkling, the probability that a sprinkled point $p\in M$ below the surface $\Sigma$ is $k$\emph{-next-to-maximal} is given by the probability that $k$ points of the sprinkling lie in the region $J^{+} (p)\cap J^{-} (\Sigma)$, the intersection of the causal future of $p$ with the causal past of the surface $\Sigma$.\footnote{For notational convenience we use the symbol $x$ to refer to the causal element $x\in\mathcal C$, to its embedding in the manifold $M$, and to its coordinates in some chart on $M$.} This region will in general be the interior of some curvy $d$-dimensional cone truncated by the surface $\Sigma$, as illustrated in Figure~\ref{fig:Nmin_Nmax}. We will refer to these regions as \emph{truncated cones}. The Poisson process assigns a probability
\be\label{Poisson}
\mathbb P\left (\text{k points in }J^{+} (p)\cap J^{-} (\Sigma)\right)=\frac{\left (\rho\: V_\blacktriangledown (p)\right)^k}{k!}e^{-\rho V_\blacktriangledown (p)}
\ee
to this event, where $V_\blacktriangledown (p)\equiv V (J^{+} (p)\cap J^{-} (\Sigma))$ is the spacetime volume of the region $J^{+} (p)\cap J^{-} (\Sigma)$. The probability of sprinkling an element into an infinitesimal four-volume $dV_p$ at $p$ is $\rho dV_p$ where $\rho=l^{-d}$ is the sprinkling density, and so the total expected number of $k$-next-to-maximal elements below $\Sigma$ is
\be\label{eq:nmax}
\left\langle \textbf{N}_{max}^{ (k)}\right\rangle =\rho\int_{J^{-} (\Sigma)}dV_p\; \frac{\left (\rho\: V_\blacktriangledown (p)\right)^k}{k!}e^{-\rho V_\blacktriangledown (p)}
\ee
Similarly the expected number of $k$-next-to-minimal elements above $\Sigma$ is
\be\label{eq:nmin}
\left\langle \textbf{N}_{min}^{ (k)}\right\rangle =\rho\int_{J^{+} (\Sigma)}dV_p\; \frac{\left (\rho\: V_\blacktriangle (p)\right)^k}{k!}e^{-\rho V_\blacktriangle (p)}
\ee
where $V_\blacktriangle (p)\equiv V (J^{+} (\Sigma)\cap J^{-} (p))$.

In the limit $\rho\rightarrow\infty$, both quantities will diverge, but if some combination of $k$-next-to-maximal/minimal terms grows slower than or at order $\rho^{1-\frac2d}$, the random variable counterpart of the proposed action~\eqref{general_boundary_sum} will tend to a finite value in the continuum limit.

Consider a set of synchronous or Gaussian Normal Coordinates (GNC) $x^\mu= (t,\mathbf x)$ adapted to $\Sigma$ such that in a neighbourhood $U_\Sigma$ of $\Sigma$ the line element is
\be
ds^2 = -dt^2 + h_{ij} (t,\mathbf x) dx^i dx^j.
\ee
In these coordinates, the surface $\Sigma$ corresponds to $t=0$, and the coordinate $t$ measures the proper time elapsed along geodesics whose tangent vector is proportional to the surface normal on $\Sigma$.
%\be\label{GNC_metric}
%g_{\mu\nu} (x)=
%\begin{pmatrix}
 %-1&0 \\
% 0&h_{ij} (x)
%\end{pmatrix}.
%\ee
%In order to find $N_{max}$, the number of maximal points below the surface, one has to use the fact that the points have been sprinkled with a Poisson distribution. 
%If we pick a point $x_0= (0,\mathbf{x})\in \Sigma$ on the surface that has the same spatial coordinates as $x= (t,\mathbf{x})$ then the coordinate time, $t$, is the proper time elapsed along the unique geodesic between $x_0$ and $x$. The volume $V (\Sigma,x)$ then only depends on $t$.~\mb{error?}  
The integrals~\eqref{eq:nmax} and ~\eqref{eq:nmin} seem intractable as they stand, since the integration is over the entire causal past/future of the surface and the volume expressions will be complicated in the presence of curvature. However, for large $\rho$ (small $l$), the integrands $\exp (-\rho V)$ will be exponentially suppressed unless the volumes are small. Consider the spacetime region $U_\varepsilon=\left\{|t|<\varepsilon\right\}$ around the surface $\Sigma$ for some $\varepsilon>0$ (with $\varepsilon$ small enough such that the GNC system is valid throughout $U_\varepsilon$). We will assume that for any such $\varepsilon$, the contribution from points with $|t|>\varepsilon$ to the integrals~\eqref{eq:nmax} and \eqref{eq:nmin} can be made arbitrarily small by setting $\rho$ large enough, since the volumes of the truncated past/future lightcones for such points will be too large. Hence, as $\rho\rightarrow\infty$, the integration ranges in \eqref{eq:nmax} and \eqref{eq:nmin} can be cut off at finite time $t=\pm\varepsilon$ with $\varepsilon$ arbitrarily small, which allows us to expand the time integration-variable about zero. These assumptions of course impose certain regularity conditions on the surface $\Sigma$, but so long as $\Sigma$ is compact and everywhere spacelike  then these conditions are satisfied.\mb{do we need to say more about this? e.g. can't have asymp. null $\Sigma$ i guess} The integrals then simplify to
\begin{gather}\label{eq:nmax_and_eq:nmin}
\begin{aligned}
\left\langle \textbf{N}_{max}^{ (k)}\right\rangle & =\rho \int_{\Sigma}d^{d-1}x\int_{-\varepsilon}^{0}dt\:
h^{\frac{1}{2}}\left (1+
\frac{1}{2}\frac{\dot{h}}{h}t+O (t^2)\right)
 \frac{\left (\rho\: V_\blacktriangledown (p)\right)^k}{k!} e^{-\rho V_\blacktriangledown (t,\mathbf x)}
\\
\left\langle \textbf{N}_{min}^{ (k)}\right\rangle & =\rho \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
h^{\frac{1}{2}}\left (1+
\frac{1}{2}\frac{\dot{h}}{h}t+O (t^2)\right) \frac{\left (\rho\: V_\blacktriangle (p)\right)^k}{k!} e^{-\rho V_\blacktriangle (t,\mathbf x)}
\end{aligned}
\end{gather}
where $h\equiv det\left (h_{ij} (0,\mathbf{x})\right)$ and $\dot{}\equiv \frac{\partial}{\partial t}$ and the metric determinant has been in expanded in small $t$.

The only truncated cones that contribute will have small volumes, as we are restricting ourselves close to the surface.
%\mb{this sounds circular w.r.t. last paragraph} 
This means that the volumes can be approximated in some scheme which will be made precise below. Any higher order corrections to these volumes, above the order we require, will vanish in the limit of $\rho\rightarrow\infty$.

\subsection{Lightcone Volumes}

In order to evaluate the volume $V_\blacktriangledown (p)$ or $V_\blacktriangle (p)$ of a truncated cone we will perform a coordinate transform to Riemann Normal Coordinates (RNCs) in a neighbourhood containing the cone. The discussion for the two volume integrals is identical so we will only outline it for $V_\blacktriangle (p)$ (i.e. for points to the future of $\Sigma$).

Fix $p\in M$ and denote its coordinate values in GNCs by $x^\mu_p= (t_p,\mathbf x_p)$. It will be convenient to use RNCs centered not at the tip, $p$, of the cone but instead at the point $p_0$ where the unique geodesic through $p$ whose tangent is normal to $\Sigma$ intersects $\Sigma$. In GNCs this simply corresponds to the point $p_0$ with coordinates $x_0^\mu=x^\mu (p_0)= (0,\mathbf x_p)$. RNCs centered at $p_0$ will be given the symbol $y^{\overline{\mu}}$.
%\mb{let's put all facts about $x^\mu$ and $y^{\overline\mu}$ for $p$ and $p_0$ here and expansion of det in RNC.} 
We need to assume that for any $p$ in $U_\varepsilon$, the Riemann normal neighbourhood $U_p\subset M$ (throughout which the RNC system centered at $p_0$ is well-defined) contains the truncated cone $J^- (p)\cap J^+ (\Sigma)$. This assumption seems reasonable given the that $U_\varepsilon$ can be made arbitrarily ``thin'' as $\rho\rightarrow\infty$. In RNCs the metric and the Christoffel symbols at $p_0$ are those of flat space:
\be\label{eq:RNCMetricTransAtPAndChris}
g_{\overline{\mu} \overline{\nu}} (p_0)=\eta_{\overline{\mu} \overline{\nu}}=A^{\mu}_{\;\overline{\mu}}A^{\nu}_{\;\overline{\nu}}g_{\mu\nu} (p_0)\;,\;\;\;\;\Gamma^{\overline{\mu}}_{\;\overline{\nu}\overline{\rho}} (p_0)=0
\ee
The $A^{\mu}_{\;\overline{\mu}}$ govern the coordinate transformation from GNCs to RNCs to linear order. %but $O (x^2)$ corrections may be required. 
To second order the coordinate transformation is given by
\be\label{eq:RNCtotaltrans}
y^{\overline{\mu}}=A^{\overline{\mu}}_{\;\nu}x^\nu+\frac{1}{2}A^{\overline{\mu}}_{\;\mu}\Gamma^{\mu}_{\;\nu\rho} (p_0)x^\nu x^\rho+O ( (x-x_0)^3).
\ee
The inverse relation to first order is 
\be\label{eq:RNCinversetrans}
x^{\mu}=A^{\mu}_{\;\overline{\nu}}y^{\overline{\nu}}+O (y^2)
\ee
and one finds that the $A^{\mu}_{\;\overline{\mu}}$ must satisfy
\be\label{eq:RNCeqnforA}
A^{\overline{\mu}}_{\;\mu}A^{\mu}_{\;\overline{\nu}}=\delta^{\overline{\mu}}_{\;\overline{\nu}}\;,\;\;\;\;A^{\mu}_{\;\overline{\mu}}A^{\overline{\mu}}_{\;\nu}=\delta^{\mu}_{\;\nu}
\ee
These relations for RNC will cover all that will be needed in this discussion.

Now the volume of the truncated cone can be found as follows. 
%Consider the past lightcone emanating at the point $p_0= (T,\mathbf x_0)$ truncated by $\Sigma$. There is a unique point $q_0= (0,\mathbf x_0)$ on $\Sigma$ associated with $p_0$, separated from $p_0$ by proper time $T$. 
The volume of the cone is given by
\be\label{eq:VolumeWithNoSimplifications}
V_\blacktriangle (p)=\int_{\mathcal{X}_p} d^d x\;\sqrt{-g}
\ee
where the integration region, $\mathcal{X}_p\equiv  J^- (p)\cap J^+ (\Sigma)$ will be a complicated expression. As we are dealing with points close to the surface we can use the  RNCs $y^{\overline{\mu}}$, defined as above, about the point $p_0$. Let us denote the time-coordinate of $p$ in GNCs by $T=x^0_p=t_p$. For the transformation from GNCs to RNCs at $p_0$ one finds that $A^{\overline 0}_{\;0}=1$, $A^{\overline 0}_{\;i}=0$ and $\delta_{\ibar\jbar}=A^i_{\;\ibar}A^j_{\;\jbar}h_{ij} (p_0)$, which in particular implies $y^{\overline{0}}_p=\overline{t}_p=t_p=T$. In RNC one can expand the metric determinant in~\eqref{eq:VolumeWithNoSimplifications} to find
\be\label{eq:VolumeWithRNC}
V_\blacktriangle (p) =\int_{\mathcal{X}_p}d^dy+\int_{\mathcal{X}_p}d^dy\left (-\frac{1}{6}R_{\overline{\mu}\overline{\nu}} (p_0)y^{\overline{\mu}}y^{\overline{\nu}} \right)+O (T^{d+3})
\ee
where $R_{\overline{\mu}\overline{\nu}} (p_0)$ is the Ricci tensor in RNC evaluated at $p_0$.\mb{justify beforehand which orders to keep?}
%The arguments of the volume function need not be changed as the volume only depends on the points in the manifold, $q_0$ and $p_0$. 
The second term comes in at $O (T^{d+2})$ so the volume we have to calculate has reduced to
\be\label{eq:VolumeToLowestOrder}
V_\blacktriangle (p) =\int_{\mathcal{X}_p}d^dy+O (T^{d+2})
\ee
Terms of $O (T^{d+2})$ can be retained till the end, but in the limit, $\rho\rightarrow\infty$, they vanish. This will be proved below.

This is now a simple volume integral but we need to find the boundaries of $\mathcal{X}_p$ to write down the integration limits. There are two parts to the boundary of the (solid) truncated cone: the lightcone emanating at $p$ and the base, given by the intersection of $\Sigma$ with the interior of the lightcone. First we look at the lightcone. Following \cite{Khetrapal_Sumati:Causal_Diamond_Volume} it can be shown that the first curvature correction to the lightcone comes in at $O (T^{d+2})$ and so can be ignored for our purposes. This means that the lightcone can be treated as effectively flat in RNCs and thus corresponding to the set $ (y^{\overline{1}})^2+ (y^{\overline{2}})^2+...+ (y^{\overline{d-1}})^2= T-\overline{t}$.\mb{what limit on $\bar t$?}

The base of the cone in GNCs is simply a part of the surface $t=0$, so we can use (\ref{eq:RNCtotaltrans}) to find the equation for the surface in RNCs. Equation (\ref{eq:RNCtotaltrans}) gives
\be\label{eq:BottomSurfaceWithGNC}
\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij} (p_0)x^i x^j+O ( x^3)
\ee
The linear part on the right of (\ref{eq:RNCtotaltrans}) vanishes as $A^{\overline{0}}_{\;\mu}x^{\mu}=x^0$ (as $A^{\overline{0}}_{\;i}=0$ and $A^{\overline{0}}_{\;0}=1$) and $x^0=t=0$ for the bottom surface. Using the inverse RNC relation (\ref{eq:RNCinversetrans}) one can find the equation for the bottom surface in RNCs:
\be\label{eq:BottomSurface}
\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij} (p_0)A^{i}_{\;\ibar}A^{j}_{\;\jbar}y^{\ibar} y^{\jbar}+O (y^3)
\ee
Let us rewrite this equation in spherically symmetric coordinates, i.e. define $r=\sqrt{\delta_{\ibar\jbar}y^\ibar y^\jbar}$ and the usual angular coordinates $\phi_1,..,\phi_{d-2}$ in terms of the spatial coordinates $y^{\overline{1}} = r \cos (\phi_1),\ldots, y^{\overline{d-1}} = r \sin (\phi_1) \cdots \sin (\phi_{d-3}) \sin (\phi_{d-2})$. Then
\be\label{eq:RadialBottomSurface}
\overline{t}=\frac{1}{2}\left (\Gamma^{0}_{\;ij} (p_0)A^{i}_{\;\ibar}A^{j}_{\;\jbar}\frac{y^{\ibar} y^{\jbar}}{r^2}\right)r^2+O (y^3)=\frac{1}{2}f (\mathbf{x}_p,{\boldsymbol\phi})r^2+O (y^3)
\ee
where $\boldsymbol\phi$ stands collectively for all the angular coordinates $\phi_1,..,\phi_{d-2}$. The function $f (\mathbf{x}_p,\boldsymbol\phi)$ depends on $\mathbf{x}_p$ since $\Gamma^{0}_{\;ij}$ and $A^{i}_{\;\ibar}$ depend on $p_0$.
%One can see that $f (\mathbf{x}_0,\boldsymbol\phi)$ depends on the angles, $\phi$, by substituting the relations below into (\ref{eq:RadialBottomSurface})
%\begin{equation}
%\begin{aligned}\label{eq:SphericalCoords}
%y^{\overline{1}} &= r \cos (\phi_1)  \\
%y^{\overline{2}} &= r \sin (\phi_1) \cos (\phi_2)  \\
%y^{\overline{2}} &= r \sin (\phi_1) \sin (\phi_2) \cos (\phi_3)  \\
%    &\vdots  \\
%y^{\overline{d-2}} &= r \sin (\phi_1) \cdots \sin (\phi_{d-3}) \cos (\phi_{d-2})  \\
%y^{\overline{d-1}} &= r \sin (\phi_1) \cdots \sin (\phi_{d-3}) \sin (\phi_{d-2})
%\end{aligned}
%\end{equation}

With the boundaries of the integration region in place, we can now write down the integral explicitly in spherical coordinates:
\be\label{eq:VolumeIntegralSpherical}
V_\blacktriangle (p)=\int_{S^{d-2}}
d\Omega_{d-2}
\int_{0}^{r_{max} (\phi)}r^{d-2}dr
\int_{\frac{1}{2}f (\mathbf{x}_p,\phi)r^2}^{-r+T}
d\overline{t}+O (T^{d+2})
\ee
where $r_{max} (\boldsymbol\phi)$ is the value of the radial coordinate for which the surface $\Sigma$ intersects the lightcone at angle $\boldsymbol\phi$, as shown in Figure \ref{fig:cone_plot}. To find this value one needs to solve 
\be
\frac{1}{2}f (\mathbf{x}_p,\boldsymbol\phi){r_{max}}^2 (\boldsymbol\phi)=-r_{max} (\boldsymbol\phi)+T
\ee 
for $r_{max} (\boldsymbol\phi)$ and take the positive solution. The solution can be expanded in $T$ and is simply $r_{max}=T+O (T^2)$, with angular dependent terms contributing at $O (T^2)$. The $O (T^2)$ term will contribute at $O (T^{d+2})$ in the volume integral and so can be ignored. Substituting $r_{max}=T$ into (\ref{eq:VolumeIntegralSpherical}) allows us to evaluate the integral, wherein we find 
%\begin{figure}[t]
 % \centering
  %  {\includegraphics[scale=0.5]{coneplot}}
   %  \caption{The size of the region inside the top and bottom bounding surfaces is the volume we want to calculate.}
    % \label{fig:cone_plot}
%\end{figure}
\be\label{eq:VolumeNoK}
V_\blacktriangle (p)
=\frac{V_{d-1}}{d}T^d\left (1-\frac{d}{2 (d+1)}\Gamma^{0}_{\;ij} (p_0)A^{i}_{\;\ibar}A^{j}_{\;\jbar}\delta^{\ibar\jbar}T\right)
+O (T^{d+2})
\ee
where the $\delta^{\ibar\jbar}$ comes from the fact that cross terms ($\ibar\neq \jbar$) vanish under the angular integration. The defining relations for $A^{i}_{\;\ibar}$ can now be rearranged to give $A^{i}_{\;\ibar}A^{j}_{\;\jbar}\delta^{\ibar\jbar}=h^{ij} (p_0)$. In GNCs the extrinsic curvature on the surface is given by
\be\label{eq:K}
K
=g^{\mu\nu }\nabla_{\mu}n_{\nu}
=-\Gamma^{0}_{\;ij}h^{ij}=-\frac{1}{2}\frac{\dot{h}}{h}.
\ee
Substituting this into~\eqref{eq:VolumeNoK} we obtain 
\begin{align}
V_\blacktriangle (T,\mathbf x)
&=\frac{V_{d-1}}{d}T^d\left (1+\frac{d}{2 (d+1)}K (0,\mathbf{x})T\right)
+O (T^{d+2}) \label{eq:TopVolumeWithK}\\
V_\blacktriangledown (-T,\mathbf x)
&=\frac{V_{d-1}}{d}T^d\left (1-\frac{d}{2 (d+1)}K (0,\mathbf{x})T\right)
+O (T^{d+2}), \label{eq:BottomVolumeWithK}
\end{align}
having dropped the subscript $p$. Given the volume expressions in GNCs we now proceed to evaluate the integrals for $\left\langle \textbf{N}_{min}^{ (k)}\right\rangle$ and $\left\langle \textbf{N}_{max}^{ (k)}\right\rangle$.

\subsection{The mean of $\textbf{S}^{ (d)}_{GHY}[\mathcal {C}^-,\mathcal{C}^+;\vec{p},\vec{q}]$}

Using the previous equations for the volumes of the truncated cones we find that~\eqref{eq:nmax_and_eq:nmin} reduces to
\begin{gather}\label{eq:nmax_and_eq:nmin_volume_expanded}
\begin{aligned}
\left\langle \textbf{N}_{max}^{ (k)}\right\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{-\varepsilon}^{0}dt\:
h^{\frac{1}{2}}\left (1+
\frac{1}{2}\frac{\dot{h}}{h}t\right)
 \\
 & \times \Big ( A (-t)^d \Big)^k 
 \Big ( 1 - B (-t) \Big)^k
 e^{-\rho A (-t)^d \left[1-B (-t) \right]} + .\,.\,.
\\
\left\langle \textbf{N}_{min}^{ (k)}\right\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
h^{\frac{1}{2}}\left (1+
\frac{1}{2}\frac{\dot{h}}{h}t\right)
 \\
 & \times \Big ( A\: t^d \Big)^k 
 \Big ( 1 + B\: t \Big)^k
 e^{-\rho A\: t^d \left[1+B\: t \right]} + .\,.\,.
\end{aligned}
\end{gather}
where we have defined
\begin{gather}\label{A_and_B_defn}
\begin{aligned}
A & := \frac{V_{d-1}}{d} \\
B & := \frac{V_{d-1}}{2 (d+1)}K (0,\mathbf{x})
\end{aligned}
\end{gather}
Once again the higher order terms that have been ignored in both equations will be shown to vanish in the limit. From~\eqref{eq:K} we can substitute $K$ in for $\frac{1}{2}\frac{\dot{h}}{h}t$. We swap the integration variable to $t\rightarrow -t$ in the first equation and expand the $O (t^{d+1})$ part of the exponentials to give
\begin{gather}\label{eq:nmax_and_eq:nmin_volume_expo_expanded}
\begin{aligned}
\left\langle \textbf{N}_{max}^{ (k)}\right\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
h^{\frac{1}{2}}\left (1+
K\: t\right)
 \\
 & \times \Big ( A\: t^d \Big)^k 
 \Big ( 1 - B\: t \Big)^k
 \Big ( 1 + \rho A B\: t^{d+1} \Big)
 e^{-\rho A\: t^d} + .\,.\,.
\\
\left\langle \textbf{N}_{min}^{ (k)}\right\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
h^{\frac{1}{2}}\left (1-
Kt\right)
 \\
 & \times \Big ( A\: t^d \Big)^k 
 \Big ( 1 + B\: t \Big)^k
  \Big ( 1 - \rho A B\: t^{d+1} \Big)
 e^{-\rho A\: t^d} + .\,.\,.
\end{aligned}
\end{gather}
The reason for expanding the exponentials is that the integrals will have to be put into the form of Gaussian integrals in order to evaluate them in the limit $\rho\rightarrow\infty$. The above expressions are almost the same apart from a few sign differences, so from now on we will just work with the expression for $\left\langle \textbf{N}_{min}^{ (k)}\right\rangle$. By expanding the brackets, and retaining only necessary orders, the integral can be split into two parts of differing powers of $\rho$.
\begin{gather}\label{eq:n_min_different_rho_terms}
\begin{aligned}
\left\langle \textbf{N}_{min}^{ (k)}\right\rangle & = \frac{\rho^{k+1}A^k}{k!}\int_{\Sigma}d^{d-1}x\: h^{\frac{1}{2}}\int_{0}^{\varepsilon}dt\:
\left\lbrace \left ( t^{dk} + \left (kB-K \right)t^{dk+1}\right) + O\left (t^{dk+2}\right) \right\rbrace e^{-\rho A\: t^d}
 \\
 & - \frac{\rho^{k+2}A^{k+1}}{k!}\int_{\Sigma}d^{d-1}x\: h^{\frac{1}{2}}\int_{0}^{\varepsilon}dt\:
\left\lbrace Bt^{dk+d+1} + O\left (t^{dk+d+2}\right) \right\rbrace e^{-\rho A\: t^d}
\end{aligned}
\end{gather}
To find out how this diverges in the limit of $\rho \rightarrow\infty$, it suffices to find the divergence of the following general expression
\be\label{eq:general_t_n_integral}
\lim_{\rho\rightarrow\infty}\rho^{p}\int_{0}^{\varepsilon}dt\
t^{q}e^{-\rho At^{d}}
\ee
where $p,q \in \mathbb{R}$. We make the substitution $z=\rho At^{d}$ to put~\eqref{eq:general_t_n_integral} into the form of an incomplete gamma function.
\be\label{eq:incomplete_gamma_function}
\lim_{\rho\rightarrow\infty}\frac{A^{-\left (\frac{q+1}{d} \right)}}{d}\rho^{p-\left (\frac{q+1}{d} \right)}\int_{0}^{\rho A \varepsilon^d}dz\
z^{\left (\frac{q+1}{d} \right)-1}e^{-z}
\ee
The pre-factor has come from the substitution. In the limit we can retain the $\rho$ outside the integral while taking the integration limit to $\infty$. In doing this the integral becomes a gamma function.
\be\label{eq:gamma_function}
\lim_{\rho\rightarrow\infty}\frac{A^{-\left (\frac{q+1}{d} \right)}}{d}\rho^{p-\left (\frac{q+1}{d} \right)}\int_{0}^{\infty}dz\
z^{\left (\frac{q+1}{d} \right)-1}e^{-z}=\lim_{\rho\rightarrow\infty}
\frac{A^{-\left (\frac{q+1}{d} \right)}}{d}\rho^{p-\left (\frac{q+1}{d} \right)}\Gamma\left ( \frac{q+1}{d} \right)
\ee
We can use this procedure to take the limit of~\eqref{eq:n_min_different_rho_terms} and formulate the answer in terms of gamma functions. The limits of both $\left\langle \textbf{N}_{max}^{ (k)}\right\rangle$ and $\left\langle \textbf{N}_{min}^{ (k)}\right\rangle$ are then given by
\begin{gather}\label{eq:nmax_nmin_final}
\begin{aligned}
\lim_{\rho\rightarrow\infty}\left\langle \textbf{N}_{max}^{ (k)}\right\rangle = \lim_{\rho\rightarrow\infty} & \left\lbrace \rho^{1-\frac{1}{d}} \left (b_d\right)^{-1} \frac{\Gamma\left (\frac{1}{d}+k\right)}{k!}
\int_{\Sigma}d^{d-1}x\: \sqrt{h} \right.
 \\
 &  \left. +\rho^{1-\frac{2}{d}} \left (c_d\right)^{-1} \frac{\Gamma\left (\frac{2}{d}+k\right)}{k!}
\int_{\Sigma}d^{d-1}x\: \sqrt{h}K + O\left (\rho^{1-\frac{3}{d}} \right) \right\rbrace
\\
\lim_{\rho\rightarrow\infty}\left\langle \textbf{N}_{min}^{ (k)}\right\rangle = \lim_{\rho\rightarrow\infty} & \left\lbrace \rho^{1-\frac{1}{d}} \left (b_d\right)^{-1} \frac{\Gamma\left (\frac{1}{d}+k\right)}{k!}
\int_{\Sigma}d^{d-1}x\: \sqrt{h} \right.
 \\
 &  \left. -\rho^{1-\frac{2}{d}} \left (c_d\right)^{-1} \frac{\Gamma\left (\frac{2}{d}+k\right)}{k!}
\int_{\Sigma}d^{d-1}x\: \sqrt{h}K + O\left (\rho^{1-\frac{3}{d}} \right) \right\rbrace
\\
\end{aligned}
\end{gather}

These objects diverge as $\rho^{1-\frac{1}{d}}$ in the limit. The factor of $\rho$ that must be included in the formula for the surface volume is exactly the inverse of this ($\rho^{\frac{1}{d}-1}$), as can be seen from~\eqref{general_area_sum} if one substitutes in $l=\rho^{-\frac{1}{d}}$. If we include this factor above, and take the limit, we see that the only remaining terms are proportional to the surface volume. The other terms all have negative powers of $\rho$ and therefore vanish as $\rho\rightarrow\infty$. This means that if we take the mean of $\textbf{A}^{ (d)}$, in the limit of $\rho\rightarrow\infty$, and substitute in the above expressions, we will be left with something proportional to the surface volume. Then it is easy to see that in order to get the surface volume exactly one must force $\vec{p}$ and $\vec{q}$ to satisfy~\eqref{area_coefficient_relation}. This concludes the proof of the spatial volume claim,~\eqref{eq:conjecture_for_area}.

If we take the mean of $\textbf{S}^{ (d)}_{GHY}$ and the limit $\rho\rightarrow\infty$, we can substitute in the above expressions to find that the only terms left that have non-negative powers of $\rho$ (and therefore non-vanishing in the limit) are that of the spatial volume and boundary term. The term proportional to the spatial volume is divergent, so the coefficients, $\vec{p}$ and $\vec{q}$, must satisfy~\eqref{coefficient_relation1} to make this term vanish. The other condition,~\eqref{coefficient_relation2}, ensures that what remains is equal to the GHY boundary term. Thus, the proof of~\eqref{eq:mainconjecture} is complete.

The form of~\eqref{eq:nmax_nmin_final} seems to suggest general series expansions of $\left\langle \textbf{N}_{max}^{ (k)}\right\rangle$ and $\left\langle \textbf{N}_{min}^{ (k)}\right\rangle$ in reducing powers of $\rho$. Written in terms of the discreteness scale, $l$, it is a Laurent series starting at $l^{1-d}$. The higher order terms, in $l$, will possibly be proportional to higher order normal derivatives of the surface volume. If this were true then one could find causal set versions for normal derivatives of the surface volume to any order. The exact form of the constant in the next term of the series is unknown and more work must be done to determine a general trend for the constants, and hence find the general series.

\section{Fluctuations}
So far we have only talked about the mean of the causal set boundary action. Let us now turn to its fluctuations (the standard deviation) $\sigma[\textbf{S}^{ (d)}_{GHY}]=\text{Var}[\textbf{S}^{ (d)}_{GHY}]^\frac12$.\footnote{Whenever we say fluctuations we refer to the standard deviation of the random variable, not to its variance.} We can make a heuristic argument to estimate the dependence of fluctuations on $\rho=l^{-d}$. In any spacetime region of fixed volume $V$ the number of causal set elements in a sprinkling is a random variable, $\textbf{N}$, with mean $\left\langle\textbf{N}\right\rangle=\rho V$, and Poisson fluctuations of order $\sqrt{\left\langle\textbf{N}\right\rangle}$. We will use $\textbf{S}^{ (d)}_{0}$ for the boundary action, as it is the simplest case in which to make this argument for the fluctuations. We believe this choice is valid as the simulations have suggested that all members of the $\textbf{S}^{ (d)}_{GHY}$ family have similar fluctions. Since $\textbf{N}_{max}^{ (0)}$ and $\textbf{N}_{min}^{ (0)}$ are random variables associated with a codimension-1 surface (they count elements ``near" $\Sigma$), we may expect their mean values to scale like $\left\langle\textbf{N}\right\rangle^\frac{d-1}{d}$ --- which agrees with the leading order behaviour of~\eqref{eq:nmax_nmin_final} --- and thus they would inherit fluctuations of order $\left\langle\textbf{N}\right\rangle^\frac{d-1}{2d}$. 
%{This comes from the following observation. In a volume corresponding to a thickening of the hypersurface $\Sigma$ by one unit of the discreteness scale $l$ (e.g. by Lie dragging the surface along its normal by an amount $l$),.... }
The action $\textbf{S}^{ (d)}_{0}$, being proportional to $\rho^\frac{2-d}{d}$ times the difference of two independent\footnote{The independence is as a feature of the Poisson process in $M$: the number of elements $\textbf{N}[R]$ in any subregion $R\subset M$ is a Poisson variable itself, and for any two disjoint regions $R_1$, $R_2$ the numbers $\textbf{N}[R_1]$ and $\textbf{N}[R_2]$ are independent random variables.} random variables with standard deviation $\left\langle\textbf{N}\right\rangle^\frac{d-1}{2d} = (\rho V)^\frac{d-1}{2d}$ should see fluctuations of order $\rho^\frac{2-d}{d}\rho^\frac{d-1}{2d}=\rho^\frac{3-d}{2d}$. This suggests that for $d=2$ these fluctuations should grow like $\rho^{\frac{1}{4}}$ as $\rho\rightarrow\infty$, for $d=3$ they should be constant, and for $d>3$ they should be damped.

In order to test this heuristic argument one may like to write down the integral expression for $\text{Var}[\textbf{S}^{ (d)}_0]^\frac12$ but it is complicated enough even in finite regions of flat space, for which it is not very illuminating to reproduce here. It also seems far less tractable than the expression for the mean because it involves terms of the type $\int d^dx\int d^dy \exp\left[-\rho V (J^+ (x)\cap J^+ (y))\right]$ for which the above methods of cutting off the time-integration range and Taylor expanding will not follow through. Instead we will show the results of computer simulations that support results of the heuristic argument.

%\begin{figure}[t!]
 % \centering
  %  {\includegraphics[scale=0.6]{GHY-fluctuation_plots}}
   %  \caption{A $\log-\log$ plot (base $2$) of the fluctuations in $S^{ (d)}_{GHY}$ for a flat ($K=0$) surface in $d=2,3$ and $4$-dimensional Minkowski space. Each data point represents the sample standard deviation in $S^{ (d)}_{GHY}$ on a sample of $R=100$ runs. Black dots, blue triangles and red squares correspond to the simulation results in $d=2,3$ and $4$ dimensions, respectively. The corresponding black, blue and red lines have gradients $\frac12$, $0$ and $-\frac18$ and best-fit intercepts of order $1$.}
    % \label{fig:fluctuations}
%\end{figure}

The simulations were carried out as follows. Denote the discreteness scale by $l$. Take a $d$-cube $[0,L]^d$ in $d$-dimensional Minkowski space with metric $ds^2=-dt^2+d{\mathbf x}^2$ and define the hypersurface $\Sigma: t=L/2$, which partitions the cube into two equal halves. Sprinkle at density $\rho=l^{-d}$ into the cube, which in any given run will place $N$ points inside the cube, where $N$ is the realisation of the Poisson random number $\textbf{N}$ with mean $\left\langle \textbf{N}\right\rangle = \rho V=  (L/l)^d$. Evaluate $S^{ (d)}_0=\rho^\frac{2-d}{d}c_d\left (2\Gamma\left (\frac{2}{d}\right)\right)^{-1} (N^{ (0)}_{\vphantom{i}max} - N_{min}^{ (0)})$ for $\Sigma$ by counting the minimal/maximal elements in the upper/lower half of the cube ($S^{ (d)}_0$, $N^{ (0)}_{\vphantom{i}max}$, and $N^{ (0)}_{min}$ of course being realisations of the random variables $\textbf{S}^{ (d)}_0$, $\textbf{N}^{ (0)}_{\vphantom{i}max}$, and $\textbf{N}^{ (0)}_{min}$), and repeat for $R$ runs. The value we expect for the mean (sum of all $S^{ (d)}_0$ values for each run divided by $R$) is zero in all cases, since the extrinsic curvature of the surface is zero.

In Figure~\ref{fig:fluctuations} we display the simulation results for $d=2,3,4$ spacetime dimensions, with $\left\langle\textbf{N}\right\rangle=\rho$ ranging up to $2^{20}\approx 1$ million, having set $V=L^d=1$. Each data point represents the fluctuations (the standard deviation) in a sample of $R=100$ runs. The solid lines have been obtained by fitting an arbitrary constant multiplier in the scaling law predicted by the argument above, $\gamma (d)\times \left\langle\textbf{N}\right\rangle^\frac{3-d}{2d}$, to the data sets. The best fit values are all of order $1$: $\gamma (2)=0.63$, $\gamma (3)=0.91$, and $\gamma (4)=1.07$.
There is clearly a good fit between the data and the scaling predicted by the heuristic argument. Although not visible on the plot, the data points were all consistent with zero mean as well.

As mentioned above, we have performed analogous simulations for other family members, such as $S^{ (d)}_+$. This expression involves the terms $N^{ (1)}_{max}$ and $N^{ (0)}_{max}$, which means that heuristic arguments of the type given above are hampered by the fact that the two random variables whose difference is taken are not independent. However, the results for the scaling of the fluctuations turn out to be identical to those obtained in the first case.

\section{A Candidate Boundary Term for an Alexandrov Interval}
\newcommand{\vol}{\mathrm{vol}}

Until now our approach has been to reconstruct the  continuum GHY term  from the information in a causal set. 
The candidate described in the previous sections for a spatial boundary has a natural expression in a causal set  in  terms of its maximal and minimal elements which are the  analogs of a  space-like boundary.  A suitable analog  of  a time-like boundary  term has been harder to  construct; there is no simple causal set  characterisation of such  a boundary. Indeed, it is difficult to give a causal set  characterisation of a  spacetime region bounded by time-like surfaces. On the other hand, this is not true of causally defined spacetime regions since  past and future sets are defined naturally in  a causal set. An important example of such a region is the  Alexandrov interval which is used extensively  in calculations of causal set kinematics.  In the continuum such a region  is bounded by null hypersurfaces and their codimension 2 joints.

Conversely in the continuum, while the GHY term  is well defined for both a space-like and time-like boundary  (including their co-dimension 2 joints)  it is an open question what the contribution from a null  boundary  is \cite{Gibbons_Hawking_Boundary, Harris}. Apart from some sketchy suggestions in the literature \cite{neiman} there is no rigourous construction of such a term; indeed the standard variational route for finding the boundary contribution to the action appears to fail for null boundaries.

Given our ignorance in this matter,  we can ask if causal set theory  might  in fact guide us in the right direction.   Our starting point is the $d$-dimensional  Benincasa-Dowker Action \cite{Benincasa_Dowker:The_Scalar_Curvature_of_a_Causal_Set, Dowker_Glaser:dAlembertians_for_Causal_Sets} which is constructed from the mean number  $N_i^{ (d)}$\footnote{ $N_i^{ (d)}$ is dimension  and geometry dependent, but for our purpose it suffices to retain the dimension label. } of $i+1$-element inclusive intervals in the causal set 
\begin{equation}   
8 \pi G {S_{BD}^{ (d)}}\left[\mathcal C\right] = -\alpha_d \rho^{\frac{2-d}{d}}\biggl ( N[\mathcal C]+ \frac{\beta_d}{ \alpha_d} \sum_{i=1}^{n_d-1}  C^{ (d)}_{i} N _i^{ (d)}[\mathcal C] \biggr) 
\label{bd} 
\end{equation}  
with   
\begin{equation}
\begin{aligned}
 \alpha_d =   
\begin{cases}
\displaystyle
-\frac1{\Gamma\left (1+\frac2d\right)}c_d ^{2/d} \;   &d \, \, \mathrm{odd} \\
\displaystyle
- \frac{2}{\Gamma\left (1+\frac2d\right)}c_d ^{2/d}  \;   &d \, \,  \mathrm{even} \\
\end{cases}
\quad\text{and}\quad 
%\end{equation}
%\begin{equation}
\beta_d =
\begin{cases}\displaystyle
\frac{d+1}{2^{d-1}\Gamma\left (1+\frac2d\right)} c_d^{2/d}   \;   &d\mathrm{ \, \, odd}\\ 
\displaystyle
\frac{\Gamma\left (\frac{d}{2}+2\right)\Gamma\left (\frac{d}{2}+1\right)}{\Gamma\left (\frac{2}{d}\right)\Gamma\left (d\right)}c_d^{2/d}  &  d\mathrm{ \, \,  even}\\ 
\end{cases} 
\end{aligned}
\end{equation}
and
\begin{equation} 
n_d = 
\begin{cases} 
\frac{d-1}{2} + 2  \quad & d\mathrm{ \, \, odd}\\ 
\frac{d}{2} + 2  \quad &d\mathrm{ \, \,  even,}\\ 
\end{cases} 
\end{equation} 
where  $2^{-d/2} c_d$ is the volume of a unit Alexandrov interval $I (x,y)$, i.e., one for which the proper time $\tau (x,y)=1$. 
%%{ {\bf \color{red} this is not correct! $c_d= \vol (S^{d-2}) \frac{ 2^{\frac{2-d}{2}}}{d (d-1)}$ and $\vol (S^{d-2})$ the volume of the homogeneous $d-2$ sphere}}.  
The coefficients $C_i^{ (d)}$ of the terms $N_i^{ (d)}$ in the sum are
\begin{equation}
\label{cid}
 C_i^{ (d)}= 
\begin{cases} 
\displaystyle\sum_{k=0}^{i-1} (-1)^k\binom{i-1}{k} \frac{\Gamma\left (d/2 (k+1)+3/2\right)}{\Gamma\left (d/2+3/2\right) \Gamma\left (1+dk/2\right)}  \quad  &d\mathrm{ \, \, odd}\\ 
\displaystyle\sum_{k=0}^{i-1} (-1)^k\binom{i-1}{k} \frac{\Gamma\left (d/2 (k+1)+2\right)}{\Gamma\left (d/2+2\right) \Gamma\left (1+dk/2\right)}  &d \mathrm{ \, \,  even.}\\ 
\end{cases} 
\end{equation} 
These coefficients can be expressed more compactly as hypergeometric functions 
\begin{equation}
C_i^d={}_{q+1}F_{q} (\{ a_1, \ldots a_q, i-1\}, \{b_1, \ldots b_q \} |1)
\label{chyp} 
\end{equation} 
with $q=d/2+1/2, a_i=\frac{d+2i}{d}, b_i=2i/d$ for $d$ odd and $q=d/2, a_i=\frac{d+2i+2}{d}, b_i=2i/d$  for $d$ even.

If $S_{BD}^{ (d)}$ were precisely the discrete analog of the Einstein Hilbert action {\it without} any boundary contributions, then in flat spacetime  $S_{BD}^{ (d)}$ should  vanish. However, as shown explicitly in \cite{bbdtwo} this is not the case for Alexandrov intervals in $d=2$  flat spacetime. Indeed, there is a residual contribution that goes to a non-vanishing constant in the continuum limit $N\rightarrow \infty$.  When the action is evaluated in rectangular regions in the same spacetime or for Alexandrov intervals in  the flat trousers spacetime, this is no longer the case. Thus, it seems plausible that the $d=2$ BD action contains boundary information. Moreover, in the case of Alexandrov intervals the asymptotic value of the action is the same  irrespective of the interval size. While this might suggest a  topological origin, we will now show that it is a part of a more general  result for $d>2$ and  has a geometrical origin. Namely it corresponds to the volume  $\vol (S^{d-2})$ of the $S^{d-2}$ joint which is independent of the interval size {\it only} in $d=2$.

Consider an Alexandrov interval $I (p,q)=I^+ (p) \cap I^- (q)$ for two events $p,q$ in a causal spacetime.   While the  boundary of $I (p,q)$ may be quite complicated in general, some of these components must necessarily be null.  In flat spacetime, for example,  the boundary of $I (p,q)$ consists of the two events $p$ and $q$, the subsets $B^\pm$ of the  two null hypersurfaces $\partial J^+ (p)$ and $\partial J^- (q)$, and their intersection or ``joint''  $S^{d-2}=\partial J^+ (p) \cap \partial J^- (q)$ which is a codimension 2 sphere.  Evaluating the BD action in an Alexandrov interval in flat spacetime should therefore yield  only boundary dependent terms.  In flat spacetime the $S^{d-2}$ joint  has radius $\tau (p,q)/2$, and thus increases with the size of the Alexandrov interval ${\mathrm{vol}} (I (p,q)) = N/\rho$, where $N$ is the mean of the number of causal set elements sprinkled into $I (p,q)$ . In what follows we take the continuum limit $\rho, N \rightarrow \infty$ while keeping ${\mathrm{vol}} (I (p,q))$ fixed. \ss{It would be nice to have a fancy picture of an Alex. interval here..}  
%% If $p$ or $q$ were to be taken to time-like past or future infinity, this volume would go to infinity as expected.

In \cite{Glaser_Sumati:Locality_in_Causal_Set} a closed form expression was obtained for the mean value of the number of $i+1$ element inclusive intervals $N_i^d$ contained in an Alexandrov interval in $d$ dimensional flat spacetime 
\begin{equation} N_i^{ (d)}= \frac{\Gamma\left (d\right)^2 N^{i+2}}{\Gamma\left (i\right)} \sum_{k=0}^\infty \frac{ (-N)^k\,\Gamma\left (\frac{d (k+i)}{2}+1\right)  \Gamma\left (\frac{d (k+i+1)}{2}+1\right)}{ (k+i+2)  (k+i+1) \Gamma\left (k+1\right)\Gamma\left (\frac{d (k+i)}{2} +d\right) \Gamma\left (\frac{d (k+i+1)}{2}+d\right)},\nonumber
\end{equation} 
where $i \geq 1$.  We use this to expand the sum in (\ref{bd}) as a power series in $N$: $\sum_{i=1}^{n_d}C_i^{ (d)} N_i^{ (d)} = \sum_{j=1}^\infty A_j^{ (d)}  N^{j+1}$.  After a rearrangement and redefinition of indices we find that 
\begin{equation}
 A_j^{ (d)} = \Gamma\left (d\right)^2\left ( \sum_{i=1}^{D+2} (-1)^i \binom{j-1}{i-1}  C_i^{ (d)}  \right) \frac{ (-1)^j}{ (j+1)!}\frac{\Gamma\left (\frac{d}{2} (j-1)+1\right)\Gamma\left (\frac{d}{2}j+1\right)}{\Gamma\left (\frac{d}{2} (j-1)+d\right) \Gamma\left (\frac{d}{2}j+d\right)}, 
\end{equation} 
where $d=2D$ for $d$ even and $d=2D+1$ for $d$ odd.   
% Using the abbreviated form $C_i^d=\sum_{l=0}^{i-1} \binom{i-1}{l}A (d,l)$ where $A (d,l)$ is defined using \ref{cid} 
% we find that 
% \begin{equation}
% \sum_{i=1}^{D+2} (-1)^i \binom{j-1}{i-1}  C_i^d = (-1)^{j-1} A (d,j-1),  
% \end{equation}      
% where we have used the binomial identities  BLAH. Thus 
% \begin{equation}
%  \sum_{i=1}^{n_d} C_i^d N_i^{ (d)}= ( (d-1)!)^2 \sum_{j=1}^\infty (-1)^j B (d,j) \frac{\Gamma (\frac{d}{2}j+1)}{\Gamma (j+2)\Gamma (\frac{d}{2} (j-1)+d)\Gamma (\frac{d}{2}j+d)}N^{j+1}
% \label{thesum} 
% \end{equation} 
% where 
% \begin{equation} 
% B (d,j) = \left \{ 
% \begin{array}{ccc} 
% \frac{\Gamma (\frac{d}{2}j+3/2)}{\Gamma\left (\frac{d}{2}+3/2)}  &\quad & d \, \, odd, \\  \frac{\Gamma\left (\frac{d}{2}j+2\right)}{\Gamma\left (\frac{d}{2}+2\right)}  &\quad & d \, \, even.  \\ 
% \end{array} 
% \right
% \end{equation} 
This can be simplified to 
\begin{equation}
A_j^d= \gamma_j^{ (d)}   \frac{\Gamma\left (d\right)^2 (-1)^{j+1}}{\Gamma\left (\frac{d}{2} (j + 1)\right) \Gamma\left (\frac{d}{2} (2 + j)\right) \Gamma\left (2 + j\right) }
\label{simplercoefft} 
\end{equation} 
where 
\begin{equation} \gamma_j^{ (d)}= 
\begin{cases} 
\displaystyle\frac{\sqrt{\pi}}{2^{1+dj} }\frac{\Gamma\left (2+dj\right)}{\Gamma\left (\frac{d-1}{2}\right)}    \quad   & d \, \, \mathrm{odd} \\
\displaystyle\frac{\Gamma\left (1+\frac{d}{2}j\right)  \Gamma\left (2+\frac{d}{2}j\right) }{\Gamma\left (\frac{d}{2}\right)}  \quad  & d\mathrm{ \, \,  even.} \\
\end{cases} 
\end{equation} 
A further simplification then allows us to include the first term in the action (\ref{bd}) in the series expansion so that 
\begin{equation} 
8 \pi G S_{BD}^{ (d)}[\mathcal C] = -\alpha_d \rho^{\frac{2-d}{d}} \sum_{j=0}^\infty (-1)^j Q_j^{ (d)} N^{j+1} 
\end{equation}
where 
\begin{equation} 
Q_j^{ (d)}  = {\widetilde \gamma}_j^{\, (d)} \, \, \frac{\Gamma\left (d\right)\Gamma\left (\frac{d}{2}\right)}{\Gamma\left (\frac{d}{2} (1+j)\right) \Gamma\left (\frac{d}{2} (2+j)\right) \Gamma\left (2+j\right)}
\label{fullcoefft} \end{equation}
and
\begin{equation}
{\widetilde \gamma}_j^{\, (d)} = 
\begin{cases} 
2^{-dj}    \quad   &{d} \, \, \mathrm{odd} \\
\Gamma\left (1+\frac{d}{2}j\right) \Gamma\left (2+\frac{d}{2}j\right)   \quad   &d\mathrm{ \, \,  even.} \\
\end{cases} 
\end{equation} 
Taking our cues from the behaviour of $8 \pi G S_{BD}^{ (d)}[\mathcal C]$ for $d=2,3$ in the  $N \rightarrow \infty$ limit,  we will consider the ratio
\begin{equation} 
\frac{ 8 \pi G S_{BD}^{ (d)}[\mathcal C]}{\vol (S^{d-2})} = \frac{b}{d (d-1) \Gamma\left (1+2/d\right) N^{\frac{d-2}{d}}} \left ( N + \frac{\beta_d}{\alpha_d} \sum_{i=1}^{n_d} C_i^{ (d)} N_i^{ (d)} \right)  
\label{ratio} 
\end{equation} 
where $b=1$ for $d$ odd and $b=2$ for $d$ even.
Using this simplification an explicit calculation in Mathematica\footnote{The above simplifications greatly aid the Mathematica calculation and allow us to extend it to higher $d$.}  for $d=2, \ldots 16$ gives 
\begin{equation} 
\lim_{N \rightarrow \infty} \frac{ 8 \pi G S_{BD}^d (C)}{\vol (S^{d-2})} = 1,  
\label{result} 
\end{equation}

Since each $N_i^{ (d)}$  can be expressed as a hypergeometric function \cite{Glaser_Sumati:Locality_in_Causal_Set} it  might seem possible to evaluate (\ref{ratio}) analytically term by term using a large $N$ expansion.  In \cite{Glaser_Sumati:Locality_in_Causal_Set} the asymptotic form of $N_i^{ (d)}$ was explicitly calculated, and for $d>2$, the leading order was shown to  $\sim  N^{2-\frac{2}{d}}$. This provides us a good consistency check for our result (\ref{result}) since this would mean that the leading order contributions from the $N_i^{ (d)}$ cancel in the action. This is indeed explicitly seen to be the case, again for $d=3, \ldots 16$. Similarly for $d=2$ where the leading order contribution is $\sim  N \log N$.  Indeed, in all cases the next to leading order terms given in \cite{Glaser_Sumati:Locality_in_Causal_Set} also do not have the requisite $ \sim N^{\frac{d-2}{d}}$ dependence required by (\ref{result}); in this case, the explicit coefficients have not been calculated  to be able   check for  a vanishing contribution to (\ref{result}).  This suggests that  we cannot rely   on the the asymptotic form of the individual terms  in the sum, and would need to look elsewhere to find an analytic proof for general $d$. While it may be possible to use the simplification (\ref{fullcoefft})  to obtain a closed form expression for the action for general $d$, it suffices for our purposes that (\ref{result}) has been explicitly checked in $2\leq d \leq 16$ dimensions.

The result we have obtained is for flat spacetime. One may speculate that the presence of curvature will not modify the essence of this result,  i.e., that the BD action for an Alexandrov interval  will include boundary contributions from the  $S^{d-2}$ at the joint, at least  when the boundary components  are of the same type as in flat spacetime.   What will differ of course is the fact that the principal radii of the $S^{d-2}$ will in general vary from point to point, hence complicating the relationship between the $d$-volume of the interval and the $ (d-2)$-volume of $S^{d-2}$. In  \cite{Khetrapal_Sumati:Causal_Diamond_Volume} it was shown that in an RNC up to first order corrections the intrinsic boundary geometry of the Alexandrov interval is still flat. This would allow a simpler relation between $\vol (S^{d-2})$ and $N$, making it possible to do an RNC version of this  calculation.

The fact that the BD action does not vanish in flat spacetime suggests  that this action  does not only contain  information about the bulk Einstein Hilbert action but also extra boundary information.  One might be tempted to speculate further that this action contains {\it all}  possible boundary or GHY terms suggesting that the null boundary GHY contribution for an Alexandrov interval in any spacetime always vanishes. Moreover, it is only the spatial $S^{d-2}$ joint  which gives  a non-vanishing contribution $\vol (S^{d-2})$.  Interestingly, this is precisely what \cite{neiman} claims is the GHY term for an Alexandrov interval.

\newpage

\bibliographystyle{jhep}

\bibliography{references}

\end{document}

%%
JHEPPUB:

\title{Gibbons-Hawking-York Boundary Term in Causal Sets}
 \author[a]{M. Buck}
 \author[a,b]{\!, F. Dowker}
 \author[a]{and I. Jubb\,}
\affiliation[a]{Theoretical Physics Group, Blackett Laboratory, Imperial College, London, SW7 2AZ, U.K.}
\affiliation[b]{Institute for Quantum Computing, University of Waterloggedo, ON, N2L 2Y5, Canada}

\abstract{ 
We show some stuff.
}

\begin{document}

\maketitle

\pagebreak

\noindent NOTES:\\
1. define s as the RV, not as its mean

\section{Introduction}

%%IOPART

\begin{document}

\title{Gibbons-Hawking-York Boundary Term in Causal Sets}

\author{Michel Buck Fay Dowker, Ian Jubb}
\address{Blackett Laboratory, Imperial College, London, SW7 2AZ, U.K.}

\begin{abstract}

We show some stuff.

\end{abstract}
%\pacs{03.67.-a, 03.65.Ta, 03.70.+k}

\maketitle
\section{Introduction}
%%

### 2014-09-18 test.tex
initial commit

--- 2014-09-18 paper_v5.tex
@@
 \newcommand{\eea}{\end{aligned}}
-\newcommand{\mb}[1]{\marginnote{\texttt{\small MB:\,#1}}}
-\renewcommand{\ij}[1]{\marginnote{\texttt{\small IJ:\,#1}}}
+\newcommand{\mb}[1]{\marginnote{\color{red}{\small MB:\,#1}}}
+\renewcommand{\ij}[1]{\marginnote{\color{red}{\small IJ:\,#1}}}
+\renewcommand{\ss}[1]{\marginnote{\color{red}{\small SS:\,#1}}}
 \newcommand{\ibar}{{\overline{\emph\i}}}
@@
 \title{Boundary terms in the action for Causal Sets}
- \author[a]{M. Buck}
- \author[a,b]{\!, F. Dowker}
- \author[a]{and I. Jubb\,}
+ \author[a]{Michel Buck}
+ \author[a,b]{\!, Fay Dowker}
+ \author[a]{\!, Ian Jubb\,}
+ \author[c]{and Sumati Surya}
 \affiliation[a]{Theoretical Physics Group, Blackett Laboratory, Imperial College, London, SW7 2AZ, U.K.}
 \affiliation[b]{Institute for Quantum Computing, University of Waterloo, ON, N2L 2Y5, Canada}
+\affiliation[c]{Raman Research Institute, CV Raman Ave, Sadashivanagar, Bangalore 560080, India}
 
@@
 This procedure, or tool, is called a \textit{Poisson sprinkling}, or just a \textit{sprinkling}.
-It is a Poisson process which provides a way to generate a causet from a $d$-dimensional Lorentzian manifold $(\mathcal{M},g)$ by selecting points in $\mathcal{M}$ to be the elements of $\mathcal{C}$, with an order relation given by the causal order of the manifold.
+It is a Poisson process which provides a way to generate a causet from a $d$-dimensional Lorentzian manifold $(M,g)$ by selecting certain points in $M$ to be the elements of $\mathcal{C}$, with an order relation given by the causal order of the manifold.
 The number of points chosen in a region of spacetime volume, $V$, is a Poisson random variable.
@@
 It is called a sprinkling as one can envisage the point selection process as a `sprinkling' of points into the manifold.
-If a causet, $\mathcal{C}$, can be generated with relatively high probability by a sprinkling into the manifold, $(\mathcal{M},g)$, then we say that the manifold is a good approximation for the causal set.
+If a causet, $\mathcal{C}$, can be generated with relatively high probability by a sprinkling into the manifold, $(M,g)$, then we say that the manifold is a good approximation for the causal set.
 
@@
 \be\label{eq:GHYBT_in_continuum}
-S_{GHY} = \pm \frac{1}{l_p^{d-2}}\int_{\Sigma} K d\Sigma
+S_{GHY} = \mp \frac{1}{l_p^{d-2}}\int_{\Sigma} K d\Sigma
 \ee
@@
 Now observe that the integral in~\eqref{eq:GHYBT_in_continuum} can be thought of as the ``volume gradient'' across the surface $\Sigma$. More formally
-\be
+\be\label{eq:normal_deriv_boundary}
 \int K d\Sigma = \frac{\partial}{\partial n}\int d\Sigma,
@@
 \mb{does the antichain analogy really stand?}
-In the causal set, the analogue of a spacelike hypersurface is an anti-chain, i.e.
-a subset $\mathcal{A}\subseteq\mathcal{C}$ in which all elements are unrelated.
-The intuitive analogue of the boundary term would then be the rate of change of the number of causal set elements below and above the antichain.
-We shall see that this intuitive idea indeed bears out.
+
+We will define the causal set analogue of a spacelike hypersurface to be a particular partition of the set $\mathcal{C}$ into parts $\mathcal{C}^+$ and $\mathcal{C}^-$.
+For this partition to make sense as a spacelike hypersurface we require that no elements in $\mathcal{C}^+$ can precede any elements in $\mathcal{C}^-$.
+The reason for this condition is that if one could faithfully embed $\mathcal{C}$ in some manifold, $M$ ($\mathcal{C}$ can be generated via a sprinkling into $M$), then given a spacelike hypersurface, $\Sigma$, we can partition the causet into two bits, denoting the future(past) part as $\mathcal{C}^+$($\mathcal{C}^-$), which is made up of those elements sprinkled into $M^+$($M^-$).
+The causal relations between the two parts immediately satisfy our requirement for the partition to be spacelike.
+We will refer to a partition as spacelike if it abides by this condition, even if the causet is not embeddable in some manifold.
+Spacetime volume in the causal set is obtained simply by counting the number of elements, and hence the volume gradient intuitively corresponds to the difference in the number of ``immediate neighbours'' to the future and to the past of the spacelike partition.
+What does it mean to be immediate neighbours to a spacelike partition?
+For immediate neighbours to the past(future) of the partition we mean elements of $\mathcal{C}^-$($\mathcal{C}^+$) that have no elements to their future(past) in $\mathcal{C}^-$($\mathcal{C}^+$) (these elements being called \textit{maximal}(\textit{minimal})).
+We shall see that this intuitive idea for the boundary term indeed bears out.
+However, one does not need to know the maximal and minimal elements to calculate the boundary term.
+It can be found from any elements that lie reasonably close to the surface (in the time coordinate on $M$) but are not necessarily maximal or minimal.
+This will be made more precise below.
+
+In the continuum the GHY term can be calculated for a spacelike hypersurface, $\Sigma$, that is either a future or past boundary of $M$.
+We wish to calculate the GHY term for future or past boundaries of causets, but there are a couple issues we must deal with.
+Firstly, if we can embed a causet $\mathcal{C}$ into $M$ then the boundary surface of $M$ will not induce a partition of $\mathcal{C}$ as it did previously (every element of $\mathcal{C}$ will lie in the bulk of $M$), thus our previous idea for the boundary term does not work any more.
+Secondly, what does it mean to say a bounding surface of a causet if it is not embeddable in some manifold?
+Previously if the causet was not embeddable we could still define the surface as a partition of the causet.
+We shall see that the first issue is not a problem, and we will show that the GHY term can be found for a future(past) boundary of a causet (that is embeddable in $M$) by taking a difference in the number of certain elements that lie close to the future(past) bounding surface of $M$.
+We state that calculating the GHY term this way must apply whether or not the causet is embeddable, but this runs us into the second issue above.
+For what boundary are we calculating our GHY term if the causet is not embeddable?
+We propose that spacelike future and past boundaries exist for all causets of finite cardinality.
+By definition these causets are finite in temporal extent, and so if they can be embedded in some manifold then that manifold can always have spacelike future and past boundaries.
+These future and past boundaries are abstract objects for which quantities such as the GHY term can be calculated for.
+They only have physical meaning when the causet is embeddable in some manifold with a future(past) spacelike boundary, denoted $\Sigma$, in which case the quantities we calculate for the abstract future(past) boundary will agree with those calculated for $\Sigma$ (calculated with the continuum definitions) in some continuum limit.
+
+One can also use partitions of a causet to define the causal set analogues of timelike and null hypersurfaces.
+If the partition is such that an element of $\mathcal{C}^-$ precedes and element of $\mathcal{C}^+$ which again precedes and element of $\mathcal{C}^-$, then the partition corresponds to a timelike hypersurface.
+If we have a partition that satisfies the requirements for a spacelike partition, we can define a null partition as that which has the following restrictions.
+For every $x_0\in\mathcal{C}^+$ one can find a sequence of elements $x_0 \succeq x_1 \succeq ...\succeq x_n$ such that $x_n$ is preceded by nothing and all the $x_i \in \mathcal{C}^+$; and for every $x_0\in\mathcal{C}^-$ one can find a sequence of elements $x_0 \preceq x_1 \preceq ...\preceq x_n$ such that $x_n$ precedes nothing and all the $x_i \in \mathcal{C}^-$.
+There is also the possibility of a surface changing between spacelike, timelike and null at different parts.
+In this case the situation is a little more complicated in causal set terms, and one must restrict to a suitable subsets of points such that one of the definitions for the different types of surfaces holds true while the others fail.
+This subset of points, with its particular relations across the particular partition, then corresponds to a portion of the surface being either spacelike, timelike or null.
+
 %This suggests a very simple analogy for a causal set with a ``spacelike hypersurface''.
@@
 
-Consider a causal set $\mathcal C$ obtained by a Poisson sprinkling at density $\rho=1/l^d$ into a $d$-dimensional spacetime $(M,g)$.
-A Cauchy surface $\Sigma\subset M$ induces a partition $\mathcal C = \mathcal C^+ \cup\, \mathcal C^-$ of the sprinkling, where $\mathcal C^+$ and $\mathcal C^-$ denote the restrictions of $\mathcal C$ to the points sprinkled to the future and past of $\Sigma$, respectively.
-Spacetime volume in the causal set is obtained simply by counting the number of elements, and hence the volume gradient corresponds to the difference in the number of ``immediate neighbours'' to the future and to the past of the surface.
-The most natural definition for the nearest neighbours to the future/past of $\Sigma$ in the sprinkling is to take the minimal/maximal elements in $\mathcal C^\pm$, respectively.
-Thus $x\in \mathcal C^-$ is a maximal element in $\mathcal C^-$ if it is maximal in the precedence relation of the causal set, i.e.
-if there exists no element $y\in\mathcal C^-$ such that $x\preceq y$.
-Figure \ref{fig:Nmin_Nmax} shows an illustration of the idea.
-The minimal and maximal elements have been highlighted.
-%in COLOUR.
+Let us define the operators $N_{\vphantom{i}max}^{(k)}\left[\mathcal{C} \right]$ and $N_{min}^{(k)}\left[\mathcal{C}^- \right]$ that, for a given causet, $\mathcal{C}$, return the number of $k$-next-to-maximal/minimal elements respectively.
+A $k$-next-to-maximal(minimal) element is one for which there are $k$ elements to its future(past).
+This means that the $0$-to-maximal(minimal) elements are the maximal(minimal) elements.
 
-We will also show that it is possible to restrict oneself to either $\mathcal C^+$ or $\mathcal C^-$ and still calculate the GHY term.
-To do this requires us to define the $k$-next-to-maximal/minimal elements in $\mathcal C^-$ or $\mathcal C^+$ respectively.
-These are the elements of the causal set that have $k$ elements to their future or past.
-This means that the $0$-to-maximal/minimal elements are the maximal/minimal elements.
-%Elements with $1$ element to their past/future have been highlighted in COLOUR in Figure \ref{fig:Nmin_Nmax}, and those with $2$ have been highlighted in COLOUR.
-
-Let us denote the number of $k$-next-to-maximal elements in $\mathcal C^-$ by $N_{max}^{(k)}$ and the number of $k$-next-to-minimal elements in $\mathcal C^+$  by $N_{min}^{(k)}$.
-We propose the following definitions for the discrete Gibbons-Hawking-York boundary term (one can adopt the appropriate definition to best suit their calculation needs):\mb{put $\Sigma$ in argument, $S[\mathcal C, \Sigma]$?}
-
-\be\label{GH_boundary_to_causet}
- \mathcal{S}^{(d)}_{GHY}[\mathcal C]=\left(l/l_p\right)^{d-2}\frac{c_{d}}{\Gamma\left(\frac{2}{d} \right)}\times
-  \begin{cases}
-   \frac{1}{2}\left(N_{max}^{(0)}-N_{min}^{(0)}\right) \\
-   d\: N_{max}^{(1)}-N_{max}^{(0)} \\
-   N_{min}^{(0)}-d\: N_{min}^{(1)}
-  \end{cases}
+Given a causet, $\mathcal{C}$, with a spacelike partition so that $\mathcal C = \mathcal C^+ \cup\, \mathcal C^-$, we propose the following family of causal set operators for calculating the GHY boundary term of this partition:
+\be\label{general_boundary_sum}
+\mathcal{S}^{(d)}_{GHY}\left[\mathcal{C}^-,\mathcal{C}^+;\mathbf{p}, \mathbf{q} \right]= \left(l/l_p\right)^{d-2} c_{d}
+\left( \sum_m p_m N_{max}^{(m)}\left[\mathcal{C}^- \right]
++  \sum_n q_n N_{min}^{(n)}\left[\mathcal{C}^+ \right]\right)
 \ee
@@
 \be\label{Cn}
-c_{d}=\frac{d(d+1)}{(d+2)}\left[\frac{A_{d-2}}{d(d-1)}\right]^{\frac{2}{d}},
+c_{d}=\frac{d(d+1)}{(d+2)}\left[\frac{V_{d-1}}{d}\right]^{\frac{2}{d}},
 \ee
-$A_d=2\pi^\frac{d+1}2/\Gamma\left(\frac{d+1}{2}\right)$ and denotes the volume of the unit $d$-sphere.
-In fact, one can write down a general expression for the boundary term as a sum over $k$  $N_{\vphantom{i}max}^{\: (k)}$ and/or $N_{min}^{\: (k)}$ terms with particular coefficients.
-This can be written as follows:
+and where $V_d=\pi^{\frac{d}{2}}/\Gamma\left(\frac{d}{2}+1\right)$ denotes the volume of the unit $d$-ball ($\Gamma(t)$ is a gamma function, defined as $\Gamma(t):=\int_0^\infty dx\: x^{t-1}\: e^{-x}$).
+We have used $\mathbf{p}$ and $\mathbf{q}$ to stand for infinite dimensional vectors $(p_0,...,p_m,..)$ and $(q_0,...,q_n,..)$ respectively, where all the $p_m,\: q_n \in \mathbb{R}$.
+These vectors contain a finite number of non-zero entries, corresponding to the coefficients of the different $N_{{max\vphantom{i}}}^{\:(m)}$ and $N_{{min}}^{\:(n)}$ terms.
+Their values can be chosen arbitrarily, so long as they satisfy the following conditions.
+\begin{align}\label{coefficient_relation1}
+& \sum_m p_m \frac{\Gamma\left(\frac{1}{d}+m \right)}{m!}  + \sum_n q_n\frac{\Gamma\left(\frac{1}{d}+n \right)}{n!}=0
+\\
+& \label{coefficient_relation2}\sum_m p_m \frac{\Gamma\left(\frac{2}{d}+m \right)}{m!}  - \sum_n q_n\frac{\Gamma\left(\frac{2}{d}+n \right)}{n!}=1
+\end{align}
+The aforementioned family simply corresponds to all the different vectors, $\mathbf{p}$ and $\mathbf{q}$, that satisfy the above criteria.
 
-\begin{gather}\label{general_boundary_sum}
-\begin{aligned}
-\mathcal{S}^{(d)}_{GHY}[\mathcal C]=\left(l/l_p\right)^{d-2} & c_{d}\left( \sum_m p_m \frac{\Gamma\left(\frac{2}{d}+m \right)}{\Gamma\left(\frac{1}{d}+m \right)}  - \sum_n q_n\frac{\Gamma\left(\frac{2}{d}+n \right)}{\Gamma\left(\frac{1}{d}+n \right)} \right)^{-1} \\
-& \times \left( \sum_m p_m \frac{m!}{\Gamma\left(\frac{1}{d}+m \right)} N_{max}^{(m)}
-+  \sum_n q_n \frac{n!}{\Gamma\left(\frac{1}{d}+n \right)} N_{min}^{(n)}\right)
-\end{aligned}
-\end{gather}
-The coefficients, $p_m , q_n \in \mathbb{R}$, must satisfy the following relation:
+Given a $d$ dimensional, spatially compact Lorentzian manifold, $M$, the family of operators $\mathcal{S}^{(d)}_{GHY}$ (for different $\mathbf{p}$ and $\mathbf{q}$) defines a family of random variables in the following way.
+The process of sprinkling points into $M$ with density $\rho=l^{-d}$ generates a random causet $\mathcal{C}$, and so the operators $N_{\vphantom{i}max}^{(k)}$ and $N_{min}^{(k)}$ acting on this random causet are now the random variables $\textbf{N}_{\vphantom{i}max}^{(k)}$ and $\textbf{N}_{min}^{(k)}$.
+These random variables can be substituted into~\eqref{general_boundary_sum} for $N_{\vphantom{i}max}^{(k)}$ and $N_{min}^{(k)}$ to give the family of random variables, $\textbf{S}^{(d)}_{GHY}$.
+These random variables will be functions of $\rho$ and the two parts of the manifold, $M^{\pm}$, where  $\textbf{N}_{\vphantom{i}max}^{(k)}$($\textbf{N}_{min}^{(k)}$) depends on $M^-$($M^+$) and $\rho$.
+The expectation value of $\textbf{S}^{(d)}_{GHY}$ in this sprinkling process will be denoted by $\overline{S}^{(d)}_{GHY}:=\mathbb{E}(\textbf{S}^{(d)}_{GHY})$.
+We claim that in the limit of infinite density the value of $\overline{S}^{(d)}_{GHY}$, for any vectors $\mathbf{p}$ and $\mathbf{q}$ satisfying~\eqref{coefficient_relation1} and ~\eqref{coefficient_relation1}, tends to the continuum GHY boundary term for a surface $\Sigma$ that partitions the manifold into $M^+$ and $M^-$:
+\be
+\lim_{l\rightarrow0}\overline{S}^{(d)}_{GHY}[M^-,M^+,\rho;\mathbf{p} , \mathbf{q}]= \frac{1}{l_p^{d-2}}\int_{\Sigma} d^{d-1}x\: \sqrt{h}\: K\label{eq:mainconjecture}
+\ee
 
-\be\label{coefficient_relation}
-\sum_m p_m + \sum_n q_n = 0
+There is a huge amount of freedom that remains in the choice of $\mathbf{p}$ and $\mathbf{q}$.
+One can see that at least two terms are necessary in order to satisfy~\eqref{coefficient_relation1} and~\eqref{coefficient_relation2}.
+For any two terms the coefficients will be unique, and for more than two terms this uniqueness is lost.
+In a finite difference method one can form a discrete derivative in one direction by taking a difference of two values at different points along that direction, and dividing by the distance between them.
+We are doing something very similar here.
+The boundary action (in the continuum) is a derivative in one direction, the normal direction (as in~\eqref{eq:normal_deriv_boundary}), and so a finite difference method should be able to approximate this with just two terms.
+This agrees with the fact that we are able write the discrete boundary action with only two terms.
+
+There are a few, two-term, members of the family that are worth mentioning, as they have simple forms that make them good candidates for computations.
+Given a causet, $\mathcal{C}$, with a spacelike partition so that $\mathcal C = \mathcal C^+ \cup\, \mathcal C^-$, we define the causal set operator $\mathcal{S}^{(d)}_{\pm}$ as a particular member of the family in~\eqref{general_boundary_sum}:
+\be\label{eq:min_max_simple_formula}
+\mathcal{S}^{(d)}_{\pm}\left[\mathcal{C}^-,\mathcal{C}^+\right]:=\mathcal{S}^{(d)}_{GHY}\left[\mathcal{C}^-,\mathcal{C}^+;\mathbf{p}, -\mathbf{p} \right]=\left(l/l_p\right)^{d-2}\frac{c_{d}}{2\Gamma\left(\frac{2}{d} \right)}\left( N_{\vphantom{i}max}^{\: (0)}\left[\mathcal{C}^- \right] - N_{min}^{\: (0)}\left[\mathcal{C}^+ \right] \right)
 \ee
-The different definitions of the boundary term in (\ref{GH_boundary_to_causet}) are just simple cases of this general formula, and the objects that one would preferably use in a calculation.
-From above one can see that at least two terms are needed in the sum to find the boundary term.
-The condition~\eqref{coefficient_relation} means that there will be one free parameter if two terms are present\footnote{For example you could include only the maximal and minimal terms (like the first case in~\eqref{GH_boundary_to_causet}) with $p_0=p$ and $q_0=-p$ where $p\in\mathbb{R}$.
-This would satisfy~\eqref{coefficient_relation} and would leave $p$ as a free parameter.
-In this case you would get a factor of $p^{-1}$ from the term after $c_d$ in~\eqref{general_boundary_sum} which will cancel the $p$ from the last term.}.
-This free parameter will be cancelled by the term after $c_d$ in~\eqref{general_boundary_sum}.
-Thus, with only two terms, the form of the discrete boundary term is fixed.
-If one chooses to include more than two terms the coefficients of the $N_{{max\vphantom{i}}}^{\:(k)}$ and/or $N_{{min}}^{\:(k)}$ terms are no longer unique.
+where $\mathbf{p}=\left(\left(2\Gamma\left(\frac{2}{d} \right)\right)^{-1},0,0,...\right)$.
+This expression is the easiest to employ computationally, and we shall use its random variable counterpart, $\textbf{S}^{(d)}_{\pm}\left[M^-,M^+,\rho\right]:=\textbf{S}^{(d)}_{GHY}\left[M^-,M^+,\rho;\mathbf{p}, -\mathbf{p} \right]$, later when we study the fluctuations of the discrete boundary actions numerically.
 
-To support this proposal we show that in the continuum limit of $l\rightarrow 0$ (i.e. $\rho\rightarrow\infty$) we obtain
-%of infinite sprinkling density into a $d$-dimensional spacetime we obtain
-\be
-\lim_{l\rightarrow0}\langle S^{(d)}_{GHY}[\mathcal C] \rangle= \frac{1}{l_p^{d-2}}\int_{\Sigma} d^{d-1}x\: \sqrt{h}\: K\label{eq:mainconjecture}
+If one wants to calculate the GHY boundary term for the future(past) boundary of a single causet, $\mathcal{C}$, then one simply inserts $\mathcal{C}$ into the $\mathcal{C}^-$($\mathcal{C}^+$) argument of $\mathcal{S}^{(d)}_{GHY}$ and sets $\mathbf{q}$($\mathbf{p}$) to the zero vector, $\mathbf{0}=(0,0,...)$.
+This renders the $\mathcal{C}^+$($\mathcal{C}^-$) argument unnecessary and so one can insert the empty set, $\emptyset$, for convenience.
+We define the operators for the future and past boundaries respectively as:
+\begin{align}\label{eq:future_past_boundary_terms1}
+\mathcal{S}^{(d)}_{+}[\mathcal{C}]:= & \mathcal{S}^{(d)}_{GHY}\left[\mathcal{C},\emptyset;\mathbf{p},\mathbf{0}\right] =\left(l/l_p\right)^{d-2}\frac{c_{d}}{\Gamma\left(\frac{2}{d} \right)}\left( d\:N_{\vphantom{i}max}^{\: (1)}\left[\mathcal{C} \right] - N_{\vphantom{i}max}^{\: (0)}\left[\mathcal{C} \right] \right)
+\\
+\label{eq:future_past_boundary_terms2}
+\mathcal{S}^{(d)}_{-}[\mathcal{C}]:= & \mathcal{S}^{(d)}_{GHY}\left[\emptyset,\mathcal{C};\mathbf{0},-\mathbf{p}\right] =\left(l/l_p\right)^{d-2}\frac{c_{d}}{\Gamma\left(\frac{2}{d} \right)}\left( N_{min}^{\: (0)}\left[\mathcal{C} \right] - d\: N_{min}^{\: (1)}\left[\mathcal{C} \right] \right)
+\end{align}
+where $\mathbf{p}=\left(-\Gamma\left(\frac{2}{d} \right)^{-1},d\:\Gamma\left(\frac{2}{d} \right)^{-1},0,0,...\right)$.
+In writing down an action for a particular causal set, $\mathcal{C}$, one would write down the Benincasa-Dowker action, $\mathcal{S}^{(d)}_{BD}[\mathcal{C}]$.
+To complete the story we propose that one must include the boundary terms $\mathcal{S}^{(d)}_{+}[\mathcal{C}]$ and $\mathcal{S}^{(d)}_{-}[\mathcal{C}]$.
+Explicitly, the total action for a causal set should be given by:
+\be\label{eq:total_causet_action}
+\mathcal{S}^{(d)}=\mathcal{S}^{(d)}_{BD}+\mathcal{S}^{(d)}_{+}-\mathcal{S}^{(d)}_{-}
 \ee
-for the most general definition above. Here $\langle\cdot\rangle$ denotes the mean over sprinklings.
+where we have omitted the causet argument, $\mathcal{C}$.
 
-We also propose a discrete definition for the spatial volume of the surface:
+There may be contributions from the causal set analogue of timelike boundaries (although there are arguments to suggest that they make no contribution), and we discuss below certain attempts we have made to find discrete GHY terms for such boundaries.
 
-\be\label{eq:surface_volume}
-\mathcal{A}^{(d)}[\mathcal{C}]=\left(l/l_p\right)^{d-1}\frac{b_{d}}{\Gamma\left(\frac{1}{d}\right)}\: N_{{\,}_{max/min}}^{\:(0)}
+We also propose a family of operators for the causal set analogue of spatial volume for the spacelike partition:
+\be\label{general_area_sum}
+\mathcal{A}^{(d)}[\mathcal{C}^-,\mathcal{C}^+;\mathbf{p},\mathbf{q}]=\left(l/l_p\right)^{d-1}b_{d}\left(\sum_m p_m N_{\vphantom{i}max}^{\: (m)}\left[\mathcal{C}^- \right] + \sum_n q_n N_{min}^{\: (n)}\left[\mathcal{C}^+ \right]\right)
 \ee
 where the constant $b_d$ is given by
+\be\label{constant_b_d}
+b_d=d\left[\frac{V_{d-1}}{d}\right]^{\frac{1}{d}}
+\ee
+and we have used similar definitions for the arguments $\mathbf{p}$ and $\mathbf{q}$. This time however, the coefficients must satisfy 
+\be\label{area_coefficient_relation}
+\sum_m p_m \frac{\Gamma\left(\frac{1}{d}+m \right)}{m!}  + \sum_n q_n\frac{\Gamma\left(\frac{1}{d}+n \right)}{n!}=1
+\ee
+From this relation we can see that only one term is necessary to determine an expression for the discrete surface volume, and that when more than one term is added the coefficients will cease to be unique.
+Once again we can define a family of random variables, $\textbf{A}^{(d)}$, and their expectation values, $\overline{A}^{(d)}:=\mathbb{E}(\textbf{A}^{(d)})$, which will depend upon $M^{\pm}$ and $\rho$.
+We make another claim that, in the limit of infinite density, $\overline{A}^{(d)}$ tends to the spatial volume of the surface $\Sigma$.
+\be\label{eq:conjecture_for_area}
+\lim_{l\rightarrow0}\overline{A}^{(d)}[M^-,M^+,\rho;\mathbf{p} , \mathbf{q}]= \frac{1}{l_p^{d-1}}\int_{\Sigma} d^{d-1}x\: \sqrt{h}
+\ee
+where the right hand side is the spatial volume of the surface $\Sigma$. In proving~\eqref{eq:mainconjecture} we will establish the results necessary to prove this claim as well.
 
-\be\label{constant_b_d}
-b_d=d\left[\frac{A_{d-2}}{d(d-1)}\right]^{\frac{1}{d}}
-\ee
-Again on can find a general expression written as a sum of different terms.
-
-\be\label{general_area_sum}
-\mathcal{A}^{(d)}[\mathcal{C}]=\left(l/l_p\right)^{d-1}b_{d}\sum_m p_m \frac{m!}{\Gamma\left(\frac{1}{d}+m \right)}\: N_{{\,}_{max/min}}^{\:(m)}
-\ee
-where the coefficients, $p_m$, must satisfy $\sum_m p_m =1$.
-This time we see that only one term is needed to find the surface volume, and that when more than one term is added the coefficients will not be unique.
-We support this claim in the same way as the boundary term, and show that
-
-\be\label{eq:conjecture_for_area}
-\lim_{l\rightarrow0}\langle \mathcal{A}^{(d)}[\mathcal{C}] \rangle= \frac{1}{l_p^{d-1}}\int_{\Sigma} d^{d-1}x\: \sqrt{h}
-\ee
-where the right hand side is the spatial volume of the surface $\Sigma$.
+By making the same restrictions on the arguments as we did for $\mathcal{S}^{(d)}_{GHY}$, we can find expressions for the causal set analogue of the spatial volume of the future or past boundaries.
+We define operators for the future and past boundaries respectively as two of the simplest members of the family:
+\begin{align}\label{eq:future_past_spatial_volume}
+\mathcal{A}^{(d)}_{+}[\mathcal{C}]:=\mathcal{A}^{(d)}[\mathcal{C},\emptyset;\mathbf{p},\mathbf{0}]= & \left(l/l_p\right)^{d-1}\frac{b_{d}}{\Gamma\left(\frac{1}{d}\right)}\: N_{\vphantom{i}max}^{\: (0)}[\mathcal{C}]
+\\
+\mathcal{A}^{(d)}_{-}[\mathcal{C}]:=\mathcal{A}^{(d)}[\emptyset,\mathcal{C};\mathbf{0},\mathbf{p}]= & \left(l/l_p\right)^{d-1}\frac{b_{d}}{\Gamma\left(\frac{1}{d}\right)}\: N_{min}^{\: (0)}[\mathcal{C}]
+\end{align}
+where $\mathbf{p}=(\Gamma\left(\frac{1}{d}\right)^{-1},0,0,...)$.
 
@@
 These approximations are justified on the grounds that if certain orders are retained throughout the calculation they will in fact vanish in the limit $\rho \rightarrow \infty$
-\begin{figure}
-  \centering
-    {\includegraphics[width=\textwidth]{minmaxplot}}
-     \caption{An illustration of the idea.
+%\begin{figure}
+  %\centering
+   % {\includegraphics[width=\textwidth]{minmaxplot}}
+    % \caption{An illustration of the idea.
 Pictured here is a causal set obtained by a sprinkling into a spacetime partitioned by a spacelike hypersurface.
 The minimal and maximal points about the surface are highlighted and the shaded regions illustrate the regions whose volumes $V_\blacktriangle$ and $V_\blacktriangledown$ are needed in the proof.}
-     \label{fig:Nmin_Nmax}
-\end{figure}
+     %\label{fig:Nmin_Nmax}
+%*\end{figure}
 
@@
 
-\subsection{Poisson Sprinklings for $\langle N_{max}^{(m)}\rangle$ and $\langle N_{min}^{(m)}\rangle$}
+\subsection{Poisson Sprinklings for $\left\langle N_{max}^{\:(m)}\right\rangle$ and $\left\langle N_{min}^{\: (n)}\right\rangle$}
 
-In order to prove~\eqref{eq:mainconjecture} and~\eqref{eq:conjecture_for_area} let us first derive expressions for the mean values of $N_{max}^{(m)}$ and $N_{min}^{(m)}$.
+In order to prove~\eqref{eq:mainconjecture} and~\eqref{eq:conjecture_for_area} let us first derive expressions for the mean values of $N_{\vphantom{i}max}^{\: (m)}$ and $N_{min}^{(n)}$.
+The proof will will follow almost automatically, as~\eqref{general_boundary_sum} and~\eqref{general_area_sum} are just linear combination of $N_{\vphantom{i}max}^{\: (m)}$ and $N_{min}^{(n)}$ terms, and thus their mean values will just be linear combinations of $\left\langle N_{\vphantom{i}max}^{\: (m)}\right\rangle$ and $\left\langle N_{min}^{\: (n)}\right\rangle$ terms.
+
 For any instance of the sprinkling, the probability that a sprinkled point $p\in M$ below the surface $\Sigma$ is $k$\emph{-next-to-maximal} is given by the probability that $k$ points of the sprinkling lie in the region $J^{+}(p)\cap J^{-}(\Sigma)$, the intersection of the causal future of $p$ with the causal past of the surface $\Sigma$.\footnote{For notational convenience we use the symbol $x$ to refer to the causal element $x\in\mathcal C$, to its embedding in the manifold $M$, and to its coordinates in some chart on $M$.} This region will in general be the interior of some curvy $d$-dimensional cone truncated by the surface $\Sigma$, as illustrated in Figure~\ref{fig:Nmin_Nmax}.
-We will refer to these regions as \emph{truncated cones} below.
+We will refer to these regions as \emph{truncated cones}.
 The Poisson process assigns a probability
@@
 
-In the limit $\rho\rightarrow\infty$, both quantities will diverge, but if some combination of maximal and minimal terms grows slower than or at order $\rho^{1-\frac2d}$, the proposed action~\eqref{general_boundary_sum} will tend to a finite value in the continuum limit.
+In the limit $\rho\rightarrow\infty$, both quantities will diverge, but if some combination of $k$-next-to-maximal/minimal terms grows slower than or at order $\rho^{1-\frac2d}$, the proposed action~\eqref{general_boundary_sum} will tend to a finite value in the continuum limit.
 
@@
 Hence, as $\rho\rightarrow\infty$, the integration ranges in \eqref{eq:nmax} and \eqref{eq:nmin} can be cut off at finite time $t=\pm\varepsilon$ with $\varepsilon$ arbitrarily small, which allows us to expand the time integration-variable about zero.
-These assumptions of course impose certain regularity conditions on the surface $\Sigma$, but we will assume that they are reasonable conditions.\mb{do we need to say more about this?
+These assumptions of course impose certain regularity conditions on the surface $\Sigma$, but so long as $\Sigma$ is compact and everywhere spacelike  then these conditions are satisfied.\mb{do we need to say more about this?
 e.g.
@@
 %\mb{this sounds circular w.r.t. last paragraph} 
-This means that the volumes can be approximated by the volumes of less curvy cones, and it can be shown that higher order corrections, in the approximation scheme, vanish in the limit of $\rho \rightarrow \infty$.
-This will all be made more precise below.
+This means that the volumes can be approximated in some scheme that will be made precise below.
+Any higher order corrections to these volumes, above the order we require, will vanish in the limit of $\rho\rightarrow\infty$.
 
@@
 This is now a simple volume integral but we need to find the boundaries of $\mathcal{X}_p$ to write down the integration limits.
-There are two boundaries to the (solid) truncated cone: the lightcone emanating at $p$ and the base, given by the intersection of $\Sigma$ with the interior of the lightcone.
+There are two parts to the boundary of the (solid) truncated cone: the lightcone emanating at $p$ and the base, given by the intersection of $\Sigma$ with the interior of the lightcone.
 First we look at the lightcone.
@@
 The $O(T^2)$ term will contribute at $O(T^{d+2})$ in the volume integral and so can be ignored.
-Substituting $r_{max}=T$ into (\ref{eq:VolumeIntegralSpherical}) lets us evaluate the integral and we find
-\begin{figure}[t]
-  \centering
-    {\includegraphics[scale=0.5]{coneplot}}
-     \caption{The size of the region inside the top and bottom bounding surfaces is the volume we want to calculate.}
-     \label{fig:cone_plot}
-\end{figure}
+Substituting $r_{max}=T$ into (\ref{eq:VolumeIntegralSpherical}) allows us to evaluate the integral, wherein we find
+%\begin{figure}[t]
+ % \centering
+  %  {\includegraphics[scale=0.5]{coneplot}}
+   %  \caption{The size of the region inside the top and bottom bounding surfaces is the volume we want to calculate.}
+    % \label{fig:cone_plot}
+%\end{figure}
 \be\label{eq:VolumeNoK}
 V_\blacktriangle(p)
-=\frac{A_{d-2}}{d(d-1)}T^d\left(1-\frac{d}{2(d+1)}\Gamma^{0}_{\;ij}(p_0)A^{i}_{\;\ibar}A^{j}_{\;\jbar}\delta^{\ibar\jbar}T\right)
+=\frac{V_{d-1}}{d}T^d\left(1-\frac{d}{2(d+1)}\Gamma^{0}_{\;ij}(p_0)A^{i}_{\;\ibar}A^{j}_{\;\jbar}\delta^{\ibar\jbar}T\right)
 +O(T^{d+2})
 \ee
-where $A_{d-2}$ is the volume of the unit $(d-2)$-sphere and the $\delta^{\ibar\jbar}$ comes from the fact that cross terms ($\ibar\neq \jbar$) vanish under the angular integration.
+where the $\delta^{\ibar\jbar}$ comes from the fact that cross terms ($\ibar\neq \jbar$) vanish under the angular integration.
 The defining relations for $A^{i}_{\;\ibar}$ can now be rearranged to give $A^{i}_{\;\ibar}A^{j}_{\;\jbar}\delta^{\ibar\jbar}=h^{ij}(p_0)$.
-Now in GNCs the extrinsic curvature on the surface is given by
+In GNCs the extrinsic curvature on the surface is given by
 
@@
 \ee
-Substituting this into~\eqref{eq:VolumeNoK} with we obtain 
+Substituting this into~\eqref{eq:VolumeNoK} we obtain 
 \begin{align}
 V_\blacktriangle(T,\mathbf x)
-&=\frac{A_{d-2}}{d(d-1)}T^d\left(1+\frac{d}{2(d+1)}K(0,\mathbf{x})T\right)
+&=\frac{V_{d-1}}{d}T^d\left(1+\frac{d}{2(d+1)}K(0,\mathbf{x})T\right)
 +O(T^{d+2}) \label{eq:TopVolumeWithK}\\
 V_\blacktriangledown(-T,\mathbf x)
-&=\frac{A_{d-2}}{d(d-1)}T^d\left(1-\frac{d}{2(d+1)}K(0,\mathbf{x})T\right)
+&=\frac{V_{d-1}}{d}T^d\left(1-\frac{d}{2(d+1)}K(0,\mathbf{x})T\right)
 +O(T^{d+2}), \label{eq:BottomVolumeWithK}
 \end{align}
-having dropped the subscript $p$. Given the volume expressions in GNCs we now proceed to evaluate the integrals for $\langle N_{min}^{(k)}\rangle $ and $\langle N_{max}^{(k)}\rangle$.
+having dropped the subscript $p$. Given the volume expressions in GNCs we now proceed to evaluate the integrals for $\langle N_{min}^{(k)}\rangle $ and $\langle N_{\vphantom{i}max}^{(k)}\rangle$.
 
-\subsection{The mean of $S^{(d)}_{GHY}[\mathcal C]$}
+\subsection{The mean of $S^{(d)}_{GHY}[\mathcal {C}^-,\mathbf{p};\mathcal{C}^+,\mathbf{q}]$}
 
@@
 \begin{aligned}
-A & := \frac{A_{d-2}}{d(d-1)} \\
-B & := \frac{A_{d-2}}{2(d-1)(d+1)}K(0,\mathbf{x})
+A & := \frac{V_{d-1}}{d} \\
+B & := \frac{V_{d-1}}{2(d+1)}K(0,\mathbf{x})
 \end{aligned}
@@
 We can use this procedure to take the limit of~\eqref{eq:n_min_different_rho_terms} and formulate the answer in terms of gamma functions.
-The limits of both $\langle N_{min}^{(k)}\rangle$ and $\langle N_{min}^{(k)}\rangle$ are then given by
+The limits of both $\langle N_{\vphantom{i}max}^{(k)}\rangle$ and $\langle N_{min}^{(k)}\rangle$ are then given by
 
@@
 
-These object diverge as $\rho^{1-\frac{1}{d}}$ in the limit.
-The factor of $\rho$ that must be included in the formula for the surface volume is exactly the inverse of this ($\rho^{\frac{1}{d}-1}$), as can be seen from~\eqref{eq:surface_volume} if one substitutes in $l=\rho^{-\frac{1}{d}}$.
-If we include this factor above and then take the limit we see that the only remaining terms are proportional to the surface volume.
+These objects diverge as $\rho^{1-\frac{1}{d}}$ in the limit.
+The factor of $\rho$ that must be included in the formula for the surface volume is exactly the inverse of this ($\rho^{\frac{1}{d}-1}$), as can be seen from~\eqref{general_area_sum} if one substitutes in $l=\rho^{-\frac{1}{d}}$.
+If we include this factor above, and take the limit, we see that the only remaining terms are proportional to the surface volume.
 The other terms all have negative powers of $\rho$ and therefore vanish as $\rho\rightarrow\infty$.
-From this it follows that the mean of any linear combination of terms like $\rho^{\frac{1}{d}-1}\langle N_{max}^{(k)}\rangle$ or $\rho^{\frac{1}{d}-1}\langle N_{min}^{(k)}\rangle$ will give something proportional to the surface volume, so long as the terms containing the surface volume don't exactly cancel.
-To simplify the coefficients in this linear sum we take out a factor of $\frac{k!}{\Gamma\left(\frac{1}{d}+k\right)}$.
-The last step in arriving at~\eqref{general_area_sum} is to enforce the condition that the coefficients sum to unity, in order for the linear sum to equal the surface volume and not just be proportional.
-This concludes the proof of the surface volume conjecture.
+This means that if we take the mean of~\eqref{general_area_sum}, in the limit of $\rho\rightarrow\infty$, and substitute in the above expressions, we will be left with something proportional to the surface volume.
+Then it is easy to see that in order to get the surface volume exactly one must force $\mathbf{p}$ and $\mathbf{q}$ to satisfy~\eqref{area_coefficient_relation}.
+This concludes the proof of the spatial volume claim,~\eqref{eq:conjecture_for_area}.
 
-If we multiply~\eqref{eq:nmax_nmin_final} by the appropriate factor of $\rho$ from the boundary term ($\rho^{\frac{2}{d}-1}$) we find that the above equation still diverges as $\rho^{\frac{1}{d}}$ in the limit.
-This divergence comes from the surface volume term, while the term proportional to the GHY boundary term is finite.
-In order to recover the boundary term we must take a linear combination of these objects such that the term proportional to the surface are vanishes, while the term proportional to the boundary term does not.
-This is where the condition~\eqref{coefficient_relation} comes in.
-This condition for the coefficients is only so simple because we have taken out a factor of $\frac{k!}{\Gamma\left(\frac{1}{d}+k\right)}$ from each coefficient.
-The remaining sum is proportional to the boundary term, and in order to get it exactly we simply have to divide out by the proportionality constant.
-At this point one arrives at~\eqref{general_boundary_sum} and the proof is concluded.
+If we take the mean of~\eqref{general_boundary_sum} and the limit $\rho\rightarrow\infty$, we can substitute in the above expressions to find that the only terms left that have non-negative powers of $\rho$ (and therefore non-vanishing in the limit) are that of the spatial volume and boundary term.
+The term proportional to the spatial volume is divergent, so the coefficients, $\mathbf{p}$ and $\mathbf{q}$, must satisfy~\eqref{coefficient_relation1} to make this term vanish.
+The other condition,~\eqref{coefficient_relation2}, ensures that what remains is equal to the GHY boundary term.
+Thus, the proof of~\eqref{eq:mainconjecture} is complete.
+
+The form of~\eqref{eq:nmax_nmin_final} seems to suggest general series expansions of $\langle N_{max}^{(k)}\rangle$ and $\langle N_{min}^{(k)}\rangle$ in reducing powers of $\rho$.
+Written in terms of the discreteness scale, $l$, it is a Laurent series starting at $l^{1-d}$.
+The higher order terms, in $l$, will possibly be proportional to higher order normal derivatives of the surface volume.
+If this were true then one could find discrete versions for normal derivatives of the surface volume to any order.
+The exact form of the constant in the next term of the series is unknown and more work must be done to determine a general trend for the constants, and hence find the general series.
 
 \section{Fluctuations}
-So far we have only talked about the mean of the causal set boundary action.
-Let us now turn to its fluctuations (the standard deviation) $\sigma[S^{(d)}_{GHY}(\mathcal C)]=\text{Var}[S^{(d)}_{GHY}(\mathcal C)]^\frac12$.\footnote{Whenever we say fluctuations we refer to the standard deviation of the random variable, not to its variance.} We can make a heuristic argument to estimate the dependence of fluctuations on $\rho=l^{-d}$.
+So far we have only talked about the mean of the causal set boundary action for sprinklings.
+Let us now turn to its fluctuations (the standard deviation) $\sigma[\mathbf S^{(d)}_{GHY}]=\text{Var}[\mathbf S^{(d)}_{GHY}]^\frac12$.\footnote{Whenever we say fluctuations we refer to the standard deviation of the random variable, not to its variance.} If we restrict ourselves for now to the particular boundary action $\mathbf S^{(d)}_{0}$, which is proportional to the difference between the random variables $\mathbf N^{(0)}_{\vphantom{i}max}$ and $\mathbf N_{min}^{(0)}$ associated with $M^-$ and $M^+$, we can give a heuristic argument to estimate the dependence of fluctuations on $\rho=l^{-d}$.
 In any spacetime region of fixed volume $V$ the number of causal set elements $N$ experiences Poisson fluctuations of order $\sqrt N$.
-We will use the first case of~\eqref{GH_boundary_to_causet} as it is the simplest case in which to make this argument for the fluctuations.
-Since $N_{max}^{(0)}$ and $N_{min}^{(0)}$ are quantities associated with a codimension-1 surface (they count elements ``near" $\Sigma$), we may expect them to scale like $N^\frac{d-1}{d}$ --- which agrees with the leading order behaviour of~\eqref{eq:nmax_nmin_final} --- and thus to inherit fluctuations of order $N^\frac{d-1}{2d}$.
+Since $\mathbf N_{max}^{(0)}$ and $\mathbf N_{min}^{(0)}$ are quantities associated with a codimension-1 surface (they count elements ``near" $\Sigma$), we may expect them to scale like $N^\frac{d-1}{d}$ (which indeed agrees with the leading order behaviour of~\eqref{eq:nmax_nmin_final}) and thus to inherit fluctuations of order $N^\frac{d-1}{2d}$.
 %{This comes from the following observation.
@@
 }
-The action $S^{(d)}_{GHY}$, being proportional to $\rho^\frac{2-d}{d}$ times the difference of two independent\footnote{The independence is as a feature of the Poisson process in $M$: the number of elements $N(R)$ in any subregion $R\subset M$ is a Poisson variable itself, and for any two disjoint regions $R_1$, $R_2$ the numbers $N(R_1)$ and $N(R_2)$ are independent random variables.} random variables with standard deviation $N^\frac{d-1}{2d} = (\rho V)^\frac{d-1}{2d}$ should see fluctuations of order $\rho^\frac{2-d}{d}\rho^\frac{d-1}{2d}=\rho^\frac{3-d}{2d}$.
+The action $\mathbf S^{(d)}_{0}$, being proportional to $\rho^\frac{2-d}{d}$ times the difference of two \emph{independent}\footnote{The independence is as a feature of the Poisson process in $M$: the number of elements $N(R)$ in any subregion $R\subset M$ is a Poisson variable itself, and for any two disjoint regions $R_1$, $R_2$ the numbers $N(R_1)$ and $N(R_2)$ are independent random variables.} random variables with standard deviation $N^\frac{d-1}{2d} = (\rho V)^\frac{d-1}{2d}$ should then see fluctuations of order $\rho^\frac{2-d}{d}\rho^\frac{d-1}{2d}=\rho^\frac{3-d}{2d}$.
 This suggests that for $d=2$ these fluctuations should grow like $\rho^{\frac{1}{4}}$ as $\rho\rightarrow\infty$, for $d=3$ they should be constant, and for $d>3$ they should be damped.
 
-In order to test this heuristic argument one may like to write down the integral expression for $\text{Var}[S^{(d)}_{GHY}(\mathcal C)]^\frac12$ but it is complicated enough even in finite regions of flat space for which it is not very illuminating to reproduce here.
+In order to test this heuristic argument one may like to write down the integral expression for $\text{Var}[\mathbf S^{(d)}_{GHY}]$, but it is complicated enough even for $\mathbf S^{(d)}_0$ in finite regions of flat space that it will not be very illuminating to reproduce here.
 It also seems far less tractable than the expression for the mean because it involves terms of the type $\int d^dx\int d^dy \exp\left[-\rho V(J^+(x)\cap J^+(y))\right]$ for which the above methods of cutting off the time-integration range and Taylor expanding will not follow through.
@@
 
-\begin{figure}[t!]
-  \centering
-    {\includegraphics[scale=0.6]{GHY-fluctuation_plots}}
-     \caption{A $\log-\log$ plot (base $2$) of the fluctuations in $S^{(d)}_{GHY}$ for a flat ($K=0$) surface in $d=2,3$ and $4$-dimensional Minkowski space.
+%\begin{figure}[t!]
+ % \centering
+  %  {\includegraphics[scale=0.6]{GHY-fluctuation_plots}}
+   %  \caption{A $\log-\log$ plot (base $2$) of the fluctuations in $S^{(d)}_{GHY}$ for a flat ($K=0$) surface in $d=2,3$ and $4$-dimensional Minkowski space.
 Each data point represents the sample standard deviation in $S^{(d)}_{GHY}$ on a sample of $R=100$ runs.
@@
 The corresponding black, blue and red lines have gradients $\frac12$, $0$ and $-\frac18$ and best-fit intercepts of order $1$.}
-     \label{fig:fluctuations}
-\end{figure}
+    % \label{fig:fluctuations}
+%\end{figure}
 
@@
 Take a $d$-cube $[0,L]^d$ in $d$-dimensional Minkowski space with metric $ds^2=-dt^2+d{\mathbf x}^2$ and define the hypersurface $\Sigma: t=L/2$, which partitions the cube into two equal halves.
-Sprinkle at density $\rho=l^{-d}$ into the cube, which in any given run will place $N$ points inside the cube where $N$ is a Poisson random number with mean $\langle N\rangle = \rho V=  (L/l)^d$.
-Evaluate $S^{(d)}_{GHY}=\rho^\frac{2-d}{d}c_d\left(2\Gamma\left(\frac{2}{d}\right)\right)^{-1}(N_{max} - N_{min})$ for $\Sigma$ by counting the minimal/maximal elements in the upper/lower half of the cube, and repeat for $R$ runs.
+Sprinkle at density $\rho=l^{-d}$ into the cube, which will place $\mathbf N$ points inside the cube, where $\mathbf N$ is a Poisson random number with mean $\left< \mathbf N\right> = \rho V=  (L/l)^d$.
+We evaluate the random variable $\mathbf S^{(d)}_{0}=\rho^\frac{2-d}{d}c_d\left(2\Gamma\left(\frac{2}{d}\right)\right)^{-1}(\mathbf N^{(0)}_{\vphantom{i}max} - \mathbf N_{min}^{(0)})$ by counting the minimal/maximal elements in the upper/lower half of the cube, and repeat for $R$ runs.
 The expected value for the mean is zero in all cases, since the extrinsic curvature of the surface is zero.
@@
 
-We have performed analogous simulations for the second formula in~\eqref{GH_boundary_to_causet}, involving $N^{(1)}_{max}$ and $N^{(0)}_{max}$.
+For completeness we have performed analogous simulations for $\mathbf S^{(d)}_-$, which involves counting $\mathbf N^{(1)}_{max}$ and $\mathbf N^{(0)}_{max}$.
 In this case, heuristic arguments of the type given above are hampered by the fact that the two random variables whose difference is taken are not independent.
-However, the results for the scaling of the fluctuations turn out to be identical to those obtained in the first case.
+However, the results for the scaling of the fluctuations turn out to be identical to those obtained for $\mathbf S^{(d)}_0$.
+
+\section{Timelike Boundaries}
+
+\section{A Candidate Boundary Term for an Alexandrov Interval}
+\newcommand{\vol}{\mathrm{vol}}
+
+Until now our approach has been to reconstruct the  continuum GHY term  from the information in a causal set. 
+The candidate described in the previous sections for a spatial boundary has a natural expression in a causal set  in  terms of its maximal and minimal elements which are the  analogs of a  space-like boundary.
+A suitable analog  of  a time-like boundary  term has been harder to  construct; there is no simple causal set  characterisation of such  a boundary.
+Indeed, it is difficult to give a causal set  characterisation of a  spacetime region bounded by time-like surfaces.
+On the other hand, this is not true of causally defined spacetime regions since  past and future sets are defined naturally in  a causal set.
+An important example of such a region is the  Alexandrov interval which is used extensively  in calculations of causal set kinematics.
+In the continuum such a region  is bounded by null hypersurfaces and their codimension 2 joints.
+
+Conversely in the continuum,   while the GHY term  is well defined for both a space-like and time-like boundary  (including their co-dimension 2 joints)  it is an open question what the contribution from a null  boundary  is \cite{Gibbons_Hawking_Boundary, Harris}.
+Apart from some sketchy suggestions in the literature \cite{neiman} there is no rigourous construction of such a term; indeed the standard variational route for finding the boundary contribution to the action appears to fail for null boundaries.
+
+Given our ignorance in this matter,  we can ask if causal set theory  might  in fact guide us in the right direction.
+Our starting point is the $d$-dimensional  Benincasa-Dowker Action \cite{Benincasa_Dowker:The_Scalar_Curvature_of_a_Causal_Set, Dowker_Glaser:dAlembertians_for_Causal_Sets} which is constructed from the mean number  $N_i^{(d)}${\footnote{ $N_i^{(d)}$ is dimension  and geometry dependent, but for our purpose it suffices to retain the dimension label.
+} of $i+1$-element inclusive intervals in the causal set
+\begin{equation}   
+8 \pi G {S_{BD}^{(d)}}[\mathcal C] = -\alpha_d \rho^{\frac{2-d}{d}}\biggl( N[\mathcal C]+ \frac{\beta_d}{ \alpha_d} \sum_{i=1}^{n_d-1}  C^{(d)}_{i} N _i^{(d)}[\mathcal C] \biggr) 
+\label{bd} 
+\end{equation}  
+with   
+\begin{equation}
+\begin{aligned}
+ \alpha_d =   
+\begin{cases}
+-\frac{1}{\Gamma(\frac{2+d}{d})}c_d ^{2/d}  \quad   &d \, \, \mathrm{odd} \\
+-2 \frac{1}{\Gamma(\frac{2+d}{d})}c_d ^{2/d}  \quad   &d \, \,  \mathrm{even} \\
+\end{cases}
+\qquad\text{and}\qquad \beta_d = 
+\begin{cases} 
+\frac{d+1}{2^{d-1}\Gamma(2/d +1)} c_d^{2/d}   \quad   &d\mathrm{ \, \, odd}\\ 
+\frac{\Gamma(d/2+2)\Gamma(d/2+1)}{\Gamma(2/d)\Gamma(d)}c_d^{2/d}  &  d\mathrm{ \, \,  even}\\ 
+\end{cases} 
+\end{aligned}
+\end{equation}
+and
+\begin{equation} 
+n_d = 
+\begin{cases} 
+\frac{d-1}{2} + 2  \quad & d\mathrm{ \, \, odd}\\ 
+\frac{d}{2} + 2  \quad &d\mathrm{ \, \,  even,}\\ 
+\end{cases} 
+\end{equation} 
+where  $2^{-d/2} c_d$ is the volume of a unit Alexandrov interval $I(x,y)$, i.e., one for which the proper time $\tau(x,y)=1$. 
+%%{ {\bf \color{red} this is not correct! $c_d= \vol(S^{d-2}) \frac{ 2^{\frac{2-d}{2}}}{d(d-1)}$ and $\vol(S^{d-2})$ the volume of the homogeneous $d-2$ sphere}}.  
+The coefficients $C_i^{(d)}$ of the terms $N_i^{(d)}$ in the sum are
+\begin{equation}
+\label{cid}
+ C_i^{(d)}= 
+\begin{cases} 
+\sum_{k=0}^{i-1} (-1)^k\binom{i-1}{k} \frac{\Gamma(d/2(k+1)+3/2)}{\Gamma(d/2+3/2) \Gamma(1+dk/2)}  \quad  &\mathrm{d \, \, odd}\\ 
+\sum_{k=0}^{i-1} (-1)^k\binom{i-1}{k} \frac{\Gamma(d/2(k+1)+2)}{\Gamma(d/2+2) \Gamma(1+dk/2)}  & \mathrm{d \, \,  even.}\\ 
+\end{cases} 
+\end{equation} 
+These coefficients can be expressed more compactly as hypergeometric functions 
+\begin{equation}
+C_i^d={}_{q+1}F_{q}(\{ a_1, \ldots a_q, i-1\}, \{b_1, \ldots b_q \} |1)
+\label{chyp} 
+\end{equation} 
+with $q=d/2+1/2, a_i=\frac{d+2i}{d}, b_i=2i/d$ for $d$ odd and $q=d/2, a_i=\frac{d+2i+2}{d}, b_i=2i/d$  for $d$ even.
+
+If $S_{BD}^{(d)}$ were precisely the discrete analog of the Einstein Hilbert action {\it without} any boundary contributions, then in flat spacetime  $S_{BD}^{(d)}$ should  vanish.
+However, as shown explicitly in \cite{bbdtwo} this is not the case for Alexandrov intervals in $d=2$  flat spacetime.
+Indeed, there is a residual contribution that goes to a non-vanishing constant in the continuum limit $N\rightarrow \infty$.
+When the action is evaluated in rectangular regions in the same spacetime or for Alexandrov intervals in  the flat trousers spacetime, this is no longer the case.
+Thus, it seems plausible that the $d=2$ BD action contains boundary information.
+Moreover, in the case of Alexandrov intervals the asymptotic value of the action is the same  irrespective of the interval size.
+While this might suggest a  topological origin, we will now show that it is a part of a more general  result for $d>2$ and  has a geometrical origin.
+Namely it corresponds to the volume  $\vol(S^{d-2})$ of the $S^{d-2}$ joint which is independent of the interval size {\it only} in $d=2$.
+
+Consider an Alexandrov interval $I(p,q)=I^+(p) \cap I^-(q)$ for two events $p,q$ in a causal spacetime.
+While the  boundary of $I(p,q)$ may be quite complicated in general, some of these components must necessarily be null.
+In flat spacetime, for example,  the boundary of $I(p,q)$ consists of the two events $p$ and $q$, the subsets $B^\pm$ of the  two null hypersurfaces $\partial J^+(p)$ and $\partial J^-(q)$, and their intersection or ``joint''  $S^{d-2}=\partial J^+(p) \cap \partial J^-(q)$ which is a codimension 2 sphere.
+Evaluating the BD action in an Alexandrov interval in flat spacetime should therefore yield  only boundary dependent terms.
+In flat spacetime the $S^{d-2}$ joint  has radius $\tau(p,q)/2$, and thus increases with the size of the Alexandrov interval ${\mathrm{vol}}(I(p,q)) = N/\rho$, where $N$ is the mean of the number of causal set elements sprinkled into $I(p,q)$ .
+In what follows we take the continuum limit $\rho, N \rightarrow \infty$ while keeping ${\mathrm{vol}}(I(p,q))$ fixed.
+\ss{It would be nice to have a fancy picture of an Alex.
+interval here..}
+%% If $p$ or $q$ were to be taken to time-like past or future infinity, this volume would go to infinity as expected.
+
+In \cite{Glaser_Sumati:Locality_in_Causal_Set} a closed form expression was obtained for the mean value of the number of $i+1$ element inclusive intervals $N_i^d$ contained in an Alexandrov interval in $d$ dimensional flat spacetime
+\begin{equation} N_i^d= \frac{\Gamma(d)^2\, \, N^{i+2}}{\Gamma(i)} \sum_{k=0}^\infty \frac{(-N)^k \, \, \, \Gamma(\frac{d(k+i)}{2}+1) \, \, \,  \Gamma(\frac{d(k+i+1)}{2}+1)}{(k+i+2) \, (k+i+1)\, \Gamma(k+1)\, \Gamma(\frac{d(k+i)}{2} +d) \,\Gamma(\frac{d(k+i+1)}{2}+d)},
+\end{equation} 
+where $i \geq 1$.
+We use this to expand the sum in (\ref{bd}) as a power series in $N$: $\sum_{i=1}^{n_d} C_i^d \, \, N_i^{(d)} = \sum_{j=1}^\infty A_j^d \, \, N^{j+1}$.
+After a rearrangement and redefinition of indices we find that
+\begin{equation}
+ A_j^d = \Gamma(d)^2\biggl( \sum_{i=1}^{D+2} (-1)^i \binom{j-1}{i-1}  C_i^d  \biggr) \frac{(-1)^j}{(j+1)!}\frac{\Gamma(\frac{d}{2}(j-1)+1)\Gamma(\frac{d}{2}j+1)}{\Gamma(\frac{d}{2}(j-1)+d) \Gamma(\frac{d}{2}j+d)},
+\end{equation} 
+where $d=2D$ for $d$ even and $d=2D+1$ for $d$ odd.   
+% Using the abbreviated form $C_i^d=\sum_{l=0}^{i-1} \binom{i-1}{l}A(d,l)$ where $A(d,l)$ is defined using \ref{cid} 
+% we find that 
+% \begin{equation}
+% \sum_{i=1}^{D+2} (-1)^i \binom{j-1}{i-1}  C_i^d = (-1)^{j-1} A(d,j-1),  
+% \end{equation}      
+% where we have used the binomial identities  BLAH. Thus 
+% \begin{equation}
+%  \sum_{i=1}^{n_d} C_i^d N_i^{(d)}= ((d-1)!)^2 \sum_{j=1}^\infty(-1)^j B(d,j) \frac{\Gamma(\frac{d}{2}j+1)}{\Gamma(j+2)\Gamma(\frac{d}{2}(j-1)+d)\Gamma(\frac{d}{2}j+d)}N^{j+1}
+% \label{thesum} 
+% \end{equation} 
+% where 
+% \begin{equation} 
+% B(d,j) = \left \{ 
+% \begin{array}{ccc} 
+% \frac{\Gamma(\frac{d}{2}j+3/2)}{\Gamma(\frac{d}{2}+3/2)}  &\quad & d \, \, odd, \\  \frac{\Gamma(\frac{d}{2}j+2)}{\Gamma(\frac{d}{2}+2)}  &\quad & d \, \, even.  \\ 
+% \end{array} 
+% \right
+% \end{equation} 
+This can be simplified to 
+\begin{equation}
+A_j^d= \gamma_j^d \, \, \Gamma(d)^2 (-1)^{j+1} \frac{1}{\Gamma(\frac{d}{2} (j + 1)) \Gamma(\frac{d}{2} (2 + j)) \Gamma(2 + j) }
+\label{simplercoefft} 
+\end{equation} 
+where 
+\begin{equation} \gamma_j^d= 
+\begin{cases} 
+\frac{\sqrt{\pi}}{2^{1+dj} }\frac{\Gamma(2+dj)}{\Gamma(\frac{d-1}{2})}    \quad   \mathrm{d} \, \, \mathrm{odd} \\
+\frac{\Gamma(1+\frac{d}{2}j)  \Gamma(2+\frac{d}{2}j) }{\Gamma(\frac{d}{2})}  \quad   \mathrm{d \, \,  even} \\
+\end{cases} 
+\end{equation} 
+A further simplification then allows us to include the first term in the action (\ref{bd}) in the series expansion so that 
+\begin{equation} 
+8 \pi G S_{BD}^d(C) = -\alpha_d \rho^{\frac{2-d}{d}} \sum_{j=0}^\infty (-1)^j Q_j^d N^{j+1} 
+\end{equation}
+where 
+\begin{eqnarray} 
+Q_j^d & =& {\widetilde \gamma}_j^d \, \, \frac{\Gamma(d)\Gamma(\frac{d}{2})}{\Gamma(\frac{d}{2}(1+j)) \Gamma(\frac{d}{2}(2+j)) \Gamma(2+j)} \\ \nonumber 
+{\widetilde \gamma}_j^d &=& 
+\begin{cases} 
+2^{-dj}    \quad   \mathrm{d} \, \, \mathrm{odd} \\
+\Gamma(1+\frac{d}{2}j) \Gamma(2+\frac{d}{2}j)   \quad   \mathrm{d \, \,  even} \\
+\end{cases} 
+\label{fullcoefft} 
+\end{eqnarray} 
+Taking our cues from the behaviour of $8 \pi G S^d(C)$ for $d=2,3$ in the  $N \rightarrow \infty$ limit,  we will consider the ratio
+\begin{equation} 
+\frac{ 8 \pi G S^d(C)}{\vol(S^{d-2})} = \frac{b}{d(d-1) \Gamma(1+2/d) N^{\frac{d-2}{d}}} \left( N + \frac{\beta_d}{\alpha_d} \sum_{i=1}^{n_d} C_i^d N_i^d \right)  
+\label{ratio} 
+\end{equation} 
+where $b=1$ for $d$ odd and $b=2$ for $d$ even.
+Using this simplification an explicit calculation in Mathematica\footnote{The above simplifications greatly aid the Mathematica calculation and allow us to extend it to higher $d$.}  for $d=2, \ldots 16$ gives
+\begin{equation} 
+\lim_{N \rightarrow \infty} \frac{ 8 \pi G S_{BD}^d(C)}{\vol(S^{d-2})} = 1,  
+\label{result} 
+\end{equation}
+
+Since each $N_i^d$  can be expressed as a hypergeometric function \cite{Glaser_Sumati:Locality_in_Causal_Set} it  might seem possible to evaluate (\ref{ratio}) analytically term by term using a large $N$ expansion.
+In \cite{Glaser_Sumati:Locality_in_Causal_Set} the asymptotic form of $N_i^d$ was explicitly calculated, and for $d>2$, the leading order was shown to  $\sim  N^{2-\frac{2}{d}}$.
+This provides us a good consistency check for our result (\ref{result}) since this would mean that the leading order contributions from the $N_i^d$ cancel in the action.
+This is indeed explicitly seen to be the case, again for $d=3, \ldots 16$.
+Similarly for $d=2$ where the leading order contribution is $\sim  N \log N$.
+Indeed, in all cases the next to leading order terms given in \cite{Glaser_Sumati:Locality_in_Causal_Set} also do not have the requisite $ \sim N^{\frac{d-2}{d}}$ dependence required by (\ref{result}); in this case, the explicit coefficients have not been calculated  to be able   check for  a vanishing contribution to (\ref{result}).
+This suggests that  we cannot rely   on the the asymptotic form of the individual terms  in the sum, and would need to look elsewhere to find an analytic proof for general $d$.
+While it may be possible to use the simplification (\ref{fullcoefft})  to obtain a closed form expression for the action for general $d$, it suffices for our purposes that (\ref{result}) has been explicitly checked in $2\leq d \leq 16$ dimensions.
+
+The result we have obtained is for flat spacetime.
+One may speculate that the presence of curvature will not modify the essence of this result,  i.e., that the BD action for an Alexandrov interval  will include boundary contributions from the  $S^{d-2}$ at the joint, at least  when the boundary components  are of the same type as in flat spacetime.
+What will differ of course is the fact that the principal radii of the $S^{d-2}$ will in general vary from point to point, hence complicating the relationship between the $d$-volume of the interval and the $(d-2)$-volume of $S^{d-2}$.
+In  \cite{Khetrapal_Sumati:Causal_Diamond_Volume} it was shown that in an RNC up to first order corrections the intrinsic boundary geometry of the Alexandrov interval is still flat.
+This would allow a simpler relation between $\vol(S^{d-2})$ and $N$, making it possible to do an RNC version of this  calculation.
+
+The fact that the BD action does not vanish in flat spacetime suggests  that this action  does not only contain  information about the bulk Einstein Hilbert action but also extra boundary information.
+One might be tempted to speculate further that this action contains {\it all}  possible boundary or GHY terms suggesting that the null boundary GHY contribution for an Alexandrov interval in any spacetime always vanishes.
+Moreover, it is only the spatial $S^{d-2}$ joint  which gives  a non-vanishing contribution $\vol(S^{d-2})$.
+Interestingly, this is precisely what \cite{neiman} claims is the GHY term for an Alexandrov interval.
+
 \newpage

### 2014-09-18 test2.tex
initial commit

--- 2014-09-18 test.tex
@@
 \ee
-and where $V_d=\pi^{\frac{d}{2}}/\Gamma\left(\frac{d}{2}+1\right)$ denotes the volume of the unit $d$-ball ($\Gamma(t)$ is a gamma function, defined as $\Gamma(t):=\int_0^\infty dx\: x^{t-1}\: e^{-x}$).
+and where $V_d=\pi^{\frac{d}{2}}/\Gamma\left(\frac{d}{2}+1\right)$ denotes the volume of the unit $d$-ball ($XXt)$ is a gamma function, defined as $\Gamma(t):=\int_0^\infty dx\: x^{t-1}\: e^{-x}$).
 We have used $\mathbf{p}$ and $\mathbf{q}$ to stand for infinite dimensional vectors $(p_0,...,p_m,..)$ and $(q_0,...,q_n,..)$ respectively, where all the $p_m,\: q_n \in \mathbb{R}$.
-These vectors contain a finite number of non-zero entries, corresponding to the coefficients of the different $N_{{max\vphantom{i}}}^{\:(m)}$ and $N_{{min}}^{\:(n)}$ terms.
+These vectors contain a finite number of non-zero entries, corresponding to the coefficients of the different $N_{{max\vphantom{i}}}^{\:(m)}$ and $N_{{min}}^{\:(n}$ terms.
 Their values can be chosen arbitrarily, so long as they satisfy the following conditions.
@@
 \begin{cases}
--\frac{1}{\Gamma(\frac{2+d}{d})}c_d ^{2/d}  \quad   &d \, \, \mathrm{odd} \\
--2 \frac{1}{\Gamma(\frac{2+d}{d})}c_d ^{2/d}  \quad   &d \, \,  \mathrm{even} \\
+-\frac{1}{XX\frac{2+d}{d}}c_d ^{2/d}  \quad   &d \, \, \mathrm{odd} \\
+-2 \frac{1}{XX\frac{2+d}{d}}c_d ^{2/d}  \quad   &d \, \,  \mathrm{even} \\
 \end{cases}
@@
 \begin{cases} 
-\frac{d+1}{2^{d-1}\Gamma(2/d +1)} c_d^{2/d}   \quad   &d\mathrm{ \, \, odd}\\ 
-\frac{\Gamma(d/2+2)\Gamma(d/2+1)}{\Gamma(2/d)\Gamma(d)}c_d^{2/d}  &  d\mathrm{ \, \,  even}\\ 
+\frac{d+1}{2^{d-1}XX2/d +1} c_d^{2/d}   \quad   &d\mathrm{ \, \, odd}\\ 
+\frac{XXd/2+2)\Gamma(d/2+1)}{\Gamma(2/d)\Gamma(d}c_d^{2/d}  &  d\mathrm{ \, \,  even}\\ 
 \end{cases} 
@@
 \begin{cases} 
-\sum_{k=0}^{i-1} (-1)^k\binom{i-1}{k} \frac{\Gamma(d/2(k+1)+3/2)}{\Gamma(d/2+3/2) \Gamma(1+dk/2)}  \quad  &\mathrm{d \, \, odd}\\ 
-\sum_{k=0}^{i-1} (-1)^k\binom{i-1}{k} \frac{\Gamma(d/2(k+1)+2)}{\Gamma(d/2+2) \Gamma(1+dk/2)}  & \mathrm{d \, \,  even.}\\ 
+\sum_{k=0}^{i-1} (-1)^k\binom{i-1}{k} \frac{XXd/2(k+1)+3/2)}{\Gamma(d/2+3/2) \Gamma(1+dk/2}  \quad  &\mathrm{d \, \, odd}\\ 
+\sum_{k=0}^{i-1} (-1)^k\binom{i-1}{k} \frac{XXd/2(k+1)+2)}{\Gamma(d/2+2) \Gamma(1+dk/2}  & \mathrm{d \, \,  even.}\\ 
 \end{cases} 
@@
 In \cite{Glaser_Sumati:Locality_in_Causal_Set} a closed form expression was obtained for the mean value of the number of $i+1$ element inclusive intervals $N_i^d$ contained in an Alexandrov interval in $d$ dimensional flat spacetime
-\begin{equation} N_i^d= \frac{\Gamma(d)^2\, \, N^{i+2}}{\Gamma(i)} \sum_{k=0}^\infty \frac{(-N)^k \, \, \, \Gamma(\frac{d(k+i)}{2}+1) \, \, \,  \Gamma(\frac{d(k+i+1)}{2}+1)}{(k+i+2) \, (k+i+1)\, \Gamma(k+1)\, \Gamma(\frac{d(k+i)}{2} +d) \,\Gamma(\frac{d(k+i+1)}{2}+d)},
+\begin{equation} N_i^d= \frac{XXd)^2\, \, N^{i+2}}{\Gamma(i)} \sum_{k=0}^\infty \frac{(-N)^k \, \, \, \Gamma(\frac{d(k+i)}{2}+1) \, \, \,  \Gamma(\frac{d(k+i+1)}{2}+1)}{(k+i+2) \, (k+i+1)\, \Gamma(k+1)\, \Gamma(\frac{d(k+i)}{2} +d) \,\Gamma(\frac{d(k+i+1)}{2}+d},
 \end{equation} 
@@
 \begin{equation}
- A_j^d = \Gamma(d)^2\biggl( \sum_{i=1}^{D+2} (-1)^i \binom{j-1}{i-1}  C_i^d  \biggr) \frac{(-1)^j}{(j+1)!}\frac{\Gamma(\frac{d}{2}(j-1)+1)\Gamma(\frac{d}{2}j+1)}{\Gamma(\frac{d}{2}(j-1)+d) \Gamma(\frac{d}{2}j+d)},
+ A_j^d = XXd)^2\biggl( \sum_{i=1}^{D+2} (-1)^i \binom{j-1}{i-1}  C_i^d  \biggr) \frac{(-1)^j}{(j+1)!}\frac{\Gamma(\frac{d}{2}(j-1)+1)\Gamma(\frac{d}{2}j+1)}{\Gamma(\frac{d}{2}(j-1)+d) \Gamma(\frac{d}{2}j+d},
 \end{equation} 
@@
 % \begin{equation}
-%  \sum_{i=1}^{n_d} C_i^d N_i^{(d)}= ((d-1)!)^2 \sum_{j=1}^\infty(-1)^j B(d,j) \frac{\Gamma(\frac{d}{2}j+1)}{\Gamma(j+2)\Gamma(\frac{d}{2}(j-1)+d)\Gamma(\frac{d}{2}j+d)}N^{j+1}
+%  \sum_{i=1}^{n_d} C_i^d N_i^{(d)}= ((d-1)!)^2 \sum_{j=1}^\infty(-1)^j B(d,j) \frac{XX\frac{d}{2}j+1)}{\Gamma(j+2)\Gamma(\frac{d}{2}(j-1)+d)\Gamma(\frac{d}{2}j+d}N^{j+1}
 % \label{thesum} 
@@
 % \begin{array}{ccc} 
-% \frac{\Gamma(\frac{d}{2}j+3/2)}{\Gamma(\frac{d}{2}+3/2)}  &\quad & d \, \, odd, \\  \frac{\Gamma(\frac{d}{2}j+2)}{\Gamma(\frac{d}{2}+2)}  &\quad & d \, \, even.  \\ 
+% \frac{XX\frac{d}{2}j+3/2)}{\Gamma(\frac{d}{2}+3/2)}  &\quad & d \, \, odd, \\  \frac{\Gamma(\frac{d}{2}j+2)}{\Gamma(\frac{d}{2}+2}  &\quad & d \, \, even.  \\ 
 % \end{array} 
@@
 \begin{equation}
-A_j^d= \gamma_j^d \, \, \Gamma(d)^2 (-1)^{j+1} \frac{1}{\Gamma(\frac{d}{2} (j + 1)) \Gamma(\frac{d}{2} (2 + j)) \Gamma(2 + j) }
+A_j^d= \gamma_j^d \, \, XXd)^2 (-1)^{j+1} \frac{1}{\Gamma(\frac{d}{2} (j + 1)) \Gamma(\frac{d}{2} (2 + j)) \Gamma(2 + j }
 \label{simplercoefft} 
@@
 \begin{cases} 
-\frac{\sqrt{\pi}}{2^{1+dj} }\frac{\Gamma(2+dj)}{\Gamma(\frac{d-1}{2})}    \quad   \mathrm{d} \, \, \mathrm{odd} \\
-\frac{\Gamma(1+\frac{d}{2}j)  \Gamma(2+\frac{d}{2}j) }{\Gamma(\frac{d}{2})}  \quad   \mathrm{d \, \,  even} \\
+\frac{\sqrt{\pi}}{2^{1+dj} }\frac{XX2+dj)}{\Gamma(\frac{d-1}{2}}    \quad   \mathrm{d} \, \, \mathrm{odd} \\
+\frac{XX1+\frac{d}{2}j)  \Gamma(2+\frac{d}{2}j) }{\Gamma(\frac{d}{2}}  \quad   \mathrm{d \, \,  even} \\
 \end{cases} 
@@
 \begin{eqnarray} 
-Q_j^d & =& {\widetilde \gamma}_j^d \, \, \frac{\Gamma(d)\Gamma(\frac{d}{2})}{\Gamma(\frac{d}{2}(1+j)) \Gamma(\frac{d}{2}(2+j)) \Gamma(2+j)} \\ \nonumber 
+Q_j^d & =& {\widetilde \gamma}_j^d \, \, \frac{XXd)\Gamma(\frac{d}{2})}{\Gamma(\frac{d}{2}(1+j)) \Gamma(\frac{d}{2}(2+j)) \Gamma(2+j} \\ \nonumber 
 {\widetilde \gamma}_j^d &=& 
@@
 2^{-dj}    \quad   \mathrm{d} \, \, \mathrm{odd} \\
-\Gamma(1+\frac{d}{2}j) \Gamma(2+\frac{d}{2}j)   \quad   \mathrm{d \, \,  even} \\
+XX1+\frac{d}{2}j) \Gamma(2+\frac{d}{2}j   \quad   \mathrm{d \, \,  even} \\
 \end{cases} 
@@
 \begin{equation} 
-\frac{ 8 \pi G S^d(C)}{\vol(S^{d-2})} = \frac{b}{d(d-1) \Gamma(1+2/d) N^{\frac{d-2}{d}}} \left( N + \frac{\beta_d}{\alpha_d} \sum_{i=1}^{n_d} C_i^d N_i^d \right)  
+\frac{ 8 \pi G S^d(C)}{\vol(S^{d-2})} = \frac{b}{d(d-1) XX1+2/d) N^{\frac{d-2}{d}}} \left( N + \frac{\beta_d}{\alpha_d} \sum_{i=1}^{n_d} C_i^d N_i^d \right  
 \label{ratio} 

### 2014-09-18 test6.tex
initial commit

--- 2014-09-18 test.tex
@@
 \ee
-and where $V_d=\pi^{\frac{d}{2}}/\Gamma\left(\frac{d}{2}+1\right)$ denotes the volume of the unit $d$-ball ($\Gamma(t)$ is a gamma function, defined as $\Gamma(t):=\int_0^\infty dx\: x^{t-1}\: e^{-x}$).
+and where $V_d=\pi^{\frac{d}{2}}/\Gamma\left(\frac{d}{2}+1\right)$ denotes the volume of the unit $d$-ball ($\Gamma\left(t)$ is a gamma function, defined as $\Gamma(t):=\int_0^\infty dx\: x^{t-1}\: e^{-x}$).
 We have used $\mathbf{p}$ and $\mathbf{q}$ to stand for infinite dimensional vectors $(p_0,...,p_m,..)$ and $(q_0,...,q_n,..)$ respectively, where all the $p_m,\: q_n \in \mathbb{R}$.
-These vectors contain a finite number of non-zero entries, corresponding to the coefficients of the different $N_{{max\vphantom{i}}}^{\:(m)}$ and $N_{{min}}^{\:(n)}$ terms.
+These vectors contain a finite number of non-zero entries, corresponding to the coefficients of the different $N_{{max\vphantom{i}}}^{\:(m)}$ and $N_{{min}}^{\:(n\right)}$ terms.
 Their values can be chosen arbitrarily, so long as they satisfy the following conditions.
@@
 \begin{cases}
--\frac{1}{\Gamma(\frac{2+d}{d})}c_d ^{2/d}  \quad   &d \, \, \mathrm{odd} \\
--2 \frac{1}{\Gamma(\frac{2+d}{d})}c_d ^{2/d}  \quad   &d \, \,  \mathrm{even} \\
+-\frac{1}{\Gamma\left(\frac{2+d}{d}\right)}c_d ^{2/d}  \quad   &d \, \, \mathrm{odd} \\
+-2 \frac{1}{\Gamma\left(\frac{2+d}{d}\right)}c_d ^{2/d}  \quad   &d \, \,  \mathrm{even} \\
 \end{cases}
@@
 \begin{cases} 
-\frac{d+1}{2^{d-1}\Gamma(2/d +1)} c_d^{2/d}   \quad   &d\mathrm{ \, \, odd}\\ 
-\frac{\Gamma(d/2+2)\Gamma(d/2+1)}{\Gamma(2/d)\Gamma(d)}c_d^{2/d}  &  d\mathrm{ \, \,  even}\\ 
+\frac{d+1}{2^{d-1}\Gamma\left(2/d +1\right)} c_d^{2/d}   \quad   &d\mathrm{ \, \, odd}\\ 
+\frac{\Gamma\left(d/2+2)\Gamma(d/2+1)}{\Gamma(2/d)\Gamma(d\right)}c_d^{2/d}  &  d\mathrm{ \, \,  even}\\ 
 \end{cases} 
@@
 \begin{cases} 
-\sum_{k=0}^{i-1} (-1)^k\binom{i-1}{k} \frac{\Gamma(d/2(k+1)+3/2)}{\Gamma(d/2+3/2) \Gamma(1+dk/2)}  \quad  &\mathrm{d \, \, odd}\\ 
-\sum_{k=0}^{i-1} (-1)^k\binom{i-1}{k} \frac{\Gamma(d/2(k+1)+2)}{\Gamma(d/2+2) \Gamma(1+dk/2)}  & \mathrm{d \, \,  even.}\\ 
+\sum_{k=0}^{i-1} (-1)^k\binom{i-1}{k} \frac{\Gamma\left(d/2(k+1)+3/2)}{\Gamma(d/2+3/2) \Gamma(1+dk/2\right)}  \quad  &\mathrm{d \, \, odd}\\ 
+\sum_{k=0}^{i-1} (-1)^k\binom{i-1}{k} \frac{\Gamma\left(d/2(k+1)+2)}{\Gamma(d/2+2) \Gamma(1+dk/2\right)}  & \mathrm{d \, \,  even.}\\ 
 \end{cases} 
@@
 In \cite{Glaser_Sumati:Locality_in_Causal_Set} a closed form expression was obtained for the mean value of the number of $i+1$ element inclusive intervals $N_i^d$ contained in an Alexandrov interval in $d$ dimensional flat spacetime
-\begin{equation} N_i^d= \frac{\Gamma(d)^2\, \, N^{i+2}}{\Gamma(i)} \sum_{k=0}^\infty \frac{(-N)^k \, \, \, \Gamma(\frac{d(k+i)}{2}+1) \, \, \,  \Gamma(\frac{d(k+i+1)}{2}+1)}{(k+i+2) \, (k+i+1)\, \Gamma(k+1)\, \Gamma(\frac{d(k+i)}{2} +d) \,\Gamma(\frac{d(k+i+1)}{2}+d)},
+\begin{equation} N_i^d= \frac{\Gamma\left(d)^2\, \, N^{i+2}}{\Gamma(i)} \sum_{k=0}^\infty \frac{(-N)^k \, \, \, \Gamma(\frac{d(k+i)}{2}+1) \, \, \,  \Gamma(\frac{d(k+i+1)}{2}+1)}{(k+i+2) \, (k+i+1)\, \Gamma(k+1)\, \Gamma(\frac{d(k+i)}{2} +d) \,\Gamma(\frac{d(k+i+1)}{2}+d\right)},
 \end{equation} 
@@
 \begin{equation}
- A_j^d = \Gamma(d)^2\biggl( \sum_{i=1}^{D+2} (-1)^i \binom{j-1}{i-1}  C_i^d  \biggr) \frac{(-1)^j}{(j+1)!}\frac{\Gamma(\frac{d}{2}(j-1)+1)\Gamma(\frac{d}{2}j+1)}{\Gamma(\frac{d}{2}(j-1)+d) \Gamma(\frac{d}{2}j+d)},
+ A_j^d = \Gamma\left(d)^2\biggl( \sum_{i=1}^{D+2} (-1)^i \binom{j-1}{i-1}  C_i^d  \biggr) \frac{(-1)^j}{(j+1)!}\frac{\Gamma(\frac{d}{2}(j-1)+1)\Gamma(\frac{d}{2}j+1)}{\Gamma(\frac{d}{2}(j-1)+d) \Gamma(\frac{d}{2}j+d\right)},
 \end{equation} 
@@
 % \begin{equation}
-%  \sum_{i=1}^{n_d} C_i^d N_i^{(d)}= ((d-1)!)^2 \sum_{j=1}^\infty(-1)^j B(d,j) \frac{\Gamma(\frac{d}{2}j+1)}{\Gamma(j+2)\Gamma(\frac{d}{2}(j-1)+d)\Gamma(\frac{d}{2}j+d)}N^{j+1}
+%  \sum_{i=1}^{n_d} C_i^d N_i^{(d)}= ((d-1)!)^2 \sum_{j=1}^\infty(-1)^j B(d,j) \frac{\Gamma\left(\frac{d}{2}j+1)}{\Gamma(j+2)\Gamma(\frac{d}{2}(j-1)+d)\Gamma(\frac{d}{2}j+d\right)}N^{j+1}
 % \label{thesum} 
@@
 % \begin{array}{ccc} 
-% \frac{\Gamma(\frac{d}{2}j+3/2)}{\Gamma(\frac{d}{2}+3/2)}  &\quad & d \, \, odd, \\  \frac{\Gamma(\frac{d}{2}j+2)}{\Gamma(\frac{d}{2}+2)}  &\quad & d \, \, even.  \\ 
+% \frac{\Gamma\left(\frac{d}{2}j+3/2)}{\Gamma(\frac{d}{2}+3/2)}  &\quad & d \, \, odd, \\  \frac{\Gamma(\frac{d}{2}j+2)}{\Gamma(\frac{d}{2}+2\right)}  &\quad & d \, \, even.  \\ 
 % \end{array} 
@@
 \begin{equation}
-A_j^d= \gamma_j^d \, \, \Gamma(d)^2 (-1)^{j+1} \frac{1}{\Gamma(\frac{d}{2} (j + 1)) \Gamma(\frac{d}{2} (2 + j)) \Gamma(2 + j) }
+A_j^d= \gamma_j^d \, \, \Gamma\left(d)^2 (-1)^{j+1} \frac{1}{\Gamma(\frac{d}{2} (j + 1)) \Gamma(\frac{d}{2} (2 + j)) \Gamma(2 + j\right) }
 \label{simplercoefft} 
@@
 \begin{cases} 
-\frac{\sqrt{\pi}}{2^{1+dj} }\frac{\Gamma(2+dj)}{\Gamma(\frac{d-1}{2})}    \quad   \mathrm{d} \, \, \mathrm{odd} \\
-\frac{\Gamma(1+\frac{d}{2}j)  \Gamma(2+\frac{d}{2}j) }{\Gamma(\frac{d}{2})}  \quad   \mathrm{d \, \,  even} \\
+\frac{\sqrt{\pi}}{2^{1+dj} }\frac{\Gamma\left(2+dj)}{\Gamma(\frac{d-1}{2}\right)}    \quad   \mathrm{d} \, \, \mathrm{odd} \\
+\frac{\Gamma\left(1+\frac{d}{2}j)  \Gamma(2+\frac{d}{2}j) }{\Gamma(\frac{d}{2}\right)}  \quad   \mathrm{d \, \,  even} \\
 \end{cases} 
@@
 \begin{eqnarray} 
-Q_j^d & =& {\widetilde \gamma}_j^d \, \, \frac{\Gamma(d)\Gamma(\frac{d}{2})}{\Gamma(\frac{d}{2}(1+j)) \Gamma(\frac{d}{2}(2+j)) \Gamma(2+j)} \\ \nonumber 
+Q_j^d & =& {\widetilde \gamma}_j^d \, \, \frac{\Gamma\left(d)\Gamma(\frac{d}{2})}{\Gamma(\frac{d}{2}(1+j)) \Gamma(\frac{d}{2}(2+j)) \Gamma(2+j\right)} \\ \nonumber 
 {\widetilde \gamma}_j^d &=& 
@@
 2^{-dj}    \quad   \mathrm{d} \, \, \mathrm{odd} \\
-\Gamma(1+\frac{d}{2}j) \Gamma(2+\frac{d}{2}j)   \quad   \mathrm{d \, \,  even} \\
+\Gamma\left(1+\frac{d}{2}j) \Gamma(2+\frac{d}{2}j\right)   \quad   \mathrm{d \, \,  even} \\
 \end{cases} 
@@
 \begin{equation} 
-\frac{ 8 \pi G S^d(C)}{\vol(S^{d-2})} = \frac{b}{d(d-1) \Gamma(1+2/d) N^{\frac{d-2}{d}}} \left( N + \frac{\beta_d}{\alpha_d} \sum_{i=1}^{n_d} C_i^d N_i^d \right)  
+\frac{ 8 \pi G S^d(C)}{\vol(S^{d-2})} = \frac{b}{d(d-1) \Gamma\left(1+2/d) N^{\frac{d-2}{d}}} \left( N + \frac{\beta_d}{\alpha_d} \sum_{i=1}^{n_d} C_i^d N_i^d \right\right)  
 \label{ratio} 

### 2014-09-29 temp.tex
changed notation for Nmaxk/Nmink to Fk/Pk

--- 2014-09-18 paper_v7.tex
@@
 \usepackage{jheppub}
-\usepackage{marginnote,xparse,changepage,caption,subfigure}
+\usepackage{marginnote,xparse,changepage,caption,graphicx,subcaption}
 
@@
 \newcommand{\eea}{\end{aligned}}
+\newcommand{\F}[1]{F_{#1}}
+\renewcommand{\P}[1]{P_{#1}}
+\newcommand{\BF}[1]{\mathbf{F}_{#1}}
+\newcommand{\BP}[1]{\mathbf{P}_{#1}}
 \newcommand{\mb}[1]{\marginnote{\texttt{\small MB:\,#1}}}
@@
 
-Consider a sufficiently well-behaved $d$-dimensional spacetime $ (M,g)$ and a surface $\Sigma$ in $M$, which for now we take to be a Cauchy surface.
-The causal past and future sets $M^\pm=J^\pm (\Sigma)$ of the surface then form a partition of $M$, and $\partial M^\pm = \Sigma$.
+Consider a sufficiently well-behaved $d$-dimensional spacetime $(M,g)$ and a surface $\Sigma$ in $M$, which we shall take to be a Cauchy surface (hence $(M,g)$ is globally hyperbolic).
+The causal past and future sets $M^\pm=J^\pm (\Sigma)$ of the surface then form a partition of $M$, and $\partial M^\pm = \Sigma$ ($J^\pm$ denoting the chronological future/past).
 The Gibbons-Hawking-York boundary term for $M^\pm$ is in this case given by
@@
 Spacetime volume in the causal set is obtained simply by counting the number of elements, and hence the volume gradient intuitively corresponds to the difference in the number of ``immediate neighbours'' to the future and to the past of a hypersurface.
-What does it mean to be immediate neighbours to a surface in a causal set?
-In order to make this notion precise, we need to define the causal set analogue of a hypersurface, as well as an appropriate notion of ``closeness''.
+In order to make this notion precise, we need to define the causal set analogue of a hypersurface, as well as an appropriate notion of ``closeness'' to it.
 This turns out to be most natural for the analogue of what in the continuum would correspond to a Cauchy hypersurface.
 
-In the causet, a (partitioning) hypersurface can be described as a particular partition of the causet.
-\mb{why say this here?}We will only use this definition for causets of finite cardinality.Define a \emph{past} \emph{stem} to be a subset $\mathcal{C}'\subseteq\mathcal{C}$ of the causet that contains its own past\IGNORE{, i.e.
-$(\exists x\in\mathcal{C}' \text{ such that } y\prec x)\implies y\in\mathcal{C}'$}.
-That is, any element in the causet that precedes at least one element in $\mathcal{C}'$ is also in $\mathcal{C}'$.
-A \emph{future stem} is defined analogously.
-Let us now define a \emph{spacelike partition} as one which partitions the causet  $\mathcal{C}$ into two parts, $\mathcal{C}^-$ and $\mathcal{C}^+$\IGNORE{($\mathcal{C}=\mathcal{C}^-\cup\mathcal{C}^+$)}, where $\mathcal{C}^-$ is a past stem and $\mathcal{C}^+$ is its complement, $\mathcal{C}^+=\mathcal{C}\setminus\mathcal{C}^-$.
-One can equivalently pick some future stem $\mathcal{C}^+$ and its complement $\mathcal{C}^- = \mathcal{C}\setminus\mathcal{C}^+$ to define a spacelike partition.
-The reason that we call some partition of this nature ``spacelike" is the following.
-Given a causet that can be faithfully embedded in some manifold, $M$, then given a spacelike hypersurface, $\Sigma$, we can partition the causet into two bits\mb{only true if $\Sigma$ is Cauchy, no?}, denoting the future (past) part as $\mathcal{C}^+$ ($\mathcal{C}^-$), which is made up of those elements that lie in $M^+$ ($M^-$).
-The causal relations between the two parts, that can be found from the causal strucutre of the manifold, immediately satisfy our requirement for the partition to be spacelike.
-We will still refer to a partition as \emph{spacelike} even if the causet is not faithfully embeddable in some manifold.
-For immediate neighbours to the past of the partition we mean elements of $\mathcal{C}^-$ that have no elements in their causal future in $\mathcal{C}^-$ (these elements being called \textit{maximal} elements in $\mathcal{C}^-$).
-We assume an analogous definition for immediate neighbours to the $future$ of the partition, which we shall refer to as minimal elements in $\mathcal{C}^-$.
-We shall see that the intuitive idea for the boundary term as the ``rate of change of spacetime atoms'' across a partition indeed bears out.
-However, one does not need to know the maximal and minimal elements to calculate the boundary term.
+Given a causet that can be faithfully embedded in a manifold $M$ with Cauchy hypersurface $\Sigma$, we can partition the causet into two sets $\mathcal{C}^+$ and $\mathcal{C}^-$, corresponding to the sets of elements that lie in $M^+=J^+(\Sigma)$ and $M^-=J^-(\Sigma)$.
+By virtue of $\Sigma$ being Cauchy, the set $\mathcal C^-$ contains its own past (i.e.
+any element in $\mathcal C$ that precedes at least one element in $\mathcal{C}^-$ is also in $\mathcal{C}^-$).
+Such a set is known as a \emph{past stem}.
+Similarly, $\mathcal C^+$ is a \emph{future stem} given the analogous definition.
+These properties capture the essential features of a partition induced by a Cauchy surface.
+Hence we propose more generally that for any causet $\mathcal C$, a \emph{Cauchy partition} is a partition that splits $\mathcal{C}$ into two parts, $\mathcal{C}^-$ and $\mathcal{C}^+=\mathcal{C}\setminus\mathcal{C}^-$ \IGNORE{($\mathcal{C}=\mathcal{C}^-\cup\mathcal{C}^+$)}, such that $\mathcal{C}^-$ is a past stem (or, equivalently, that $\mathcal{C}^+$ be a future stem and $\mathcal{C}^- = \mathcal{C}\setminus\mathcal{C}^+$).
+A sprinkling into a manifold with Cauchy surface will of course produce a causal set with Cauchy partition, but we shall refer to a partition of a causal set satisfying the above definition as Cauchy even when the causet is not obtained via sprinkling faithfully embeddable into some manifold.
+
+Given a causal set with Cauchy partition, there follows a natural definition for the ``immediate neighbours'' in the past and future sets of the partition.
+The ``immediate neighbours in the past'' can be defined as those elements of $\mathcal{C}^-$ that have no elements in their causal future in $\mathcal{C}^-$.
+Such elements will be referred to as \textit{maximal elements} in $\mathcal{C}^-$.
+We assume an analogous definition for immediate neighbours to the $future$ of the partition, which we shall refer to as \emph{minimal elements} in $\mathcal{C}^+$.
+
+Given these definitions, we will see that the intuitive idea for the boundary term as the ``rate of change of spacetime atoms'' across a partition indeed bears out. \\
+%However, one does not need to know the maximal and minimal elements to calculate the boundary term.
 It can be found from any elements that lie reasonably close to the surface (in the time coordinate on $M$) but are not necessarily maximal or minimal.
@@
 \begin{figure}
-\IGNORE{
         \centering
-        \begin{subfigure}[b]{0.5\textwidth}
+        \begin{subfigure}[b]{0.49\textwidth}
                 \includegraphics[width=\textwidth]{low_density_sprinkling_with_null_and_spacelike}
@@
         \end{subfigure}
-        \begin{subfigure}[b]{0.5\textwidth}
+        \begin{subfigure}[b]{0.49\textwidth}
                 \includegraphics[width=\textwidth]{high_density_sprinkling_with_null_and_spacelike}
@@
         \end{subfigure}
-        }
         \caption{The partition induced by the sprinkling is unchanged after tilting the surface between null and spacelike.
@@
 
-Let us define the functions $N_{\vphantom{i}max}^{ (k)}\left[\mathcal{C} \right]$ and $N_{min}^{ (k)}\left[\mathcal{C}^- \right]$ that, for a given finite causet, $\mathcal{C}$, return the number of $k$\emph{-next-to-maximal}/\emph{minimal} elements respectively.
-A $k$-next-to-maximal (minimal) element is one for which there are $k$ elements to its future (past).
-This means that the $0$-next-to-maximal (minimal) elements are the maximal (minimal) elements.
+For a given causet $\mathcal C$, let us define the function $\F{k}\left[\mathcal{C} \right]$, which counts the number of elements in $\mathcal C$ that have exactly $k$ elements in their future, and the function $\P{k}\left[\mathcal{C}\right]$, which counts the number of elements which have exactly $k$ elements in their past.
+An $F_0$ element is then a \emph{maximal} element, a $P_0$ element is a \emph{minimal} element, an $F_1$ element is (in some sense) next-to-maximal, and so on.\mb{Ian: your plot here?}
 
-Given a causet, $\mathcal{C}$, with a spacelike partition so that $\mathcal C = \mathcal C^+ \cup\, \mathcal C^-$, we propose the following family of functions for calculating the GHY boundary term of this partition:
+Given a causet, $\mathcal{C}$, with a spacelike partition so that $\mathcal C = \mathcal C^+ \cup\, \mathcal C^-$, we propose the following family of functions as the generalised bounary action:
 \be\label{general_boundary_sum}
 S^{ (d)}_{GHY}\left[\mathcal{C}^-,\mathcal{C}^+;\vec{p}, \vec{q} \right]= \left (l/l_p\right)^{d-2} c_{d}
-\left ( \sum_m p_m N_{\vphantom{i}max}^{ (m)}\left[\mathcal{C}^- \right]
-+  \sum_n q_n N_{min}^{ (n)}\left[\mathcal{C}^+ \right]\right)
+\left ( \sum_m p_m \F{m}\left[\mathcal{C}^- \right]
++  \sum_n q_n \P{m}\left[\mathcal{C}^+ \right]\right)
 \ee
@@
 These vectors contain a finite number of non-zero entries, corresponding to the coefficients of the different $N_{{max\vphantom{i}}}^{\: (m)}$ and $N_{{min}}^{\: (n)}$ terms.
-We use the summation $\sum_m$ ($\sum_n$) to denote a sum that runs from $0$ to $m_{max}$ ($n_{max}$), where $m_{max}$ ($n_{max}$) is some integer, the value of which depends on the causet $\mathcal{C}^-$ ($\mathcal{C}^+$), for which all $N_{\vphantom{i}max}^{ (m)}\left[\mathcal{C}^- \right]=0$ ($N_{min}^{ (n)}\left[\mathcal{C}^+ \right]=0$) for any $m>m_{max}$ ($n>n_{max}$).
+We use the summation $\sum_m$ ($\sum_n$) to denote a sum that runs from $0$ to $m_{max}$ ($n_{max}$), where $m_{max}$ ($n_{max}$) is some integer, the value of which depends on the causet $\mathcal{C}^-$ ($\mathcal{C}^+$), for which all $\F{m}\left[\mathcal{C}^- \right]=0$ ($\P{m}\left[\mathcal{C}^+ \right]=0$) for any $m>m_{max}$ ($n>n_{max}$).
 A finite value for $m_{max}$ ($n_{max}$) will always exist for any finite causet, as it will not be large enough or be structured in such a way to have any $m$-next-to-maximal ($n$-next-to-minimal) elements for $m>m_{max}$ ($n>n_{max}$).
@@
 Given a $d$ dimensional, spatially compact Lorentzian manifold, $M$, the family of functions $S^{ (d)}_{GHY}$ (for different $\vec{p}$ and $\vec{q}$) defines a family of random variables in the following way.
-The process of sprinkling points into $M$ with density $\rho=l^{-d}$ generates a random causet $\mathcal{C}$, and so the functions $N_{\vphantom{i}max}^{ (k)}$ and $N_{min}^{ (k)}$ acting on this random causet are now the random variables $\textbf{N}_{\vphantom{i}max}^{ (k)}$ and $\textbf{N}_{min}^{ (k)}$.
-These random variables can be substituted into~\eqref{general_boundary_sum} for $N_{\vphantom{i}max}^{ (k)}$ and $N_{min}^{ (k)}$ to give the family of random variables, $\textbf{S}^{ (d)}_{GHY}$.
-These random variables will be functions of $\rho$ and the manifold, $M^{\pm}$, where the dependence of $\textbf{N}_{\vphantom{i}max}^{ (k)}$ ($\textbf{N}_{min}^{ (k)}$) on the manifold is restricted to $M^-$ ($M^+$).
+The process of sprinkling points into $M$ with density $\rho=l^{-d}$ generates a random causet $\mathcal{C}$, and so the functions $\F{k}$ and $\P{k}$ acting on this random causet are now the random variables $\BF{k}$ and $\BP{k}$.
+These random variables can be substituted into~\eqref{general_boundary_sum} for $\F{k}$ and $\P{k}$ to give the family of random variables, $\textbf{S}^{ (d)}_{GHY}$.
+These random variables will be functions of $\rho$ and the manifold, $M^{\pm}$, where the dependence of $\BF{k}$ ($\BP{k}$) on the manifold is restricted to $M^-$ ($M^+$).
 We claim that in the limit of infinite density the expectation value, in the sprinkling process, of $\textbf{S}^{ (d)}_{GHY}$ (for any vectors $\vec{p}$ and $\vec{q}$ satisfying~\eqref{coefficient_relation1} and ~\eqref{coefficient_relation1}) tends to the continuum GHY boundary term for a surface $\Sigma$ that partitions the manifold into $M^+$ and $M^-$:
@@
 In order to prove~\eqref{eq:mainconjecture} and~\eqref{eq:conjecture_for_area} let us first derive expressions for the mean values of $\textbf{N}_{\vphantom{i}max}^{\: (m)}$ and $\textbf{N}_{min}^{ (n)}$.
-The proof will will follow almost automatically, as~\eqref{general_boundary_sum} and~\eqref{general_area_sum} are just linear combination of $N_{max}^{\: (m)}$ and $N_{min}^{ (n)}$ terms, and thus the mean values of their random variable counterparts will just be linear combinations of $\left\langle \textbf{N}_{\vphantom{i}max}^{\: (m)}\right\rangle$ and $\left\langle \textbf{N}_{min}^{\: (n)}\right\rangle$ terms.
+The proof will will follow almost automatically, as~\eqref{general_boundary_sum} and~\eqref{general_area_sum} are just linear combination of $N_{max}^{\: (m)}$ and $\P{m}$ terms, and thus the mean values of their random variable counterparts will just be linear combinations of $\left\langle \textbf{N}_{\vphantom{i}max}^{\: (m)}\right\rangle$ and $\left\langle \textbf{N}_{min}^{\: (n)}\right\rangle$ terms.
 
@@
 
-In \cite{Glaser_Sumati:Locality_in_Causal_Set} a closed form expression was obtained for the mean value of the number of $i+1$ element inclusive intervals $N_i^d$ contained in an Alexandrov interval in $d$ dimensional flat spacetime
+In \cite{Glaser_Sumati:Locality_in_Causal_Set} a closed form expression was obtained for the mean value of the number of $i+1$ element inclusive intervals $N_i^{(d)}$ contained in an Alexandrov interval in $d$ dimensional flat spacetime
 \begin{equation} N_i^{ (d)}= \frac{\Gamma\left (d\right)^2 N^{i+2}}{\Gamma\left (i\right)} \sum_{k=0}^\infty \frac{ (-N)^k\,\Gamma\left (\frac{d (k+i)}{2}+1\right)  \Gamma\left (\frac{d (k+i+1)}{2}+1\right)}{ (k+i+2)  (k+i+1) \Gamma\left (k+1\right)\Gamma\left (\frac{d (k+i)}{2} +d\right) \Gamma\left (\frac{d (k+i+1)}{2}+d\right)},\nonumber

### 2014-09-29 paper_v7.tex
included some packages in current v7
minor aesthetic edits

@@
 
-Let us define the functions $N_{\vphantom{i}max}^{ (k)}\left[\mathcal{C} \right]$ and $N_{min}^{ (k)}\left[\mathcal{C}^- \right]$ that, for a given finite causet, $\mathcal{C}$, return the number of $k$\emph{-next-to-maximal}/\emph{minimal} elements respectively.
-A $k$-next-to-maximal (minimal) element is one for which there are $k$ elements to its future (past).
-This means that the $0$-next-to-maximal (minimal) elements are the maximal (minimal) elements.
+For a given causet $\mathcal C$, let us define the function $\F{k}\left[\mathcal{C} \right]$, which counts the number of elements in $\mathcal C$ that have exactly $k$ elements in their future, and the function $\P{k}\left[\mathcal{C}\right]$, which counts the number of elements which have exactly $k$ elements in their past.
+An ``$F_0$ element'' is then a \emph{maximal} element, a ``$P_0$ element'' is a \emph{minimal} element, an ``$F_1$ element'' is (in some sense) next-to-maximal, and so on.\mb{Ian: your plot here?}
 
-Given a causet, $\mathcal{C}$, with a spacelike partition so that $\mathcal C = \mathcal C^+ \cup\, \mathcal C^-$, we propose the following family of functions for calculating the GHY boundary term of this partition:
+Given a causet, $\mathcal{C}$, with a spacelike partition so that $\mathcal C = \mathcal C^+ \cup\, \mathcal C^-$, we propose the following family of functions as the generalised bounary action:
 \be\label{general_boundary_sum}
 S^{ (d)}_{GHY}\left[\mathcal{C}^-,\mathcal{C}^+;\vec{p}, \vec{q} \right]= \left (l/l_p\right)^{d-2} c_{d}
-\left ( \sum_m p_m N_{\vphantom{i}max}^{ (m)}\left[\mathcal{C}^- \right]
-+  \sum_n q_n N_{min}^{ (n)}\left[\mathcal{C}^+ \right]\right)
+\left ( \sum_m p_m \F{m}\left[\mathcal{C}^- \right]
++  \sum_n q_n \P{n}\left[\mathcal{C}^+ \right]\right)
 \ee
@@
 These vectors contain a finite number of non-zero entries, corresponding to the coefficients of the different $N_{{max\vphantom{i}}}^{\: (m)}$ and $N_{{min}}^{\: (n)}$ terms.
-We use the summation $\sum_m$ ($\sum_n$) to denote a sum that runs from $0$ to $m_{max}$ ($n_{max}$), where $m_{max}$ ($n_{max}$) is some integer, the value of which depends on the causet $\mathcal{C}^-$ ($\mathcal{C}^+$), for which all $N_{\vphantom{i}max}^{ (m)}\left[\mathcal{C}^- \right]=0$ ($N_{min}^{ (n)}\left[\mathcal{C}^+ \right]=0$) for any $m>m_{max}$ ($n>n_{max}$).
-A finite value for $m_{max}$ ($n_{max}$) will always exist for any finite causet, as it will not be large enough or be structured in such a way to have any $m$-next-to-maximal ($n$-next-to-minimal) elements for $m>m_{max}$ ($n>n_{max}$).
+We use the summation $\sum_m$ ($\sum_n$) to denote a sum that runs from $0$ to $m_{max}$ ($n_{max}$), where $m_{max}$ ($n_{max}$) is some integer, the value of which depends on the causet $\mathcal{C}^-$ ($\mathcal{C}^+$), for which all $\F{m}\left[\mathcal{C}^- \right]=0$ ($\P{m}\left[\mathcal{C}^+ \right]=0$) for any $m>m_{max}$ ($n>n_{max}$).
+A finite value for $m_{max}$ ($n_{max}$) will always exist for any finite causet, as it will not be large enough or be structured in such a way to have any $\F{m}$ ($\P{n}$) elements for $m>m_{max}$ ($n>n_{max}$).
 The vectors must be chosen such that they satisfy the following conditions.
@@
 \be\label{eq:min_max_simple_formula}
-S^{ (d)}_{0}\left[\mathcal{C}^-,\mathcal{C}^+\right]:=S^{ (d)}_{GHY}\left[\mathcal{C}^-,\mathcal{C}^+;\vec{p}, -\vec{p} \right]=\left (l/l_p\right)^{d-2}\frac{c_{d}}{2\Gamma\left (\frac{2}{d} \right)}\left ( N_{\vphantom{i}max}^{\: (0)}\left[\mathcal{C}^- \right] - N_{min}^{\: (0)}\left[\mathcal{C}^+ \right] \right)
+S^{ (d)}_{0}\left[\mathcal{C}^-,\mathcal{C}^+\right]:=S^{ (d)}_{GHY}\left[\mathcal{C}^-,\mathcal{C}^+;\vec{p}, -\vec{p} \right]=\left (l/l_p\right)^{d-2}\frac{c_{d}}{2\Gamma\left (\frac{2}{d} \right)}\left ( \F{0}\left[\mathcal{C}^- \right] - \P{0}\left[\mathcal{C}^+ \right] \right)
 \ee
@@
 
-If one wants to calculate the GHY boundary term for the future (past) boundary of a single causet, $\mathcal{C}$, then one simply inserts $\mathcal{C}$ into the $\mathcal{C}^-$ ($\mathcal{C}^+$) argument of $S^{ (d)}_{GHY}$ and sets $\vec{q}$ ($\vec{p}$) to the zero vector, $\mathbf{0}= (0,0,...)$.
+If one wants to calculate the GHY boundary term for the future (past) boundary of a single causet, $\mathcal{C}$, then one simply inserts $\mathcal{C}$ into the $\mathcal{C}^-$ ($\mathcal{C}^+$) argument of $S^{ (d)}_{GHY}$ and sets $\vec{q}$ ($\vec{p}$) to the zero vector, $\vec{0}= (0,0,...)$.
 This renders the $\mathcal{C}^+$ ($\mathcal{C}^-$) argument unnecessary and so one can insert the empty set, $\emptyset$, for convenience.
@@
 \begin{align}\label{eq:future_past_boundary_terms1}
-S^{ (d)}_{+}[\mathcal{C}]:= & S^{ (d)}_{GHY}\left[\mathcal{C},\emptyset;\vec{p},\mathbf{0}\right] =\left (l/l_p\right)^{d-2}\frac{c_{d}}{\Gamma\left (\frac{2}{d} \right)}\left ( d\:N_{\vphantom{i}max}^{\: (1)}\left[\mathcal{C} \right] - N_{\vphantom{i}max}^{\: (0)}\left[\mathcal{C} \right] \right)
+S^{ (d)}_{+}[\mathcal{C}]:= & S^{ (d)}_{GHY}\left[\mathcal{C},\emptyset;\vec{p},\vec{0}\right] =\left (l/l_p\right)^{d-2}\frac{c_{d}}{\Gamma\left (\frac{2}{d} \right)}\left ( d\:\F{1}\left[\mathcal{C} \right] - \F{0}\left[\mathcal{C} \right] \right)
 \\
 \label{eq:future_past_boundary_terms2}
-S^{ (d)}_{-}[\mathcal{C}]:= & S^{ (d)}_{GHY}\left[\emptyset,\mathcal{C};\mathbf{0},-\vec{p}\right] =\left (l/l_p\right)^{d-2}\frac{c_{d}}{\Gamma\left (\frac{2}{d} \right)}\left ( N_{min}^{\: (0)}\left[\mathcal{C} \right] - d\: N_{min}^{\: (1)}\left[\mathcal{C} \right] \right)
+S^{ (d)}_{-}[\mathcal{C}]:= & S^{ (d)}_{GHY}\left[\emptyset,\mathcal{C};\vec{0},-\vec{p}\right] =\left (l/l_p\right)^{d-2}\frac{c_{d}}{\Gamma\left (\frac{2}{d} \right)}\left ( \P{0}\left[\mathcal{C} \right] - d\: \P{1}\left[\mathcal{C} \right] \right)
 \end{align}
@@
 \be\label{general_area_sum}
-A^{ (d)}[\mathcal{C}^-,\mathcal{C}^+;\vec{p},\vec{q}]=\left (l/l_p\right)^{d-1}b_{d}\left (\sum_m p_m N_{\vphantom{i}max}^{\: (m)}\left[\mathcal{C}^- \right] + \sum_n q_n N_{min}^{\: (n)}\left[\mathcal{C}^+ \right]\right)
+A^{ (d)}[\mathcal{C}^-,\mathcal{C}^+;\vec{p},\vec{q}]=\left (l/l_p\right)^{d-1}b_{d}\left (\sum_m p_m \F{m}\left[\mathcal{C}^- \right] + \sum_n q_n \P{n}\left[\mathcal{C}^+ \right]\right)
 \ee
@@
 \be\label{constant_b_d}
-b_d=d\left[\frac{V_{d-1}}{d}\right]^{\frac{1}{d}}
+b_d=d\left(\frac{V_{d-1}}{d}\right)^{\frac{1}{d}}
 \ee
@@
 \begin{align}\label{eq:future_past_spatial_volume}
-A^{ (d)}_{+}[\mathcal{C}]:=A^{ (d)}[\mathcal{C},\emptyset;\vec{p},\mathbf{0}]= & \left (l/l_p\right)^{d-1}\frac{b_{d}}{\Gamma\left (\frac{1}{d}\right)}\: N_{\vphantom{i}max}^{\: (0)}[\mathcal{C}]
+A^{ (d)}_{+}[\mathcal{C}]:=A^{ (d)}[\mathcal{C},\emptyset;\vec{p},\vec{0}]= & \left (l/l_p\right)^{d-1}\frac{b_{d}}{\Gamma\left (\frac{1}{d}\right)}\: \F{0}[\mathcal{C}]
 \\
-A^{ (d)}_{-}[\mathcal{C}]:=A^{ (d)}[\emptyset,\mathcal{C};\mathbf{0},\vec{p}]= & \left (l/l_p\right)^{d-1}\frac{b_{d}}{\Gamma\left (\frac{1}{d}\right)}\: N_{min}^{\: (0)}[\mathcal{C}]
+A^{ (d)}_{-}[\mathcal{C}]:=A^{ (d)}[\emptyset,\mathcal{C};\vec{0},\vec{p}]= & \left (l/l_p\right)^{d-1}\frac{b_{d}}{\Gamma\left (\frac{1}{d}\right)}\: \P{0}[\mathcal{C}]
 \end{align}
@@
 
-\subsection{Poisson Sprinklings for $\left\langle \textbf{N}_{max}^{\: (m)}\right\rangle$ and $\left\langle \textbf{N}_{min}^{\: (n)}\right\rangle$}
+\subsection{Poisson Sprinklings for $\left\langle \BF{m}\right\rangle$ and $\left\langle \BP{m}\right\rangle$}
 
-In order to prove~\eqref{eq:mainconjecture} and~\eqref{eq:conjecture_for_area} let us first derive expressions for the mean values of $\textbf{N}_{\vphantom{i}max}^{\: (m)}$ and $\textbf{N}_{min}^{ (n)}$.
-The proof will will follow almost automatically, as~\eqref{general_boundary_sum} and~\eqref{general_area_sum} are just linear combination of $N_{max}^{\: (m)}$ and $N_{min}^{ (n)}$ terms, and thus the mean values of their random variable counterparts will just be linear combinations of $\left\langle \textbf{N}_{\vphantom{i}max}^{\: (m)}\right\rangle$ and $\left\langle \textbf{N}_{min}^{\: (n)}\right\rangle$ terms.
+In order to prove~\eqref{eq:mainconjecture} and~\eqref{eq:conjecture_for_area} let us first derive expressions for the mean values of $\BF{m}$ and $\BP{m}$.
+The proof will will follow almost automatically, as~\eqref{general_boundary_sum} and~\eqref{general_area_sum} are just linear combination of $\F{m}$ and $\P{m}$ terms, and thus the mean values of their random variable counterparts will just be linear combinations of $\left\langle \BF{m}\right\rangle$ and $\left\langle \BP{m}\right\rangle$ terms.
 
-For any instance of the sprinkling, the probability that a sprinkled point $p\in M$ below the surface $\Sigma$ is $k$\emph{-next-to-maximal} is given by the probability that $k$ points of the sprinkling lie in the region $J^{+} (p)\cap J^{-} (\Sigma)$, the intersection of the causal future of $p$ with the causal past of the surface $\Sigma$.\footnote{For notational convenience we use the symbol $x$ to refer to the causal element $x\in\mathcal C$, to its embedding in the manifold $M$, and to its coordinates in some chart on $M$.} This region will in general be the interior of some curvy $d$-dimensional cone truncated by the surface $\Sigma$, as illustrated in Figure~\ref{fig:Nmin_Nmax}.
+For any instance of the sprinkling, the probability that a sprinkled point $p\in M$ below the surface $\Sigma$ is $F_k$ is given by the probability that $k$ points of the sprinkling lie in the region $J^{+} (p)\cap J^{-} (\Sigma)$, the intersection of the causal future of $p$ with the causal past of the surface $\Sigma$.\footnote{For notational convenience we use the symbol $x$ to refer to the causal element $x\in\mathcal C$, to its embedding in the manifold $M$, and to its coordinates in some chart on $M$.} This region will in general be the interior of some curvy $d$-dimensional cone truncated by the surface $\Sigma$, as illustrated in Figure~\ref{fig:Nmin_Nmax}.
 We will refer to these regions as \emph{truncated cones}.
@@
 to this event, where $V_\blacktriangledown (p)\equiv V (J^{+} (p)\cap J^{-} (\Sigma))$ is the spacetime volume of the region $J^{+} (p)\cap J^{-} (\Sigma)$.
-The probability of sprinkling an element into an infinitesimal four-volume $dV_p$ at $p$ is $\rho dV_p$ where $\rho=l^{-d}$ is the sprinkling density, and so the total expected number of $k$-next-to-maximal elements below $\Sigma$ is
+The probability of sprinkling an element into an infinitesimal four-volume $dV_p$ at $p$ is $\rho dV_p$ where $\rho=l^{-d}$ is the sprinkling density, and so the total expected number of $\F{k}$ elements below $\Sigma$ is
 \be\label{eq:nmax}
-\left\langle \textbf{N}_{max}^{ (k)}\right\rangle =\rho\int_{J^{-} (\Sigma)}dV_p\; \frac{\left (\rho\: V_\blacktriangledown (p)\right)^k}{k!}e^{-\rho V_\blacktriangledown (p)}
+\left\langle \BF{k}\right\rangle =\rho\int_{J^{-} (\Sigma)}dV_p\; \frac{\left (\rho\: V_\blacktriangledown (p)\right)^k}{k!}e^{-\rho V_\blacktriangledown (p)}
 \ee
-Similarly the expected number of $k$-next-to-minimal elements above $\Sigma$ is
+Similarly the expected number of $\P{k}$ elements above $\Sigma$ is
 \be\label{eq:nmin}
-\left\langle \textbf{N}_{min}^{ (k)}\right\rangle =\rho\int_{J^{+} (\Sigma)}dV_p\; \frac{\left (\rho\: V_\blacktriangle (p)\right)^k}{k!}e^{-\rho V_\blacktriangle (p)}
+\left\langle \BP{k}\right\rangle =\rho\int_{J^{+} (\Sigma)}dV_p\; \frac{\left (\rho\: V_\blacktriangle (p)\right)^k}{k!}e^{-\rho V_\blacktriangle (p)}
 \ee
@@
 
-In the limit $\rho\rightarrow\infty$, both quantities will diverge, but if some combination of $k$-next-to-maximal/minimal terms grows slower than or at order $\rho^{1-\frac2d}$, the random variable counterpart of the proposed action~\eqref{general_boundary_sum} will tend to a finite value in the continuum limit.
+In the limit $\rho\rightarrow\infty$, both quantities will diverge, but if some combination of $\P{k}$ and $\F{k}$ grows slower than or at order $\rho^{1-\frac2d}$, the random variable counterpart of the proposed action~\eqref{general_boundary_sum} will tend to a finite value in the continuum limit.
 
@@
 \begin{aligned}
-\left\langle \textbf{N}_{max}^{ (k)}\right\rangle & =\rho \int_{\Sigma}d^{d-1}x\int_{-\varepsilon}^{0}dt\:
+\left\langle \BF{k}\right\rangle & =\rho \int_{\Sigma}d^{d-1}x\int_{-\varepsilon}^{0}dt\:
 h^{\frac{1}{2}}\left (1+
@@
 \\
-\left\langle \textbf{N}_{min}^{ (k)}\right\rangle & =\rho \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
+\left\langle \BP{k}\right\rangle & =\rho \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
 h^{\frac{1}{2}}\left (1+
@@
 \end{align}
-having dropped the subscript $p$.
-Given the volume expressions in GNCs we now proceed to evaluate the integrals for $\left\langle \textbf{N}_{min}^{ (k)}\right\rangle$ and $\left\langle \textbf{N}_{max}^{ (k)}\right\rangle$.
+having dropped the subscript $p$. Given the volume expressions in GNCs we now proceed to evaluate the integrals for $\left\langle \BP{k}\right\rangle$ and $\left\langle \BF{k}\right\rangle$.
 
@@
 \begin{aligned}
-\left\langle \textbf{N}_{max}^{ (k)}\right\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{-\varepsilon}^{0}dt\:
+\left\langle \BF{k}\right\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{-\varepsilon}^{0}dt\:
 h^{\frac{1}{2}}\left (1+
@@
 \\
-\left\langle \textbf{N}_{min}^{ (k)}\right\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
+\left\langle \BP{k}\right\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
 h^{\frac{1}{2}}\left (1+
@@
 \begin{aligned}
-\left\langle \textbf{N}_{max}^{ (k)}\right\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
+\left\langle \BF{k}\right\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
 h^{\frac{1}{2}}\left (1+
@@
 \\
-\left\langle \textbf{N}_{min}^{ (k)}\right\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
+\left\langle \BP{k}\right\rangle = \frac{\rho^{k+1}}{k!} & \int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
 h^{\frac{1}{2}}\left (1-
@@
 The reason for expanding the exponentials is that the integrals will have to be put into the form of Gaussian integrals in order to evaluate them in the limit $\rho\rightarrow\infty$.
-The above expressions are almost the same apart from a few sign differences, so from now on we will just work with the expression for $\left\langle \textbf{N}_{min}^{ (k)}\right\rangle$.
+The above expressions are almost the same apart from a few sign differences, so from now on we will just work with the expression for $\left\langle \BP{k}\right\rangle$.
 By expanding the brackets, and retaining only necessary orders, the integral can be split into two parts of differing powers of $\rho$.
@@
 \begin{aligned}
-\left\langle \textbf{N}_{min}^{ (k)}\right\rangle & = \frac{\rho^{k+1}A^k}{k!}\int_{\Sigma}d^{d-1}x\: h^{\frac{1}{2}}\int_{0}^{\varepsilon}dt\:
+\left\langle \BP{k}\right\rangle & = \frac{\rho^{k+1}A^k}{k!}\int_{\Sigma}d^{d-1}x\: h^{\frac{1}{2}}\int_{0}^{\varepsilon}dt\:
 \left\lbrace \left ( t^{dk} + \left (kB-K \right)t^{dk+1}\right) + O\left (t^{dk+2}\right) \right\rbrace e^{-\rho A\: t^d}
@@
 We can use this procedure to take the limit of~\eqref{eq:n_min_different_rho_terms} and formulate the answer in terms of gamma functions.
-The limits of both $\left\langle \textbf{N}_{max}^{ (k)}\right\rangle$ and $\left\langle \textbf{N}_{min}^{ (k)}\right\rangle$ are then given by
+The limits of both $\left\langle \BF{k}\right\rangle$ and $\left\langle \BP{k}\right\rangle$ are then given by
 \begin{gather}\label{eq:nmax_nmin_final}
 \begin{aligned}
-\lim_{\rho\rightarrow\infty}\left\langle \textbf{N}_{max}^{ (k)}\right\rangle = \lim_{\rho\rightarrow\infty} & \left\lbrace \rho^{1-\frac{1}{d}} \left (b_d\right)^{-1} \frac{\Gamma\left (\frac{1}{d}+k\right)}{k!}
+\lim_{\rho\rightarrow\infty}\left\langle \BF{k}\right\rangle = \lim_{\rho\rightarrow\infty} & \left\lbrace \rho^{1-\frac{1}{d}} \left (b_d\right)^{-1} \frac{\Gamma\left (\frac{1}{d}+k\right)}{k!}
 \int_{\Sigma}d^{d-1}x\: \sqrt{h} \right.
@@
 \\
-\lim_{\rho\rightarrow\infty}\left\langle \textbf{N}_{min}^{ (k)}\right\rangle = \lim_{\rho\rightarrow\infty} & \left\lbrace \rho^{1-\frac{1}{d}} \left (b_d\right)^{-1} \frac{\Gamma\left (\frac{1}{d}+k\right)}{k!}
+\lim_{\rho\rightarrow\infty}\left\langle \BP{k}\right\rangle = \lim_{\rho\rightarrow\infty} & \left\lbrace \rho^{1-\frac{1}{d}} \left (b_d\right)^{-1} \frac{\Gamma\left (\frac{1}{d}+k\right)}{k!}
 \int_{\Sigma}d^{d-1}x\: \sqrt{h} \right.
@@
 
-The form of~\eqref{eq:nmax_nmin_final} seems to suggest general series expansions of $\left\langle \textbf{N}_{max}^{ (k)}\right\rangle$ and $\left\langle \textbf{N}_{min}^{ (k)}\right\rangle$ in reducing powers of $\rho$.
+The form of~\eqref{eq:nmax_nmin_final} seems to suggest general series expansions of $\left\langle \BF{k}\right\rangle$ and $\left\langle \BP{k}\right\rangle$ in reducing powers of $\rho$.
 Written in terms of the discreteness scale, $l$, it is a Laurent series starting at $l^{1-d}$.
@@
 We believe this choice is valid as the simulations have suggested that all members of the $\textbf{S}^{ (d)}_{GHY}$ family have similar fluctions.
-Since $\textbf{N}_{max}^{ (0)}$ and $\textbf{N}_{min}^{ (0)}$ are random variables associated with a codimension-1 surface (they count elements ``near" $\Sigma$), we may expect their mean values to scale like $\left\langle\textbf{N}\right\rangle^\frac{d-1}{d}$ --- which agrees with the leading order behaviour of~\eqref{eq:nmax_nmin_final} --- and thus they would inherit fluctuations of order $\left\langle\textbf{N}\right\rangle^\frac{d-1}{2d}$.
+Since $\BF{0}$ and $\BP{0}$ are random variables associated with a codimension-1 surface (they count elements ``near" $\Sigma$), we may expect their mean values to scale like $\left\langle\textbf{N}\right\rangle^\frac{d-1}{d}$ --- which agrees with the leading order behaviour of~\eqref{eq:nmax_nmin_final} --- and hence that they inherit fluctuations of order $\left\langle\textbf{N}\right\rangle^\frac{d-1}{2d} = (\rho V)^\frac{d-1}{2d}$.
 %{This comes from the following observation.
@@
 }
-The action $\textbf{S}^{ (d)}_{0}$, being proportional to $\rho^\frac{2-d}{d}$ times the difference of two independent\footnote{The independence is as a feature of the Poisson process in $M$: the number of elements $\textbf{N}[R]$ in any subregion $R\subset M$ is a Poisson variable itself, and for any two disjoint regions $R_1$, $R_2$ the numbers $\textbf{N}[R_1]$ and $\textbf{N}[R_2]$ are independent random variables.} random variables with standard deviation $\left\langle\textbf{N}\right\rangle^\frac{d-1}{2d} = (\rho V)^\frac{d-1}{2d}$ should see fluctuations of order $\rho^\frac{2-d}{d}\rho^\frac{d-1}{2d}=\rho^\frac{3-d}{2d}$.
+The action $\textbf{S}^{ (d)}_{0}$, being proportional to $\rho^\frac{2-d}{d}$ times the difference of two independent\footnote{The independence is as a feature of the Poisson process in $M$: the number of elements $\textbf{N}[R]$ in any subregion $R\subset M$ is a Poisson variable itself, and for any two disjoint regions $R_1$, $R_2$ the numbers $\textbf{N}[R_1]$ and $\textbf{N}[R_2]$ are independent random variables.} random variables with standard deviation of order $\rho^\frac{d-1}{2d}$  should see fluctuations of order $\rho^\frac{2-d}{d}\rho^\frac{d-1}{2d}=\rho^\frac{3-d}{2d}$.
 This suggests that for $d=2$ these fluctuations should grow like $\rho^{\frac{1}{4}}$ as $\rho\rightarrow\infty$, for $d=3$ they should be constant, and for $d>3$ they should be damped.
@@
 Sprinkle at density $\rho=l^{-d}$ into the cube, which in any given run will place $N$ points inside the cube, where $N$ is the realisation of the Poisson random number $\textbf{N}$ with mean $\left\langle \textbf{N}\right\rangle = \rho V=  (L/l)^d$.
-Evaluate $S^{ (d)}_0=\rho^\frac{2-d}{d}c_d\left (2\Gamma\left (\frac{2}{d}\right)\right)^{-1} (N^{ (0)}_{\vphantom{i}max} - N_{min}^{ (0)})$ for $\Sigma$ by counting the minimal/maximal elements in the upper/lower half of the cube ($S^{ (d)}_0$, $N^{ (0)}_{\vphantom{i}max}$, and $N^{ (0)}_{min}$ of course being realisations of the random variables $\textbf{S}^{ (d)}_0$, $\textbf{N}^{ (0)}_{\vphantom{i}max}$, and $\textbf{N}^{ (0)}_{min}$), and repeat for $R$ runs.
+Evaluate $S^{ (d)}_0=\rho^\frac{2-d}{d}c_d\left (2\Gamma\left (\frac{2}{d}\right)\right)^{-1} (\F{0} - \P{0})$ for $\Sigma$ by counting the minimal/maximal elements in the upper/lower half of the cube ($S^{ (d)}_0$, $\F{0}$, and $\P{0}$ of course being realisations of the random variables $\textbf{S}^{ (d)}_0$, $\BF{0}$, and $\BP{0}$), and repeat for $R$ runs.
 The value we expect for the mean (sum of all $S^{ (d)}_0$ values for each run divided by $R$) is zero in all cases, since the extrinsic curvature of the surface is zero.
@@
 
-As mentioned above, we have performed analogous simulations for other family members, such as $S^{ (d)}_+$.
-This expression involves the terms $N^{ (1)}_{max}$ and $N^{ (0)}_{max}$, which means that heuristic arguments of the type given above are hampered by the fact that the two random variables whose difference is taken are not independent.
+As mentioned above, we have performed analogous simulations for other family members, such as $\mathbf S^{ (d)}_+$.
+This expression involves the terms $\BF{0}$ and $\BF{1}$, which means that heuristic arguments of the type given above are hampered by the fact that the two random variables whose difference is taken are not independent.
 However, the results for the scaling of the fluctuations turn out to be identical to those obtained in the first case.
@@
 Given our ignorance in this matter,  we can ask if causal set theory  might  in fact guide us in the right direction.
-Our starting point is the $d$-dimensional  Benincasa-Dowker Action \cite{Benincasa_Dowker:The_Scalar_Curvature_of_a_Causal_Set, Dowker_Glaser:dAlembertians_for_Causal_Sets} which is constructed from the mean number  $N_i^{ (d)}$\footnote{ $N_i^{ (d)}$ is dimension  and geometry dependent, but for our purpose it suffices to retain the dimension label.
-} of $i+1$-element inclusive intervals in the causal set
+Our starting point is the $d$-dimensional  Benincasa-Dowker Action \cite{Benincasa_Dowker:The_Scalar_Curvature_of_a_Causal_Set, Dowker_Glaser:dAlembertians_for_Causal_Sets} which is constructed from the mean number  $N_i^{ (d)}$ of $i+1$-element inclusive intervals in the causal set\footnote{ $N_i^{ (d)}$ is dimension  and geometry dependent, but for our purpose it suffices to retain the dimension label.
+}
 \begin{equation}   

### 2014-09-29 paper_v7.tex
minor aesthetic edits

@@
 It is (i) reflexive: $x\preceq x$, (ii) acyclic: $x\preceq y\preceq x \Rightarrow x=y$, and (iii) transitive: $x\preceq y\preceq z \Rightarrow x\preceq z$ for all points $x, y, z \in \mathcal{C}$.
-We define an inclusive order interval as the set $I (x,y)\equiv \lbrace z\in\mathcal{C}|x\preceq z\preceq y\rbrace$ for any $x, y\in\mathcal{C}$.
+We define an inclusive order interval as the set $I (x,y)\equiv [ z\in\mathcal{C}|x\preceq z\preceq y]$ for any $x, y\in\mathcal{C}$.
 The \textit{locally finite} condition is simply that the cardinality of any order interval is finite, that is $|I (x,y)|<\infty$.
@@
 
-\subsection{The mean of $\textbf{S}^{ (d)}_{GHY}[\mathcal {C}^-,\mathcal{C}^+;\vec{p},\vec{q}]$}
+\subsection{The mean of $\textbf{S}^{ (d)}_{GHY}\left[\mathcal {C}^-,\mathcal{C}^+;\vec{p},\vec{q}\,\right]$}
 
@@
 \left\langle \BP{k}\right\rangle & = \frac{\rho^{k+1}A^k}{k!}\int_{\Sigma}d^{d-1}x\: h^{\frac{1}{2}}\int_{0}^{\varepsilon}dt\:
-\left\lbrace \left ( t^{dk} + \left (kB-K \right)t^{dk+1}\right) + O\left (t^{dk+2}\right) \right\rbrace e^{-\rho A\: t^d}
+\left[ \left ( t^{dk} + \left (kB-K \right)t^{dk+1}\right) + O\left (t^{dk+2}\right) \right] e^{-\rho A\: t^d}
  \\
  & - \frac{\rho^{k+2}A^{k+1}}{k!}\int_{\Sigma}d^{d-1}x\: h^{\frac{1}{2}}\int_{0}^{\varepsilon}dt\:
-\left\lbrace Bt^{dk+d+1} + O\left (t^{dk+d+2}\right) \right\rbrace e^{-\rho A\: t^d}
+\left[ Bt^{dk+d+1} + O\left (t^{dk+d+2}\right) \right] e^{-\rho A\: t^d}
 \end{aligned}
@@
 \begin{aligned}
-\lim_{\rho\rightarrow\infty}\left\langle \BF{k}\right\rangle = \lim_{\rho\rightarrow\infty} & \left\lbrace \rho^{1-\frac{1}{d}} \left (b_d\right)^{-1} \frac{\Gamma\left (\frac{1}{d}+k\right)}{k!}
+\lim_{\rho\rightarrow\infty}\left\langle \BF{k}\right\rangle = \lim_{\rho\rightarrow\infty} & \left[ \rho^{1-\frac{1}{d}} \left (b_d\right)^{-1} \frac{\Gamma\left (\frac{1}{d}+k\right)}{k!}
 \int_{\Sigma}d^{d-1}x\: \sqrt{h} \right.
@@
  &  \left. +\rho^{1-\frac{2}{d}} \left (c_d\right)^{-1} \frac{\Gamma\left (\frac{2}{d}+k\right)}{k!}
-\int_{\Sigma}d^{d-1}x\: \sqrt{h}K + O\left (\rho^{1-\frac{3}{d}} \right) \right\rbrace
+\int_{\Sigma}d^{d-1}x\: \sqrt{h}K + O\left (\rho^{1-\frac{3}{d}} \right) \right]
 \\
-\lim_{\rho\rightarrow\infty}\left\langle \BP{k}\right\rangle = \lim_{\rho\rightarrow\infty} & \left\lbrace \rho^{1-\frac{1}{d}} \left (b_d\right)^{-1} \frac{\Gamma\left (\frac{1}{d}+k\right)}{k!}
+\lim_{\rho\rightarrow\infty}\left\langle \BP{k}\right\rangle = \lim_{\rho\rightarrow\infty} & \left[ \rho^{1-\frac{1}{d}} \left (b_d\right)^{-1} \frac{\Gamma\left (\frac{1}{d}+k\right)}{k!}
 \int_{\Sigma}d^{d-1}x\: \sqrt{h} \right.
@@
  &  \left. -\rho^{1-\frac{2}{d}} \left (c_d\right)^{-1} \frac{\Gamma\left (\frac{2}{d}+k\right)}{k!}
-\int_{\Sigma}d^{d-1}x\: \sqrt{h}K + O\left (\rho^{1-\frac{3}{d}} \right) \right\rbrace
+\int_{\Sigma}d^{d-1}x\: \sqrt{h}K + O\left (\rho^{1-\frac{3}{d}} \right) \right]
 \\
@@
 \end{cases}
-\quad\text{and}\quad 
-%\end{equation}
-%\begin{equation}
+%\quad\text{and}\quad 
+\end{aligned}
+\end{equation}
+and
+\begin{equation}
+\begin{aligned}
 \beta_d =
@@
 \begin{cases} 
-\displaystyle\sum_{k=0}^{i-1} (-1)^k\binom{i-1}{k} \frac{\Gamma\left (d/2 (k+1)+3/2\right)}{\Gamma\left (d/2+3/2\right) \Gamma\left (1+dk/2\right)}  \quad  &d\mathrm{ \, \, odd}\\ 
-\displaystyle\sum_{k=0}^{i-1} (-1)^k\binom{i-1}{k} \frac{\Gamma\left (d/2 (k+1)+2\right)}{\Gamma\left (d/2+2\right) \Gamma\left (1+dk/2\right)}  &d \mathrm{ \, \,  even.}\\ 
+\displaystyle\sum_{k=0}^{i-1} (-1)^k\binom{i-1}{k} \frac{\Gamma\left(\frac{d}2(k+1)+\frac32\right)}{\Gamma\left (\frac{d+3}{2}\right) \Gamma\left (1+\frac{d}2k\right)}  \quad  &d\mathrm{ \, \, odd}\\ 
+\displaystyle\sum_{k=0}^{i-1} (-1)^k\binom{i-1}{k} \frac{\Gamma\left (\frac{d}2 (k+1)+2\right)}{\Gamma\left (\frac{d}2+2\right) \Gamma\left (1+\frac{d}2k\right)}  &d \mathrm{ \, \,  even.}\\ 
 \end{cases} 
@@
 \begin{equation}
-C_i^d={}_{q+1}F_{q} (\{ a_1, \ldots a_q, i-1\}, \{b_1, \ldots b_q \} |1)
+C_i^{(d)}={}_{q+1}F_{q} (\{ a_1, \ldots, a_q, i-1\}, \{b_1, \ldots b_q \} |1)
 \label{chyp} 
 \end{equation} 
-with $q=d/2+1/2, a_i=\frac{d+2i}{d}, b_i=2i/d$ for $d$ odd and $q=d/2, a_i=\frac{d+2i+2}{d}, b_i=2i/d$  for $d$ even.
+with $q=\frac{d+1}{2}$, $a_i=\frac{d+2i}{d}$, $b_i=\frac{2i}{d}$ for $d$ odd and $q=\frac{d}{2}$, $a_i=\frac{d+2i+2}{d}$, and $b_i=\frac{2i}{d}$  for $d$ even.
 
@@
 In \cite{Glaser_Sumati:Locality_in_Causal_Set} a closed form expression was obtained for the mean value of the number of $i+1$ element inclusive intervals $N_i^{(d)}$ contained in an Alexandrov interval in $d$ dimensional flat spacetime
-\begin{equation} N_i^{ (d)}= \frac{\Gamma\left (d\right)^2 N^{i+2}}{\Gamma\left (i\right)} \sum_{k=0}^\infty \frac{ (-N)^k\,\Gamma\left (\frac{d (k+i)}{2}+1\right)  \Gamma\left (\frac{d (k+i+1)}{2}+1\right)}{ (k+i+2)  (k+i+1) \Gamma\left (k+1\right)\Gamma\left (\frac{d (k+i)}{2} +d\right) \Gamma\left (\frac{d (k+i+1)}{2}+d\right)},\nonumber
+\begin{equation} N_i^{ (d)}= \frac{\Gamma\left (d\right)^2 N^{i+2}}{\Gamma\left (i\right)} \sum_{k=0}^\infty \frac{ (-N)^k\,\Gamma\left (\frac{d}2 (k+i)+1\right)  \Gamma\left (\frac{d}2 (k+i+1)+1\right)}{ (k+i+2)  (k+i+1) \Gamma\left (k+1\right)\Gamma\left (\frac{d}2 (k+i) +d\right) \Gamma\left (\frac{d}2 (k+i+1)+d\right)},\nonumber
 \end{equation} 
@@
 \begin{equation}
- A_j^{ (d)} = \Gamma\left (d\right)^2\left ( \sum_{i=1}^{D+2} (-1)^i \binom{j-1}{i-1}  C_i^{ (d)}  \right) \frac{ (-1)^j}{ (j+1)!}\frac{\Gamma\left (\frac{d}{2} (j-1)+1\right)\Gamma\left (\frac{d}{2}j+1\right)}{\Gamma\left (\frac{d}{2} (j-1)+d\right) \Gamma\left (\frac{d}{2}j+d\right)},
+ A_j^{ (d)} = \Gamma\left (d\right)^2\frac{ (-1)^j}{ (j+1)!}\frac{\Gamma\left (\frac{d}{2} (j-1)+1\right)\Gamma\left (\frac{d}{2}j+1\right)}{\Gamma\left (\frac{d}{2} (j-1)+d\right) \Gamma\left (\frac{d}{2}j+d\right)}\sum_{i=1}^{D+2} (-1)^i \binom{j-1}{i-1}  C_i^{ (d)},
 \end{equation} 
@@
 \begin{equation}
-A_j^d= \gamma_j^{ (d)}   \frac{\Gamma\left (d\right)^2 (-1)^{j+1}}{\Gamma\left (\frac{d}{2} (j + 1)\right) \Gamma\left (\frac{d}{2} (2 + j)\right) \Gamma\left (2 + j\right) }
+A_j^{(d)}= \gamma_j^{ (d)}   \frac{\Gamma\left (d\right)^2 (-1)^{j+1}}{\Gamma\left (\frac{d}{2} (j + 1)\right) \Gamma\left (\frac{d}{2} (2 + j)\right) \Gamma\left (2 + j\right) }
 \label{simplercoefft} 

### 2014-09-30 paper_v7.tex
some edits

@@
 
-Consider a sufficiently well-behaved $d$-dimensional spacetime $(M,g)$ and a surface $\Sigma$ in $M$, which we shall take to be a Cauchy surface (hence $(M,g)$ is globally hyperbolic).
-The causal past and future sets $M^\pm=J^\pm (\Sigma)$ of the surface then form a partition of $M$, and $\partial M^\pm = \Sigma$ ($J^\pm$ denoting the chronological future/past).
+Consider a sufficiently well-behaved $d$-dimensional spacetime $(M,g)$ and a surface $\Sigma$ in $M$, which we shall take to be a Cauchy surface\mb{maybe too strong.
+what about boxes then?
+maybe partitioning surface?}.
+The causal past and future sets $M^\pm=J^\pm (\Sigma)$ of the surface then form a partition of $M$, and $\partial M^\pm = \Sigma$.
 The Gibbons-Hawking-York boundary term for $M^\pm$ is in this case given by
 \be\label{eq:GHYBT_in_continuum}
-\mathcal{S}_{GHY} = \mp \frac{1}{l_p^{d-2}}\int_{\Sigma} K d\Sigma
+{S}^{(d)}_{GHY}\left[M^\pm\right]= \mp \frac{1}{l_p^{d-2}}\int_{\Sigma} K d\Sigma
 \ee
@@
 
-Given a causet that can be faithfully embedded in a manifold $M$ with Cauchy hypersurface $\Sigma$, we can partition the causet into two sets $\mathcal{C}^+$ and $\mathcal{C}^-$, corresponding to the sets of elements that lie in $M^+=J^+(\Sigma)$ and $M^-=J^-(\Sigma)$.
-By virtue of $\Sigma$ being Cauchy, the set $\mathcal C^-$ contains its own past (i.e.
-any element in $\mathcal C$ that precedes at least one element in $\mathcal{C}^-$ is also in $\mathcal{C}^-$).
+Given a causet that can be faithfully embedded in a manifold $M$ with Cauchy hypersurface $\Sigma$, we obtain a natural partition of the causet into two parts $\mathcal{C}^+$ and $\mathcal{C}^-$, corresponding to the sets of elements that lie in $M^+=J^+(\Sigma)$ and $M^-=J^-(\Sigma)$.
+By virtue of $\Sigma$ being Cauchy, the set $\mathcal C^-$ contains its own past, that is, any element in $\mathcal C$ that precedes at least one element in $\mathcal{C}^-$ is also in $\mathcal{C}^-$.
 Such a set is known as a \emph{past stem}.
-Similarly, $\mathcal C^+$ is a \emph{future stem} given the analogous definition.
-These properties capture the essential features of a partition induced by a Cauchy surface.
-Hence we propose more generally that for any causet $\mathcal C$, a \emph{Cauchy partition} is a partition that splits $\mathcal{C}$ into two parts, $\mathcal{C}^-$ and $\mathcal{C}^+=\mathcal{C}\setminus\mathcal{C}^-$ \IGNORE{($\mathcal{C}=\mathcal{C}^-\cup\mathcal{C}^+$)}, such that $\mathcal{C}^-$ is a past stem (or, equivalently, that $\mathcal{C}^+$ be a future stem and $\mathcal{C}^- = \mathcal{C}\setminus\mathcal{C}^+$).
+Similarly, $\mathcal C^+$ is a \emph{future stem} given an analogous definition.
+These two properties capture the essential features of a partition induced by a Cauchy surface.
+Hence we propose more generally that for any causet $\mathcal C$, a \emph{Cauchy partition} is a partition that splits $\mathcal{C}$ into two parts, $\mathcal{C}^-$ and $\mathcal{C}^+=\mathcal{C}\setminus\mathcal{C}^-$ \IGNORE{($\mathcal{C}=\mathcal{C}^-\cup\mathcal{C}^+$)}, such that $\mathcal{C}^-$ is a past stem (or, equivalently, that $\mathcal{C}^+$ is a future stem and $\mathcal{C}^- = \mathcal{C}\setminus\mathcal{C}^+$).
 A sprinkling into a manifold with Cauchy surface will of course produce a causal set with Cauchy partition, but we shall refer to a partition of a causal set satisfying the above definition as Cauchy even when the causet is not obtained via sprinkling faithfully embeddable into some manifold.
@@
 Given a causal set with Cauchy partition, there follows a natural definition for the ``immediate neighbours'' in the past and future sets of the partition.
-The ``immediate neighbours in the past'' can be defined as those elements of $\mathcal{C}^-$ that have no elements in their causal future in $\mathcal{C}^-$.
+The ``immediate neighbours in the past'' of the partition can be defined as those elements of $\mathcal{C}^-$ that have no elements in their causal future in $\mathcal{C}^-$.
 Such elements will be referred to as \textit{maximal elements} in $\mathcal{C}^-$.
@@
 
-For a given causet $\mathcal C$, let us define the function $\F{k}\left[\mathcal{C} \right]$, which counts the number of elements in $\mathcal C$ that have exactly $k$ elements in their future, and the function $\P{k}\left[\mathcal{C}\right]$, which counts the number of elements which have exactly $k$ elements in their past.
+For a given causet $\mathcal C$, let us define the function $\F{k}\left[\mathcal{C}\,\right]$, which counts the number of elements in $\mathcal C$ that have exactly $k$ elements in their future, and the function $\P{k}\left[\mathcal{C}\,\right]$, which counts the number of elements which have exactly $k$ elements in their past.
 An ``$F_0$ element'' is then a \emph{maximal} element, a ``$P_0$ element'' is a \emph{minimal} element, an ``$F_1$ element'' is (in some sense) next-to-maximal, and so on.\mb{Ian: your plot here?}
 
-Given a causet, $\mathcal{C}$, with a spacelike partition so that $\mathcal C = \mathcal C^+ \cup\, \mathcal C^-$, we propose the following family of functions as the generalised bounary action:
+Given a causet, $\mathcal{C}$, with a spacelike partition so that $\mathcal C = \mathcal C^+ \cup\, \mathcal C^-$, we propose the following family of functions as the generalised causal set boundary action:
 \be\label{general_boundary_sum}
-S^{ (d)}_{GHY}\left[\mathcal{C}^-,\mathcal{C}^+;\vec{p}, \vec{q} \right]= \left (l/l_p\right)^{d-2} c_{d}
+S^{ (d)}_{GHY}\left[\mathcal{C}^-,\mathcal{C}^+;\vec{p}, \vec{q}\,\right]= \left (l/l_p\right)^{d-2} c_{d}
 \left ( \sum_m p_m \F{m}\left[\mathcal{C}^- \right]
@@
 \be\label{Cn}
-c_{d}=\frac{d (d+1)}{ (d+2)}\left[\frac{V_{d-1}}{d}\right]^{\frac{2}{d}},
+c_{d}=\frac{d (d+1)}{ (d+2)}\left(\frac{V_{d-1}}{d}\right)^{\frac{2}{d}},
 \ee
-and where $V_d=\pi^{\frac{d}{2}}/\Gamma\left (\frac{d}{2}+1\right)$ denotes the volume of the unit $d$-ball ($\Gamma (t)$ is a gamma function, defined as $\Gamma (t):=\int_0^\infty dx\: x^{t-1}\: e^{-x}$).
+
+$V_d=\pi^{\frac{d}{2}}/\Gamma\left (\frac{d}{2}+1\right)$ denotes the volume of the unit $d$-ball and $\Gamma (t)$ is the gamma function.
 We have used $\vec{p}$ and $\vec{q}$ to stand for infinite dimensional vectors $ (p_0,...,p_m,..)$ and $ (q_0,...,q_n,..)$ respectively, where all the $p_m,\: q_n \in \mathbb{R}$.
-These vectors contain a finite number of non-zero entries, corresponding to the coefficients of the different $N_{{max\vphantom{i}}}^{\: (m)}$ and $N_{{min}}^{\: (n)}$ terms.
+These vectors contain a finite number of non-zero entries, corresponding to the coefficients of the different $\F{m}$ and $\P{m}$ terms.
 We use the summation $\sum_m$ ($\sum_n$) to denote a sum that runs from $0$ to $m_{max}$ ($n_{max}$), where $m_{max}$ ($n_{max}$) is some integer, the value of which depends on the causet $\mathcal{C}^-$ ($\mathcal{C}^+$), for which all $\F{m}\left[\mathcal{C}^- \right]=0$ ($\P{m}\left[\mathcal{C}^+ \right]=0$) for any $m>m_{max}$ ($n>n_{max}$).
@@
 \be\label{eq:min_max_simple_formula}
-S^{ (d)}_{0}\left[\mathcal{C}^-,\mathcal{C}^+\right]:=S^{ (d)}_{GHY}\left[\mathcal{C}^-,\mathcal{C}^+;\vec{p}, -\vec{p} \right]=\left (l/l_p\right)^{d-2}\frac{c_{d}}{2\Gamma\left (\frac{2}{d} \right)}\left ( \F{0}\left[\mathcal{C}^- \right] - \P{0}\left[\mathcal{C}^+ \right] \right)
+S^{ (d)}_{0}\left[\mathcal{C}^-,\mathcal{C}^+\right]:=S^{ (d)}_{GHY}\left[\mathcal{C}^-,\mathcal{C}^+;\vec{p}, -\vec{p}\,\right]=\left (l/l_p\right)^{d-2}\frac{c_{d}}{2\Gamma\left (\frac{2}{d} \right)}\left ( \F{0}\left[\mathcal{C}^- \right] - \P{0}\left[\mathcal{C}^+ \right] \right)
 \ee
 where $\vec{p}=\left (\left (2\Gamma\left (\frac{2}{d} \right)\right)^{-1},0,0,...\right)$.
-This expression is the easiest to employ computationally, and we shall use its random variable counterpart, $\textbf{S}^{ (d)}_{0}\left[M^-,M^+,\rho\right]:=\textbf{S}^{ (d)}_{GHY}\left[M^-,M^+,\rho;\vec{p}, -\vec{p} \right]$, later when we study the fluctuations of the discrete boundary actions numerically.
+This expression is the easiest to employ computationally, and we shall use its random variable counterpart, $\textbf{S}^{ (d)}_{0}\left[M^-,M^+,\rho\right]:=\textbf{S}^{ (d)}_{GHY}\left[M^-,M^+,\rho;\vec{p}, -\vec{p}\,\right]$, later when we study the fluctuations of the discrete boundary actions numerically.
 
@@
 \begin{align}\label{eq:future_past_boundary_terms1}
-S^{ (d)}_{+}[\mathcal{C}]:= & S^{ (d)}_{GHY}\left[\mathcal{C},\emptyset;\vec{p},\vec{0}\right] =\left (l/l_p\right)^{d-2}\frac{c_{d}}{\Gamma\left (\frac{2}{d} \right)}\left ( d\:\F{1}\left[\mathcal{C} \right] - \F{0}\left[\mathcal{C} \right] \right)
+S^{ (d)}_{+}[\mathcal{C}]:= & S^{ (d)}_{GHY}\left[\mathcal{C},\emptyset;\vec{p},\vec{0}\,\right] =\left (l/l_p\right)^{d-2}\frac{c_{d}}{\Gamma\left (\frac{2}{d} \right)}\left ( d\:\F{1}\left[\mathcal{C}\,\right] - \F{0}\left[\mathcal{C}\,\right] \right)
 \\
 \label{eq:future_past_boundary_terms2}
-S^{ (d)}_{-}[\mathcal{C}]:= & S^{ (d)}_{GHY}\left[\emptyset,\mathcal{C};\vec{0},-\vec{p}\right] =\left (l/l_p\right)^{d-2}\frac{c_{d}}{\Gamma\left (\frac{2}{d} \right)}\left ( \P{0}\left[\mathcal{C} \right] - d\: \P{1}\left[\mathcal{C} \right] \right)
+S^{ (d)}_{-}[\mathcal{C}]:= & S^{ (d)}_{GHY}\left[\emptyset,\mathcal{C};\vec{0},-\vec{p}\,\right] =\left (l/l_p\right)^{d-2}\frac{c_{d}}{\Gamma\left (\frac{2}{d} \right)}\left ( \P{0}\left[\mathcal{C}\,\right] - d\: \P{1}\left[\mathcal{C}\,\right] \right)
 \end{align}
@@
 
-%MB1: (too vague) During the calculation approximations will be made.
+MB1: (too vague) During the calculation approximations will be made.
 These approximations are justified on the grounds that if certain orders are retained throughout the calculation they will in fact vanish in the limit $\rho \rightarrow \infty$
-%\begin{figure}
-  %\centering
-   % {\includegraphics[width=\textwidth]{minmaxplot}}
-    % \caption{An illustration of the idea.
+\begin{figure}
+  \centering
+    {\includegraphics[width=\textwidth]{minmaxplot}}
+     \caption{An illustration of the idea.
 Pictured here is a causal set obtained by a sprinkling into a spacetime partitioned by a spacelike hypersurface.
 The minimal and maximal points about the surface are highlighted and the shaded regions illustrate the regions whose volumes $V_\blacktriangle$ and $V_\blacktriangledown$ are needed in the proof.}
-     %\label{fig:Nmin_Nmax}
-%*\end{figure}
+     \label{fig:Nmin_Nmax}
+*\end{figure}
 
@@
 Substituting $r_{max}=T$ into (\ref{eq:VolumeIntegralSpherical}) allows us to evaluate the integral, wherein we find
-%\begin{figure}[t]
- % \centering
-  %  {\includegraphics[scale=0.5]{coneplot}}
-   %  \caption{The size of the region inside the top and bottom bounding surfaces is the volume we want to calculate.}
-    % \label{fig:cone_plot}
-%\end{figure}
+\begin{figure}[t]
+  \centering
+    {\includegraphics[scale=0.5]{coneplot}}
+     \caption{The size of the region inside the top and bottom bounding surfaces is the volume we want to calculate.}
+     \label{fig:cone_plot}
+\end{figure}
 \be\label{eq:VolumeNoK}
@@
 
-%\begin{figure}[t!]
- % \centering
-  %  {\includegraphics[scale=0.6]{GHY-fluctuation_plots}}
-   %  \caption{A $\log-\log$ plot (base $2$) of the fluctuations in $S^{ (d)}_{GHY}$ for a flat ($K=0$) surface in $d=2,3$ and $4$-dimensional Minkowski space.
+\begin{figure}[t!]
+  \centering
+    {\includegraphics[scale=0.6]{GHY-fluctuation_plots}}
+     \caption{A $\log-\log$ plot (base $2$) of the fluctuations in $S^{ (d)}_{GHY}$ for a flat ($K=0$) surface in $d=2,3$ and $4$-dimensional Minkowski space.
 Each data point represents the sample standard deviation in $S^{ (d)}_{GHY}$ on a sample of $R=100$ runs.
@@
 The corresponding black, blue and red lines have gradients $\frac12$, $0$ and $-\frac18$ and best-fit intercepts of order $1$.}
-    % \label{fig:fluctuations}
-%\end{figure}
+     \label{fig:fluctuations}
+\end{figure}
 
@@
 \begin{equation}
-A_j^{(d)}= \gamma_j^{ (d)}   \frac{\Gamma\left (d\right)^2 (-1)^{j+1}}{\Gamma\left (\frac{d}{2} (j + 1)\right) \Gamma\left (\frac{d}{2} (2 + j)\right) \Gamma\left (2 + j\right) }
+A_j^{(d)}=   \frac{\Gamma\left (d\right)^2 (-1)^{j+1}}{\Gamma\left (\frac{d}{2} (j + 1)\right) \Gamma\left (\frac{d}{2} (2 + j)\right) \Gamma\left (2 + j\right) } \gamma_j^{ (d)}
 \label{simplercoefft} 
@@
 \begin{equation} 
-\frac{ 8 \pi G S_{BD}^{ (d)}[\mathcal C]}{\vol (S^{d-2})} = \frac{b}{d (d-1) \Gamma\left (1+2/d\right) N^{\frac{d-2}{d}}} \left ( N + \frac{\beta_d}{\alpha_d} \sum_{i=1}^{n_d} C_i^{ (d)} N_i^{ (d)} \right)
+\frac{ 8 \pi G S_{BD}^{ (d)}[\mathcal C]}{\vol (S^{d-2})} = \frac{b}{d (d-1) \Gamma\left (1+\frac2d\right) N^{\frac{d-2}{d}}} \left ( N + \frac{\beta_d}{\alpha_d} \sum_{i=1}^{n_d} C_i^{ (d)} N_i^{ (d)} \right)
 \label{ratio} 
@@
 where $b=1$ for $d$ odd and $b=2$ for $d$ even.
-Using this simplification an explicit calculation in Mathematica\footnote{The above simplifications greatly aid the Mathematica calculation and allow us to extend it to higher $d$.}  for $d=2, \ldots 16$ gives
+Using this simplification an explicit calculation in Mathematica\footnote{The above simplifications greatly aid the Mathematica calculation and allow us to extend it to higher $d$.}  for $d=2, \ldots, 16$ gives
 \begin{equation} 

### 2014-10-08 paper_v7.tex
some edits

@@
 The action of a given causal set (also causet) is a crucial piece of the kinematics that would be extremely useful to know and understand.
-Proposals for the action of a causal set are available \cite{Benincasa_Dowker:The_Scalar_Curvature_of_a_Causal_Set} and these hold analytically in some cases and numerically in many more.\mb{rephrase} These cases being when the causal set is embeddable in some existing spacetime.
-One can show that the action of the causal set then agrees with the Einstein-Hilbert action of the spacetime in some continuum limit.
+Proposals for the action of a causal set are available \cite{Benincasa_Dowker:The_Scalar_Curvature_of_a_Causal_Set} and these hold analytically in some cases and numerically in many more.\mb{rephrase} One can show that the action of the causal set then agrees with the Einstein-Hilbert action of the spacetime in some continuum limit.
 How this limiting procedure is carried out will be described in more detail below.
 
-It is well known that the Einstein-Hilbert action is not the full story in the continuum.
+It is well known that the Einstein-Hilbert action, $S_{EH}$, is not the full story in the continuum.
 In the presence of spacetime boundaries the gravitational action must include a boundary term $S_{GHY}$, the Gibbons-Hawking-York action, in order to yield a well-defined variational principle~\cite{Gibbons_Hawking_Boundary}.
@@
 The \textit{locally finite} condition is simply that the cardinality of any order interval is finite, that is $|I (x,y)|<\infty$.
-This condition ensures we are dealing with a discrete structure, as all the other properties of the relation would apply to an order relation between points on a continuous manifold.
+This condition ensures we are dealing with a discrete structure, as all the other properties of the relation would apply to an order relation between points on a continuous Lorentzian manifold.
 
@@
 It is called a sprinkling as one can envisage the point selection process as a `sprinkling' of points into the manifold.
-If a causet, $\mathcal{C}$, can be generated with relatively high probability by a sprinkling into the manifold, $ (M,g)$, then we say that the manifold is a good approximation for the causal set.
+If a causet, $\mathcal{C}$, can be generated with relatively high probability by a sprinkling into the manifold, $ (M,g)$, then we say that the manifold is a good approximation for the causal set, and that the causet can be \emph{faithfully embedded} in $M$.
 
@@
 The causal past and future sets $M^\pm=J^\pm (\Sigma)$ of the surface then form a partition of $M$, and $\partial M^\pm = \Sigma$.
-The Gibbons-Hawking-York boundary term for $M^\pm$ is in this case given by
+The Gibbons-Hawking-York boundary term for $M^\pm$ is in natural units given by
 \be\label{eq:GHYBT_in_continuum}
@@
 \ee
-where $l_p$ is the rationalised Planck length and $K$ is the trace of the extrinsic curvature $K_{\mu\nu}=h_{\mu}^\rho h_\nu^\sigma \nabla_\rho n_\sigma$ of $\Sigma$ defined with future-pointing timelike unit normal covector $n_{\mu}=\partial_\mu S/\sqrt{g^{\mu\nu}\partial_\mu S\partial_\nu S}$.
+where $l_p$ is the rationalised Planck length for a general dimension, $l_p=(8\pi G)^{\frac{1}{d-2}}$ (where $G$ is the gravitational constant for a general dimension) and $K$ is the trace of the extrinsic curvature $K_{\mu\nu}=h_{\mu}^\rho h_\nu^\sigma \nabla_\rho n_\sigma$ of $\Sigma$ defined with future-pointing timelike unit normal covector $n_{\mu}=\partial_\mu S/\sqrt{g^{\mu\nu}\partial_\mu S\partial_\nu S}$.
 (We work with a mostly plus convention for the metric, in which this translates into a past-pointing normal vector $n^{\mu}$).\mb{ok or more detail on conventions needed?}
@@
 Hence we propose more generally that for any causet $\mathcal C$, a \emph{Cauchy partition} is a partition that splits $\mathcal{C}$ into two parts, $\mathcal{C}^-$ and $\mathcal{C}^+=\mathcal{C}\setminus\mathcal{C}^-$ \IGNORE{($\mathcal{C}=\mathcal{C}^-\cup\mathcal{C}^+$)}, such that $\mathcal{C}^-$ is a past stem (or, equivalently, that $\mathcal{C}^+$ is a future stem and $\mathcal{C}^- = \mathcal{C}\setminus\mathcal{C}^+$).
-A sprinkling into a manifold with Cauchy surface will of course produce a causal set with Cauchy partition, but we shall refer to a partition of a causal set satisfying the above definition as Cauchy even when the causet is not obtained via sprinkling faithfully embeddable into some manifold.
+A sprinkling into a manifold with Cauchy surface will of course produce a causal set with Cauchy partition, but we shall refer to a partition of a causal set satisfying the above definition as Cauchy even when the causet is not obtained via a sprinkling into some manifold.
 
@@
 We assume an analogous definition for immediate neighbours to the $future$ of the partition, which we shall refer to as \emph{minimal elements} in $\mathcal{C}^+$.
-
-Given these definitions, we will see that the intuitive idea for the boundary term as the ``rate of change of spacetime atoms'' across a partition indeed bears out. \\
+Given these definitions, we will see that the intuitive idea for the boundary term as the ``rate of change of spacetime atoms'' across a partition indeed bears out.
+We will also see that the causal set admits other expressions for the boundary action, the one just stated being the most natural.
 %However, one does not need to know the maximal and minimal elements to calculate the boundary term.
@@
 Previously if the causet was not faithfully embeddable we could still define the causal set analogue of the surface as a partition of the causet.
-We shall see that the first issue is not a problem, and that the GHY term can be found for a future (past) boundary of $M$ (the manifold that the causet has been faithfully embedded in) by taking a difference in the number of certain elements that lie close to the future (past) bounding surface of $M$.
-Calculating the GHY term in this way must apply whether or not the causet is faithfully embeddable, but this runs us into the second issue above.
-For what boundary are we calculating our GHY term if the causet is not faithfully embeddable?
-Given a causet with maximal and/or minimal elements, certain quantities can be calculated for that causet, like the GHY term.
+The first issue is not a problem, and we will see that a causal set GHY term can be found for a future (past) boundary of $M$ (the manifold that the causet has been faithfully embedded in) via some other expression in terms of the causet.
+This expression applies to any causet, and so must apply whether or not the causet is faithfully embeddable, but this runs us into the second issue above.
+What do we mean by a boundary of a causet?
+If one is given a causet with maximal and/or minimal elements, certain quantities can be calculated for that causet, like the GHY term.
 We will not provide any physical meaning to a \emph{future} or \emph{past boundary} of a causet, and shall merely state that they are abstract objects for which certain causal set quantities can be calculated.
@@
 The situation for a null hypersurface is as follows.
-For any finite density sprinkling into a manifold with a null hypersurface, that null hypersurface will partition the causal set in such a way that agrees with the definition of a spacelike partition.
-We should really refer to such a partition as ``non-timelike", but as we are dealing with spacelike hypersurfaces for most of this paper we shall continue to call such a partition spacelike.
-To aid in the understanding of why our definition falls short in distinguishing a spacelike and null hypersurface imagine a finite density sprinkling into a patch of $2$-dimensional Minkowski spacetime.
+For any finite density sprinkling into a manifold (that admits a Cauchy surface) with a null hypersurface, $\Sigma$, that null hypersurface will partition the causal set in such a way that agrees with our definition of a Cauchy partition.\footnote{This fact means we should really refer to such a partition as ``non-timelike", but as we are dealing with Cauchy hypersurfaces for most of this paper we shall continue to call such a partition Cauchy} The reason for this is that a causal set cannot distinguish between a null and spacelike hypersurface.
+To see why this is imagine a finite density sprinkling into a patch of $2$-dimensional Minkowski spacetime.
 Figure~\ref{fig:low_density_sprinkling_with_null_and_spacelike} shows how, given a null surface that partitions the causet into two parts, we can tilt the surface slightly so that it is now spacelike, while still preserving the partition.
 Figure~\ref{fig:high_density_sprinkling_with_null_and_spacelike} shows that even as we increase the density of the sprinkling we can tilt the surface by a smaller amount such that the partition remains unchanged.
+Hence, for any finite density sprinkling one can always find some slight deviation from a partitioning null hypersurface to one that is spacelike, while still preserving the partition.
+As the partition does not change we have no way of knowing whether the hypersurface is null or spacelike.
 
@@
         \end{subfigure}
-        \caption{The partition induced by the sprinkling is unchanged after tilting the surface between null and spacelike.
+        \caption{The partition induced by the sprinkling is unchanged after tilting the surface between null (solid line) and spacelike (dashed line).
 With an even higher density, as in (b), one can simply titl the surface by some smaller amount to keep the partition unaltered.}\label{fig:animals}
@@
 For a given causet $\mathcal C$, let us define the function $\F{k}\left[\mathcal{C}\,\right]$, which counts the number of elements in $\mathcal C$ that have exactly $k$ elements in their future, and the function $\P{k}\left[\mathcal{C}\,\right]$, which counts the number of elements which have exactly $k$ elements in their past.
-An ``$F_0$ element'' is then a \emph{maximal} element, a ``$P_0$ element'' is a \emph{minimal} element, an ``$F_1$ element'' is (in some sense) next-to-maximal, and so on.\mb{Ian: your plot here?}
+An ``$F_0$ element'' is then a \emph{maximal} element and a ``$P_0$ element'' is a \emph{minimal} element.
 
@@
 \ee
-
 $V_d=\pi^{\frac{d}{2}}/\Gamma\left (\frac{d}{2}+1\right)$ denotes the volume of the unit $d$-ball and $\Gamma (t)$ is the gamma function.
 We have used $\vec{p}$ and $\vec{q}$ to stand for infinite dimensional vectors $ (p_0,...,p_m,..)$ and $ (q_0,...,q_n,..)$ respectively, where all the $p_m,\: q_n \in \mathbb{R}$.
-These vectors contain a finite number of non-zero entries, corresponding to the coefficients of the different $\F{m}$ and $\P{m}$ terms.
-We use the summation $\sum_m$ ($\sum_n$) to denote a sum that runs from $0$ to $m_{max}$ ($n_{max}$), where $m_{max}$ ($n_{max}$) is some integer, the value of which depends on the causet $\mathcal{C}^-$ ($\mathcal{C}^+$), for which all $\F{m}\left[\mathcal{C}^- \right]=0$ ($\P{m}\left[\mathcal{C}^+ \right]=0$) for any $m>m_{max}$ ($n>n_{max}$).
-A finite value for $m_{max}$ ($n_{max}$) will always exist for any finite causet, as it will not be large enough or be structured in such a way to have any $\F{m}$ ($\P{n}$) elements for $m>m_{max}$ ($n>n_{max}$).
+These vectors contain a finite number of non-zero entries, corresponding to the coefficients of the different $\F{m}$ and $\P{n}$ terms.
+We use the summation $\sum_m$ ($\sum_n$) to denote a sum that runs from $0$ to $m_{max}$ ($n_{max}$), where $m_{max}$ ($n_{max}$) is some integer, the value of which depends on the causet $\mathcal{C}^-$ ($\mathcal{C}^+$), for which all $\F{m}\left[\mathcal{C}^- \right]=0$ ($\P{n}\left[\mathcal{C}^+ \right]=0$) for any $m>m_{max}$ ($n>n_{max}$).
+A finite value for $m_{max}$ ($n_{max}$) will always exist for any finite causet, as the causet will not be large enough or be structured in such a way to have any $\F{m}$ ($\P{n}$) elements for $m>m_{max}$ ($n>n_{max}$).
 The vectors must be chosen such that they satisfy the following conditions.
@@
 where $\langle\cdot\rangle$ denotes the mean over sprinklings and we have explicitly stated the arguments of $\textbf{S}^{ (d)}_{GHY}$.
+Technically the right hand side of~\eqref{eq:mainconjecture} is equal to ${S}^{(d)}_{GHY}$ ($-{S}^{(d)}_{GHY}$) evaluated for the manifold $M^-$ ($M^+$) with future (past) bounding surface $\Sigma$.
 
@@
 \be\label{eq:min_max_simple_formula}
-S^{ (d)}_{0}\left[\mathcal{C}^-,\mathcal{C}^+\right]:=S^{ (d)}_{GHY}\left[\mathcal{C}^-,\mathcal{C}^+;\vec{p}, -\vec{p}\,\right]=\left (l/l_p\right)^{d-2}\frac{c_{d}}{2\Gamma\left (\frac{2}{d} \right)}\left ( \F{0}\left[\mathcal{C}^- \right] - \P{0}\left[\mathcal{C}^+ \right] \right)
+S^{ (d)}_{0}\left[\mathcal{C}^-,\mathcal{C}^+\right]:=S^{ (d)}_{GHY}\left[\mathcal{C}^-,\mathcal{C}^+;\vec{p}_0, -\vec{p}_0\,\right]=\left (l/l_p\right)^{d-2}\frac{c_{d}}{2\Gamma\left (\frac{2}{d} \right)}\left ( \F{0}\left[\mathcal{C}^- \right] - \P{0}\left[\mathcal{C}^+ \right] \right)
 \ee
-where $\vec{p}=\left (\left (2\Gamma\left (\frac{2}{d} \right)\right)^{-1},0,0,...\right)$.
-This expression is the easiest to employ computationally, and we shall use its random variable counterpart, $\textbf{S}^{ (d)}_{0}\left[M^-,M^+,\rho\right]:=\textbf{S}^{ (d)}_{GHY}\left[M^-,M^+,\rho;\vec{p}, -\vec{p}\,\right]$, later when we study the fluctuations of the discrete boundary actions numerically.
+where $\vec{p}_0=\left (\left (2\Gamma\left (\frac{2}{d} \right)\right)^{-1},0,0,...\right)$.
+This expression is the easiest to employ computationally, and we shall use its random variable counterpart, $\textbf{S}^{ (d)}_{0}\left[M^-,M^+,\rho\right]:=\textbf{S}^{ (d)}_{GHY}\left[M^-,M^+,\rho;\vec{p}_0, -\vec{p}_0\right]$, later when we study the fluctuations of the discrete boundary actions numerically.
 
@@
 \begin{align}\label{eq:future_past_boundary_terms1}
-S^{ (d)}_{+}[\mathcal{C}]:= & S^{ (d)}_{GHY}\left[\mathcal{C},\emptyset;\vec{p},\vec{0}\,\right] =\left (l/l_p\right)^{d-2}\frac{c_{d}}{\Gamma\left (\frac{2}{d} \right)}\left ( d\:\F{1}\left[\mathcal{C}\,\right] - \F{0}\left[\mathcal{C}\,\right] \right)
+S^{ (d)}_{+}[\mathcal{C}]:= & S^{ (d)}_{GHY}\left[\mathcal{C},\emptyset;\vec{p}_1,\vec{0}\,\right] =\left (l/l_p\right)^{d-2}\frac{c_{d}}{\Gamma\left (\frac{2}{d} \right)}\left ( d\:\F{1}\left[\mathcal{C}\,\right] - \F{0}\left[\mathcal{C}\,\right] \right)
 \\
 \label{eq:future_past_boundary_terms2}
-S^{ (d)}_{-}[\mathcal{C}]:= & S^{ (d)}_{GHY}\left[\emptyset,\mathcal{C};\vec{0},-\vec{p}\,\right] =\left (l/l_p\right)^{d-2}\frac{c_{d}}{\Gamma\left (\frac{2}{d} \right)}\left ( \P{0}\left[\mathcal{C}\,\right] - d\: \P{1}\left[\mathcal{C}\,\right] \right)
+S^{ (d)}_{-}[\mathcal{C}]:= & S^{ (d)}_{GHY}\left[\emptyset,\mathcal{C};\vec{0},-\vec{p}_1\right] =\left (l/l_p\right)^{d-2}\frac{c_{d}}{\Gamma\left (\frac{2}{d} \right)}\left ( \P{0}\left[\mathcal{C}\,\right] - d\: \P{1}\left[\mathcal{C}\,\right] \right)
 \end{align}
-where $\vec{p}=\left (-\Gamma\left (\frac{2}{d} \right)^{-1},d\:\Gamma\left (\frac{2}{d} \right)^{-1},0,0,...\right)$.
+where $\vec{p}_1=\left (-\Gamma\left (\frac{2}{d} \right)^{-1},d\:\Gamma\left (\frac{2}{d} \right)^{-1},0,0,...\right)$.
 In writing down an action for a particular causal set, $\mathcal{C}$, one would write down the Benincasa-Dowker action, $S^{ (d)}_{BD}[\mathcal{C}]$.
@@
 \ee
-where the right hand side is the spatial volume of the surface $\Sigma$. In proving~\eqref{eq:mainconjecture} we will establish the results necessary to prove this claim as well.
+where the right hand side is the spatial volume of the surface $\Sigma$, oriented such that its normal vector points backwards in time.
+In proving~\eqref{eq:mainconjecture} we will establish the results necessary to prove this claim as well.
 
@@
 \begin{align}\label{eq:future_past_spatial_volume}
-A^{ (d)}_{+}[\mathcal{C}]:=A^{ (d)}[\mathcal{C},\emptyset;\vec{p},\vec{0}]= & \left (l/l_p\right)^{d-1}\frac{b_{d}}{\Gamma\left (\frac{1}{d}\right)}\: \F{0}[\mathcal{C}]
+A^{ (d)}_{+}[\mathcal{C}]:=A^{ (d)}[\mathcal{C},\emptyset;\vec{p}_2,\vec{0}]= & \left (l/l_p\right)^{d-1}\frac{b_{d}}{\Gamma\left (\frac{1}{d}\right)}\: \F{0}[\mathcal{C}]
 \\
-A^{ (d)}_{-}[\mathcal{C}]:=A^{ (d)}[\emptyset,\mathcal{C};\vec{0},\vec{p}]= & \left (l/l_p\right)^{d-1}\frac{b_{d}}{\Gamma\left (\frac{1}{d}\right)}\: \P{0}[\mathcal{C}]
+A^{ (d)}_{-}[\mathcal{C}]:=A^{ (d)}[\emptyset,\mathcal{C};\vec{0},\vec{p}_2]= & \left (l/l_p\right)^{d-1}\frac{b_{d}}{\Gamma\left (\frac{1}{d}\right)}\: \P{0}[\mathcal{C}]
 \end{align}
-where $\vec{p}= (\Gamma\left (\frac{1}{d}\right)^{-1},0,0,...)$.
+where $\vec{p}_2= (\Gamma\left (\frac{1}{d}\right)^{-1},0,0,...)$.
 
-MB1: (too vague) During the calculation approximations will be made.
-These approximations are justified on the grounds that if certain orders are retained throughout the calculation they will in fact vanish in the limit $\rho \rightarrow \infty$
 \begin{figure}
@@
 In order to prove~\eqref{eq:mainconjecture} and~\eqref{eq:conjecture_for_area} let us first derive expressions for the mean values of $\BF{m}$ and $\BP{m}$.
-The proof will will follow almost automatically, as~\eqref{general_boundary_sum} and~\eqref{general_area_sum} are just linear combination of $\F{m}$ and $\P{m}$ terms, and thus the mean values of their random variable counterparts will just be linear combinations of $\left\langle \BF{m}\right\rangle$ and $\left\langle \BP{m}\right\rangle$ terms.
+The proof will follow almost automatically, as~\eqref{general_boundary_sum} and~\eqref{general_area_sum} are just linear combination of $\F{m}$ and $\P{m}$ terms, and thus the mean values of their random variable counterparts will just be linear combinations of $\left\langle \BF{m}\right\rangle$ and $\left\langle \BP{m}\right\rangle$ terms.
 
-For any instance of the sprinkling, the probability that a sprinkled point $p\in M$ below the surface $\Sigma$ is $F_k$ is given by the probability that $k$ points of the sprinkling lie in the region $J^{+} (p)\cap J^{-} (\Sigma)$, the intersection of the causal future of $p$ with the causal past of the surface $\Sigma$.\footnote{For notational convenience we use the symbol $x$ to refer to the causal element $x\in\mathcal C$, to its embedding in the manifold $M$, and to its coordinates in some chart on $M$.} This region will in general be the interior of some curvy $d$-dimensional cone truncated by the surface $\Sigma$, as illustrated in Figure~\ref{fig:Nmin_Nmax}.
+For any instance of the sprinkling, the probability that a sprinkled point $p\in M$ below the surface $\Sigma$ is a $F_k$ element is given by the probability that $k$ points of the sprinkling lie in the region $J^{+} (p)\cap J^{-} (\Sigma)$, the intersection of the causal future of $p$ with the causal past of the surface $\Sigma$.\footnote{For notational convenience we use the symbol $x$ to refer to the causal element $x\in\mathcal C$, to its embedding in the manifold $M$, and to its coordinates in some chart on $M$.} This region will in general be the interior of some curvy $d$-dimensional cone truncated by the surface $\Sigma$, as illustrated in Figure~\ref{fig:Nmin_Nmax}.
 We will refer to these regions as \emph{truncated cones}.
@@
 \ee
-to this event, where $V_\blacktriangledown (p)\equiv V (J^{+} (p)\cap J^{-} (\Sigma))$ is the spacetime volume of the region $J^{+} (p)\cap J^{-} (\Sigma)$.
+to this event, where $V_\blacktriangledown (p):= V (J^{+} (p)\cap J^{-} (\Sigma))$ is the spacetime volume of the region $J^{+} (p)\cap J^{-} (\Sigma)$.
 The probability of sprinkling an element into an infinitesimal four-volume $dV_p$ at $p$ is $\rho dV_p$ where $\rho=l^{-d}$ is the sprinkling density, and so the total expected number of $\F{k}$ elements below $\Sigma$ is
@@
 \ee
-where $V_\blacktriangle (p)\equiv V (J^{+} (\Sigma)\cap J^{-} (p))$.
+where $V_\blacktriangle (p):= V (J^{+} (\Sigma)\cap J^{-} (p))$.
 
@@
 The volume $V (\Sigma,x)$ then only depends on $t$.~\mb{error?}
-The integrals~\eqref{eq:nmax} and ~\eqref{eq:nmin} seem intractable as they stand, since the integration is over the entire causal past/future of the surface and the volume expressions will be complicated in the presence of curvature.
+The integrals~\eqref{eq:nmax} and~\eqref{eq:nmin} seem intractable as they stand, since the integration is over the entire causal past/future of the surface and the volume expressions will be complicated in the presence of curvature.
 However, for large $\rho$ (small $l$), the integrands $\exp (-\rho V)$ will be exponentially suppressed unless the volumes are small.
@@
 Hence, as $\rho\rightarrow\infty$, the integration ranges in \eqref{eq:nmax} and \eqref{eq:nmin} can be cut off at finite time $t=\pm\varepsilon$ with $\varepsilon$ arbitrarily small, which allows us to expand the time integration-variable about zero.
-These assumptions of course impose certain regularity conditions on the surface $\Sigma$, but so long as $\Sigma$ is compact and everywhere spacelike  then these conditions are satisfied.\mb{do we need to say more about this?
-e.g.
-can't have asymp.
-null $\Sigma$ i guess} The integrals then simplify to
+These assumptions of course impose certain regularity conditions on the surface $\Sigma$, but so long as $\Sigma$ is compact and everywhere spacelike ($\Sigma$ is a Cauchy surface) then all we require is that the radius of extrinsic curvature, $R$, satisfies $R\gg l$.
+The integrals then simplify to
 \begin{gather}\label{eq:nmax_and_eq:nmin}
@@
 \end{gather}
-where $h\equiv det\left (h_{ij} (0,\mathbf{x})\right)$ and $\dot{}\equiv \frac{\partial}{\partial t}$ and the metric determinant has been in expanded in small $t$.
-
-The only truncated cones that contribute will have small volumes, as we are restricting ourselves close to the surface.
-%\mb{this sounds circular w.r.t. last paragraph} 
-This means that the volumes can be approximated in some scheme which will be made precise below.
-Any higher order corrections to these volumes, above the order we require, will vanish in the limit of $\rho\rightarrow\infty$.
+where $h:= det\left (h_{ij} (0,\mathbf{x})\right)$ and $\dot{}\: := \frac{\partial}{\partial t}$ and the metric determinant has been in expanded in small $t$.
 
@@
 
-In order to evaluate the volume $V_\blacktriangledown (p)$ or $V_\blacktriangle (p)$ of a truncated cone we will perform a coordinate transform to Riemann Normal Coordinates (RNCs) in a neighbourhood containing the cone.
+Any truncated cone that contributes to the integral will be small, and so we can approximate the volumes, $V_\blacktriangledown (p)$ and $V_\blacktriangle (p)$, with a coordinate transformation to Riemann Normal Coordinates (RNCs) in a neighbourhood containing the cone.
+This will result in the volumes being able to be expressed as a series expansion in $t$.
+We will only need to retain up to a certain order in $t$ as higher order terms will be shown to vanish when we take the limit of $\rho\rightarrow\infty$ on the mean values of~\eqref{general_boundary_sum} and~\eqref{general_area_sum}.
+Henceforth, whenever we state that a term of some order vanishes in the limit we mean that it vanishes in the way just stated.
 The discussion for the two volume integrals is identical so we will only outline it for $V_\blacktriangle (p)$ (i.e.
@@
 In GNCs this simply corresponds to the point $p_0$ with coordinates $x_0^\mu=x^\mu (p_0)= (0,\mathbf x_p)$.
-RNCs centered at $p_0$ will be given the symbol $y^{\overline{\mu}}$.
+RNCs centered at $p_0$ will be given the symbol $y^{\overline{\mu}}$, or in terms of coordinates as $(y^{\overline{0}},\mathbf{y})=(\overline{t},\mathbf{y})$.
 %\mb{let's put all facts about $x^\mu$ and $y^{\overline\mu}$ for $p$ and $p_0$ here and expansion of det in RNC.} 
@@
 \ee
-These relations for RNC will cover all that will be needed in this discussion.
+Let us denote the time-coordinate of $p$ in GNCs by $T=x^0_p=t_p$.
+For the transformation from GNCs to RNCs at $p_0$ one finds that $A^{\overline 0}_{\;0}=1$, $A^{\overline 0}_{\;i}=0$ and $\delta_{\ibar\jbar}=A^i_{\;\ibar}A^j_{\;\jbar}h_{ij} (p_0)$, which in particular implies $y^{\overline{0}}_p=\overline{t}_p=t_p=T$.
 
@@
 \ee
-where the integration region, $\mathcal{X}_p\equiv  J^- (p)\cap J^+ (\Sigma)$ will be a complicated expression.
-As we are dealing with points close to the surface we can use the  RNCs $y^{\overline{\mu}}$, defined as above, about the point $p_0$.
-Let us denote the time-coordinate of $p$ in GNCs by $T=x^0_p=t_p$.
-For the transformation from GNCs to RNCs at $p_0$ one finds that $A^{\overline 0}_{\;0}=1$, $A^{\overline 0}_{\;i}=0$ and $\delta_{\ibar\jbar}=A^i_{\;\ibar}A^j_{\;\jbar}h_{ij} (p_0)$, which in particular implies $y^{\overline{0}}_p=\overline{t}_p=t_p=T$.
+where the integration region, $\mathcal{X}_p:= J^- (p)\cap J^+ (\Sigma)$ will be a complicated expression.
 In RNC one can expand the metric determinant in~\eqref{eq:VolumeWithNoSimplifications} to find
@@
 %The arguments of the volume function need not be changed as the volume only depends on the points in the manifold, $q_0$ and $p_0$. 
-The second term comes in at $O (T^{d+2})$ so the volume we have to calculate has reduced to
+The second integral comes in at $O (T^{d+2})$ so the volume may be written as
 \be\label{eq:VolumeToLowestOrder}
@@
 \ee
-Terms of $O (T^{d+2})$ can be retained till the end, but in the limit, $\rho\rightarrow\infty$, they vanish. This will be proved below.
+Terms of $O (T^{d+2})$ can be retained till the end, but we will prove that they vanish in the limit.
 
 This is now a simple volume integral but we need to find the boundaries of $\mathcal{X}_p$ to write down the integration limits.
-There are two parts to the boundary of the (solid) truncated cone: the lightcone emanating at $p$ and the base, given by the intersection of $\Sigma$ with the interior of the lightcone.
-First we look at the lightcone.
-Following \cite{Khetrapal_Sumati:Causal_Diamond_Volume} it can be shown that the first curvature correction to the lightcone comes in at $O (T^{d+2})$ and so can be ignored for our purposes.
-This means that the lightcone can be treated as effectively flat in RNCs and thus corresponding to the set $ (y^{\overline{1}})^2+ (y^{\overline{2}})^2+...+ (y^{\overline{d-1}})^2= T-\overline{t}$.\mb{what limit on $\bar t$?}
+There are two parts to the boundary of the (solid) truncated cone: the \emph{base}, defined as $B(p):=\Sigma\cap J^-(p)$, and the \emph{top} defined as the set corresponding to the lightcone emanating at $p$ with all points below $\Sigma$ removed, written as $T(p):=\partial J^-(p)\setminus\partial(J^-(p)\cap (J^-(\Sigma)\setminus\Sigma))$.
+Now the boundary of $\mathcal{X}_p$ is simply $\partial\mathcal{X}_p=B(p)\cup T(p)$.
+First we look at the top.
+Following \cite{Khetrapal_Sumati:Causal_Diamond_Volume} it can be shown that the first curvature correction to the lightcone volume comes in at $O (T^{d+2})$ and so can be ignored for our purposes.
+This means that the top can be treated as an effectively flat lightcone in RNCs and thus it corresponds to the set of points satisfying $ (y^{\overline{1}})^2+ (y^{\overline{2}})^2+...+ (y^{\overline{d-1}})^2= T-\overline{t}$ where the coordinates  $y^{\overline{i}}$ run from $0$ up to some maximum at the points where the flat cone meets the surface $\Sigma$.
 
@@
 Let us rewrite this equation in spherically symmetric coordinates, i.e.
-define $r=\sqrt{\delta_{\ibar\jbar}y^\ibar y^\jbar}$ and the usual angular coordinates $\phi_1,..,\phi_{d-2}$ in terms of the spatial coordinates $y^{\overline{1}} = r \cos (\phi_1),\ldots, y^{\overline{d-1}} = r \sin (\phi_1) \cdots \sin (\phi_{d-3}) \sin (\phi_{d-2})$.
+define $r:=\sqrt{\delta_{\ibar\jbar}y^\ibar y^\jbar}$ and the usual angular coordinates $\phi_1,..,\phi_{d-2}$ in terms of the spatial coordinates $y^{\overline{1}} = r \cos (\phi_1),\ldots, y^{\overline{d-1}} = r \sin (\phi_1) \cdots \sin (\phi_{d-3}) \sin (\phi_{d-2})$.
 Then
@@
 \ee
-where $r_{max} (\boldsymbol\phi)$ is the value of the radial coordinate for which the surface $\Sigma$ intersects the lightcone at angle $\boldsymbol\phi$, as shown in Figure \ref{fig:cone_plot}.
-To find this value one needs to solve
+where $r_{max} (\boldsymbol\phi)$ is the value of the radial coordinate for which the base intersects the top at an angle $\boldsymbol\phi$, as shown in Figure \ref{fig:cone_plot}.
+To find this value we equate the time coordinates of the top and base surfaces to give
 \be
@@
 \ee 
-for $r_{max} (\boldsymbol\phi)$ and take the positive solution.
+We then solve this for $r_{max} (\boldsymbol\phi)$ and take the positive solution.
 The solution can be expanded in $T$ and is simply $r_{max}=T+O (T^2)$, with angular dependent terms contributing at $O (T^2)$.
@@
     {\includegraphics[scale=0.5]{coneplot}}
-     \caption{The size of the region inside the top and bottom bounding surfaces is the volume we want to calculate.}
+     \caption{A $3$-dimensional representation of the region $\mathcal{X}_p$ in RNCs.
+The top part of $\partial\mathcal{X}_p$ can be approximated as a flat cone, and the base surface intersects the top part at a radial coordinate, $r_{max} (\boldsymbol\phi)$, which will in general be a function of the angles $\boldsymbol\phi$ (in $3$-dimensions there is one angle $\phi$).
+This function is found by projecting down from the intersection to the $\overline{t}$ plane.
+The size of the region inside the top and bottom bounding surfaces is the volume we want to calculate.}
      \label{fig:cone_plot}
@@
 Once again the higher order terms that have been ignored in both equations will be shown to vanish in the limit.
-From~\eqref{eq:K} we can substitute $K$ in for $\frac{1}{2}\frac{\dot{h}}{h}t$.
-We swap the integration variable to $t\rightarrow -t$ in the first equation and expand the $O (t^{d+1})$ part of the exponentials to give
+From~\eqref{eq:K} we can substitute $K$ in for $-\frac{1}{2}\frac{\dot{h}}{h}t$.
+We swap the integration variable from $t\rightarrow -t$ in the first equation and expand the $O (t^{d+1})$ part of the exponentials to give
 \begin{gather}\label{eq:nmax_and_eq:nmin_volume_expo_expanded}
@@
 These objects diverge as $\rho^{1-\frac{1}{d}}$ in the limit.
-The factor of $\rho$ that must be included in the formula for the surface volume is exactly the inverse of this ($\rho^{\frac{1}{d}-1}$), as can be seen from~\eqref{general_area_sum} if one substitutes in $l=\rho^{-\frac{1}{d}}$.
+The factor of $\rho$ that must be included in the formula for the surface volume is exactly the inverse of this ($\rho^{\frac{1}{d}-1}$), as can be seen from~\eqref{general_area_sum}, if one substitutes in $l=\rho^{-\frac{1}{d}}$.
 If we include this factor above, and take the limit, we see that the only remaining terms are proportional to the surface volume.
@@
 
-The form of~\eqref{eq:nmax_nmin_final} seems to suggest general series expansions of $\left\langle \BF{k}\right\rangle$ and $\left\langle \BP{k}\right\rangle$ in reducing powers of $\rho$.
+The form of~\eqref{eq:nmax_nmin_final} seems to suggest that $\left\langle \BF{k}\right\rangle$ and $\left\langle \BP{k}\right\rangle$ can be written in terms of a series expansion in reducing powers of $\rho$.
 Written in terms of the discreteness scale, $l$, it is a Laurent series starting at $l^{1-d}$.
@@
 Until now our approach has been to reconstruct the  continuum GHY term  from the information in a causal set. 
-The candidate described in the previous sections for a spatial boundary has a natural expression in a causal set  in  terms of its maximal and minimal elements which are the  analogs of a  space-like boundary.
-A suitable analog  of  a time-like boundary  term has been harder to  construct; there is no simple causal set  characterisation of such  a boundary.
-Indeed, it is difficult to give a causal set  characterisation of a  spacetime region bounded by time-like surfaces.
+The candidate described in the previous sections for a spatial boundary has a natural expression in a causal set  in  terms of its maximal and minimal elements.
+A suitable analog of a timelike boundary  term has been harder to  construct; there is no simple causal set  characterisation of such  a boundary.
+Indeed, it is difficult to give a causal set  characterisation of a spacetime region bounded by timelike surfaces.
 On the other hand, this is not true of causally defined spacetime regions since  past and future sets are defined naturally in  a causal set.
 An important example of such a region is the  Alexandrov interval which is used extensively  in calculations of causal set kinematics.
-In the continuum such a region  is bounded by null hypersurfaces and their codimension 2 joints.
+In the continuum such a region is bounded by null hypersurfaces and their codimension-2 joints.
 
-Conversely in the continuum, while the GHY term  is well defined for both a space-like and time-like boundary  (including their co-dimension 2 joints)  it is an open question what the contribution from a null  boundary  is \cite{Gibbons_Hawking_Boundary, Harris}.
+Conversely in the continuum, while the GHY term  is well defined for both a space-like and time-like boundary  (including their codimension-2 joints)  it is an open question what the contribution from a null  boundary  is \cite{Gibbons_Hawking_Boundary, Harris}.
 Apart from some sketchy suggestions in the literature \cite{neiman} there is no rigourous construction of such a term; indeed the standard variational route for finding the boundary contribution to the action appears to fail for null boundaries.
@@
 Given our ignorance in this matter,  we can ask if causal set theory  might  in fact guide us in the right direction.
-Our starting point is the $d$-dimensional  Benincasa-Dowker Action \cite{Benincasa_Dowker:The_Scalar_Curvature_of_a_Causal_Set, Dowker_Glaser:dAlembertians_for_Causal_Sets} which is constructed from the mean number  $N_i^{ (d)}$ of $i+1$-element inclusive intervals in the causal set\footnote{ $N_i^{ (d)}$ is dimension  and geometry dependent, but for our purpose it suffices to retain the dimension label.
-}
+Our starting point is the $d$-dimensional  Benincasa-Dowker Action \cite{Benincasa_Dowker:The_Scalar_Curvature_of_a_Causal_Set, Dowker_Glaser:dAlembertians_for_Causal_Sets}, $S_{BD}^{(d)}$, which is a function on a causal set constructed from the functions $N_i^{ (d)}$ which return the number of $i+1$-element inclusive intervals in the causal set, and the function $N$ which simply returns the number of causet elements.
+The Benincasa-Dowker action can be written as
 \begin{equation}
@@
 
-If $S_{BD}^{ (d)}$ were precisely the discrete analog of the Einstein Hilbert action {\it without} any boundary contributions, then in flat spacetime  $S_{BD}^{ (d)}$ should  vanish.
+As above we can define the random variable counterpart through a sprinkling, denoted by $\textbf{S}^{ (d)}_{BD}$.
+We can take the mean over sprinklings and the limit of $\rho\rightarrow\infty$.
+In this limit we find the mean is equal to $2 S_{EH}$ evaluated on the spacetime that we sprinkled into.
+This result has been proved only for certain cases.
+
+If $S_{BD}^{ (d)}$ were precisely the discrete analog of twice the Einstein Hilbert action {\it without} any boundary contributions, then in flat spacetime  $S_{BD}^{ (d)}$ should  vanish.
 However, as shown explicitly in \cite{bbdtwo} this is not the case for Alexandrov intervals in $d=2$  flat spacetime.

### 2014-10-09 lozenges/limits.tex
added notes on lozenge calculation

\documentclass{article}
\usepackage{amsmath,amssymb}
\begin{document}
We calculate the GHY boundary term for the flat causal diamond in $D$ dimensions via two limiting procedures. Take standard polar coordinates $(t,r,\boldsymbol\theta)$. The tips of the diamond are at $(\pm T,\mathbf0)$. Consider two $\epsilon$-families of surfaces:
\begin{itemize}
\item Timelike lozenges (TL): 	$$t^2 - (T-r)^2 + \epsilon^2 = 0$$
\item Spacelike lozenges  (SL):	$$r^2 - (T-|t|)^2 + \epsilon^2  = 0$$
\end{itemize}
both approaching the diamond as $\epsilon\rightarrow0$. If we calculate the boundary action associated with these surfaces for non-zero $\epsilon$ and take the limit of that expression as $\epsilon\rightarrow0$ we obtain the following results. For the TL family we get
$$
S_{GHY} \sim2 V_{D-2}T^{D-2}\left[2(D-2)-\log\left(\frac{2T}{\epsilon}\right)\right]
$$
and for the SL family we obtain
$$
S_{GHY} \sim
\begin{cases}
\displaystyle4\log\left(\frac{2T}{\epsilon}\right) \qquad&\text{for }D=2\\
\displaystyle2 V_{D-2}T^{D-2}\frac{D-1}{D-2}\qquad&\text{for }D>2
\end{cases}
$$
where $V_n$ is the volume of the $n$-sphere. So the limits agree for $D=2$ but disagree for $D>2$.  The SL limit gives a finite constant boundary term for $D>2$ whereas the TL limit gives a logarithmic divergence in all $D$.

\end{document}

### 2014-10-13 paper_v7.tex
major changes to our section

@@
 
-Consider a sufficiently well-behaved $d$-dimensional spacetime $(M,g)$ and a surface $\Sigma$ in $M$, which we shall take to be a Cauchy surface\mb{maybe too strong.
-what about boxes then?
-maybe partitioning surface?}.
-The causal past and future sets $M^\pm=J^\pm (\Sigma)$ of the surface then form a partition of $M$, and $\partial M^\pm = \Sigma$.
+\subsection{Intuition and Relation to the Continuum}
+
+Consider a sufficiently well-behaved $d$-dimensional spacetime $(M,g)$ which admits a \emph{spacelike partitioning} surface, $\Sigma$, defined as a spacelike closed compact surface in $M$, where the causal past and future sets, $M^\pm=J^\pm (\Sigma)$, of the surface then satisfy $M^+\cap M^-=\Sigma$.
+With this condition it follows that no points in $M^+\setminus\Sigma$ precede any points in $M^-\setminus\Sigma$.
+If there does exist a $p\in M^+\setminus\Sigma$ and a $q\in M^-\setminus\Sigma$ such that $p\preceq q$, then as $q\in J^-(\Sigma)$ we find that $p\in J^-(\Sigma)$, and hence it is in the set $M^-=J^-(\Sigma)$ by definition.
+The intersection, $M^+\cap M^-=\Sigma$, is now only true if $p\in\Sigma$ which it does not, and hence there cannot exist such a $p$.
+We also have that $\partial M^\pm = \Sigma$.
 The Gibbons-Hawking-York boundary term for $M^\pm$ is in natural units given by
@@
 where $l_p$ is the rationalised Planck length for a general dimension, $l_p=(8\pi G)^{\frac{1}{d-2}}$ (where $G$ is the gravitational constant for a general dimension) and $K$ is the trace of the extrinsic curvature $K_{\mu\nu}=h_{\mu}^\rho h_\nu^\sigma \nabla_\rho n_\sigma$ of $\Sigma$ defined with future-pointing timelike unit normal covector $n_{\mu}=\partial_\mu S/\sqrt{g^{\mu\nu}\partial_\mu S\partial_\nu S}$.
-(We work with a mostly plus convention for the metric, in which this translates into a past-pointing normal vector $n^{\mu}$).\mb{ok or more detail on conventions needed?}
-
+(We work with a mostly plus convention for the metric, in which this translates into a past-pointing normal vector $n^{\mu}$).
 Now observe that the integral in~\eqref{eq:GHYBT_in_continuum} can be thought of as the ``volume gradient'' across the surface $\Sigma$:
@@
 where the right hand side is the derivative of the area $\int d\Sigma$ as each point of $\Sigma$ is moved an equal distance along the unit normal vector, $n^{\mu}$ (corresponding to the rate of change of the surface volume \emph{backwards} in time, as $n^{\mu}$ is past-pointing).
-\mb{does the antichain analogy really stand?}
 
@@
 In order to make this notion precise, we need to define the causal set analogue of a hypersurface, as well as an appropriate notion of ``closeness'' to it.
-This turns out to be most natural for the analogue of what in the continuum would correspond to a Cauchy hypersurface.
 
-Given a causet that can be faithfully embedded in a manifold $M$ with Cauchy hypersurface $\Sigma$, we obtain a natural partition of the causet into two parts $\mathcal{C}^+$ and $\mathcal{C}^-$, corresponding to the sets of elements that lie in $M^+=J^+(\Sigma)$ and $M^-=J^-(\Sigma)$.
-By virtue of $\Sigma$ being Cauchy, the set $\mathcal C^-$ contains its own past, that is, any element in $\mathcal C$ that precedes at least one element in $\mathcal{C}^-$ is also in $\mathcal{C}^-$.
+Given a causet that can be faithfully embedded in a manifold $M$ with a spacelike partitioning surface $\Sigma$, we obtain a natural partition of the causet into two parts $\mathcal{C}^+$ and $\mathcal{C}^-$, corresponding to the sets of elements that lie in $M^+=J^+(\Sigma)$ and $M^-=J^-(\Sigma)$.
+By virtue of $\Sigma$ being spacelike partitioning, the set $\mathcal C^-$ contains its own past, that is, any element in $\mathcal C$ that precedes at least one element in $\mathcal{C}^-$ is also in $\mathcal{C}^-$.
 Such a set is known as a \emph{past stem}.
 Similarly, $\mathcal C^+$ is a \emph{future stem} given an analogous definition.
-These two properties capture the essential features of a partition induced by a Cauchy surface.
-Hence we propose more generally that for any causet $\mathcal C$, a \emph{Cauchy partition} is a partition that splits $\mathcal{C}$ into two parts, $\mathcal{C}^-$ and $\mathcal{C}^+=\mathcal{C}\setminus\mathcal{C}^-$ \IGNORE{($\mathcal{C}=\mathcal{C}^-\cup\mathcal{C}^+$)}, such that $\mathcal{C}^-$ is a past stem (or, equivalently, that $\mathcal{C}^+$ is a future stem and $\mathcal{C}^- = \mathcal{C}\setminus\mathcal{C}^+$).
-A sprinkling into a manifold with Cauchy surface will of course produce a causal set with Cauchy partition, but we shall refer to a partition of a causal set satisfying the above definition as Cauchy even when the causet is not obtained via a sprinkling into some manifold.
-
-Given a causal set with Cauchy partition, there follows a natural definition for the ``immediate neighbours'' in the past and future sets of the partition.
-The ``immediate neighbours in the past'' of the partition can be defined as those elements of $\mathcal{C}^-$ that have no elements in their causal future in $\mathcal{C}^-$.
-Such elements will be referred to as \textit{maximal elements} in $\mathcal{C}^-$.
-We assume an analogous definition for immediate neighbours to the $future$ of the partition, which we shall refer to as \emph{minimal elements} in $\mathcal{C}^+$.
-Given these definitions, we will see that the intuitive idea for the boundary term as the ``rate of change of spacetime atoms'' across a partition indeed bears out.
-We will also see that the causal set admits other expressions for the boundary action, the one just stated being the most natural.
-%However, one does not need to know the maximal and minimal elements to calculate the boundary term.
-It can be found from any elements that lie reasonably close to the surface (in the time coordinate on $M$) but are not necessarily maximal or minimal.
-This will be made more precise below.
-
-In the continuum the GHY term can be calculated for a spacelike hypersurface, $\Sigma$, that is either a future or past boundary of $M$.
-We wish to calculate the GHY term for future or past boundaries of causets, but there are a couple issues we must deal with.
-First, if we can faithfully embed a causet $\mathcal{C}$ into $M$ then the boundary surface of $M$ will not induce a partition of $\mathcal{C}$ as it did previously (every element of $\mathcal{C}$ will lie in the bulk of $M$), thus our previous idea for the boundary term does not work.
-Second, what does it mean to say a bounding surface of a causet if it is not faithfully embeddable in some manifold?
-Previously if the causet was not faithfully embeddable we could still define the causal set analogue of the surface as a partition of the causet.
-The first issue is not a problem, and we will see that a causal set GHY term can be found for a future (past) boundary of $M$ (the manifold that the causet has been faithfully embedded in) via some other expression in terms of the causet.
-This expression applies to any causet, and so must apply whether or not the causet is faithfully embeddable, but this runs us into the second issue above.
-What do we mean by a boundary of a causet?
-If one is given a causet with maximal and/or minimal elements, certain quantities can be calculated for that causet, like the GHY term.
-We will not provide any physical meaning to a \emph{future} or \emph{past boundary} of a causet, and shall merely state that they are abstract objects for which certain causal set quantities can be calculated.
-The reason we say that these quantities are being evaluated for the future (past) boundary of the causet is that given a causet that is faithfully embeddable in some manifold with future (past) boundary, the values of the quantities in question will agree with those of certain geometrical quantities for the continuum boundary, in some continuum limit of the causal set.
-We will henceforth mention future/past boundaries of causets only in the context of some quantity being evaluated for them.
+We propose more generally that for any causet $\mathcal C$, a \emph{spacelike partition} is a non-trivial partition that splits $\mathcal{C}$ into two parts, $\mathcal{C}^-$ and $\mathcal{C}^+=\mathcal{C}\setminus\mathcal{C}^-$ \IGNORE{($\mathcal{C}=\mathcal{C}^-\cup\mathcal{C}^+$)}, such that $\mathcal{C}^-$ is a past stem (or, equivalently, that $\mathcal{C}^+$ is a future stem and $\mathcal{C}^- = \mathcal{C}\setminus\mathcal{C}^+$).
+A sprinkling into a manifold with a surface such as $\Sigma$ will of course produce a causal set with spacelike partition, but we shall refer to a partition of a causal set satisfying the above definition as spacelike even when the causet is not obtained via a sprinkling into some manifold.
 
@@
 The situation for a null hypersurface is as follows.
-For any finite density sprinkling into a manifold (that admits a Cauchy surface) with a null hypersurface, $\Sigma$, that null hypersurface will partition the causal set in such a way that agrees with our definition of a Cauchy partition.\footnote{This fact means we should really refer to such a partition as ``non-timelike", but as we are dealing with Cauchy hypersurfaces for most of this paper we shall continue to call such a partition Cauchy} The reason for this is that a causal set cannot distinguish between a null and spacelike hypersurface.
+For any finite density sprinkling into a manifold (that admits a spacelike partitioning surface) with a null hypersurface, $\Sigma$, that null hypersurface will partition the causal set in such a way that agrees with our definition of a spacelike partition.\footnote{This fact means we should really refer to such a partition as ``non-timelike", but as we are dealing with spacelike hypersurfaces for most of this paper we shall continue to call such a partition spacelike}~The reason for this is that a causal set cannot distinguish between a null and spacelike hypersurface.
 To see why this is imagine a finite density sprinkling into a patch of $2$-dimensional Minkowski spacetime.
@@
         \end{subfigure}
-        \caption{The partition induced by the sprinkling is unchanged after tilting the surface between null (solid line) and spacelike (dashed line).
-With an even higher density, as in (b), one can simply titl the surface by some smaller amount to keep the partition unaltered.}\label{fig:animals}
+        \caption{The partition of the causet induced by the surface is unchanged after tilting the surface between null (solid line) and spacelike (dashed line).
+With an even higher density, as in (b), one can simply tilt the surface by some smaller amount to keep the partition unaltered.}\label{fig:animals}
 \end{figure}
+
+Given a causal set with a spacelike partition, there follows a natural definition for the ``immediate neighbours'' in the past and future sets of the partition.
+The ``immediate neighbours in the past'' of the partition can be defined as those elements of $\mathcal{C}^-$ that have no elements in their causal future in $\mathcal{C}^-$.
+Such elements will be referred to as \textit{maximal elements} in $\mathcal{C}^-$.
+We assume an analogous definition for immediate neighbours to the $future$ of the partition, which we shall refer to as \emph{minimal elements} in $\mathcal{C}^+$.
+Given these definitions, we will see that the intuitive idea for the boundary term as the ``rate of change of spacetime atoms'' across a partition indeed bears out.
+
+In the continuum the GHY term is usually calculated for the future or past bounding surfaces of a manifold, not for a partitioning surface.
+We wish to calculate the GHY term for future or past boundaries of causets, but there are a couple issues we must deal with.
+First, if we can faithfully embed a causet $\mathcal{C}$ into $M$ then the boundary surface of $M$ will induce a trivial partition of $\mathcal{C}$ (every element of $\mathcal{C}$ will lie in the bulk of $M$), thus the previous notion of the boundary term as a rate of change of spacetime atoms across a partition does not work.
+Second, what does it mean to say a bounding surface of a causet if it is not faithfully embeddable in some manifold?
+Previously if the causet was not faithfully embeddable we could still define the causal set analogue of the surface as a partition of the causet, but now we have that either $\mathcal{C}^+$ or $\mathcal{C}^-$ will be the empty set.
+The first issue is not a problem, and we will see that a causal set GHY term can be found for a future (past) boundary of $M$ (the manifold that the causet has been faithfully embedded in) via some other expression in terms of the causet.
+This expression applies to any causet, and so must apply whether or not the causet is faithfully embeddable, but this runs us into the second issue above.
+What do we mean by a boundary of a causet?
+If one is given a causet with maximal and/or minimal elements, certain quantities can be calculated for that causet, like the GHY term we will propose in the next section.
+We will not provide any physical meaning to a \emph{future} or \emph{past boundary} of a causet, and shall merely state that they are abstract objects for which certain causal set quantities can be calculated.
+The reason we say that these quantities are being evaluated for the future (past) boundary of the causet is that given a causet that is faithfully embeddable in some manifold with future (past) boundary, the values of the quantities in question will agree with those of certain geometrical quantities for the continuum boundary, in some continuum limit of the causal set.
+We will henceforth mention future/past boundaries of causets only in the context of some quantity being evaluated for them.
+
+\subsection{The Boundary Action Family}
 
@@
 We use the summation $\sum_m$ ($\sum_n$) to denote a sum that runs from $0$ to $m_{max}$ ($n_{max}$), where $m_{max}$ ($n_{max}$) is some integer, the value of which depends on the causet $\mathcal{C}^-$ ($\mathcal{C}^+$), for which all $\F{m}\left[\mathcal{C}^- \right]=0$ ($\P{n}\left[\mathcal{C}^+ \right]=0$) for any $m>m_{max}$ ($n>n_{max}$).
-A finite value for $m_{max}$ ($n_{max}$) will always exist for any finite causet, as the causet will not be large enough or be structured in such a way to have any $\F{m}$ ($\P{n}$) elements for $m>m_{max}$ ($n>n_{max}$).
+Finite values for $m_{max}$ and $n_{max}$ will always exist for any finite causet.
 The vectors must be chosen such that they satisfy the following conditions.
@@
 \end{align}
-The aforementioned family simply corresponds to all the different vectors, $\vec{p}$ and $\vec{q}$, that satisfy the above criteria.
+Any member of the aforementioned family simply corresponds to a choice of the vectors $\vec{p}$ and $\vec{q}$ satisfying the above criteria.
 
@@
 These random variables can be substituted into~\eqref{general_boundary_sum} for $\F{k}$ and $\P{k}$ to give the family of random variables, $\textbf{S}^{ (d)}_{GHY}$.
-These random variables will be functions of $\rho$ and the manifold, $M^{\pm}$, where the dependence of $\BF{k}$ ($\BP{k}$) on the manifold is restricted to $M^-$ ($M^+$).
+These random variables will be functions of $\rho$ and the two parts of the partitioned manifold, $M^{\pm}$, where the dependence of $\BF{k}$ ($\BP{k}$) on the manifold is restricted to $M^-$ ($M^+$).
 We claim that in the limit of infinite density the expectation value, in the sprinkling process, of $\textbf{S}^{ (d)}_{GHY}$ (for any vectors $\vec{p}$ and $\vec{q}$ satisfying~\eqref{coefficient_relation1} and ~\eqref{coefficient_relation1}) tends to the continuum GHY boundary term for a surface $\Sigma$ that partitions the manifold into $M^+$ and $M^-$:
@@
 There is a huge amount of freedom that remains in the choice of $\vec{p}$ and $\vec{q}$.
-One can see that at least two terms are necessary in order to satisfy~\eqref{coefficient_relation1} and~\eqref{coefficient_relation2}.
-For any two terms the coefficients will be unique, and for more than two terms this uniqueness is lost.
+One can see that at least two non-zero entries in $\vec{p}$ or $\vec{q}$ are necessary in order to satisfy~\eqref{coefficient_relation1} and~\eqref{coefficient_relation2}.
+The values of these non-zero entries will unique, but if more than two entries are non-zero this uniqueness is lost.
 In a finite difference method one can form a discrete derivative in one direction by taking a difference of two values at different points along that direction, and dividing by the distance between them.
@@
 The boundary action (in the continuum) is a derivative in one direction, the normal direction (as in~\eqref{eq:normal_deriv_boundary}), and so a finite difference method should be able to approximate this with just two terms.
-This agrees with the fact that we are able write the discrete boundary action with only two terms.
+This agrees with the fact that we are able write the discrete boundary action with only two non-zero entries.
 
-There are a few, two-term, members of the family that are worth mentioning, as they have simple forms that make them good candidates for computations.
-Given a causet, $\mathcal{C}$, with a spacelike partition so that $\mathcal C = \mathcal C^+ \cup\, \mathcal C^-$, we define the function $S^{ (d)}_0$ as a particular member of the family in~\eqref{general_boundary_sum}:
+We will now mention a few of the simplest looking members of the family that arise from having only two non-zero entries in the vectors $\vec{p}$ and $\vec{q}$.
+Given a causet, $\mathcal{C}$, with a spacelike partition into $\mathcal C^+$ and $\mathcal C^-$ we define the function $S^{ (d)}_0$ as a particular member of the family in~\eqref{general_boundary_sum}:
 \be\label{eq:min_max_simple_formula}
@@
 where $\vec{p}_0=\left (\left (2\Gamma\left (\frac{2}{d} \right)\right)^{-1},0,0,...\right)$.
+This is precisely the rate of change of spacetime atoms previously mentioned, as it is just the difference in minimal elements of $\mathcal{C}^+$ and maximal elements of $\mathcal{C}^-$.
 This expression is the easiest to employ computationally, and we shall use its random variable counterpart, $\textbf{S}^{ (d)}_{0}\left[M^-,M^+,\rho\right]:=\textbf{S}^{ (d)}_{GHY}\left[M^-,M^+,\rho;\vec{p}_0, -\vec{p}_0\right]$, later when we study the fluctuations of the discrete boundary actions numerically.
@@
 In writing down an action for a particular causal set, $\mathcal{C}$, one would write down the Benincasa-Dowker action, $S^{ (d)}_{BD}[\mathcal{C}]$.
-To complete the story we propose that one must include the boundary terms $S^{ (d)}_{+}[\mathcal{C}]$ and $S^{ (d)}_{-}[\mathcal{C}]$.
-Explicitly, the total action for a causal set should be given by:
+In addition to this we propose that one could include the boundary terms $S^{ (d)}_{+}[\mathcal{C}]$ and $S^{ (d)}_{-}[\mathcal{C}]$.
+Explicitly, the total action for a causal set is then given by:
 \be\label{eq:total_causet_action}
@@
 There may be contributions from the causal set analogue of timelike boundaries but some results suggest that these are included in $S^{ (d)}_{BD}$.
-We will see below that $S^{ (d)}_{BD}$ also includes a contribution from a different kind of boundary.
+
+\subsection{The Surface Volume Family}
 
@@
 From this relation we can see that only one term is necessary to determine an expression for the discrete surface volume, and that when more than one term is added the coefficients will cease to be unique.
-Once again we can define a family of random variables, $\textbf{A}^{ (d)}$ which will depend upon $M^{\pm}$ and $\rho$.
+Once again we can define a family of random variables, $\textbf{A}^{ (d)}$, which will depend upon the parts of the partitioned manifold, $M^{\pm}$, and the sprinkling density, $\rho$.
 We make another claim that, in the limit of infinite density, the expectation value of $\textbf{A}^{ (d)}$ in the sprinkling process tends to the spatial volume of the surface $\Sigma$.
@@
 \ee
-where the right hand side is the spatial volume of the surface $\Sigma$, oriented such that its normal vector points backwards in time.
-In proving~\eqref{eq:mainconjecture} we will establish the results necessary to prove this claim as well.
+where the right hand side is the spatial volume of the surface $\Sigma$. In proving~\eqref{eq:mainconjecture} we will establish the results necessary to prove this claim as well.
 
@@
 
-\begin{figure}
-  \centering
-    {\includegraphics[width=\textwidth]{minmaxplot}}
-     \caption{An illustration of the idea.
-Pictured here is a causal set obtained by a sprinkling into a spacetime partitioned by a spacelike hypersurface.
-The minimal and maximal points about the surface are highlighted and the shaded regions illustrate the regions whose volumes $V_\blacktriangle$ and $V_\blacktriangledown$ are needed in the proof.}
-     \label{fig:Nmin_Nmax}
-*\end{figure}
+%\begin{figure}
+%  \centering
+%    {\includegraphics[width=\textwidth]{minmaxplot}}
+%     \caption{An illustration of the idea.
+Pictured %here is a causal set obtained by a sprinkling into a %spacetime partitioned by a spacelike hypersurface.
+%The minimal and maximal points about the surface are %highlighted and the shaded regions illustrate the %regions whose volumes $V_\blacktriangle$ and %$V_\blacktriangledown$ are needed in the proof.}
+%     \label{fig:Nmin_Nmax}
+%*\end{figure}
 
@@
 
-\subsection{Poisson Sprinklings for $\left\langle \BF{m}\right\rangle$ and $\left\langle \BP{m}\right\rangle$}
+\subsection{$\left\langle \BF{m}\right\rangle$ and $\left\langle \BP{m}\right\rangle$ for Poisson Sprinklings}
 
 In order to prove~\eqref{eq:mainconjecture} and~\eqref{eq:conjecture_for_area} let us first derive expressions for the mean values of $\BF{m}$ and $\BP{m}$.
-The proof will follow almost automatically, as~\eqref{general_boundary_sum} and~\eqref{general_area_sum} are just linear combination of $\F{m}$ and $\P{m}$ terms, and thus the mean values of their random variable counterparts will just be linear combinations of $\left\langle \BF{m}\right\rangle$ and $\left\langle \BP{m}\right\rangle$ terms.
+The proof will follow almost automatically, as~\eqref{general_boundary_sum} and~\eqref{general_area_sum} are just linear combinations of $\F{m}$ and $\P{m}$ terms, and thus the mean values of the random variable counterparts for~\eqref{general_boundary_sum} and~\eqref{general_area_sum} will just be linear combinations of $\left\langle \BF{m}\right\rangle$ and $\left\langle \BP{m}\right\rangle$ terms.
 
-For any instance of the sprinkling, the probability that a sprinkled point $p\in M$ below the surface $\Sigma$ is a $F_k$ element is given by the probability that $k$ points of the sprinkling lie in the region $J^{+} (p)\cap J^{-} (\Sigma)$, the intersection of the causal future of $p$ with the causal past of the surface $\Sigma$.\footnote{For notational convenience we use the symbol $x$ to refer to the causal element $x\in\mathcal C$, to its embedding in the manifold $M$, and to its coordinates in some chart on $M$.} This region will in general be the interior of some curvy $d$-dimensional cone truncated by the surface $\Sigma$, as illustrated in Figure~\ref{fig:Nmin_Nmax}.
+For any instance of the sprinkling, the probability that a sprinkled point $p\in M$ below the surface $\Sigma$ is an $F_k$ element is given by the probability that $k$ points of the sprinkling lie in the region $J^{+} (p)\cap J^{-} (\Sigma)$, the intersection of the causal future of $p$ with the causal past of the surface $\Sigma$.\footnote{For notational convenience we use the symbol $x$ to refer to the causal element $x\in\mathcal C$, to its embedding in the manifold $M$, and to its coordinates in some chart on $M$.} This region, for some general spacetime, will be the interior of some curvy $d$-dimensional lightcone truncated by the surface $\Sigma$.
 We will refer to these regions as \emph{truncated cones}.
@@
 
-Consider a set of synchronous or Gaussian Normal Coordinates (GNC) $x^\mu= (t,\mathbf x)$ adapted to $\Sigma$ such that in a neighbourhood $U_\Sigma$ of $\Sigma$ the line element is
+Consider a set of synchronous or Gaussian Normal Coordinates (GNC), $x^\mu= (t,\mathbf x)$, adapted to $\Sigma$ such that in a neighbourhood $U_\Sigma$ of $\Sigma$ the line element is
 \be
@@
 Hence, as $\rho\rightarrow\infty$, the integration ranges in \eqref{eq:nmax} and \eqref{eq:nmin} can be cut off at finite time $t=\pm\varepsilon$ with $\varepsilon$ arbitrarily small, which allows us to expand the time integration-variable about zero.
-These assumptions of course impose certain regularity conditions on the surface $\Sigma$, but so long as $\Sigma$ is compact and everywhere spacelike ($\Sigma$ is a Cauchy surface) then all we require is that the radius of extrinsic curvature, $R$, satisfies $R\gg l$.
+These assumptions of course impose certain regularity conditions on the surface $\Sigma$, but so long as $\Sigma$ is a closed compact manifold and is everywhere spacelike then all we require is that the radius of extrinsic curvature, $R$, satisfies $R\gg l$.
 The integrals then simplify to
@@
 
-Any truncated cone that contributes to the integral will be small, and so we can approximate the volumes, $V_\blacktriangledown (p)$ and $V_\blacktriangle (p)$, with a coordinate transformation to Riemann Normal Coordinates (RNCs) in a neighbourhood containing the cone.
-This will result in the volumes being able to be expressed as a series expansion in $t$.
-We will only need to retain up to a certain order in $t$ as higher order terms will be shown to vanish when we take the limit of $\rho\rightarrow\infty$ on the mean values of~\eqref{general_boundary_sum} and~\eqref{general_area_sum}.
+In order to evaluate the volume $V_\blacktriangledown (p)$ or $V_\blacktriangle (p)$ of a truncated cone we will perform a coordinate transformation to Riemann Normal Coordinates (RNCs) in a neighbourhood containing the cone.
+The volumes can be expressed by series expansions in $t$.
+We will only need to retain up to order $O(t^{d+1})$ as higher order terms will be shown to vanish when we take the limit of $\rho\rightarrow\infty$ on the mean values of~\eqref{general_boundary_sum} and~\eqref{general_area_sum}.
 Henceforth, whenever we state that a term of some order vanishes in the limit we mean that it vanishes in the way just stated.
@@
 \ee
-where $R_{\overline{\mu}\overline{\nu}} (p_0)$ is the Ricci tensor in RNC evaluated at $p_0$.\mb{justify beforehand which orders to keep?}
-%The arguments of the volume function need not be changed as the volume only depends on the points in the manifold, $q_0$ and $p_0$. 
-The second integral comes in at $O (T^{d+2})$ so the volume may be written as
+where $R_{\overline{\mu}\overline{\nu}} (p_0)$ is the Ricci tensor in RNC evaluated at $p_0$. The second integral comes in at $O (T^{d+2})$ so the volume may be written as
 \be\label{eq:VolumeToLowestOrder}
@@
 The other terms all have negative powers of $\rho$ and therefore vanish as $\rho\rightarrow\infty$.
-This means that if we take the mean of $\textbf{A}^{ (d)}$, in the limit of $\rho\rightarrow\infty$, and substitute in the above expressions, we will be left with something proportional to the surface volume.
-Then it is easy to see that in order to get the surface volume exactly one must force $\vec{p}$ and $\vec{q}$ to satisfy~\eqref{area_coefficient_relation}.
+We can take the mean and limit of $\textbf{A}^{ (d)}$ and substitute in the above expressions to find
+\begin{gather}\label{eq:area_final_proof}
+\begin{aligned}
+\lim_{\rho\rightarrow\infty} & \left\langle \textbf{A}^{ (d)}\right\rangle = \lim_{\rho\rightarrow\infty} \frac{\rho^{\frac{1}{d}-1}}{l_p^{d-1}} b_{d}\left (\sum_m p_m \left\langle\textbf{F}_m\right\rangle  + \sum_n q_n \left\langle\textbf{P}_n\right\rangle\right)
+\\
+= & \lim_{\rho\rightarrow\infty}\Bigg[ \frac{1}{l_p^{d-1}} \left(\sum_m p_m \frac{\Gamma\left (\frac{1}{d}+m \right)}{m!}  + \sum_n q_n\frac{\Gamma\left (\frac{1}{d}+n \right)}{n!} \right) \int_{\Sigma}d^{d-1}x\: \sqrt{h}+O(\rho^{-\frac{1}{d}})\Bigg]
+\\
+= & \frac{1}{l_p^{d-1}}\int_{\Sigma}d^{d-1}x\: \sqrt{h}
+\end{aligned}
+\end{gather}
+Where the last line is only true if the vectors, $\vec{p}$ and $\vec{q}$ satisfy~\eqref{area_coefficient_relation}.
 This concludes the proof of the spatial volume claim,~\eqref{eq:conjecture_for_area}.
 
-If we take the mean of $\textbf{S}^{ (d)}_{GHY}$ and the limit $\rho\rightarrow\infty$, we can substitute in the above expressions to find that the only terms left that have non-negative powers of $\rho$ (and therefore non-vanishing in the limit) are that of the spatial volume and boundary term.
-The term proportional to the spatial volume is divergent, so the coefficients, $\vec{p}$ and $\vec{q}$, must satisfy~\eqref{coefficient_relation1} to make this term vanish.
-The other condition,~\eqref{coefficient_relation2}, ensures that what remains is equal to the GHY boundary term.
-Thus, the proof of~\eqref{eq:mainconjecture} is complete.
+We can also take the mean and limit of $\textbf{S}^{ (d)}_{GHY}$ to find
+\begin{gather}\label{eq:boundary_final_proof}
+\begin{aligned}
+\lim_{\rho\rightarrow\infty} & \left\langle \textbf{S}^{ (d)}_{GHY}\right\rangle = \lim_{\rho\rightarrow\infty} \frac{\rho^{\frac{2}{d}-1}}{l_p^{d-2}} c_{d}\left (\sum_m p_m \left\langle\textbf{F}_m\right\rangle  + \sum_n q_n \left\langle\textbf{P}_n\right\rangle\right)
+\\
+= & \lim_{\rho\rightarrow\infty}\Bigg[\frac{\rho^{\frac{1}{d}}}{l_p^{d-1}}\frac{c_d}{b_d}\left(\sum_m p_m \frac{\Gamma\left (\frac{1}{d}+m \right)}{m!}  + \sum_n q_n\frac{\Gamma\left (\frac{1}{d}+n \right)}{n!} \right) \int_{\Sigma}d^{d-1}x\: \sqrt{h}
+\\
+& +\frac{1}{l_p^{d-2}}\left(\sum_m\frac{\Gamma\left (\frac{2}{d}+m \right)}{m!}  - \sum_n q_n\frac{\Gamma\left (\frac{2}{d}+n \right)}{n!} \right) \int_{\Sigma}d^{d-1}x\: \sqrt{h}K + O(\rho^{-\frac{1}{d}})\Bigg]
+\\
+& = \frac{1}{l_p^{d-2}}\int_{\Sigma}d^{d-1}x\: \sqrt{h}K
+\end{aligned}
+\end{gather}
+Where the last line is true only if the vectors $\vec{p}$ and $\vec{q}$ satisfy~\eqref{coefficient_relation1} and~\eqref{coefficient_relation2}. The proof of~\eqref{eq:mainconjecture} is now complete.
 
@@
 The higher order terms, in $l$, will possibly be proportional to higher order normal derivatives of the surface volume.
-If this were true then one could find causal set versions for normal derivatives of the surface volume to any order.
+If this were true then one could find causal set versions for the normal derivatives of the surface volume to any order.
 The exact form of the constant in the next term of the series is unknown and more work must be done to determine a general trend for the constants, and hence find the general series.

### 2014-10-13 lozenges/limits.tex
corrected mistakes in lozenge calculation
there was a missing constant factor in the TL calculation

@@
 $$
-S_{GHY} \sim2 V_{D-2}T^{D-2}\left[2(D-2)-\log\left(\frac{2T}{\epsilon}\right)\right]
+S_{GHY} \sim
+\begin{cases}
+\displaystyle4\log\left(\frac{2T}{\epsilon}\right) \qquad&\text{for }D=2\\
+\displaystyle2 V_{D-2}T^{D-2}\left[-\log\left(\frac{2T}{\epsilon}\right)+\frac{D-1}{D-2}+H_{D-3}\right]\qquad&\text{for }D>2
+\end{cases}
 $$
+where $H_n$ is the $n^{th}$harmonic number
 and for the SL family we obtain

### 2014-10-13 paper_v7.tex
some edits to paper

@@
 
-Consider a sufficiently well-behaved $d$-dimensional spacetime $(M,g)$ which admits a \emph{spacelike partitioning} surface, $\Sigma$, defined as a spacelike closed compact surface in $M$, where the causal past and future sets, $M^\pm=J^\pm (\Sigma)$, of the surface then satisfy $M^+\cap M^-=\Sigma$.
+Consider a sufficiently well-behaved $d$-dimensional spacetime $(M,g)$, of finite volume, which admits a \emph{spacelike partitioning} surface, $\Sigma$, defined as a spacelike closed compact surface in $M$, where the causal past and future sets, $M^\pm=J^\pm (\Sigma)$, of the surface then satisfy $M^+\cap M^-=\Sigma$.
 With this condition it follows that no points in $M^+\setminus\Sigma$ precede any points in $M^-\setminus\Sigma$.
@@
 
-We can also take the mean and limit of $\textbf{S}^{ (d)}_{GHY}$ to find
+We can also take the mean and limit of $\textbf{S}^{ (d)}_{GHY}$ and subsitute in~\eqref{eq:nmax_nmin_final} to find
 \begin{gather}\label{eq:boundary_final_proof}
@@
 \\
-& +\frac{1}{l_p^{d-2}}\left(\sum_m\frac{\Gamma\left (\frac{2}{d}+m \right)}{m!}  - \sum_n q_n\frac{\Gamma\left (\frac{2}{d}+n \right)}{n!} \right) \int_{\Sigma}d^{d-1}x\: \sqrt{h}K + O(\rho^{-\frac{1}{d}})\Bigg]
++ &\frac{1}{l_p^{d-2}}\left(\sum_m\frac{\Gamma\left (\frac{2}{d}+m \right)}{m!}  - \sum_n q_n\frac{\Gamma\left (\frac{2}{d}+n \right)}{n!} \right) \int_{\Sigma}d^{d-1}x\: \sqrt{h}K + O(\rho^{-\frac{1}{d}})\Bigg]
 \\
-& = \frac{1}{l_p^{d-2}}\int_{\Sigma}d^{d-1}x\: \sqrt{h}K
+= & \frac{1}{l_p^{d-2}}\int_{\Sigma}d^{d-1}x\: \sqrt{h}K
 \end{aligned}

### 2014-10-13 paper_v7.tex
some edits to paper

@@
 
-Consider a sufficiently well-behaved $d$-dimensional spacetime $(M,g)$, of finite volume, which admits a \emph{spacelike partitioning} surface, $\Sigma$, defined as a spacelike closed compact surface in $M$, where the causal past and future sets, $M^\pm=J^\pm (\Sigma)$, of the surface then satisfy $M^+\cap M^-=\Sigma$.
+Consider a sufficiently well-behaved $d$-dimensional spacetime $(M,g)$ of finite volume, which admits a \emph{spacelike partitioning} surface, $\Sigma$, defined as a spacelike closed compact surface in $M$, where the causal past and future sets, $M^\pm=J^\pm (\Sigma)$, of the surface then satisfy $M^+\cap M^-=\Sigma$.
 With this condition it follows that no points in $M^+\setminus\Sigma$ precede any points in $M^-\setminus\Sigma$.

### 2014-10-13 paper_v7.tex
edits in section 1-2

@@
 In furthering causal set theory it is crucial that we understand the kinematics of the theory.
-The action of a given causal set (also causet) is a crucial piece of the kinematics that would be extremely useful to know and understand.
-Proposals for the action of a causal set are available \cite{Benincasa_Dowker:The_Scalar_Curvature_of_a_Causal_Set} and these hold analytically in some cases and numerically in many more.\mb{rephrase} One can show that the action of the causal set then agrees with the Einstein-Hilbert action of the spacetime in some continuum limit.
-How this limiting procedure is carried out will be described in more detail below.
+One avenue in this direction is the attempt to define a discrete counterpart of the gravitational action $S[\mathcal C]$ associated with a causal set $\mathcal C$, which furnishes the weight $e^{iS[\mathcal C]}$ in the graviational sum over histories.
+A proposal for the action of a causal set have recently been made \cite{Benincasa_Dowker:The_Scalar_Curvature_of_a_Causal_Set}, and for a number of special cases, it has been shown that for causal sets that are well-approximated by continuum spacetimes (in a sense that will be made clear below), the proposed action agrees with the Einstein-Hilbert action of the spacetime in the appropriate continuum limit.\mb{given what conditions?}
 
-It is well known that the Einstein-Hilbert action, $S_{EH}$, is not the full story in the continuum.
+In the continuum, it is well known that the Einstein-Hilbert action, $S_{EH}$, is not the full story.
 In the presence of spacetime boundaries the gravitational action must include a boundary term $S_{GHY}$, the Gibbons-Hawking-York action, in order to yield a well-defined variational principle~\cite{Gibbons_Hawking_Boundary}.
@@
 In order to say we have a causal set analogue for a continuous expression we need a well defined procedure for relating the discrete theory to the continuum.
-This procedure, or tool, is called a \textit{Poisson sprinkling}, or just a \textit{sprinkling}.
-It is a Poisson process which provides a way to generate a causet from a $d$-dimensional Lorentzian manifold $ (M,g)$ by selecting certain points in $M$ to be the elements of $\mathcal{C}$, with an order relation given by the causal order of the manifold.
-The number of points chosen in a region of spacetime volume, $V$, is a Poisson random variable.
+This procedure is called a \textit{sprinkling}.
+A sprinkling is a Poisson process which provides a way to generate a causet from a $d$-dimensional Lorentzian manifold $ (M,g)$ by selecting certain points in $M$ to be the elements of $\mathcal{C}$, with an order relation given by the causal order of the manifold.\mb{certain points?} The number of points chosen in a region of spacetime volume, $V$, is a Poisson random variable.
 This means that the expected number of points in some region will be $\rho V$, where $\rho$ is the density of the sprinkling.
@@
 
-Consider a sufficiently well-behaved $d$-dimensional spacetime $(M,g)$ of finite volume, which admits a \emph{spacelike partitioning} surface, $\Sigma$, defined as a spacelike closed compact surface in $M$, where the causal past and future sets, $M^\pm=J^\pm (\Sigma)$, of the surface then satisfy $M^+\cap M^-=\Sigma$.
-With this condition it follows that no points in $M^+\setminus\Sigma$ precede any points in $M^-\setminus\Sigma$.
-If there does exist a $p\in M^+\setminus\Sigma$ and a $q\in M^-\setminus\Sigma$ such that $p\preceq q$, then as $q\in J^-(\Sigma)$ we find that $p\in J^-(\Sigma)$, and hence it is in the set $M^-=J^-(\Sigma)$ by definition.
-The intersection, $M^+\cap M^-=\Sigma$, is now only true if $p\in\Sigma$ which it does not, and hence there cannot exist such a $p$.
-We also have that $\partial M^\pm = \Sigma$.
-The Gibbons-Hawking-York boundary term for $M^\pm$ is in natural units given by
+In order to give intuition to the ideas laid out in this paper it will be most convenient to consider spacetimes with boundary obtained by partitioning a manifold (without boundary) into two parts as follows.
+Let $(M,g)$ be a sufficiently well-behaved $d$-dimensional Lorentzian manifold.
+We assume that $(M,g)$ has finite volume and admits a \emph{spacelike partitioning} surface, $\Sigma$, defined as a spacelike closed compact surface in $M$, such that the causal past and future sets, $M^\pm=J^\pm (\Sigma)$ satisfy $M^+\cap M^-=\Sigma$.
+It then follows that $\partial M^\pm = \Sigma$, and further that no points in $M^+\setminus\Sigma$ precede any points in $M^-\setminus\Sigma$.~\!\footnote{If there did exist a $p\in M^+\setminus\Sigma$ and $q\in M^-\setminus\Sigma$ such that $p\preceq q$, then we would have $(q\in J^-(\Sigma) \text{ and } p\preceq q)\implies p\in J^-(\Sigma)\implies p\in M^-$, which would contradict $M^+\cap M^-=\Sigma$.} The Gibbons-Hawking-York boundary term for $M^\pm$ is then given by
 \be\label{eq:GHYBT_in_continuum}
@@
 \ee
-where $l_p$ is the rationalised Planck length for a general dimension, $l_p=(8\pi G)^{\frac{1}{d-2}}$ (where $G$ is the gravitational constant for a general dimension) and $K$ is the trace of the extrinsic curvature $K_{\mu\nu}=h_{\mu}^\rho h_\nu^\sigma \nabla_\rho n_\sigma$ of $\Sigma$ defined with future-pointing timelike unit normal covector $n_{\mu}=\partial_\mu S/\sqrt{g^{\mu\nu}\partial_\mu S\partial_\nu S}$.
-(We work with a mostly plus convention for the metric, in which this translates into a past-pointing normal vector $n^{\mu}$).
-Now observe that the integral in~\eqref{eq:GHYBT_in_continuum} can be thought of as the ``volume gradient'' across the surface $\Sigma$:
+where $l_p$ is the rationalised Planck length $l_p=(8\pi G)^{\frac{1}{d-2}}$ ($G$ denoting the gravitational constant in $d$ dimensions) and where $K$ is the trace of the extrinsic curvature $K_{\mu\nu}=h_{\mu}^\rho h_\nu^\sigma \nabla_\rho n_\sigma$ of $\Sigma$, defined here with future-pointing timelike unit normal covector $n_{\mu}=\partial_\mu S/\sqrt{g^{\mu\nu}\partial_\mu S\partial_\nu S}$.
+(We will work with a mostly plus convention for the metric, in which this translates into a past-pointing normal vector $n^{\mu}$).
+
+Observe now that the integral in~\eqref{eq:GHYBT_in_continuum} can be thought of as the ``volume gradient'' across the surface $\Sigma$:
 \be\label{eq:normal_deriv_boundary}
@@
 
+This observation suggests a natural candidate for the discrete analogue of the boundary action associated with the surface $\Sigma$, for a causet that can be faithfully embedded in $M$.
 Spacetime volume in the causal set is obtained simply by counting the number of elements, and hence the volume gradient intuitively corresponds to the difference in the number of ``immediate neighbours'' to the future and to the past of a hypersurface.
-In order to make this notion precise, we need to define the causal set analogue of a hypersurface, as well as an appropriate notion of ``closeness'' to it.
+In order to make this notion precise, we need to define the discrete analogue of the hypersurface $\Sigma$, and also what it means for a causal set element to be an ``immediate neighbour'' to $\Sigma$.
 
-Given a causet that can be faithfully embedded in a manifold $M$ with a spacelike partitioning surface $\Sigma$, we obtain a natural partition of the causet into two parts $\mathcal{C}^+$ and $\mathcal{C}^-$, corresponding to the sets of elements that lie in $M^+=J^+(\Sigma)$ and $M^-=J^-(\Sigma)$.
+Given a causet $\mathcal C$ that can be faithfully embedded in a manifold $M$ with a spacelike partitioning surface $\Sigma$, we obtain a natural partition of the causet into two parts $\mathcal{C}^+$ and $\mathcal{C}^-$, corresponding to the sets of elements that lie in $M^+=J^+(\Sigma)$ and $M^-=J^-(\Sigma)$.
 By virtue of $\Sigma$ being spacelike partitioning, the set $\mathcal C^-$ contains its own past, that is, any element in $\mathcal C$ that precedes at least one element in $\mathcal{C}^-$ is also in $\mathcal{C}^-$.
-Such a set is known as a \emph{past stem}.
-Similarly, $\mathcal C^+$ is a \emph{future stem} given an analogous definition.
+A set with this property is known as a \emph{past stem}.
+Similarly, $\mathcal C^+$ is a \emph{future stem} given the analogous definition.
 We propose more generally that for any causet $\mathcal C$, a \emph{spacelike partition} is a non-trivial partition that splits $\mathcal{C}$ into two parts, $\mathcal{C}^-$ and $\mathcal{C}^+=\mathcal{C}\setminus\mathcal{C}^-$ \IGNORE{($\mathcal{C}=\mathcal{C}^-\cup\mathcal{C}^+$)}, such that $\mathcal{C}^-$ is a past stem (or, equivalently, that $\mathcal{C}^+$ is a future stem and $\mathcal{C}^- = \mathcal{C}\setminus\mathcal{C}^+$).
@@
 
-Determining rules for a partition to correspond to a timelike hypersurface in the continuum is extremely difficult and no satisfactory answer has been found yet.
-The situation for a null hypersurface is as follows.
-For any finite density sprinkling into a manifold (that admits a spacelike partitioning surface) with a null hypersurface, $\Sigma$, that null hypersurface will partition the causal set in such a way that agrees with our definition of a spacelike partition.\footnote{This fact means we should really refer to such a partition as ``non-timelike", but as we are dealing with spacelike hypersurfaces for most of this paper we shall continue to call such a partition spacelike}~The reason for this is that a causal set cannot distinguish between a null and spacelike hypersurface.
-To see why this is imagine a finite density sprinkling into a patch of $2$-dimensional Minkowski spacetime.
-Figure~\ref{fig:low_density_sprinkling_with_null_and_spacelike} shows how, given a null surface that partitions the causet into two parts, we can tilt the surface slightly so that it is now spacelike, while still preserving the partition.
+We restrict our discussion to spacelike hypersurfaces, because the definition of the causet counterpart to the boundary action associated with a \emph{timelike} hypersurface is much less straight-forward, so much so that no satisfactory candidate has yet been found.
+The special case for null hypersurfaces deserves some comments.
+First, it appears that while the GHY boundary term is well defined for space-like and time-like boundaries in the continuum, it is an open question what the contribution from a null boundary is~\cite{Gibbons_Hawking_Boundary, Harris}.\mb{copied this from sumati's section before seeing her newest edits, so beware of repetition.} Second, in the discrete setting, the distinction between a null and a spacelike hypersurface loses its meaning to some extent, since any null hypersurface is ``infinitesimally close'' to a spacelike hypersurface.
+Hence a finite causet cannot distinguish a null hypersurface from a spacelike hypersurface.
+To see why this is, imagine a causal set obtained via sprinkling into a patch of $2$-dimensional Minkowski space.
+Figure~\ref{fig:low_density_sprinkling_with_null_and_spacelike} shows how, given a null surface that partitions the causet into two parts, we can tilt the surface ever so slightly to make it spacelike, while still preserving the partition.
 Figure~\ref{fig:high_density_sprinkling_with_null_and_spacelike} shows that even as we increase the density of the sprinkling we can tilt the surface by a smaller amount such that the partition remains unchanged.
-Hence, for any finite density sprinkling one can always find some slight deviation from a partitioning null hypersurface to one that is spacelike, while still preserving the partition.
-As the partition does not change we have no way of knowing whether the hypersurface is null or spacelike.
+A finite causet faithfully embeddable into a manifold with a null hypersurface therefore has no way of knowing whether the hypersurface is null or spacelike.
+With these comments in mind, we return to the spacelike case but we shall return to the question of null boundaries below.
 
@@
 
-In the continuum the GHY term is usually calculated for the future or past bounding surfaces of a manifold, not for a partitioning surface.
-We wish to calculate the GHY term for future or past boundaries of causets, but there are a couple issues we must deal with.
-First, if we can faithfully embed a causet $\mathcal{C}$ into $M$ then the boundary surface of $M$ will induce a trivial partition of $\mathcal{C}$ (every element of $\mathcal{C}$ will lie in the bulk of $M$), thus the previous notion of the boundary term as a rate of change of spacetime atoms across a partition does not work.
-Second, what does it mean to say a bounding surface of a causet if it is not faithfully embeddable in some manifold?
+\mb{rephrase}In the continuum, the GHY term is usually calculated for the future or past \emph{boundary} of a manifold, not for a partitioning hypersurface in it.
+If we wish to calculate the GHY term for future or past boundaries of causets, there are a couple issues we must deal with.
+First, if we can faithfully embed a causet $\mathcal{C}$ into a manifold $M$ with spacelike boundary, then the boundary of $M$ will induce a trivial partition of $\mathcal{C}$ (every element of $\mathcal{C}$ will lie in the bulk of $M$), thus the previous notion of the boundary term as a rate of change of spacetime atoms across a partition does not work.
+Second, it is necessary to to say a bounding surface of a causet if it is not faithfully embeddable in some manifold?
 Previously if the causet was not faithfully embeddable we could still define the causal set analogue of the surface as a partition of the causet, but now we have that either $\mathcal{C}^+$ or $\mathcal{C}^-$ will be the empty set.
@@
 
-For a given causet $\mathcal C$, let us define the function $\F{k}\left[\mathcal{C}\,\right]$, which counts the number of elements in $\mathcal C$ that have exactly $k$ elements in their future, and the function $\P{k}\left[\mathcal{C}\,\right]$, which counts the number of elements which have exactly $k$ elements in their past.
+For a given causet $\mathcal C$, let us define the functions $\F{k}\left[\mathcal{C}\,\right]$, which count the number of elements in $\mathcal C$ that have exactly $k$ elements in their future, and the functions $\P{k}\left[\mathcal{C}\,\right]$, which count the number of elements which have exactly $k$ elements in their past.
 An ``$F_0$ element'' is then a \emph{maximal} element and a ``$P_0$ element'' is a \emph{minimal} element.
@@
 We have used $\vec{p}$ and $\vec{q}$ to stand for infinite dimensional vectors $ (p_0,...,p_m,..)$ and $ (q_0,...,q_n,..)$ respectively, where all the $p_m,\: q_n \in \mathbb{R}$.
-These vectors contain a finite number of non-zero entries, corresponding to the coefficients of the different $\F{m}$ and $\P{n}$ terms.
+These vectors contain a finite number of non-zero entries corresponding to the coefficients of the $\F{m}$ and $\P{n}$ terms.
 We use the summation $\sum_m$ ($\sum_n$) to denote a sum that runs from $0$ to $m_{max}$ ($n_{max}$), where $m_{max}$ ($n_{max}$) is some integer, the value of which depends on the causet $\mathcal{C}^-$ ($\mathcal{C}^+$), for which all $\F{m}\left[\mathcal{C}^- \right]=0$ ($\P{n}\left[\mathcal{C}^+ \right]=0$) for any $m>m_{max}$ ($n>n_{max}$).
-Finite values for $m_{max}$ and $n_{max}$ will always exist for any finite causet.
+Finite values for $m_{max}$ and $n_{max}$ will always exist for a finite causet.
 The vectors must be chosen such that they satisfy the following conditions.

### 2014-10-13 paper_v7.tex
edits in section 1-2

@@
 The process of sprinkling points into $M$ with density $\rho=l^{-d}$ generates a random causet $\mathcal{C}$, and so the functions $\F{k}$ and $\P{k}$ acting on this random causet are now the random variables $\BF{k}$ and $\BP{k}$.
-These random variables can be substituted into~\eqref{general_boundary_sum} for $\F{k}$ and $\P{k}$ to give the family of random variables, $\textbf{S}^{ (d)}_{GHY}$.
-These random variables will be functions of $\rho$ and the two parts of the partitioned manifold, $M^{\pm}$, where the dependence of $\BF{k}$ ($\BP{k}$) on the manifold is restricted to $M^-$ ($M^+$).
+These random variables can be substituted into~\eqref{general_boundary_sum} for $\F{k}$ and $\P{k}$ to give the family of random variables $\textbf{S}^{ (d)}_{GHY}$.
+These random variables will be functions of $\rho$ and of the two parts of the partitioned manifold, $M^{\pm}$, where the dependence of $\BF{k}$ ($\BP{k}$) on the manifold is restricted to $M^-$ ($M^+$).
 We claim that in the limit of infinite density the expectation value, in the sprinkling process, of $\textbf{S}^{ (d)}_{GHY}$ (for any vectors $\vec{p}$ and $\vec{q}$ satisfying~\eqref{coefficient_relation1} and ~\eqref{coefficient_relation1}) tends to the continuum GHY boundary term for a surface $\Sigma$ that partitions the manifold into $M^+$ and $M^-$:
@@
 where $\langle\cdot\rangle$ denotes the mean over sprinklings and we have explicitly stated the arguments of $\textbf{S}^{ (d)}_{GHY}$.
-Technically the right hand side of~\eqref{eq:mainconjecture} is equal to ${S}^{(d)}_{GHY}$ ($-{S}^{(d)}_{GHY}$) evaluated for the manifold $M^-$ ($M^+$) with future (past) bounding surface $\Sigma$.
+Technically, the right hand side of~\eqref{eq:mainconjecture} is equal to ${S}^{(d)}_{GHY}$ when evaluated for the manifold $M^-$ that has $\Sigma$ as a future boundary, but equal to  $-{S}^{(d)}_{GHY}$ when evaluated for the manifold $M^+$ that has $\Sigma$ as a past boundary.
 
@@
 
-For any instance of the sprinkling, the probability that a sprinkled point $p\in M$ below the surface $\Sigma$ is an $F_k$ element is given by the probability that $k$ points of the sprinkling lie in the region $J^{+} (p)\cap J^{-} (\Sigma)$, the intersection of the causal future of $p$ with the causal past of the surface $\Sigma$.\footnote{For notational convenience we use the symbol $x$ to refer to the causal element $x\in\mathcal C$, to its embedding in the manifold $M$, and to its coordinates in some chart on $M$.} This region, for some general spacetime, will be the interior of some curvy $d$-dimensional lightcone truncated by the surface $\Sigma$.
+For any instance of the sprinkling, the probability that a sprinkled point $p\in M$ below the surface $\Sigma$ is an $F_k$ element is given by the probability that $k$ points of the sprinkling lie in the region $J^{+} (p)\cap J^{-} (\Sigma)$, the intersection of the causal future of $p$ with the causal past of the surface $\Sigma$.\footnote{For notational convenience we use the same symbol $x$ to refer to the causal element $x\in\mathcal C$, to its embedding in the manifold $M$, and to its coordinates in some chart on $M$.} This region, for some general spacetime, will be the interior of some curvy $d$-dimensional lightcone truncated by the surface $\Sigma$.
 We will refer to these regions as \emph{truncated cones}.

### 2014-10-13 paper_v7.tex
explained bound on contributions outside nbh of sigma in 3.1

@@
 
-Consider a set of synchronous or Gaussian Normal Coordinates (GNC), $x^\mu= (t,\mathbf x)$, adapted to $\Sigma$ such that in a neighbourhood $U_\Sigma$ of $\Sigma$ the line element is
+Consider a set of synchronous or Gaussian Normal Coordinates (GNC), $x^\mu= (t,\mathbf x)$, adapted to $\Sigma$ such that in neighbourhood $U_\Sigma$ of $\Sigma$ the line element is
 \be
@@
 The integrals~\eqref{eq:nmax} and~\eqref{eq:nmin} seem intractable as they stand, since the integration is over the entire causal past/future of the surface and the volume expressions will be complicated in the presence of curvature.
-However, for large $\rho$ (small $l$), the integrands $\exp (-\rho V)$ will be exponentially suppressed unless the volumes are small.
-Consider the spacetime region $U_\varepsilon=\left\{|t|<\varepsilon\right\}$ around the surface $\Sigma$ for some $\varepsilon>0$ (with $\varepsilon$ small enough such that the GNC system is valid throughout $U_\varepsilon$).
-We will assume that for any such $\varepsilon$, the contribution from points with $|t|>\varepsilon$ to the integrals~\eqref{eq:nmax} and \eqref{eq:nmin} can be made arbitrarily small by setting $\rho$ large enough, since the volumes of the truncated past/future lightcones for such points will be too large.
-Hence, as $\rho\rightarrow\infty$, the integration ranges in \eqref{eq:nmax} and \eqref{eq:nmin} can be cut off at finite time $t=\pm\varepsilon$ with $\varepsilon$ arbitrarily small, which allows us to expand the time integration-variable about zero.
+However, due to $\Sigma$ being closed compact and $M^+$ and $M^-$ of finite volume, we can always find a subneighbourhood of $U_\Sigma$ such that the contribution to the boundary term from points outside of that neighbourhood tends to zero exponentially quickly as $\rho\rightarrow\infty$.\footnote{To see this, let us restrict ourselves to points to the future of $\Sigma$ and hence to the contributions to $\mathbf F_k$.
+There will be some $t_{max}>0$ such that all points $(t,\mathbf x)$ with $|t|<t_{max}$ lie in $U_\Sigma$.
+Choose some $t'<t_{max}$, and consider the subneighbourhood $U'_\Sigma=\left\{(t,\mathbf x)\in U_\Sigma: |t|<t'\right\}$.
+Then given any $\Lambda\in\mathbb R^+$, there exists, for each $\mathbf x$, a value $\rho_\Lambda(\mathbf x)$ such that $V_\blacktriangle(\mathbf x, t)\rho_\Lambda(\mathbf x)>\Lambda$ for all points outside of $U'_\Sigma$.
+Since $\Sigma$ is compact, $\rho_\Lambda(\mathbf x)$ achieves its maximum on $\Sigma$ and we can choose a finite constant value $\rho_\Lambda\equiv\max_{\mathbf{x}}\rho_\Lambda(\mathbf x)$ such that $\rho_\Lambda V(\mathbf x, t)>\Lambda$ for all $x\in M^+\setminus U'_\Sigma$.
+Since $M^+$ has finite volume, the total contribution from those points to $\left<\mathbf P_k\right>$ is then bounded from above by $\rho_\Lambda V(M^+\setminus U'_\Sigma)\frac{\Lambda^k}{k!}e^{-\Lambda}$ for large enough $\Lambda$, a bound that can be made arbitrarily small.}
+Hence, the integration ranges in \eqref{eq:nmax} and \eqref{eq:nmin} can be cut off at finite time $t=\pm\varepsilon$ with $\varepsilon$ arbitrarily small, which allows us to expand the time integration-variable about zero.
 These assumptions of course impose certain regularity conditions on the surface $\Sigma$, but so long as $\Sigma$ is a closed compact manifold and is everywhere spacelike then all we require is that the radius of extrinsic curvature, $R$, satisfies $R\gg l$.

### 2014-10-13 paper_v7.tex
explained bound on contributions outside nbh of sigma in 3.1
added coneplot2.pdf

@@
 However, due to $\Sigma$ being closed compact and $M^+$ and $M^-$ of finite volume, we can always find a subneighbourhood of $U_\Sigma$ such that the contribution to the boundary term from points outside of that neighbourhood tends to zero exponentially quickly as $\rho\rightarrow\infty$.\footnote{To see this, let us restrict ourselves to points to the future of $\Sigma$ and hence to the contributions to $\mathbf F_k$.
-There will be some $t_{max}>0$ such that all points $(t,\mathbf x)$ with $|t|<t_{max}$ lie in $U_\Sigma$.
+In the GNC system, there will be some $t_{max}>0$ such that all points $(t,\mathbf x)$ with $|t|<t_{max}$ lie in $U_\Sigma$.
 Choose some $t'<t_{max}$, and consider the subneighbourhood $U'_\Sigma=\left\{(t,\mathbf x)\in U_\Sigma: |t|<t'\right\}$.
@@
   \centering
-    {\includegraphics[scale=0.5]{coneplot}}
+    {\includegraphics[scale=0.5]{coneplot2}}
      \caption{A $3$-dimensional representation of the region $\mathcal{X}_p$ in RNCs.
@@
 \end{gather}
-Where the last line is only true if the vectors, $\vec{p}$ and $\vec{q}$ satisfy~\eqref{area_coefficient_relation}.
-This concludes the proof of the spatial volume claim,~\eqref{eq:conjecture_for_area}.
+Where the last line is only true if the vectors, $\vec{p}$ and $\vec{q}$ satisfy~\eqref{area_coefficient_relation}. This concludes the proof of the spatial volume claim~\eqref{eq:conjecture_for_area}.
 

### 2014-10-14 paper_v7.tex
more changes

@@
 The integrals~\eqref{eq:nmax} and~\eqref{eq:nmin} seem intractable as they stand, since the integration is over the entire causal past/future of the surface and the volume expressions will be complicated in the presence of curvature.
-However, due to $\Sigma$ being closed compact and $M^+$ and $M^-$ of finite volume, we can always find a subneighbourhood of $U_\Sigma$ such that the contribution to the boundary term from points outside of that neighbourhood tends to zero exponentially quickly as $\rho\rightarrow\infty$.\footnote{To see this, let us restrict ourselves to points to the future of $\Sigma$ and hence to the contributions to $\mathbf F_k$.
+However, due to $\Sigma$ being closed compact and $M^+$ and $M^-$ of finite volume, we can always find a subneighbourhood of $U_\Sigma$ such that the contribution to the boundary term from points outside of that subneighbourhood tends to zero exponentially quickly as $\rho\rightarrow\infty$.\footnote{To see this, let us restrict ourselves to points to the future of $\Sigma$ and hence to the contributions to $\mathbf F_k$.
 In the GNC system, there will be some $t_{max}>0$ such that all points $(t,\mathbf x)$ with $|t|<t_{max}$ lie in $U_\Sigma$.
@@
 Since $\Sigma$ is compact, $\rho_\Lambda(\mathbf x)$ achieves its maximum on $\Sigma$ and we can choose a finite constant value $\rho_\Lambda\equiv\max_{\mathbf{x}}\rho_\Lambda(\mathbf x)$ such that $\rho_\Lambda V(\mathbf x, t)>\Lambda$ for all $x\in M^+\setminus U'_\Sigma$.
-Since $M^+$ has finite volume, the total contribution from those points to $\left<\mathbf P_k\right>$ is then bounded from above by $\rho_\Lambda V(M^+\setminus U'_\Sigma)\frac{\Lambda^k}{k!}e^{-\Lambda}$ for large enough $\Lambda$, a bound that can be made arbitrarily small.}
+Since $M^+$ has finite volume, the total contribution from those points to $\left<\mathbf F_k\right>$ is then bounded from above by $\rho_\Lambda V(M^+\setminus U'_\Sigma)\frac{\Lambda^k}{k!}e^{-\Lambda}$ for large enough $\Lambda$, a bound that can be made arbitrarily small.}
 Hence, the integration ranges in \eqref{eq:nmax} and \eqref{eq:nmin} can be cut off at finite time $t=\pm\varepsilon$ with $\varepsilon$ arbitrarily small, which allows us to expand the time integration-variable about zero.
@@
 \be\label{eq:RNCtotaltrans}
-y^{\overline{\mu}}=A^{\overline{\mu}}_{\;\nu}x^\nu+\frac{1}{2}A^{\overline{\mu}}_{\;\mu}\Gamma^{\mu}_{\;\nu\rho} (p_0)x^\nu x^\rho+O ( (x-x_0)^3).
+y^{\overline{\mu}}=A^{\overline{\mu}}_{\;\nu}(x^\nu -x_0^{\nu})+\frac{1}{2}A^{\overline{\mu}}_{\;\mu}\Gamma^{\mu}_{\;\nu\rho} (p_0)(x^\nu -x_0^{\nu}) (x^\rho -x_0^{\rho})+O ( (x-x_0)^3).
 \ee
-The inverse relation to first order is 
+Where we can see that when $x^{\mu}=x_0^{\mu}$ we have that $y^{\overline{\mu}}=0$, and hence that our coordinates are centred at $p_0$. The inverse relation to first order is
 \be\label{eq:RNCinversetrans}
-x^{\mu}=A^{\mu}_{\;\overline{\nu}}y^{\overline{\nu}}+O (y^2)
+x^{\mu}=x_0^{\mu}+A^{\mu}_{\;\overline{\nu}}y^{\overline{\nu}}+O (y^3)
 \ee
-and one finds that the $A^{\mu}_{\;\overline{\mu}}$ must satisfy
+where the $O(y^2)$ term vanishes due to the fact that $\Gamma^{\overline{\mu}}_{\;\overline{\nu}\overline{\rho}} (p_0)=0$. One finds that the $A^{\mu}_{\;\overline{\mu}}$ must satisfy
 \be\label{eq:RNCeqnforA}
@@
 Let us denote the time-coordinate of $p$ in GNCs by $T=x^0_p=t_p$.
-For the transformation from GNCs to RNCs at $p_0$ one finds that $A^{\overline 0}_{\;0}=1$, $A^{\overline 0}_{\;i}=0$ and $\delta_{\ibar\jbar}=A^i_{\;\ibar}A^j_{\;\jbar}h_{ij} (p_0)$, which in particular implies $y^{\overline{0}}_p=\overline{t}_p=t_p=T$.
+For the transformation from GNCs to RNCs at $p_0$ one finds that $A^{\overline 0}_{\;0}=1$, $A^{\overline 0}_{\;i}=0$ and $\delta_{\ibar\jbar}=A^i_{\;\ibar}A^j_{\;\jbar}h_{ij} (p_0)$, which in particular implies $y^{\overline{0}}_p=\overline{t}_p=T+O(T^3)$.
 
@@
 This is now a simple volume integral but we need to find the boundaries of $\mathcal{X}_p$ to write down the integration limits.
-There are two parts to the boundary of the (solid) truncated cone: the \emph{base}, defined as $B(p):=\Sigma\cap J^-(p)$, and the \emph{top} defined as the set corresponding to the lightcone emanating at $p$ with all points below $\Sigma$ removed, written as $T(p):=\partial J^-(p)\setminus\partial(J^-(p)\cap (J^-(\Sigma)\setminus\Sigma))$.
+There are two parts to the boundary of the (solid) truncated cone: the \emph{base}, defined as $B(p):=\Sigma\cap J^-(p)$, and the \emph{top} defined as the set corresponding to the lightcone emanating at $p$ with all points below $\Sigma$ removed, written as $T(p):=\partial J^-(p)\cap J^+(\Sigma)$.
 Now the boundary of $\mathcal{X}_p$ is simply $\partial\mathcal{X}_p=B(p)\cup T(p)$.
@@
 Following \cite{Khetrapal_Sumati:Causal_Diamond_Volume} it can be shown that the first curvature correction to the lightcone volume comes in at $O (T^{d+2})$ and so can be ignored for our purposes.
-This means that the top can be treated as an effectively flat lightcone in RNCs and thus it corresponds to the set of points satisfying $ (y^{\overline{1}})^2+ (y^{\overline{2}})^2+...+ (y^{\overline{d-1}})^2= T-\overline{t}$ where the coordinates  $y^{\overline{i}}$ run from $0$ up to some maximum at the points where the flat cone meets the surface $\Sigma$.
+This means that the top can be treated as an effectively flat lightcone in RNCs, emanating from the point on the time axis at coordinate $T$, and thus it corresponds to the set of points satisfying $ (y^{\overline{1}})^2+ (y^{\overline{2}})^2+...+ (y^{\overline{d-1}})^2= T-\overline{t}$ where the coordinates  $y^{\overline{i}}$ run from $0$ up to some maximum at the points where the flat cone meets the surface $\Sigma$.
 
@@
 \be\label{eq:BottomSurfaceWithGNC}
-\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij} (p_0)x^i x^j+O ( x^3)
+\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij} (p_0)(x^i-x_0^i) (x^j-x_0^j)+O ( (x-x_0)^3)
 \ee
-The linear part on the right of (\ref{eq:RNCtotaltrans}) vanishes as $A^{\overline{0}}_{\;\mu}x^{\mu}=x^0$ (as $A^{\overline{0}}_{\;i}=0$ and $A^{\overline{0}}_{\;0}=1$) and $x^0=t=0$ for the bottom surface.
+The linear part on the right of (\ref{eq:RNCtotaltrans}) vanishes as $A^{\overline{0}}_{\;\mu}(x^{\mu}-x_0^{\mu})=t$ (as $A^{\overline{0}}_{\;i}=0$ and $A^{\overline{0}}_{\;0}=1$) and $t=0$ for the bottom surface.
 Using the inverse RNC relation (\ref{eq:RNCinversetrans}) one can find the equation for the bottom surface in RNCs:
@@
   \centering
-    {\includegraphics[scale=0.5]{coneplot2}}
+    {\includegraphics[scale=0.5,trim={1cm 5cm 1cm 6cm},clip]{coneplot3}}
      \caption{A $3$-dimensional representation of the region $\mathcal{X}_p$ in RNCs.

### 2014-10-16 paper_v7.tex
changes to start of chapter two and to rnc part

@@
 
-In order to give intuition to the ideas laid out in this paper it will be most convenient to consider spacetimes with boundary obtained by partitioning a manifold (without boundary) into two parts as follows.
 Let $(M,g)$ be a sufficiently well-behaved $d$-dimensional Lorentzian manifold.
-We assume that $(M,g)$ has finite volume and admits a \emph{spacelike partitioning} surface, $\Sigma$, defined as a spacelike closed compact surface in $M$, such that the causal past and future sets, $M^\pm=J^\pm (\Sigma)$ satisfy $M^+\cap M^-=\Sigma$.
-It then follows that $\partial M^\pm = \Sigma$, and further that no points in $M^+\setminus\Sigma$ precede any points in $M^-\setminus\Sigma$.~\!\footnote{If there did exist a $p\in M^+\setminus\Sigma$ and $q\in M^-\setminus\Sigma$ such that $p\preceq q$, then we would have $(q\in J^-(\Sigma) \text{ and } p\preceq q)\implies p\in J^-(\Sigma)\implies p\in M^-$, which would contradict $M^+\cap M^-=\Sigma$.} The Gibbons-Hawking-York boundary term for $M^\pm$ is then given by
+We assume that $(M,g)$ has no boundary, is of finite volume, and admits a \emph{spacelike partitioning} surface, $\Sigma$, defined as a spacelike closed compact surface in $M$, such that the causal past and future sets, $M^\pm=J^\pm (\Sigma)$, satisfy $M^+\cap M^-=\Sigma$.~\!\footnote{We do not require that $M^+\cup M^-=M$.} It then follows that $\partial M^\pm = \Sigma$, and further that no points in $M^+\setminus\Sigma$ precede any points in $M^-\setminus\Sigma$.~\!\footnote{If there did exist a $p\in M^+\setminus\Sigma$ and $q\in M^-\setminus\Sigma$ such that $p\preceq q$, then we would have $(q\in J^-(\Sigma) \text{ and } p\preceq q)\implies p\in J^-(\Sigma)\implies p\in M^-$, which would contradict $M^+\cap M^-=\Sigma$.} The Gibbons-Hawking-York boundary term for $M^\pm$ is then given by
 \be\label{eq:GHYBT_in_continuum}
@@
 \ee
-Where we can see that when $x^{\mu}=x_0^{\mu}$ we have that $y^{\overline{\mu}}=0$, and hence that our coordinates are centred at $p_0$. The inverse relation to first order is
+Where we can see that when $x^{\mu}=x_0^{\mu}$ we have that $y^{\overline{\mu}}=0$, and hence that our coordinates are centred at $p_0$. The inverse coordinate transformation to first order is
 \be\label{eq:RNCinversetrans}
@@
 Following \cite{Khetrapal_Sumati:Causal_Diamond_Volume} it can be shown that the first curvature correction to the lightcone volume comes in at $O (T^{d+2})$ and so can be ignored for our purposes.
-This means that the top can be treated as an effectively flat lightcone in RNCs, emanating from the point on the time axis at coordinate $T$, and thus it corresponds to the set of points satisfying $ (y^{\overline{1}})^2+ (y^{\overline{2}})^2+...+ (y^{\overline{d-1}})^2= T-\overline{t}$ where the coordinates  $y^{\overline{i}}$ run from $0$ up to some maximum at the points where the flat cone meets the surface $\Sigma$.
+This means that the top can be treated effectively as a flat lightcone in RNCs, emanating from the point on the time axis at coordinate $T$, and thus it corresponds to the set of points satisfying $ (y^{\overline{1}})^2+ (y^{\overline{2}})^2+...+ (y^{\overline{d-1}})^2= T-\overline{t}$ where the coordinates  $y^{\overline{i}}$ run from $0$ up to some maximum at the points where the flat cone meets the surface $\Sigma$.
 

### 2014-10-17 paper_v7.tex
changes to wording in RNC part to make the origin of the relations for the A matrix more obvious

@@
 Let us denote the time-coordinate of $p$ in GNCs by $T=x^0_p=t_p$.
-For the transformation from GNCs to RNCs at $p_0$ one finds that $A^{\overline 0}_{\;0}=1$, $A^{\overline 0}_{\;i}=0$ and $\delta_{\ibar\jbar}=A^i_{\;\ibar}A^j_{\;\jbar}h_{ij} (p_0)$, which in particular implies $y^{\overline{0}}_p=\overline{t}_p=T+O(T^3)$.
+Equation~\eqref{eq:RNCMetricTransAtPAndChris} is satisfied if we set $A^{\overline 0}_{\;0}=1$, $A^{\overline 0}_{\;i}=0$ and $\delta_{\ibar\jbar}=A^i_{\;\ibar}A^j_{\;\jbar}h_{ij} (p_0)$.
+These relations, along with~\eqref{eq:RNCtotaltrans}, imply that $y^{\overline{0}}_p=\overline{t}_p=T+O(T^3)$.
 

### 2014-10-22 paper_v7.tex
added some comments made some minor edits

--- 2014-10-14 paper_v7.tex
@@
 We assume that $(M,g)$ has finite volume and admits a \emph{spacelike partitioning} surface, $\Sigma$, defined as a spacelike closed compact surface in $M$, such that the causal past and future sets, $M^\pm=J^\pm (\Sigma)$ satisfy $M^+\cap M^-=\Sigma$.
-It then follows that $\partial M^\pm = \Sigma$, and further that no points in $M^+\setminus\Sigma$ precede any points in $M^-\setminus\Sigma$.~\!\footnote{If there did exist a $p\in M^+\setminus\Sigma$ and $q\in M^-\setminus\Sigma$ such that $p\preceq q$, then we would have $(q\in J^-(\Sigma) \text{ and } p\preceq q)\implies p\in J^-(\Sigma)\implies p\in M^-$, which would contradict $M^+\cap M^-=\Sigma$.} The Gibbons-Hawking-York boundary term for $M^\pm$ is then given by
+It then follows that $\partial M^\pm = \Sigma$, and further that no points in $M^+\setminus\Sigma$ precede any points in $M^-\setminus\Sigma$.~\!\footnote{If there did exist $p\in M^+\setminus\Sigma$ and $q\in M^-\setminus\Sigma$ such that $p\preceq q$, then it would follow that $(q\in J^-(\Sigma) \text{ and } p\preceq q)\implies p\in J^-(\Sigma)\implies p\in M^-$, which contradicts $M^+\cap M^-=\Sigma$.} The Gibbons-Hawking-York boundary term for $M^\pm$ is then given by
 \be\label{eq:GHYBT_in_continuum}
@@
 The integrals~\eqref{eq:nmax} and~\eqref{eq:nmin} seem intractable as they stand, since the integration is over the entire causal past/future of the surface and the volume expressions will be complicated in the presence of curvature.
-However, due to $\Sigma$ being closed compact and $M^+$ and $M^-$ of finite volume, we can always find a subneighbourhood of $U_\Sigma$ such that the contribution to the boundary term from points outside of that subneighbourhood tends to zero exponentially quickly as $\rho\rightarrow\infty$.\footnote{To see this, let us restrict ourselves to points to the future of $\Sigma$ and hence to the contributions to $\mathbf F_k$.
+However, due to $\Sigma$ being closed compact and $M^+$ and $M^-$ of finite volume, we can always find a subneighbourhood $U^{\varepsilon_{}}_\Sigma$ of $U_\Sigma$ such that the contribution to the boundary term from points outside of that neighbourhood tends to zero exponentially quickly as $\rho\rightarrow\infty$.\footnote{To see this, let us restrict ourselves to points to the future of $\Sigma$ and hence to the contributions to $\mathbf F_k$.
 In the GNC system, there will be some $t_{max}>0$ such that all points $(t,\mathbf x)$ with $|t|<t_{max}$ lie in $U_\Sigma$.
-Choose some $t'<t_{max}$, and consider the subneighbourhood $U'_\Sigma=\left\{(t,\mathbf x)\in U_\Sigma: |t|<t'\right\}$.
-Then given any $\Lambda\in\mathbb R^+$, there exists, for each $\mathbf x$, a value $\rho_\Lambda(\mathbf x)$ such that $V_\blacktriangle(\mathbf x, t)\rho_\Lambda(\mathbf x)>\Lambda$ for all points outside of $U'_\Sigma$.
-Since $\Sigma$ is compact, $\rho_\Lambda(\mathbf x)$ achieves its maximum on $\Sigma$ and we can choose a finite constant value $\rho_\Lambda\equiv\max_{\mathbf{x}}\rho_\Lambda(\mathbf x)$ such that $\rho_\Lambda V(\mathbf x, t)>\Lambda$ for all $x\in M^+\setminus U'_\Sigma$.
-Since $M^+$ has finite volume, the total contribution from those points to $\left<\mathbf F_k\right>$ is then bounded from above by $\rho_\Lambda V(M^+\setminus U'_\Sigma)\frac{\Lambda^k}{k!}e^{-\Lambda}$ for large enough $\Lambda$, a bound that can be made arbitrarily small.}
+Choose some $\epsilon<t_{max}$, and consider the subneighbourhood $U^{\varepsilon_{}}_\Sigma=\left\{(t,\mathbf x)\in U_\Sigma: |t|<\epsilon\right\}$.
+Then given any $\Lambda\in\mathbb R^+$, there exists, for each $\mathbf x$, a value $\rho_\Lambda(\mathbf x)$ such that $V_\blacktriangle(\mathbf x, t)\rho_\Lambda(\mathbf x)>\Lambda$ for all points outside of $U^{\varepsilon_{}}_\Sigma$.
+Since $\Sigma$ is compact, $\rho_\Lambda(\mathbf x)$ achieves its maximum on $\Sigma$ and we can choose a finite constant value $\rho_\Lambda\equiv\max_{\mathbf{x}}\rho_\Lambda(\mathbf x)$ such that $\rho_\Lambda V(\mathbf x, t)>\Lambda$ for all $x\in M^+\setminus U^{\varepsilon_{}}_\Sigma$.
+Since $M^+$ has finite volume, the total contribution from the points outside $U^{\varepsilon_{}}_\Sigma$ to $\left<\mathbf P_k\right>$ is then bounded from above by $\rho_\Lambda V(M^+\setminus U^{\varepsilon_{}}_\Sigma)\frac{\Lambda^k}{k!}e^{-\Lambda}$ for large enough $\Lambda$, a bound that can be made arbitrarily small.}
 Hence, the integration ranges in \eqref{eq:nmax} and \eqref{eq:nmin} can be cut off at finite time $t=\pm\varepsilon$ with $\varepsilon$ arbitrarily small, which allows us to expand the time integration-variable about zero.
@@
 The volumes can be expressed by series expansions in $t$.
-We will only need to retain up to order $O(t^{d+1})$ as higher order terms will be shown to vanish when we take the limit of $\rho\rightarrow\infty$ on the mean values of~\eqref{general_boundary_sum} and~\eqref{general_area_sum}.
+We will only need to retain terms up to order $O(t^{d+1})$, as higher order terms will be shown to vanish when we take the limit of $\rho\rightarrow\infty$ on the mean values of~\eqref{general_boundary_sum} and~\eqref{general_area_sum}.
 Henceforth, whenever we state that a term of some order vanishes in the limit we mean that it vanishes in the way just stated.
-The discussion for the two volume integrals is identical so we will only outline it for $V_\blacktriangle (p)$ (i.e.
+The discussion for the two volume integrals is identical, so we will only outline it for $V_\blacktriangle (p)$ (i.e.
 for points to the future of $\Sigma$).
 
-Fix $p\in M$ and denote its coordinate values in GNCs by $x^\mu_p= (t_p,\mathbf x_p)$.
+Fix $p\in M^+$ and denote its coordinate values in GNCs by $x^\mu_p= (t_p,\mathbf x_p)$.
 It will be convenient to use RNCs centered not at the tip, $p$, of the cone but instead at the point $p_0$ where the unique geodesic through $p$ whose tangent is normal to $\Sigma$ intersects $\Sigma$.
@@
 %\mb{let's put all facts about $x^\mu$ and $y^{\overline\mu}$ for $p$ and $p_0$ here and expansion of det in RNC.} 
-We need to assume that for any $p$ in $U_\varepsilon$, the Riemann normal neighbourhood $U_p\subset M$ (throughout which the RNC system centered at $p_0$ is well-defined) contains the truncated cone $J^- (p)\cap J^+ (\Sigma)$.
-This assumption seems reasonable given the that $U_\varepsilon$ can be made arbitrarily ``thin'' as $\rho\rightarrow\infty$.
+We need to assume that for any $p$ in $U^{\varepsilon_{}}_\Sigma$, the Riemann normal neighbourhood throughout which the RNC system centered at $p_0$ is well-defined contains the truncated cone $J^- (p)\cap J^+ (\Sigma)$.
+This assumption seems reasonable given the that $U^{\varepsilon_{}}_\Sigma$ can be made arbitrarily ``thin'' as $\rho\rightarrow\infty$.
 In RNCs the metric and the Christoffel symbols at $p_0$ are those of flat space:
@@
 \be\label{eq:RNCtotaltrans}
-y^{\overline{\mu}}=A^{\overline{\mu}}_{\;\nu}(x^\nu -x_0^{\nu})+\frac{1}{2}A^{\overline{\mu}}_{\;\mu}\Gamma^{\mu}_{\;\nu\rho} (p_0)(x^\nu -x_0^{\nu}) (x^\rho -x_0^{\rho})+O ( (x-x_0)^3).
+y^{\overline{\mu}}=A^{\overline{\mu}}_{\;\nu}x^\nu+\frac{1}{2}A^{\overline{\mu}}_{\;\mu}\Gamma^{\mu}_{\;\nu\rho} (p_0)x^\nu x^\rho+O ( (x-x_0)^3).
 \ee
-Where we can see that when $x^{\mu}=x_0^{\mu}$ we have that $y^{\overline{\mu}}=0$, and hence that our coordinates are centred at $p_0$. The inverse relation to first order is
+The inverse relation to first order is 
 \be\label{eq:RNCinversetrans}
-x^{\mu}=x_0^{\mu}+A^{\mu}_{\;\overline{\nu}}y^{\overline{\nu}}+O (y^3)
+x^{\mu}=A^{\mu}_{\;\overline{\nu}}y^{\overline{\nu}}+O (y^2)
 \ee
-where the $O(y^2)$ term vanishes due to the fact that $\Gamma^{\overline{\mu}}_{\;\overline{\nu}\overline{\rho}} (p_0)=0$. One finds that the $A^{\mu}_{\;\overline{\mu}}$ must satisfy
+and one finds that the $A^{\mu}_{\;\overline{\mu}}$ must satisfy
 \be\label{eq:RNCeqnforA}
-A^{\overline{\mu}}_{\;\mu}A^{\mu}_{\;\overline{\nu}}=\delta^{\overline{\mu}}_{\;\overline{\nu}}\;,\;\;\;\;A^{\mu}_{\;\overline{\mu}}A^{\overline{\mu}}_{\;\nu}=\delta^{\mu}_{\;\nu}
+A^{\overline{\mu}}_{\;\mu}A^{\mu}_{\;\overline{\nu}}=\delta^{\overline{\mu}}_{\;\overline{\nu}}\;,\;\;\;\;A^{\mu}_{\;\overline{\mu}}A^{\overline{\mu}}_{\;\nu}=\delta^{\mu}_{\;\nu}.
 \ee
 Let us denote the time-coordinate of $p$ in GNCs by $T=x^0_p=t_p$.
-For the transformation from GNCs to RNCs at $p_0$ one finds that $A^{\overline 0}_{\;0}=1$, $A^{\overline 0}_{\;i}=0$ and $\delta_{\ibar\jbar}=A^i_{\;\ibar}A^j_{\;\jbar}h_{ij} (p_0)$, which in particular implies $y^{\overline{0}}_p=\overline{t}_p=T+O(T^3)$.
+For the transformation from GNCs to RNCs at $p_0$ one finds that $A^{\overline 0}_{\;0}=1$, $A^{\overline 0}_{\;i}=0$ and $\delta_{\ibar\jbar}=A^i_{\;\ibar}A^j_{\;\jbar}h_{ij} (p_0)$, which in particular implies $y^{\overline{0}}_p=\overline{t}_p=t_p=T$.
 
@@
 This is now a simple volume integral but we need to find the boundaries of $\mathcal{X}_p$ to write down the integration limits.
-There are two parts to the boundary of the (solid) truncated cone: the \emph{base}, defined as $B(p):=\Sigma\cap J^-(p)$, and the \emph{top} defined as the set corresponding to the lightcone emanating at $p$ with all points below $\Sigma$ removed, written as $T(p):=\partial J^-(p)\cap J^+(\Sigma)$.
+There are two parts to the boundary of the (solid) truncated cone: the \emph{base}, defined as $B(p):=\Sigma\cap J^-(p)$, and the \emph{top} defined as the set corresponding to part of the past lightcone emanating at $p$ that lies to the future of $\Sigma$, $T(p):=\partial J^-(p)\cap J^+(\Sigma)$\mb{should we setminus $\Sigma$ from $T(p)$?}.
 Now the boundary of $\mathcal{X}_p$ is simply $\partial\mathcal{X}_p=B(p)\cup T(p)$.
@@
 Following \cite{Khetrapal_Sumati:Causal_Diamond_Volume} it can be shown that the first curvature correction to the lightcone volume comes in at $O (T^{d+2})$ and so can be ignored for our purposes.
-This means that the top can be treated as an effectively flat lightcone in RNCs, emanating from the point on the time axis at coordinate $T$, and thus it corresponds to the set of points satisfying $ (y^{\overline{1}})^2+ (y^{\overline{2}})^2+...+ (y^{\overline{d-1}})^2= T-\overline{t}$ where the coordinates  $y^{\overline{i}}$ run from $0$ up to some maximum at the points where the flat cone meets the surface $\Sigma$.
+This means that the top can be treated as an effectively flat lightcone in RNCs and thus it corresponds to the set of points satisfying $ (y^{\overline{1}})^2+ (y^{\overline{2}})^2+...+ (y^{\overline{d-1}})^2= T-\overline{t}$ where the coordinates  $y^{\overline{i}}$ run from $0$ up to some maximum at the points where the flat cone meets the surface $\Sigma$.
 
@@
 \be\label{eq:BottomSurfaceWithGNC}
-\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij} (p_0)(x^i-x_0^i) (x^j-x_0^j)+O ( (x-x_0)^3)
+\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij} (p_0)x^i x^j+O ( x^3)
 \ee
-The linear part on the right of (\ref{eq:RNCtotaltrans}) vanishes as $A^{\overline{0}}_{\;\mu}(x^{\mu}-x_0^{\mu})=t$ (as $A^{\overline{0}}_{\;i}=0$ and $A^{\overline{0}}_{\;0}=1$) and $t=0$ for the bottom surface.
+The linear part on the right of (\ref{eq:RNCtotaltrans}) vanishes as $A^{\overline{0}}_{\;\mu}x^{\mu}=x^0$ (as $A^{\overline{0}}_{\;i}=0$ and $A^{\overline{0}}_{\;0}=1$) and $x^0=t=0$ for the bottom surface.
 Using the inverse RNC relation (\ref{eq:RNCinversetrans}) one can find the equation for the bottom surface in RNCs:
@@
   \centering
-    {\includegraphics[scale=0.5,trim={1cm 5cm 1cm 6cm},clip]{coneplot3}}
+    {\includegraphics[scale=0.5]{coneplot2}}
      \caption{A $3$-dimensional representation of the region $\mathcal{X}_p$ in RNCs.
-The top part of $\partial\mathcal{X}_p$ can be approximated as a flat cone, and the base surface intersects the top part at a radial coordinate, $r_{max} (\boldsymbol\phi)$, which will in general be a function of the angles $\boldsymbol\phi$ (in $3$-dimensions there is one angle $\phi$).
+The top part of $\partial\mathcal{X}_p$ can be approximated as a flat cone, and the base surface intersects the top part at a radial coordinate, $r_{max} (\boldsymbol\phi)$, which will in general be a function of the angles $\boldsymbol\phi$ (in three dimensions there is one angle $\phi$).
 This function is found by projecting down from the intersection to the $\overline{t}$ plane.

### 2014-10-22 paper_v7.tex
changed T(p) and B(p) to T_p and B_p to match \mathcal{X}_p

--- 2014-10-17 paper_v7.tex
@@
 This is now a simple volume integral but we need to find the boundaries of $\mathcal{X}_p$ to write down the integration limits.
-There are two parts to the boundary of the (solid) truncated cone: the \emph{base}, defined as $B(p):=\Sigma\cap J^-(p)$, and the \emph{top} defined as the set corresponding to the lightcone emanating at $p$ with all points below $\Sigma$ removed, written as $T(p):=\partial J^-(p)\cap J^+(\Sigma)$.
-Now the boundary of $\mathcal{X}_p$ is simply $\partial\mathcal{X}_p=B(p)\cup T(p)$.
+There are two parts to the boundary of the (solid) truncated cone: the \emph{base}, defined as $B_p:=\Sigma\cap J^-(p)$, and the \emph{top} defined as the set corresponding to the lightcone emanating at $p$ with all points below $\Sigma$ removed, written as $T_p:=\partial J^-(p)\cap J^+(\Sigma)$.
+Now the boundary of $\mathcal{X}_p$ is simply $\partial\mathcal{X}_p=B_p\cup T_p$.
 First we look at the top.

### 2014-10-22 paper_v7.tex
minor corrections in 3.2

@@
 In GNCs this simply corresponds to the point $p_0$ with coordinates $x_0^\mu=x^\mu (p_0)= (0,\mathbf x_p)$.
-RNCs centered at $p_0$ will be given the symbol $y^{\overline{\mu}}$, or in terms of coordinates as $(y^{\overline{0}},\mathbf{y})=(\overline{t},\mathbf{y})$.
+RNCs centered at $p_0$ will be denoted as $y^{\overline{\mu}}=(y^{\overline{0}},\mathbf{y})=(\overline{t},\mathbf{y})$.
 %\mb{let's put all facts about $x^\mu$ and $y^{\overline\mu}$ for $p$ and $p_0$ here and expansion of det in RNC.} 
-We need to assume that for any $p$ in $U_\varepsilon$, the Riemann normal neighbourhood $U_p\subset M$ (throughout which the RNC system centered at $p_0$ is well-defined) contains the truncated cone $J^- (p)\cap J^+ (\Sigma)$.
+We need to assume that for any $p$ in $U_\varepsilon$, the Riemann normal neighbourhood $U_p\subset M$ throughout which the RNC system centered at $p_0$ is well-defined contains the truncated cone $J^- (p)\cap J^+ (\Sigma)$.
 This assumption seems reasonable given the that $U_\varepsilon$ can be made arbitrarily ``thin'' as $\rho\rightarrow\infty$.
-In RNCs the metric and the Christoffel symbols at $p_0$ are those of flat space:
+RNCs are constructed such that the metric and the Christoffel symbols at $p_0$ are those of flat space:
 \be\label{eq:RNCMetricTransAtPAndChris}
-g_{\overline{\mu} \overline{\nu}} (p_0)=\eta_{\overline{\mu} \overline{\nu}}=A^{\mu}_{\;\overline{\mu}}A^{\nu}_{\;\overline{\nu}}g_{\mu\nu} (p_0)\;,\;\;\;\;\Gamma^{\overline{\mu}}_{\;\overline{\nu}\overline{\rho}} (p_0)=0
+g_{\overline{\mu} \overline{\nu}} (p_0)=\eta_{\overline{\mu} \overline{\nu}}=A^{\mu}_{\;\overline{\mu}}A^{\nu}_{\;\overline{\nu}}g_{\mu\nu} (p_0)\;,\;\;\;\;\Gamma^{\overline{\mu}}_{\;\overline{\nu}\overline{\rho}} (p_0)=0.
 \ee
@@
 \be\label{eq:RNCtotaltrans}
-y^{\overline{\mu}}=A^{\overline{\mu}}_{\;\nu}(x^\nu -x_0^{\nu})+\frac{1}{2}A^{\overline{\mu}}_{\;\mu}\Gamma^{\mu}_{\;\nu\rho} (p_0)(x^\nu -x_0^{\nu}) (x^\rho -x_0^{\rho})+O ( (x-x_0)^3).
+y^{\overline{\mu}}=A^{\overline{\mu}}_{\;\nu}(x^\nu -x_0^{\nu})+\frac{1}{2}A^{\overline{\mu}}_{\;\mu}\Gamma^{\mu}_{\;\nu\rho} (p_0)(x^\nu -x_0^{\nu}) (x^\rho -x_0^{\rho})+O ( (x-x_0)^3),
 \ee
-Where we can see that when $x^{\mu}=x_0^{\mu}$ we have that $y^{\overline{\mu}}=0$, and hence that our coordinates are centred at $p_0$. The inverse coordinate transformation to first order is
+where we can see that when $x^{\mu}=x_0^{\mu}$ we have that $y^{\overline{\mu}}=0$, and hence that our coordinates are centred at $p_0$. The inverse coordinate transformation to first order is
 \be\label{eq:RNCinversetrans}
@@
 \ee
-The linear part on the right of (\ref{eq:RNCtotaltrans}) vanishes as $A^{\overline{0}}_{\;\mu}(x^{\mu}-x_0^{\mu})=t$ (as $A^{\overline{0}}_{\;i}=0$ and $A^{\overline{0}}_{\;0}=1$) and $t=0$ for the bottom surface.
+The linear part on the right hand side of (\ref{eq:RNCtotaltrans}) vanishes, since we have $A^{\overline{0}}_{\;\mu}(x^{\mu}-x_0^{\mu})=t$ (given that $A^{\overline{0}}_{\;i}=0$ and $A^{\overline{0}}_{\;0}=1$) and $t=0$ on the bottom surface.
 Using the inverse RNC relation (\ref{eq:RNCinversetrans}) one can find the equation for the bottom surface in RNCs:
@@
 \ee
-Let us rewrite this equation in spherically symmetric coordinates, i.e.
+Let us rewrite this equation in spherical polar coordinates, i.e.
 define $r:=\sqrt{\delta_{\ibar\jbar}y^\ibar y^\jbar}$ and the usual angular coordinates $\phi_1,..,\phi_{d-2}$ in terms of the spatial coordinates $y^{\overline{1}} = r \cos (\phi_1),\ldots, y^{\overline{d-1}} = r \sin (\phi_1) \cdots \sin (\phi_{d-3}) \sin (\phi_{d-2})$.

### 2014-10-22 paper_v7.tex
minor corrections in 3.2 - under construction

@@
 This assumption seems reasonable given the that $U_\varepsilon$ can be made arbitrarily ``thin'' as $\rho\rightarrow\infty$.
-RNCs are constructed such that the metric and the Christoffel symbols at $p_0$ are those of flat space:
-\be\label{eq:RNCMetricTransAtPAndChris}
-g_{\overline{\mu} \overline{\nu}} (p_0)=\eta_{\overline{\mu} \overline{\nu}}=A^{\mu}_{\;\overline{\mu}}A^{\nu}_{\;\overline{\nu}}g_{\mu\nu} (p_0)\;,\;\;\;\;\Gamma^{\overline{\mu}}_{\;\overline{\nu}\overline{\rho}} (p_0)=0.
-\ee
-The $A^{\mu}_{\;\overline{\mu}}$ govern the coordinate transformation from GNCs to RNCs to linear order. %but $O (x^2)$ corrections may be required. 
-To second order the coordinate transformation is given by
+RNCs are constructed such that the metric and the Christoffel symbols at $p_0$ are those of flat space.
+Following the usual procedure, we define the coordinates $y^{\overline\mu}$ to second order by\mb{sign?}
 \be\label{eq:RNCtotaltrans}
@@
 \ee
-where we can see that when $x^{\mu}=x_0^{\mu}$ we have that $y^{\overline{\mu}}=0$, and hence that our coordinates are centred at $p_0$. The inverse coordinate transformation to first order is
+which satisfy $y^{\overline{\mu}}=0$ when $x^{\mu}=x_0^{\mu}$ and hence that our coordinates are centred at $p_0$
+we obtain
+\be\bea
+g_{\overline{\mu} \overline{\nu}} (p_0)&=A^{\mu}_{\;\overline{\mu}}A^{\nu}_{\;\overline{\nu}}g_{\mu\nu} (p_0)\\\Gamma^{\overline{\mu}}_{\;\overline{\nu}\overline{\rho}} (p_0)&=0
+\eea\ee
+The inverse coordinate transformation to first order is
 \be\label{eq:RNCinversetrans}
@@
 \be\label{eq:RNCeqnforA}
-A^{\overline{\mu}}_{\;\mu}A^{\mu}_{\;\overline{\nu}}=\delta^{\overline{\mu}}_{\;\overline{\nu}}\;,\;\;\;\;A^{\mu}_{\;\overline{\mu}}A^{\overline{\mu}}_{\;\nu}=\delta^{\mu}_{\;\nu}
+A^{\overline{\mu}}_{\;\mu}A^{\mu}_{\;\overline{\nu}}=\delta^{\overline{\mu}}_{\;\overline{\nu}}\;,\;\;\;\;A^{\mu}_{\;\overline{\mu}}A^{\overline{\mu}}_{\;\nu}=\delta^{\mu}_{\;\nu}.
 \ee

### 2014-10-22 paper_v7.tex
minor corrections in 3.2 mostly in the intro of RNCs

@@
 This assumption seems reasonable given the that $U_\varepsilon$ can be made arbitrarily ``thin'' as $\rho\rightarrow\infty$.
-RNCs are constructed such that the metric and the Christoffel symbols at $p_0$ are those of flat space.
-Following the usual procedure, we define the coordinates $y^{\overline\mu}$ to second order by\mb{sign?}
+The RNCs are constructed such that the metric and the Christoffel symbols at $p_0$ are those of flat space.
+Following the usual procedure, we define the coordinates $y^{\overline\mu}$ to second order as\mb{sign?}
 \be\label{eq:RNCtotaltrans}
-y^{\overline{\mu}}=A^{\overline{\mu}}_{\;\nu}(x^\nu -x_0^{\nu})+\frac{1}{2}A^{\overline{\mu}}_{\;\mu}\Gamma^{\mu}_{\;\nu\rho} (p_0)(x^\nu -x_0^{\nu}) (x^\rho -x_0^{\rho})+O ( (x-x_0)^3),
+y^{\overline{\mu}}=A^{\overline{\mu}}_{\;\nu}(x^\nu -x_0^{\nu})+\frac{1}{2}A^{\overline{\mu}}_{\;\mu}\Gamma^{\mu}_{\;\nu\rho} (p_0)(x^\nu -x_0^{\nu}) (x^\rho -x_0^{\rho})+O ( (x-x_0)^3).
 \ee
-which satisfy $y^{\overline{\mu}}=0$ when $x^{\mu}=x_0^{\mu}$ and hence that our coordinates are centred at $p_0$
-we obtain
+These coordinates are correctly centred at $p_0$, satisfying $y^{\overline{\mu}}=0$ when $x^{\mu}=x_0^{\mu}$. If the constant matrix $A^{\overline\mu}_{\;\mu}$ obeys 
+\be\label{eq:RNCMetricTransAtPAndChris}
+A^{\overline\mu}_{\;\mu} A^{\overline\nu}_{\;\nu}\eta_{\overline\mu\overline\nu}=g_{\mu\nu}(p_0)
+\ee
+the metric and Christoffel symbols take the form
 \be\bea
-g_{\overline{\mu} \overline{\nu}} (p_0)&=A^{\mu}_{\;\overline{\mu}}A^{\nu}_{\;\overline{\nu}}g_{\mu\nu} (p_0)\\\Gamma^{\overline{\mu}}_{\;\overline{\nu}\overline{\rho}} (p_0)&=0
+g_{\overline{\mu} \overline{\nu}} (p_0)&=\eta_{\overline\mu\overline\nu}\\\Gamma^{\overline{\mu}}_{\;\overline{\nu}\overline{\rho}} (p_0)&=0.
 \eea\ee
-The inverse coordinate transformation to first order is
+The inverse coordinate transformation is
 \be\label{eq:RNCinversetrans}
-x^{\mu}=x_0^{\mu}+A^{\mu}_{\;\overline{\nu}}y^{\overline{\nu}}+O (y^3)
+x^{\mu}=x_0^{\mu}+A^{\mu}_{\;\overline{\mu}}y^{\overline{\mu}}+O (y^3)
 \ee
-where the $O(y^2)$ term vanishes due to the fact that $\Gamma^{\overline{\mu}}_{\;\overline{\nu}\overline{\rho}} (p_0)=0$. One finds that the $A^{\mu}_{\;\overline{\mu}}$ must satisfy
-\be\label{eq:RNCeqnforA}
-A^{\overline{\mu}}_{\;\mu}A^{\mu}_{\;\overline{\nu}}=\delta^{\overline{\mu}}_{\;\overline{\nu}}\;,\;\;\;\;A^{\mu}_{\;\overline{\mu}}A^{\overline{\mu}}_{\;\nu}=\delta^{\mu}_{\;\nu}.
-\ee
+where $A^{\mu}_{\;\overline{\mu}}$ is the inverse matrix of $A^{\overline\mu}_{\;{\mu}}$, i.e.
+$A^{\overline{\mu}}_{\;\mu}A^{\mu}_{\;\overline{\nu}}=\delta^{\overline{\mu}}_{\;\overline{\nu}}$ and $A^{\mu}_{\;\overline{\mu}}A^{\overline{\mu}}_{\;\nu}=\delta^{\mu}_{\;\nu}$.
+There is no $O(y^2)$ term in~\eqref{eq:RNCinversetrans} due to the fact that $\Gamma^{\overline{\mu}}_{\;\overline{\nu}\overline{\rho}} (p_0)=0$.
 Let us denote the time-coordinate of $p$ in GNCs by $T=x^0_p=t_p$.
-Equation~\eqref{eq:RNCMetricTransAtPAndChris} is satisfied if we set $A^{\overline 0}_{\;0}=1$, $A^{\overline 0}_{\;i}=0$ and $\delta_{\ibar\jbar}=A^i_{\;\ibar}A^j_{\;\jbar}h_{ij} (p_0)$.
+Equation~\eqref{eq:RNCMetricTransAtPAndChris} is then satisfied if we set $A^{\overline 0}_{\;0}=1$, $A^{\overline 0}_{\;i}=0$ and $A^{\ibar}_{\;i}A^{\jbar}_{\;j}\delta_{ij} =h_{ij}(p_0)\iff\delta_{\ibar\jbar}=A^i_{\;\ibar}A^j_{\;\jbar}h_{ij} (p_0)$.
 These relations, along with~\eqref{eq:RNCtotaltrans}, imply that $y^{\overline{0}}_p=\overline{t}_p=T+O(T^3)$.

### 2014-10-22 paper_v7.tex
minor corrections in 3.2 mostly in the intro of RNCs

@@
 $V_d=\pi^{\frac{d}{2}}/\Gamma\left (\frac{d}{2}+1\right)$ denotes the volume of the unit $d$-ball and $\Gamma (t)$ is the gamma function.
-We have used $\vec{p}$ and $\vec{q}$ to stand for infinite dimensional vectors $ (p_0,...,p_m,..)$ and $ (q_0,...,q_n,..)$ respectively, where all the $p_m,\: q_n \in \mathbb{R}$.
+We have used $\vec{p}$ and $\vec{q}$ to stand for infinite dimensional vectors $ (p_0,\ldots,p_m,\ldots)$ and $ (q_0,\ldots,q_n,\ldots)$ respectively, where all the $p_m,\: q_n \in \mathbb{R}$.
 These vectors contain a finite number of non-zero entries corresponding to the coefficients of the $\F{m}$ and $\P{n}$ terms.

### 2014-10-22 paper_v7.tex
minor corrections in 3.2 mostly in the intro of RNCs

@@
 Following \cite{Khetrapal_Sumati:Causal_Diamond_Volume} it can be shown that the first curvature correction to the lightcone volume comes in at $O (T^{d+2})$ and so can be ignored for our purposes.
-This means that the top can be treated effectively as a flat lightcone in RNCs, emanating from the point on the time axis at coordinate $T$, and thus it corresponds to the set of points satisfying $ (y^{\overline{1}})^2+ (y^{\overline{2}})^2+...+ (y^{\overline{d-1}})^2= T-\overline{t}$ where the coordinates  $y^{\overline{i}}$ run from $0$ up to some maximum at the points where the flat cone meets the surface $\Sigma$.
+This means that the top can be treated effectively as a flat lightcone in RNCs, emanating from the point on the time axis at coordinate $T$, and thus it corresponds to the set of points satisfying $\sum_{i=1}^{d-1}\left(y^\ibar\right)^2= T-\overline{t}$ where the coordinates  $y^{\overline{i}}$ run from $0$ up to some maximum at the points where the flat cone meets the surface $\Sigma$.
 

### 2014-10-22 paper_v7.tex
minor corrections in 3.2 mostly in the intro of RNCs

@@
 Following \cite{Khetrapal_Sumati:Causal_Diamond_Volume} it can be shown that the first curvature correction to the lightcone volume comes in at $O (T^{d+2})$ and so can be ignored for our purposes.
-This means that the top can be treated effectively as a flat lightcone in RNCs, emanating from the point on the time axis at coordinate $T$, and thus it corresponds to the set of points satisfying $\sum_{i=1}^{d-1}\left(y^\ibar\right)^2= T-\overline{t}$ where the coordinates  $y^{\overline{i}}$ run from $0$ up to some maximum at the points where the flat cone meets the surface $\Sigma$.
+This means that the top can be treated effectively as a flat lightcone in RNCs, emanating from the point on the time axis at coordinate $T$, and thus it corresponds to the set of points satisfying $\sum_{\emph\i=1}^{d-1}\left(y^\ibar\right)^2= T-\overline{t}$ where the coordinates  $y^{\overline{i}}$ run from $0$ up to some maximum at the points where the flat cone meets the surface $\Sigma$.
 

### 2014-10-22 paper_v7.tex
corrected typo

--- 2014-10-22 paper_v7.tex
@@
 Following \cite{Khetrapal_Sumati:Causal_Diamond_Volume} it can be shown that the first curvature correction to the lightcone volume comes in at $O (T^{d+2})$ and so can be ignored for our purposes.
-This means that the top can be treated effectively as a flat lightcone in RNCs, emanating from the point on the time axis at coordinate $T$, and thus it corresponds to the set of points satisfying $\sum_{i=1}^{d-1}\left(y^\ibar\right)^2= T-\overline{t}$ where the coordinates  $y^{\overline{i}}$ run from $0$ up to some maximum at the points where the flat cone meets the surface $\Sigma$.
+This means that the top can be treated effectively as a flat lightcone in RNCs, emanating from the point on the time axis at coordinate $T$, and thus it corresponds to the set of points satisfying $\sum_{\emph\i=1}^{d-1}\left(y^\ibar\right)^2= T-\overline{t}$ where the coordinates  $y^{\ibar}$ run from $0$ up to some maximum at the points where the flat cone meets the surface $\Sigma$.
 

### 2014-10-28 paper_v7.tex
put in sumatis changes to RNC stuff and components of A matrix

@@
 Let us denote the time-coordinate of $p$ in GNCs by $T=x^0_p=t_p$.
-Equation~\eqref{eq:RNCMetricTransAtPAndChris} is then satisfied if we set $A^{\overline 0}_{\;0}=1$, $A^{\overline 0}_{\;i}=0$ and $A^{\ibar}_{\;i}A^{\jbar}_{\;j}\delta_{ij} =h_{ij}(p_0)\iff\delta_{\ibar\jbar}=A^i_{\;\ibar}A^j_{\;\jbar}h_{ij} (p_0)$.
-These relations, along with~\eqref{eq:RNCtotaltrans}, imply that $y^{\overline{0}}_p=\overline{t}_p=T+O(T^3)$.
+From \cite{Khetrapal_Sumati:Causal_Diamond_Volume} we have that the tip of the cone in RNC's is $y^{\overline{0}}_p=\overline{t}_p=T+O(T^3)$, which, along with~\eqref{eq:RNCMetricTransAtPAndChris}, sets the components of $A^{\overline{\mu}}_{\mu}$ to be $A^{\overline 0}_{\;0}=1$, $A^{\overline 0}_{\;i}=0$ and $A^{\ibar}_{\;i}A^{\jbar}_{\;j}\delta_{ij} =h_{ij}(p_0)\iff\delta_{\ibar\jbar}=A^i_{\;\ibar}A^j_{\;\jbar}h_{ij} (p_0)$.
 
@@
 
-The base of the cone in GNCs is simply a part of the surface $t=0$, so we can use (\ref{eq:RNCtotaltrans}) to find the equation for the surface in RNCs. Equation (\ref{eq:RNCtotaltrans}) gives
+The base of the cone in GNCs is simply a part of the surface $t=0$, so we can use (\ref{eq:RNCtotaltrans}) to find the equation for the surface in RNCs. This gives
 \be\label{eq:BottomSurfaceWithGNC}
